### 2021-06-12 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

\documentclass[10pt,twocolumn]{IEEEtran}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\usepackage{amsmath,graphicx,amsfonts,amssymb,amsthm,epsfig,mathrsfs,balance}
\usepackage{subcaption,tikz,color,multirow,algorithm,algpseudocode,algorithmicx}
\usepackage{cite,url,framed,bm,dsfont,lipsum}

\setlength{\tabcolsep}{1.1pt}

\newtheorem{assumption}{Assumption}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{property}{Property}
\newtheorem{example}{Example}

\newcommand{\differential}{{\rm{d}}}

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\newcommand{\red}{\color{red}}

\begin{document}

\title{\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions}

\author{Abhishek~Halder,~\IEEEmembership{Member,~IEEE,} Authors}
%\author{Abhishek~Halder,~\IEEEmembership{Member,~IEEE,}
%Xinbo~Geng,~\IEEEmembership{Member,~IEEE,}
%        Kenneth~F. Caluya,
%        and~Pegah~Ojaghi% <-this % stops a space
%\thanks{A. Halder and K.F. Caluya are with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder,kcaluya\}@ucsc.edu}}.}% <-this % stops a space
%\thanks{X. Halder is with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder\}@ucsc.edu}}. This research was partially supported by the NSF award 1923278.}}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
% The paper headers
%\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%

\maketitle

\begin{abstract}
We present a proximal algorithm that performs a variational recursion on the space of joint probability measures to propagate the stochastic uncertainties in power system dynamics over high dimensional state space. The proposed algorithm takes advantage of the exact nonlinearity structures in the trajectory-level dynamics of the networked power systems, and is nonparametric. Lifting the dynamics to the space of probability measures allows us to design a scalable algorithm that obviates gridding the underlying high dimensional state space or function approximation. The proximal recursion implements a generalized gradient flow, and evolves probability weighted scattered point clouds. We provide the theoretical details, convergence guarantees, and numerical examples on realistic test systems.
\end{abstract}

\begin{IEEEkeywords}
Uncertainty propagation, power system dynamics, optimal transport, proximal operator.
\end{IEEEkeywords}

\section{Introduction}\label{sec:intro}
%\IEEEPARstart{T}{he} networked dynamics of electrical generators and loads include different types of uncertainties: intermittencies in renewable generation, load level random fluctuations, parametric uncertainties in transmission lines, uncertainties in weather, unexpected outages, noise in phasor measurement unit sensing data, among others. While significant progress has been made in statistically quantifying these individual sources of uncertainties, it remains challenging \cite{schwalbe2015mathematical} to accurately predict the joint effect of these uncertainties on the safe and robust operation of the power system over time.

%The study of power system dynamics with uncertainties is not new. However, 
\IEEEPARstart{S}{tochastic} variabilities in power grid have increased significantly in recent years both in the generation side (e.g., due to growing penetration of renewables) as well as in the load side (e.g., due to widespread adoption of plug-in electric vehicles). Several studies \cite{timko1983monte,nwankpa1992stochastic,odun2012structure,ghanavati2016identifying,apostolopoulou2016assessment} have reported that even small stochastic effects can significantly alter the assessment of transient stability, or the performance of automatic generation control. However, the lack of a rigorous yet scalable stochastic computational framework continues to impede \cite{schwalbe2015mathematical} our ability to perform transient analysis involving time varying joint probability density functions (PDFs) over the states of a large power system network. In this paper, we present a new algorithm to address this computational need.

\begin{figure}
\centering
\includegraphics[width=\linewidth]{IEEE14BusWithUnc.pdf}
\caption{\small{A schematic of the IEEE 14 bus test system with stochastic uncertainties. The Uncertainty sources may include stochastic forcing and parametric uncertainties at some generators, random variabilities at some loads, and parametric uncertainties along some transmission lines. For depiction purposes, we indicated the parametric uncertainties as PDFs, and stochastic forcing as intermittent signals.}}
\label{fig:Motivating}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{UncPropBlockDiagm.pdf}
\caption{\small{Block diagram for joint state PDF propagation.}}
\label{fig:UncertaintyPropagation}
\end{figure}

Given a networked power system, one can envisage at least three types of uncertainties affecting the dynamics: initial condition uncertainties in the state variables (e.g., rotor phase angles and angular velocities), parametric uncertainties (e.g., inertia and damping coefficients of the generators, reactance associated with different transmission lines), and stochastic forcing (e.g., intermittencies in renewable power generation, load and ambient temperature fluctuations). Fig. \ref{fig:Motivating} depicts a representative scenario. In addition, one could consider uncertainties due to random change in transmission topology resulting from unexpected outage, and uncertainties due to unmodeled dynamics. Given a statistical description of these uncertainties, our approach is to directly solve the \emph{macroscopic flow} of the joint PDFs governing the probabilistic evolution of the state as summarized in Fig. \ref{fig:UncertaintyPropagation}.

\subsubsection{Related works}
Even though the need for quantifying uncertainties in power systems simulations has been long-recognized \cite{allan8791,allan9296}, early studies were limited to statistical reliability assessment. Dynamic simulations with stochastic uncertainties for purposes such as transient stability analysis have been investigated via \emph{Monte Carlo simulations} \cite{timko1983monte,dong2012numerical,perninge2012importance,odun2012structure}.

 As is well known, the Monte Carlo techniques are easy to apply but the computational cost scales exponentially with the number of dimensions, thus making it prohibitive for realistic power systems dynamic simulation. As an alternative, \emph{probabilistic small signal analysis} \cite{nwankpa1992stochastic,rueda2009assessment,huang2013quasi,dhople2013analysis} have appeared in the power systems literature, albeit at the expense of the additional assumption that the random perturbations remain ``small". \emph{Polynomial chaos and related stochastic collocation methods} \cite{hockenberry2004evaluation,xu2019propagating} can do away with the ``small stochastic perturbation" assumption but due to the finite-dimensional approximation of the probability space, computational performance degrades if the long-term statistics are desired. Furthermore, to cope with the stochasticity, these techniques require simulating a higher-dimensional system than the dimension of the physical state space, which further limits the scalability for nonlinear simulation. More recently, approximation methods such as \emph{stochastic averaging} \cite{ju2018analytical,ju2018stochastic} based on certain energy function \cite{pai1989energy,chang1995direct,sauerpai1998} have appeared. In \cite{maldonado2018uncertainty}, an algorithm for \emph{propagating first few statistical moments} was proposed. However, this leads to moment closure problems since the dimensions of the time-varying sufficient statistics associated with the corresponding transient joint state PDFs are not known in general.

In a different vein, \emph{deterministic bounded uncertainty models} for power flow simulations have been used \cite{dimitrovski2004boundary,hiskens2006sensitivity,chen2012method,althoff2014formal} for set-valued analysis. These, however, require approximating the underlying nonlinear differential algebraic equations (DAEs) appearing in power system dynamic simulation, and thus lead to conservative analysis. For example, the method in \cite{althoff2014formal} requires converting the nonlinear DAEs to linear DAEs in such a way that guarantees set-valued over-approximation of the reachable sets. Likewise, the convex optimization-based bounded uncertainty propagation methods in \cite{choi2017propagating,choi2018propagating} require second order approximation of the power flow state variables as function of the uncertainties.

\subsubsection{Technical challenges}\label{subsubsec:technicalchallenges}
Several technical challenges need to be overcome to achieve scalable computation enabling the prediction of the joint PDFs over a time horizon of interest. \emph{First}, the trajectory level dynamic models for power systems are inherently nonlinear, which do not preserve Gaussianity, thereby requiring nonparametric prediction of the joint PDFs. \emph{Second}, the joint PDFs for realistic power systems dynamic simulation must evolve over a high dimensional state space, i.e., the joint PDF at any given time has high dimensional support. This necessitates spatial discretization-free algorithms since standard function approximation or interpolation approaches would, in general, be met with ``curse-of-dimensionality" \cite{bellman1957}. The numerical challenges aside, one cannot theoretically guarantee to find a suite of basis functions for the manifold of nonparametric PDFs. \emph{Third}, prediction based on first few statistical moments is challenging since it is not possible to \emph{a priori} guarantee a fixed or even finite dimensional sufficient statistic. For instance, propagating only the mean and covariance could be misleading when the underlying joint PDFs are multi-modal.

\subsubsection{Contributions of this paper}
Our contributions are twofold. {\red{TBD}}

\subsubsection{Notation and organization}
We use boldfaced small letters for vectors, and boldfaced capital letters for matrices. The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$. Thus, $\nabla_{\bm{x}}\cdot$ stands for the divergence, and $\Delta_{\bm{x}}$ stands for the Laplacian w.r.t. vector $\bm{x}$. We use $\langle\cdot,\cdot\rangle$ to denote the standard Euclidean inner product. The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively. The symbol $|\cdot|$ denotes the absolute value, $\otimes$ denotes the Kronecker product. The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.

The rest of this paper is structured as follows. {\red{TBD}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Models}\label{sec:models}

\subsection{Sample Path Dynamics}\label{subsec:samplepathdynamics}
We consider the coupled stochastic differential equations (SDEs) associated with the networked-reduced power systems model \cite[Ch. 7]{sauerpai1998}. Specifically, for a power network with $n$ generators, the stochastic dynamics for the $i$-th generator is given by the It\^{o} SDEs
\begin{subequations}
\begin{align}
&{\mathrm{d}}\theta_{i} = \omega_{i}\:{\mathrm{d}}t,\label{RotAngleGen}\\
&m_{i}\:{\mathrm{d}}\omega_{i} =	\left(\!P_{i} - \gamma_{i}\omega_{i} - \displaystyle\sum_{j=1}^{n} k_{ij}\sin\left(\theta_{i}-\theta_{j}-\varphi_{ij}\right)\!\right){\mathrm{d}}t \nonumber\\
&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad+\sigma_{i}\:{\mathrm{d}}w_{i},
\end{align}
\label{ItoSDEcomponentlevel}
\end{subequations}
where the state variables are the rotor angles $\theta_{i}\in[0,2\pi)$ and the rotor angular velocities $\omega_{i}\in\mathbb{R}$, for $i\in\{1,\hdots,n\}$. The stochastic forcing is modeled through the standard Wiener process $w_{i}(t)$, and the diffusion coefficient $\sigma_{i}>0$ denotes the intensity of stochastic forcing at the $i$-th generator.

With the $i$-th generator, we associate its inertia $m_{i}>0$ and damping coefficient $\gamma_{i}>0$. The other parameters: the \emph{effective power input} $P_{i}$, the \emph{phase shift} $\varphi_{ij}\in[0,\frac{\pi}{2})$, and the \emph{coupling coefficients} $k_{ij}\geq 0$, depend on the network reduced admittance matrix $\bm{Y}\equiv[Y_{ij}]_{i,j=1}^{n}\in\mathbb{C}^{n\times n}$. Specifically,
\begin{subequations}
\begin{align}
 P_{i} &= P_{i}^{\text{mech}} - E_{i}^{2} \Re\left(Y_{ii}\right), \\
 \varphi_{ij} &= \begin{cases} -\arctan\left(\dfrac{\Re\left(Y_{ij}\right)}{\Im\left(Y_{ij}\right)}\right), & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}\\
 k_{ij} &= \begin{cases} E_{i}E_{j} \vert Y_{ij}\vert, & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}
\end{align}
\end{subequations}
where $P_{i}^{\text{mech}}$ is the mechanical power input, and $E_{i}$ is the internal voltage (magnitude) for generator $i$.

% The coupling coefficients $k_{ij}$, roughly speaking, denote the maximum power transferred between generators $i$ and $j$ with $k_{ii}=0$ for $i\in\{1,\hdots,n\}$. In general, $k_{ij}$ is a function of the internal voltages at the $i$-th and $j$-th generator, and the magnitude of the $(i,j)$-th entry of the network reduced admittance matrix. We suppose that the elements $k_{ij}$ comprise a coupling matrix $\bm{K}$ of appropriate dimension.

One can view (\ref{ItoSDEcomponentlevel}) as the noisy version of the second order nonuniform Kuramoto oscillator model \cite{dorfler2012synchronization,rodrigues2016kuramoto}, also known as the structure preserving power network model \cite{odun2012structure}, given by
\begin{align}
m_{i}\ddot{\theta}_{i} + \gamma_{i}\dot{\theta}_{i} &= P_{i} - \displaystyle\sum_{j=1}^{n} k_{ij}\sin\left(\theta_{i} - \theta_{j} - \varphi_{ij}\right) \nonumber\\
&\qquad\qquad+ \sigma_{i} \times \text{stochastic forcing},
\label{LangevinForm}	
\end{align}
where the stochastic forcing is standard Gaussian white noise.

We define the positive diagonal matrices
\begin{align*}
\bm{M} &:= {\rm{diag}}\left(m_{1},\hdots,m_{n}\right), \nonumber\\
\bm{\Gamma} &:= {\rm{diag}}\left(\gamma_{1},\hdots,\gamma_{n}\right),\nonumber\\
\bm{\Sigma} &:= {\rm{diag}}\left(\sigma_{1},\hdots,\sigma_{n}\right),
%\label{PosDiagMatrices}	
\end{align*}
and rewrite (\ref{ItoSDEcomponentlevel}) as a mixed conservative-dissipative SDE in state vector $\bm{x} := (\bm{\theta},\bm{\omega})^{\top} \in \mathbb{T}^{n} \times \mathbb{R}^{n}$ as
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\theta}\\
{\mathrm{d}}\bm{\omega}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\omega}\\
-\bm{M}^{-1}\nabla_{\bm{\theta}}V(\bm{\theta}) -\bm{M}^{-1}\bm{\Gamma}\bm{\omega}  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{M}^{-1}\bm{\Sigma}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{ItoSDEvectorlevel}
\end{align}
where 
%$\odot$ denotes element-wise product, $\oslash$ denotes element-wise division, the , the rotor inertia vector $\bm{m} = (m_{1},\hdots,m_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the rotor damping coefficient vector $\bm{\gamma} = (\gamma_{1}, \hdots, \gamma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, the diffusion coefficient vector $\bm{\sigma} = (\sigma_{1}, \hdots, \sigma_{n})^{\top}\in\mathbb{R}^{n}_{>0}$, and 
$\bm{w}\in\mathbb{R}^{n}$ is the standard vector Wiener process, $\mathbb{T}^{n}$ denotes the $n$-torus $[0,2\pi)^{n}$, and the potential function $V : \mathbb{T}^{n} \mapsto \mathbb{R}$ is given by
\begin{align}
V(\bm{\theta}) := \displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{(i,j)\in\mathcal{E}}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right),
\label{potential}	
\end{align}
wherein $\mathcal{E}$ is the set of edges for the underlying graph modeling the power system network. The potential (\ref{potential}) has a natural energy function interpretation and can also be motivated by a mechanical mass-spring-damper analogy \cite{dorfler2013synchronization,ishizaki2018}. The function $V$ is continuously differentiable, bounded below, and goes to $+\infty$ as $\|\bm{x}\|_{2}\rightarrow +\infty$.
%In the development below, an important role will be played by the ``Hamiltonian-like" function
%\begin{align}
%H(\bm{x})\equiv H(\bm{\theta},\bm{\omega}) := \frac{1}{2}\bm{\omega}^{\top}{\rm{diag}}\left(\bm{\gamma}\oslash\bm{m}\right)\bm{\omega} \: + \: V(\bm{\theta}).
%\label{hamiltonian}	
%\end{align}

\subsection{Macroscopic Dynamics}\label{subsec:macroscopic}
Given the sample path dynamics \eqref{ItoSDEcomponentlevel} or equivalently \eqref{ItoSDEvectorlevel}, a prescribed initial joint state PDF 
\begin{align}
\rho_{0}(\bm{x})\equiv\rho(t=t_{0},\bm{\theta}(t_{0}),\bm{\omega}(t_{0}))
\label{FPKInitCond}	
\end{align} 
denoting initial condition uncertainties at time $t=t_{0}$, and prescribed parametric uncertainties given by the joint parameter PDF $\rho_{\text{param}}$, the uncertainty propagation problem calls for computing the transient joint state PDFs $\rho(t,\bm{x})\equiv\rho(t,\bm{\theta},\bm{\omega})$ for any desired time $t \geq t_{0}$, which is a nonnegative function supported on the state space $\mathbb{T}^{n} \times \mathbb{R}^{n}$ satisfying $\int \rho = 1$ for all $t\geq t_{0}$.

The corresponding macroscopic dynamics governing the flow of the joint state PDF $\rho(t,\bm{\theta},\bm{\omega})$ is given by a kinetic Fokker-Planck or Kolmogorov's forward partial differential equation (PDE)
\begin{align}
\frac{\partial \rho}{\partial t} &= - \langle \bm{\omega},\nabla_{\bm{\theta}}\rho \rangle  + \nabla_{\bm{\omega}} \cdot \left( \rho \left( \bm{M}^{-1}\bm{\Gamma}\bm{\omega} + \bm{M}^{-1}\nabla_{\bm{\theta}}V(\bm{\theta}) \right.\right.\nonumber\\
&\qquad\qquad\qquad \left.\left. +\frac{1}{2} \bm{M}^{-1}\bm{\Sigma}\bm{\Sigma}^{\top}\bm{M}^{-1} \nabla_{\bm{\omega}} \log \rho \right) \right),
\label{KineticFPK}	
\end{align}
subject to the initial condition \eqref{FPKInitCond} and the joint parameter PDF $\rho_{\text{param}}$. A direct numerical solution of this PDE initial value problem using conventional discretization (e.g., finite difference) or function approximation techniques will not be scalable in general, as explained in Sec. \ref{subsubsec:technicalchallenges}. In the next Section, we discuss how a measure-valued variational recursion proposed in our recent works \cite{caluya2019ACC,caluya2019TAC,halder2020hopfield,caluya2021TAC} can be employed to address this challenge.

We mention here that (\ref{ItoSDEvectorlevel}) has been used in \cite{ju2018analytical,ju2018stochastic} for uncertainty propagation via stochastic averaging approximation where the univariate energy PDF was proposed as a ``proxy" for the entire joint PDF. Most relevant to our approach in the power systems literature is the work in \cite{wang2013fokker}, which indeed voiced the need for computing the transient joint PDFs but only dealt with the single-machine-infinite-bus case -- simplest instance of (\ref{ItoSDEvectorlevel}). The resulting bivariate Fokker-Planck PDE in \cite{wang2013fokker} was solved via finite element discretization, and revealed rich stochastic dynamics and nontrivial transient stability aspects even in this simple case. However, it is unreasonable to expect that a finite element discretization, or in fact any spatial discretization scheme to solve (\ref{KineticFPK}) for moderately large $n$ in minutes of computational time, thereby limiting our current ability for realistic power systems simulation with stochastic variability. This calls for fundamentally re-thinking what does it mean to solve the PDE (\ref{KineticFPK}) for dynamics (\ref{ItoSDEvectorlevel}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Measure-valued Proximal Recursion}\label{sec:prox}
\subsection{Generalized Gradient Descent}\label{subsec:GenGradDescent}
Let $\mathcal{P}_{2}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right)$ denote the manifold of all joint PDFs supported over the state space $\mathbb{T}^{n} \times \mathbb{R}^{n}$, with finite second raw moments. Symbolically,
\begin{align}
\mathcal{P}_{2}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right) := &\bigg\{\rho : \mathbb{T}^{n} \times \mathbb{R}^{n} \mapsto \mathbb{R}_{\geq 0} \mid \int \rho = 1,\nonumber\\
&\hspace*{-0.8in}\int\bm{x}^{\top}\bm{x}\:\rho(\bm{x})\:\differential\bm{x} <\infty\; \text{for all}\;\bm{x}\equiv(\bm{\theta},\bm{\omega})^{\top}\in\mathbb{T}^{n} \times \mathbb{R}^{n} \bigg\}.
\label{DefP2}	
\end{align}
We propose to solve the initial value problem for the PDE \eqref{KineticFPK} by viewing its flow $\rho(t,\bm{\theta},\bm{\omega})$ as the gradient descent of some functional $\Phi:\mathcal{P}_{2}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right)\mapsto \mathbb{R}_{\geq 0}$ w.r.t. some distance 
\[{\rm{dist}}:\mathcal{P}_{2}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right)\times\mathcal{P}_{2}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right)\mapsto\mathbb{R}_{\geq 0}.\]
We now explain this idea in detail.

For $k\in\mathbb{N}$, and for some chosen step size $h>0$, we discretize time as $t_{k} := kh$, and define the infinite dimensional proximal operator of the functional $h\Phi$ w.r.t. the distance ${\rm{dist}}(\cdot,\cdot)$, given by
\begin{eqnarray}
{\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}) := \underset{\varrho\in\mathcal{P}_{2}} {{\rm{arg\:inf}}}\: \frac{1}{2} {\rm{dist}}^{2}\left(\varrho,\varrho_{k-1}\right) + h\: \Phi(\varrho).
\label{InfiniteDimProxOpDef}
\end{eqnarray}
Now consider a proximal recursion over the manifold $\mathcal{P}_{2}$ as
\begin{eqnarray}
\varrho_{k} = {\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}), \quad k\in\mathbb{N}, \quad \varrho_{0}(\bm{x}) := \rho_{0}(\bm{x}).
\label{ProxRecursionInfiniteDim}	
\end{eqnarray}
Given the PDE \eqref{KineticFPK}, we would like to design the functional pair $\left(\Phi,{\rm{dist}}\right)$ such that the sequence of functions $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by the proximal recursion \eqref{ProxRecursionInfiniteDim}, in the small time step limit, converges to the flow $\rho(t=kh,\bm{\theta},\bm{\omega})$ generated by the PDE initial value problem of interest. In particular,
\begin{align}
\varrho_{k}(\bm{\theta},\bm{\omega}) \xrightarrow{h\downarrow 0} \rho(t=kh,\bm{\theta},\bm{\omega}) \;\text{in}\;L^{1}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right).
\label{ConvergenceGuarantee}	
\end{align}
We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in therms of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous.

Notice that the proximal recursions given by \eqref{InfiniteDimProxOpDef}-\eqref{ProxRecursionInfiniteDim} define an infinite dimensional gradient descent of the functional $h\Phi$ over $\mathcal{P}_{2}$ w.r.t. the distance ${\rm{dist}}$. This is reminiscent of the finite dimensional gradient descent, where a gradient flow generated by an ordinary differential equation initial value problem can be recovered as the small time step limit of the sequence of vectors generated by a standard Euclidean proximal recursion; see e.g., \cite[Sec. I]{caluya2019TAC}.

That the flow generated by a Fokker-Planck PDE initial value problem can be recovered from a variational recursion of the form \eqref{ProxRecursionInfiniteDim} was first proposed in \cite{jordan1998variational}, showing that when the drift in the sample path dynamics is a gradient vector field and the diffusion is a scalar multiple of identity matrix, then ${\rm{dist}}(\cdot,\cdot)$ can be taken as the Wasserstein-2 metric arising in the theory of optimal transport \cite{villani2003topics} with $\Phi(\cdot)$ as the free energy functional. In particular, the functional $\Phi$ serves as a Lyapunov functional in the sense $\frac{\differential}{\differential t}\Phi < 0$ along the transient solution of the Fokker-Planck PDE initial value problem. This idea has since been generalized to many other types of PDE initial value problems, see e.g., \cite{ambrosio2008gradient,santambrogio2017euclidean}.

The algorithmic appeal of the proximal recursion \eqref{ProxRecursionInfiniteDim} is that it opens up the possibility to compute the solution of the PDE initial value problem via recursive convex minimization. A point cloud-based proximal algorithm was proposed in \cite{caluya2019ACC,caluya2019TAC} which was reported to have very fast runtime. Notice that even though the drift in \eqref{ItoSDEvectorlevel} is \emph{not} a gradient vector field, the algorithm in \cite[Sec. V.B]{caluya2019TAC} constructed a pair $\left(\Phi,{\rm{dist}}\right)$ such that \eqref{ProxRecursionInfiniteDim} provably approximates the transient solution of the corresponding kinetic Fokker-Planck PDE with guarantee \eqref{ConvergenceGuarantee}. However, that algorithm cannot be applied to \eqref{KineticFPK} as is. The reasons are explained next.

\subsection{Statistical Mechanics Issues}
A new difficulty for our SDE \eqref{ItoSDEvectorlevel} is that we have \emph{anisotropic} degenerate diffusion, i.e., the strengths of the noise acting in the last $n$ components of \eqref{ItoSDEvectorlevel} are nonuniform since $\bm{M}^{-1}\bm{\Sigma}$ is not identity. This complicates the matter because the construction of the functional $\Phi$ is usually motivated via free energy considerations utilizing the structure of the \emph{stationary} PDF $\rho_{\infty}(\bm{\theta},\bm{\omega})$ for \eqref{KineticFPK}. The $\rho_{\infty}$ is, in turn, guaranteed to be a unique \emph{Boltzmann distribution} of the form
\begin{subequations}
\begin{align}
\rho_{\infty}\left(\bm{\theta},\bm{\omega}\right) &= \dfrac{1}{Z}\exp\left(-\beta H\right), \\
H(\bm{\theta},\bm{\omega}) &:= V(\bm{\theta}) + \frac{1}{2}\langle\bm{\omega},\bm{M\omega}\rangle,	
\end{align}	
\label{BoltzmannPDF}
\end{subequations}
where $Z$ is a normalizing constant (known as the ``partition function"), \emph{if and only if} the so-called \emph{Einstein relation} \cite{hernandez1989equilibrium,chen2015fast} holds:
\begin{align}
\bm{\Sigma}\bm{\Sigma}^{\top} = \beta^{-1}\left(\bm{\Gamma} + \bm{\Gamma}^{\top}\right) \quad\text{for some}\;\beta > 0.
\label{EinsteinRelation}	
\end{align}
In our case, $\bm{\Sigma,\Gamma}$ are positive diagonal, and \eqref{EinsteinRelation} requires $\sigma_{i}^{2} \propto \gamma_{i}$ for all $i=1,\hdots,n$.

In the power systems context, we cannot relate the damping coefficients $\gamma_i$ with the squared intensities of stochastic forcing $\sigma_{i}^{2}$ for the generators. Thus, \eqref{EinsteinRelation} will not hold in practice, meaning either we cannot guarantee existence-uniqueness for $\rho_{\infty}$, or even if $\rho_{\infty}$ exists, it will not be of the form \eqref{BoltzmannPDF}. On one hand, this implies that our construction of $\Phi$ may not be guided by free energy considerations. On the other hand, since we are only interested in computing the transient joint PDFs, i.e., \emph{non-equilibrium} statistical mechanics, the lack of a fluctuation-dissipation relation like \eqref{EinsteinRelation} should not be a fundamental impediment in setting up a recursion such as \eqref{ProxRecursionInfiniteDim}. We next show that a simple change of variable can indeed circumvent this issue.

\subsection{From Anisotropic to Isotropic Degenerate Diffusion}\label{subsec:FromAnisoToIso}
Let the $2n\times 2n$ matrix
\begin{align}
\bm{\Psi}:= \bm{I}_{2} \otimes \bm{M}\bm{\Sigma}^{-1}, \label{DefPsi}	
\end{align}
and define the invertible linear map
\begin{align}
\begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix} \mapsto \begin{pmatrix}
\bm{\xi}\\
\bm{\eta}	
\end{pmatrix} := \bm{\Psi} \begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix}. \label{ChangeOfVar}
\end{align}
Applying It\^{o}'s lemma, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\xi}\\
{\mathrm{d}}\bm{\eta}	
\end{pmatrix}
\! = \!\begin{pmatrix}
\bm{\eta}\\
-\nabla_{\bm{\xi}}U(\bm{\xi}) -\nabla_{\bm{\eta}}F\left(\bm{\eta}\right)  	
\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
 \bm{0}_{n\times n}\\
 \bm{I}_{n}	
 \end{pmatrix}{\mathrm{d}}\bm{w},
\label{XiEtaVectorSDE}	
\end{align}
where the potentials
{\small{\begin{subequations}
\begin{align}
U(\bm{\xi}) &:= \!\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\!\! \displaystyle\sum_{(i,j)\in\mathcal{E}}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} + \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right), \label{defU}\\
F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
\end{align}
\label{NewPotentials}	
\end{subequations}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}\label{sec:conclusion}
The conclusion goes here.

%\appendices
%\section{Proof of the First Zonklar Equation}
%Appendix one text goes here.

\begin{figure}
\centering
\includegraphics[width=.8\linewidth]{BlockDiagm.pdf}
\caption{\small{Schematic of the proposed algorithmic setup for propagating the joint state PDF as probability weighted scattered point cloud $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$. The location of the points $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ can be updated by Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\varrho_{k}^{i}\}_{i=1}^{N}$ can be updated via discrete version of the proximal recursion (\ref{ProxRecursionInfiniteDim}).}}
\label{fig:BlockDiagm}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{thebibliography}{99}

\bibitem{schwalbe2015mathematical}
M. Schwalbe, \emph{Mathematical Sciences Research Challenges for the Next-generation Electric Grid: Summary of a Workshop}, National Academies Press, 2015.

% lit review starts here
%========================
\bibitem{allan8791}
R.N. Allan, R. Billinton, A.M. Breipohl, and C.H. Grigg, ``Bibliography On the Application of Probability Methods in Power Systems Reliability Evaluation 1987--1991", \emph{IEEE Transactions on Power Systems}, Vol. 9, pp. 41--49, 1994.

\bibitem{allan9296}
------, ``Bibliography On the Application of Probability Methods in Power Systems Reliability Evaluation 1992--1996", \emph{IEEE Transactions on Power Systems}, Vol. 14, pp. 51--57, 1999.

%==== MC =====

\bibitem{timko1983monte}
K.J. Timko, A. Bose, and P.M. Anderson, ``Monte Carlo Simulation of Power System Stability", 
\emph{IEEE Transactions on Power Apparatus and Systems}, Vol. PAS-102, No. 10, pp. 3453--3459, 1983.

\bibitem{dong2012numerical}
Z.Y. Dong, J.H. Zhao, and D.J. Hill, ``Numerical Simulation for Stochastic Transient Stability Assessment", \emph{IEEE Transactions on Power Systems}, Vol. 27, No. 4, pp. 1741--1749, 2012.

\bibitem{perninge2012importance}
M. Perninge, F. Lindskog, and L. Soder, ``Importance Sampling of Injected Powers for Electric Power System Security Analysis", \emph{IEEE Transactions on Power Systems}, Vol. 27, No. 1, pp. 3--11, 2012.

\bibitem{odun2012structure}
T. Odun-Ayo, and M.L. Crow, ``Structure-preserved Power System Transient Stability using Stochastic Energy Functions", \emph{IEEE Transactions on Power Systems}, Vol. 27, No. 3, pp. 1450--1458, 2012.

% ==== AGC ======

\bibitem{ghanavati2016identifying}
G. Ghanavati, P.D.H. Hines, and T.I. Lakoba, ``Identifying Useful Statistical Indicators of Proximity to Instability in Stochastic Power Systems", \emph{IEEE Transactions on Power Systems}, Vol. 31, No. 2, pp. 1360--1368, 2016.

\bibitem{apostolopoulou2016assessment}
D. Apostolopoulou, A.D. Dom{\'\i}nguez-Garc{\'\i}a, P.W. Sauer, ``An Assessment of the Impact of Uncertainty on Automatic Generation Control Systems", \emph{IEEE Transactions on Power Systems}, Vol. 31, No. 4, pp. 2657--2665, 2016.

 % == Probabilistic small signal analysis ===

\bibitem{nwankpa1992stochastic}
C.O. Nwankpa, S.M. Shahidehpour, and Z. Schuss, ``A Stochastic Approach to Small Disturbance Stability Analysis", \emph{IEEE Transactions on Power Systems}, Vol. 7, No. 4, pp. 1519--1528, 1992.

\bibitem{rueda2009assessment}
J.L. Rueda, D.G. Colom{\'e}, and I. Erlich, ``Assessment and Enhancement of Small Signal Stability Considering Uncertainties", \emph{IEEE Transactions on Power Systems}, Vol. 24, No. 1, pp. 198--207, 2009.

\bibitem{huang2013quasi}
H. Huang, C. Chung, K.W. Chan, and H. Chen, ``Quasi-Monte Carlo Based Probabilistic Small Signal Stability Analysis for Power Systems with Plug-in Electric Vehicle and Wind Power Integration", \emph{IEEE Transactions on Power Systems}, Vol. 28, No. 3, pp. 3335--3343, 2013.

\bibitem{dhople2013analysis}
S.V. Dhople, Y.C. Chen, L. DeVille, A.D. Dom{\'\i}nguez-Garc{\'\i}a, ``Analysis of Power System Dynamics Subject to Stochastic Power Injections", \emph{IEEE Transactions on Circuits and Systems I: Regular Papers}, Vol. 60, No. 12, pp. 3341--3353, 2013.

% === PC ===========

\bibitem{hockenberry2004evaluation}
J.R. Hockenberry, and B.C. Lesieutre, ``Evaluation of Uncertainty in Dynamic Simulations of Power System Models: The Probabilistic Collocation Method", \emph{IEEE Transactions on Power Systems}, Vol. 28, No. 3, pp. 3335--3343, 2013.

\bibitem{xu2019propagating}
Y. Xu, L. Mili, A. Sandu, M.R. von Spakovsky, and J. Zhao, ``Propagating Uncertainty in Power System Dynamic Simulations using Polynomial Chaos", \emph{IEEE Transactions on Power Systems}, Vol. 34, No. 1, pp. 338--348, 2019.

% === stochastic averaging based on energy function ====

\bibitem{ju2018analytical}
P. Ju, H. Li, C. Gan, Y. Liu, Y. Yu, and Y. Liu, ``Analytical Assessment for Transient Stability Under Stochastic Continuous Disturbances", \emph{IEEE Transactions on Power Systems}, Vol. 33, No. 2, pp. 2004--2014, 2018.

\bibitem{ju2018stochastic}
P. Ju, H. Li, X. Pan, C. Gan, Y. Liu, and Y. Liu, ``Stochastic Dynamic Analysis for Power Systems Under Uncertain Variability", \emph{IEEE Transactions on Power Systems}, Vol. 33, No. 4, pp. 3789--3799, 2018.

\bibitem{pai1989energy}
A. Pai, \emph{Energy Function Analysis for Power System Stability}, Springer Science \& Business Media, 1989.

\bibitem{chang1995direct} 
H-D. Chang, C-C. Chu, and G. Cauley, ``Direct Stability Analysis of Electric Power Systems using Energy Functions: Theory, Applications, and Perspective", \emph{Proceedings of the IEEE}, Vol. 83, No. 11, pp. 1497--1529, 1995.

\bibitem{sauerpai1998}
P.W. Sauer, and M.A. Pai, \emph{Power System Dynamics and Stability}, Prentice-Hall, 1998.

% === statistical moments =====

\bibitem{maldonado2018uncertainty}
D.A. Maldonado, M. Schanen, and M. Anitescu, ``Uncertainty Propagation in Power System Dynamics with the Method of Moments", \emph{2018 IEEE Power \& Energy Society General Meeting (PESGM)}, pp. 1--5, 2018. doi: 10.1109/PESGM.2018.8586023

% === deterministic bounded uncertainty ====

\bibitem{dimitrovski2004boundary}
A. Dimitrovski, and K. Tomsovic, ``Boundary Load Flow Solutions", \emph{IEEE Transactions on Power Systems}, Vol. 19, No. 1, pp. 348--355, 2004.

\bibitem{hiskens2006sensitivity}
I.A. Hiskens, and J. Alseddiqui, ``Sensitivity, Approximation, and Uncertainty in Power System Dynamic Simulation", \emph{IEEE Transactions on Power Systems}, Vol. 21, No. 4, pp. 1808--1820, 2006.

\bibitem{chen2012method}
Y.C. Chen, and A.D. Dominguez-Garcia, ``A Method to Study the Effect of Renewable Resource Variability on Power System Dynamics", \emph{IEEE Transactions on Power Systems}, Vol. 27, No. 4, pp. 1978--1989, 2012.

\bibitem{althoff2014formal}
M. Althoff, ``Formal and Compositional Analysis of Power Systems using Reachable Sets", \emph{IEEE Transactions on Power Systems}, Vol. 29, No. 5, pp. 2270--2280, 2014.

\bibitem{choi2017propagating}
H. Choi, P.J. Seiler, and S.V. Dhople, ``Propagating Uncertainty in Power-System DAE Models with Semidefinite Programming", \emph{IEEE Transactions on Power Systems}, Vol. 32, No. 4, pp. 3146--3156, 2017.

\bibitem{choi2018propagating}
H. Choi, P.J. Seiler, and S.V. Dhople, ``Propagating Uncertainty in Power Flow With the Alternating Direction Method of Multipliers", \emph{IEEE Transactions on Power Systems}, Vol. 33, No. 4, pp. 4124--4133, 2018.

% Bellman: curse of dimensionality
% =====================================
\bibitem{bellman1957}
R.E. Bellman, \emph{Dynamic Programming}, Courier Dover Publications, 1957.

% == Kuramoto second order ===

\bibitem{dorfler2012synchronization}
F. D\"{o}rfler, and F. Bullo, ``Synchronization and Transient Stability in Power Networks and Nonuniform Kuramoto Oscillators", \emph{SIAM Journal on Control and Optimization}, Vol. 50, No. 3, pp. 1616--1642, 2012.

\bibitem{rodrigues2016kuramoto}
F.A. Rodrigues, T.K. Peron, P. Ji, and J. Kurths, ``The Kuramoto Model in Complex Networks", \emph{Physics Reports}, Vol. 610, pp. 1--98, 2016.

\bibitem{dorfler2013synchronization}
F. D{\"o}rfler, M. Chertkov, and F. Bullo, ``Synchronization in Complex Oscillator Networks and Smart Grids", \emph{Proceedings of the National Academy of Sciences}, Vol. 100, No. 6, pp. 2005--2010, 2013.

\bibitem{ishizaki2018}
T. Ishizaki, A. Chakrabortty, and J-I. Imura. ``Graph-Theoretic Analysis of Power Systems", \emph{Proceedings of the IEEE}, Vol. 106, No. 5, pp. 931--952, 2018.

% == Our prox papers ===

\bibitem{caluya2019ACC}
K.F. Caluya, and A. Halder, ``Proximal Recursion for Solving the Fokker-Planck Equation", \emph{Proceedings of the 2019 American Control Conference}, pp. 4098--4103, 2019.

\bibitem{caluya2019TAC}
K.F. Caluya, and A. Halder, ``Gradient Flow Algorithms for Density Propagation in Stochastic Systems", \emph{IEEE Transactions on Automatic Control}, Vol. 65, No. 10, pp. 3991--4004, 2019.

\bibitem{halder2020hopfield}
A. Halder, K.F. Caluya, B. Travacca, and S.J. Moura, ``Hopfield Neural Network Flow: A Geometric Viewpoint", \emph{IEEE Transactions on Neural Networks and Learning Systems}, Vol. 31, No. 11, pp. 4869--4880, 2020.

\bibitem{caluya2021TAC}
K.F. Caluya, and A. Halder, ``Wasserstein Proximal Algorithms for the Schr\"{o}dinger Bridge Problem: Density Control with Nonlinear Drift", \emph{IEEE Transactions on Automatic Control}, 2021.

% === FPK for transient stability ====

\bibitem{wang2013fokker}
K. Wang, and M.L. Crow, ``The Fokker-Planck Equation for Power System Stability Probability Density Function Evolution", \emph{IEEE Transactions on Power Systems}, Vol. 28, No. 3, pp. 2994--3001, 2013.

%== Wasserstein gradient flow ====

\bibitem{jordan1998variational}
R. Jordan, D. Kinderlehrer, and F. Otto, ``The Variational Formulation of the Fokker--Planck Equation", \emph{SIAM journal on mathematical analysis}, Vol. 29, pp. 1--17, 1998.

\bibitem{villani2003topics}
C. Villani, \emph{Topics in optimal transportation}, No. 58, American Mathematical Society, 2003.

\bibitem{ambrosio2008gradient}
L.~Ambrosio, N. Gigli, and G. Savar{\'e}, \emph{Gradient Flows: in Metric Spaces and in the Space of Probability Measures}, Springer Science \& Business Media, 2008.

\bibitem{santambrogio2017euclidean}
F.~Santambrogio, ``$\{$Euclidean, Metric, and Wasserstein$\}$ Gradient Flows: An Overview", \emph{Bulletin of Mathematical Sciences}, Vol. 7, No. 1, pp. 87--154, 2017.

% Einstein relation

\bibitem{hernandez1989equilibrium}
D.B.~Hernandez, and M.~Pavon, ``Equilibrium Description of A Particle System in A Heat Bath", \emph{Acta Applicandae Mathematica}, Vol. 14, No. 3, pp. 239--258, 1989.

\bibitem{chen2015fast}
Y.~Chen, T.T.~Georgiou, and M.~Pavon, ``Fast Cooling for A System of Stochastic Oscillators", \emph{Journal of Mathematical Physics}, Vol. 56, No. 11, pp. 113302, 2015.

\end{thebibliography}

\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiography}[{\includegraphics[height=1.25in]{Abhishek-Halder.jpg}}]{Abhishek
Halder}
(S'10-M'14) is an Assistant Professor in the Department of Applied Mathematics,
and is an affiliated faculty in the Department of Electrical and Computer Engineering at
University of California, Santa Cruz. Before that he held postdoctoral
positions in the Department of Mechanical and Aerospace Engineering at
University of California, Irvine, and in the Department of Electrical
and Computer Engineering at Texas A\&M University. He obtained his
Bachelors and Masters from Indian Institute of Technology Kharagpur in
2008, and Ph.D. from Texas A\&M University in 2014, all in Aerospace
Engineering. His research interests are in stochastic systems, control
and optimization with application focus on large scale cyber-physical
systems.
\end{IEEEbiography}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2021-06-13 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively.
-The symbol $|\cdot|$ denotes the absolute value, $\otimes$ denotes the Kronecker product.
+The symbol $|\cdot|$ denotes the absolute value, $\otimes$ denotes the Kronecker product, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map.
 The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.
@@
 The potential (\ref{potential}) has a natural energy function interpretation and can also be motivated by a mechanical mass-spring-damper analogy \cite{dorfler2013synchronization,ishizaki2018}.
-The function $V$ is continuously differentiable, bounded below, and goes to $+\infty$ as $\|\bm{x}\|_{2}\rightarrow +\infty$.
+The function $V$ is continuously differentiable, bounded below, and goes to $+\infty$ as $\|\bm{\theta}\|_{2}\rightarrow +\infty$.
 %In the development below, an important role will be played by the ``Hamiltonian-like" function
@@
 
-\subsection{Statistical Mechanics Issues}
+\subsection{Statistical Mechanics Perspective}
 A new difficulty for our SDE \eqref{ItoSDEvectorlevel} is that we have \emph{anisotropic} degenerate diffusion, i.e., the strengths of the noise acting in the last $n$ components of \eqref{ItoSDEvectorlevel} are nonuniform since $\bm{M}^{-1}\bm{\Sigma}$ is not identity.
-This complicates the matter because the construction of the functional $\Phi$ is usually motivated via free energy considerations utilizing the structure of the \emph{stationary} PDF $\rho_{\infty}(\bm{\theta},\bm{\omega})$ for \eqref{KineticFPK}.
+This complicates the matter because the construction of the functional $\Phi$ in \eqref{ProxRecursionInfiniteDim} is usually motivated via free energy considerations utilizing the structure of the \emph{stationary} PDF $\rho_{\infty}(\bm{\theta},\bm{\omega})$ for \eqref{KineticFPK}.
 The $\rho_{\infty}$ is, in turn, guaranteed to be a unique \emph{Boltzmann distribution} of the form
@@
 \end{align}
-In our case, $\bm{\Sigma,\Gamma}$ are positive diagonal, and \eqref{EinsteinRelation} requires $\sigma_{i}^{2} \propto \gamma_{i}$ for all $i=1,\hdots,n$.
+In our case, $\bm{\Sigma,\Gamma}$ are positive diagonal, and \eqref{EinsteinRelation} is equivalent to $\sigma_{i}^{2} \propto \gamma_{i}$ for all $i=1,\hdots,n$.
 
@@
 \subsection{From Anisotropic to Isotropic Degenerate Diffusion}\label{subsec:FromAnisoToIso}
-Let the $2n\times 2n$ matrix
+Consider the $2n\times 2n$ matrix
 \begin{align}
@@
 \end{subequations}}}
+
+We notice that \eqref{XiEtaVectorSDE} is a mixed-conservative SDE with \emph{isotropic} degenerate diffusion.
+In particular, the pushforward of the known initial joint PDF \eqref{FPKInitCond} via $\bm{\Psi}$, is given by
+\begin{align}
+\tilde{\rho}_{0}(\bm{\xi},\bm{\eta}) := \bm{\Psi}_{\sharp} \rho_{0} 
+&= \dfrac{\rho_{0}\left(\bm{\Psi}^{-1} \begin{pmatrix}
+\bm{\xi}\\
+\bm{\eta}	
+\end{pmatrix}\right)}{|\det\left(\bm{\Psi}\right)|} \nonumber\\
+&= \dfrac{\rho_{0}\left(\bm{\Sigma}\bm{M}^{-1}\bm{\xi},\bm{\Sigma}\bm{M}^{-1}\bm{\eta}\right)}{\left(\displaystyle\prod_{i=1}^{n}m_{i}/\sigma_{i}\right)^{\!2}},
+\label{PushforwardOfIC}	
+\end{align}
+where we used the standard properties of the Kronecker product.
+The transient joint state PDF $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ corresponding to \eqref{XiEtaVectorSDE} solves the PDE initial value problem
+\begin{subequations}
+\begin{align}
+&\dfrac{\partial\tilde{\rho}}{\partial t} = -\langle\bm{\eta},\nabla_{\bm{\xi}}\tilde{\rho}\rangle + \nabla_{\bm{\eta}}\cdot\left(\tilde{\rho}\left(\nabla_{\bm{
+\xi}}U(\bm{\xi})+\nabla_{\bm{\eta}}F(\bm{\eta})\right)\right) + \dfrac{1}{2}\Delta_{\bm{\eta}}\tilde{\rho}, \label{NewPDE}\\
+&\tilde{\rho}(t=t_{0},\bm{\xi},\bm{\eta}) = \underbrace{\tilde{\rho}_{0}(\bm{\xi},\bm{\eta})}_{\text{from}\;\eqref{PushforwardIC}}. \label{NewIC}	
+\end{align}
+\label{TransformedPDEivp}	
+\end{subequations}
+In other words, \eqref{TransformedPDEivp} is the macroscopic dynamics corresponding to the sample path dynamics \eqref{XiEtaVectorSDE}.
+
+Since \eqref{NewPDE} is a standard kinetic Fokker-Planck PDE with isotropic degnerate diffusion, our strategy is to perform a proximal recursion of the form \eqref{ProxRecursionInfiniteDim} for \eqref{TransformedPDEivp}, and then to pushforward the resulting joint PDFs via $\bm{\Psi}^{-1}$ to the original $(\bm{\theta},\bm{\omega})$ state space.
+
+\subsection{Proximal Update}\label{subsec:ProxUpdate}
 

### 2021-06-13 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 The resulting bivariate Fokker-Planck PDE in \cite{wang2013fokker} was solved via finite element discretization, and revealed rich stochastic dynamics and nontrivial transient stability aspects even in this simple case.
-However, it is unreasonable to expect that a finite element discretization, or in fact any spatial discretization scheme to solve (\ref{KineticFPK}) for moderately large $n$ in minutes of computational time, thereby limiting our current ability for realistic power systems simulation with stochastic variability.
+However, it is unreasonable to expect that a finite element discretization, or in fact any spatial discretization scheme to solve (\ref{KineticFPK}) for moderately large $n$ in {\red{minutes}} of computational time, thereby limiting our current ability for realistic power systems simulation with stochastic variability.
 This calls for fundamentally re-thinking what does it mean to solve the PDE (\ref{KineticFPK}) for dynamics (\ref{ItoSDEvectorlevel}).
@@
 This complicates the matter because the construction of the functional $\Phi$ in \eqref{ProxRecursionInfiniteDim} is usually motivated via free energy considerations utilizing the structure of the \emph{stationary} PDF $\rho_{\infty}(\bm{\theta},\bm{\omega})$ for \eqref{KineticFPK}.
-The $\rho_{\infty}$ is, in turn, guaranteed to be a unique \emph{Boltzmann distribution} of the form
+The $\rho_{\infty}$ is, in turn, guaranteed to be a unique \emph{Boltzmann distribution} of the form\footnote{here $Z$ is a normalizing constant known as the ``partition function".}
 \begin{subequations}
@@
 \end{subequations}
-where $Z$ is a normalizing constant (known as the ``partition function"), \emph{if and only if} the so-called \emph{Einstein relation} \cite{hernandez1989equilibrium,chen2015fast} holds:
+\emph{if and only if} the so-called \emph{Einstein relation} \cite{hernandez1989equilibrium,chen2015fast} holds:
 \begin{align}
@@
 
-We notice that \eqref{XiEtaVectorSDE} is a mixed-conservative SDE with \emph{isotropic} degenerate diffusion.
+We notice that \eqref{XiEtaVectorSDE} is a mixed conservative-dissipative SDE with \emph{isotropic} degenerate diffusion.
 In particular, the pushforward of the known initial joint PDF \eqref{FPKInitCond} via $\bm{\Psi}$, is given by
@@
 \xi}}U(\bm{\xi})+\nabla_{\bm{\eta}}F(\bm{\eta})\right)\right) + \dfrac{1}{2}\Delta_{\bm{\eta}}\tilde{\rho}, \label{NewPDE}\\
-&\tilde{\rho}(t=t_{0},\bm{\xi},\bm{\eta}) = \underbrace{\tilde{\rho}_{0}(\bm{\xi},\bm{\eta})}_{\text{from}\;\eqref{PushforwardIC}}. \label{NewIC}	
+&\tilde{\rho}(t=t_{0},\bm{\xi},\bm{\eta}) = \underbrace{\tilde{\rho}_{0}(\bm{\xi},\bm{\eta})}_{\text{from}\;\eqref{PushforwardOfIC}}. \label{NewIC}	
 \end{align}
@@
 
-Since \eqref{NewPDE} is a standard kinetic Fokker-Planck PDE with isotropic degnerate diffusion, our strategy is to perform a proximal recursion of the form \eqref{ProxRecursionInfiniteDim} for \eqref{TransformedPDEivp}, and then to pushforward the resulting joint PDFs via $\bm{\Psi}^{-1}$ to the original $(\bm{\theta},\bm{\omega})$ state space.
+Since \eqref{NewPDE} is a kinetic Fokker-Planck PDE with isotropic degenerate diffusion, our strategy is to perform a proximal recursion of the form \eqref{ProxRecursionInfiniteDim} for \eqref{TransformedPDEivp}, and then to pushforward the resulting joint PDFs via $\bm{\Psi}^{-1}$ to the original state space.
+This is what we detail next.
 
 \subsection{Proximal Update}\label{subsec:ProxUpdate}
+Looking at \eqref{NewPotentials} and \eqref{NewPDE}, it is natural to consider the energy functional 
+\begin{align}
+\Phi(\tilde{\rho}) := \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\left(U(\bm{\xi})+F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\right)\tilde{\rho}\:\differential\bm{\xi}\differential\bm{\eta},
+\label{LyapButNotProx}	
+\end{align}
+which is the sum of a potential energy (expected value of $U$), a kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy).
+Indeed, the functional \eqref{LyapButNotProx} is a Lyapunov functional along the solution of \eqref{NewPDE}.
+To see this, ...
+
+However, it
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Proximal Algorithm}\label{sec:ProxAlgo}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Numerical Simulations}\label{sec:NumericalSimulations}
 

### 2021-06-14 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 \end{align}
-Applying It\^{o}'s lemma, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
+Applying It\^{o}'s lemma \cite[Ch. 4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
 \begin{align}
@@
 
-However, it
+However, it is not possible
 
@@
 
-% Einstein relation
+% ==== Einstein relation =====
 
@@
 Y.~Chen, T.T.~Georgiou, and M.~Pavon, ``Fast Cooling for A System of Stochastic Oscillators", \emph{Journal of Mathematical Physics}, Vol. 56, No. 11, pp. 113302, 2015.
+
+% ==== Ito's lemma ========
+
+\bibitem{oksendal2013stochastic}
+B.~Oksendal, \emph{Stochastic Differential Equations: An Introduction with Applications}, \emph{Springer Science \& Business Media}, 2013.
 

### 2021-06-14 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 \bibitem{oksendal2013stochastic}
-B.~Oksendal, \emph{Stochastic Differential Equations: An Introduction with Applications}, \emph{Springer Science \& Business Media}, 2013.
+B.~\O{}ksendal, \emph{Stochastic Differential Equations: An Introduction with Applications}, \emph{Springer Science \& Business Media}, 2013.
 

### 2021-06-14 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 \end{align}
-Applying It\^{o}'s lemma \cite[Ch. 4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
+Applying It\^{o}'s lemma \cite[Ch.
+4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, and using \eqref{XiEtaVectorSDE}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE
 \begin{align}
@@
 
-Since \eqref{NewPDE} is a kinetic Fokker-Planck PDE with isotropic degenerate diffusion, our strategy is to perform a proximal recursion of the form \eqref{ProxRecursionInfiniteDim} for \eqref{TransformedPDEivp}, and then to pushforward the resulting joint PDFs via $\bm{\Psi}^{-1}$ to the original state space.
+Since \eqref{NewPDE} is a kinetic Fokker-Planck PDE with isotropic degenerate diffusion, our strategy is to perform a proximal recursion of the form \eqref{ProxRecursionInfiniteDim} for \eqref{TransformedPDEivp} in $(\bm{\xi},\bm{\eta})$ coordinates, and then to pushforward the resulting joint PDFs via $\bm{\Psi}^{-1}$ to the original state space.
 This is what we detail next.

### 2021-06-14 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 We use boldfaced small letters for vectors, and boldfaced capital letters for matrices.
+The set of natural numbers is denoted as $\mathbb{N}$.
 The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$.
@@
 Applying It\^{o}'s lemma \cite[Ch.
-4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, and using \eqref{XiEtaVectorSDE}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE
+4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, and using \eqref{ItoSDEvectorlevel}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE
 \begin{align}

### 2021-06-15 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 
-We notice that \eqref{XiEtaVectorSDE} is a mixed conservative-dissipative SDE with \emph{isotropic} degenerate diffusion.
+Notice that \eqref{XiEtaVectorSDE} is a mixed conservative-dissipative SDE with \emph{isotropic} degenerate diffusion.
 In particular, the pushforward of the known initial joint PDF \eqref{FPKInitCond} via $\bm{\Psi}$, is given by
@@
 \end{align}
-which is the sum of a potential energy (expected value of $U$), a kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy).
+which is the sum of a potential energy (expected value of $U$), a kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy $-\int\tilde{\rho}\log\tilde{\rho}$).
 Indeed, the functional \eqref{LyapButNotProx} is a Lyapunov functional along the solution of \eqref{NewPDE}.

### 2021-06-15 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 
+Our technical approach juxtaposes with respect to the existing literature discussed above, in that we approximate neither the dynamical nonlinearity nor the statistics.
+Instead of treating the exact nonlinearity as bane, we exploit the geometry induced by the power system's dynamical nonlinearity
+
 \subsubsection{Notation and organization}
@@
 
-The corresponding macroscopic dynamics governing the flow of the joint state PDF $\rho(t,\bm{\theta},\bm{\omega})$ is given by a kinetic Fokker-Planck or Kolmogorov's forward partial differential equation (PDE)
+The corresponding macroscopic dynamics governing the flow of the joint state PDF $\rho(t,\bm{\theta},\bm{\omega})$ is given by a kinetic Fokker-Planck partial differential equation (PDE)
 \begin{align}
@@
 \begin{align}
-\bm{\Psi}:= \bm{I}_{2} \otimes \bm{M}\bm{\Sigma}^{-1}, \label{DefPsi}	
+\bm{\Psi}:= \bm{I}_{2} \otimes \left(\bm{M}\bm{\Sigma}^{-1}\right), \label{DefPsi}	
 \end{align}
@@
 \end{align}
-which is the sum of a potential energy (expected value of $U$), a kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy $-\int\tilde{\rho}\log\tilde{\rho}$).
+which is the sum of a potential energy (expected value of $U$), a kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$).
 Indeed, the functional \eqref{LyapButNotProx} is a Lyapunov functional along the solution of \eqref{NewPDE}.

### 2021-06-15 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 
-Our technical approach juxtaposes with respect to the existing literature discussed above, in that we approximate neither the dynamical nonlinearity nor the statistics.
-Instead of treating the exact nonlinearity as bane, we exploit the geometry induced by the power system's dynamical nonlinearity
+Our technical approach juxtaposes with the existing literature discussed above, in that we approximate neither the dynamical nonlinearity nor the statistics.
+Instead of treating the exact nonlinearity as bane, we exploit the geometry induced by the power system's dynamical nonlinearity over the manifold of time-varying joint state probability densities.
 

### 2021-06-15 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 Our technical approach juxtaposes with the existing literature discussed above, in that we approximate neither the dynamical nonlinearity nor the statistics.
-Instead of treating the exact nonlinearity as bane, we exploit the geometry induced by the power system's dynamical nonlinearity over the manifold of time-varying joint state probability densities.
+Instead of treating the exact nonlinearity as bane, we exploit the geometry induced by the power system's dynamical nonlinearity over the manifold of time-varying joint state PDFs, thereby enabling a new proximal algorithm to compute the transient joint PDFs.
 

### 2021-06-16 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 
-We define the positive diagonal matrices
+We define the \emph{positive diagonal} matrices
 \begin{align*}
@@
 \begin{align}
-\Phi(\tilde{\rho}) := \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\left(U(\bm{\xi})+F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\right)\tilde{\rho}\:\differential\bm{\xi}\differential\bm{\eta},
+\Phi(\tilde{\rho}) := \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\left(U(\bm{\xi})+ \frac{1}{2}\|\bm{\eta}\|_{2}^{2} + \frac{1}{2}\log\tilde{\rho}\right)\tilde{\rho}\:\differential\bm{\xi}\differential\bm{\eta},
 \label{LyapButNotProx}	
 \end{align}
-which is the sum of a potential energy (expected value of $U$), a kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$).
-Indeed, the functional \eqref{LyapButNotProx} is a Lyapunov functional along the solution of \eqref{NewPDE}.
-To see this, ...
+which is the sum of a potential energy (expected value of $U$), a kinetic energy (expected value of $\frac{1}{2}\|\bm{\eta}\|_{2}^{2}$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$).
+Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov functional, i.e., $\frac{\differential}{\differential t}\Phi \leq 0$ along the solution of \eqref{NewPDE}.
 
@@
 
-%\appendices
-%\section{Proof of the First Zonklar Equation}
-%Appendix one text goes here.
+\appendices
+\section{Showing $\frac{\differential}{\differential t}\Phi(\tilde{\rho})$ along \eqref{NewPDE}}\label{AppLyap}
+The ma
 

### 2021-06-16 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 \end{align}
-In our case, $\bm{\Sigma,\Gamma}$ are positive diagonal, and \eqref{EinsteinRelation} is equivalent to $\sigma_{i}^{2} \propto \gamma_{i}$ for all $i=1,\hdots,n$.
+In our case, $\bm{\Sigma,\Gamma}$ are positive diagonal, and \eqref{EinsteinRelation} is equivalent to the proportionality constraint: $\sigma_{i}^{2} \propto \gamma_{i}$ for all $i=1,\hdots,n$.
 

### 2021-06-16 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 \subsection{Proximal Update}\label{subsec:ProxUpdate}
-Looking at \eqref{NewPotentials} and \eqref{NewPDE}, it is natural to consider the energy functional 
+Looking at \eqref{NewPotentials} and \eqref{NewPDE}, it is natural to consider an energy functional of the form
 \begin{align}
-\Phi(\tilde{\rho}) := \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\left(U(\bm{\xi})+ \frac{1}{2}\|\bm{\eta}\|_{2}^{2} + \frac{1}{2}\log\tilde{\rho}\right)\tilde{\rho}\:\differential\bm{\xi}\differential\bm{\eta},
+\Phi(\tilde{\rho}) := \!\!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\!\!\left(\!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\!\right)\tilde{\rho}\:\differential\bm{\xi}\differential\bm{\eta},
 \label{LyapButNotProx}	
 \end{align}
-which is the sum of a potential energy (expected value of $U$), a kinetic energy (expected value of $\frac{1}{2}\|\bm{\eta}\|_{2}^{2}$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$).
-Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov functional, i.e., $\frac{\differential}{\differential t}\Phi \leq 0$ along the solution of \eqref{NewPDE}.
+which is the sum of a potential energy (expected value of scaled $U$), a weighted kinetic energy (expected value of $F\left(\bm{\eta}\right)$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$).
+Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.
 
@@
 \appendices
-\section{Showing $\frac{\differential}{\differential t}\Phi(\tilde{\rho})$ along \eqref{NewPDE}}\label{AppLyap}
-The ma
+\section{Showing $\Phi(\tilde{\rho})$ is Decreasing along the Flow Generated by \eqref{TransformedPDEivp}}\label{AppLyap}
+Let $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ be the transient solution of \eqref{TransformedPDEivp}. Define
+$$\bm{\ell}_{\bm{\xi}} := \frac{1}{2}\nabla_{\bm{\xi}}\log\tilde{\rho}, \quad \bm{\ell}_{\bm{\eta}} := \frac{1}{2}\nabla_{\bm{\eta}}\log\tilde{\rho},$$
+and rewrite \eqref{NewPDE} as
+$$\dfrac{\partial\tilde{\rho}}{\partial t} = - \nabla\cdot\left(\tilde{\rho}\bm{v}\right)$$
+where $\nabla$ is the $2n\times 1$ Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and the vector field
+$$\bm{v} := \begin{pmatrix}
+ \bm{\eta}\\
+ -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}\bm{\Gamma}^{-1}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
+ \end{pmatrix}
+.$$
+
+Now consider the time derivative of the functional \eqref{LyapButNotProx} along \eqref{NewPDE}. Following \cite[Proposition 1]{krichene2020global}, 
+\begin{align}
+\frac{\differential}{\differential t}\Phi = \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}} \bigg\langle\nabla\dfrac{\delta\Phi}{\delta\tilde{\rho}}, \tilde{\rho}\bm{v}\bigg\rangle\:\differential\bm{\xi}\differential\bm{\eta},
+\label{dPhidt}	
+\end{align}
+wherein the functional derivative
+\begin{align}
+\dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
+\end{align}
+Recalling $F(\bm{\eta})$ from \eqref{defF}, the relation \eqref{dPhidt} simplifies to
+\begin{subequations}
+\begin{align}
+\frac{\differential}{\differential t}\Phi &= \!\!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\!\bigg\langle\!\!\begin{pmatrix}
+\bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
+\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}	
+\end{pmatrix}, \begin{pmatrix}
+ \bm{\eta}\\
+ -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}\bm{\Gamma}^{-1}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
+ \end{pmatrix}\!\!\bigg\rangle\tilde{\rho}\:\differential\bm{\xi}\differential\bm{\eta}	\nonumber\\
+&= 
+\end{align}
+\label{ExplicitdPhidt}	
+\end{subequations}
 
@@
 
+% ==== Walid neural network ========
+
+\bibitem{krichene2020global}
+W. Krichene, K.F.~Caluya, and A.~Halder, ``Global Convergence of Second-order Dynamics in Two-layer Neural Networks", \emph{arXiv preprint arXiv:2007.06852}, 2020.
+
 \end{thebibliography}

### 2021-06-16 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 \usepackage{cite,url,framed,bm,dsfont,lipsum}
+\usepackage{cuted}
+\setlength\stripsep{3pt plus 1pt minus 1pt}
 
@@
 \begin{align}
-\Phi(\tilde{\rho}) := \!\!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\!\!\left(\!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\!\right)\tilde{\rho}\:\differential\bm{\xi}\differential\bm{\eta},
+\Phi(\tilde{\rho}) := \!\!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\!\!\left(\!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\!\right)\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta},
 \label{LyapButNotProx}	
@@
  \bm{\eta}\\
- -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}\bm{\Gamma}^{-1}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
+ -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
  \end{pmatrix}
@@
 \dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
+\label{FunctionalDerivative}
 \end{align}
-Recalling $F(\bm{\eta})$ from \eqref{defF}, the relation \eqref{dPhidt} simplifies to
+Recalling the explicit form of $F(\bm{\eta})$ from \eqref{defF}, and using \eqref{FunctionalDerivative}, we simplify \eqref{dPhidt} as
+\begin{strip}
+\hrule
 \begin{subequations}
 \begin{align}
-\frac{\differential}{\differential t}\Phi &= \!\!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\!\bigg\langle\!\!\begin{pmatrix}
+\frac{\differential}{\differential t}\Phi &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\langle\begin{pmatrix}
 \bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
@@
  \bm{\eta}\\
- -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}\bm{\Gamma}^{-1}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
- \end{pmatrix}\!\!\bigg\rangle\tilde{\rho}\:\differential\bm{\xi}\differential\bm{\eta}	\nonumber\\
-&= 
+ -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
+ \end{pmatrix}\bigg\rangle\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta}	\nonumber\\
+&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
+&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
 \end{align}
@@
 \end{subequations}
+\hrule	
+\end{strip}
+where \eqref{lastequality} follows from the fact that the integrals involving $\langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle$ and $\langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle$ in \eqref{lastbutoneequality} are zero, see \cite[Appendix D.3]{krichene2020global}.
+
+From \eqref{ExplicitdPhidt}, it follows that $\frac{\differential}{\differential t}\Phi \leq 0$, i.e., $\Phi$ is decreasing along the flow generated by \eqref{TransformedPDEivp}.
 

### 2021-06-16 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updated TPS manuscript

@@
 
-From \eqref{ExplicitdPhidt}, it follows that $\frac{\differential}{\differential t}\Phi \leq 0$, i.e., $\Phi$ is decreasing along the flow generated by \eqref{TransformedPDEivp}.
+From \eqref{ExplicitdPhidt}, it follows that $\frac{\differential}{\differential t}\Phi \leq 0$, i.e., $\Phi$ is non-increasing along the flow generated by \eqref{TransformedPDEivp}.
+Furthermore, \eqref{lastequality} tells us that $\frac{\differential}{\differential t}\Phi = 0$ if and only if
+$$\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}} = 0$$
 

### 2021-06-17 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

@@
 Furthermore, \eqref{lastequality} tells us that $\frac{\differential}{\differential t}\Phi = 0$ if and only if
-$$\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}} = 0$$
+$$\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}} = 0,$$
+i.e., when
 

### 2021-06-17 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

--- 2021-06-16 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 Furthermore, \eqref{lastequality} tells us that $\frac{\differential}{\differential t}\Phi = 0$ if and only if
-$$\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}} = 0$$
+$$\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}} = 0,$$
+i.e., $\Phi$ is stationary if and only if $\tilde{\rho}$ has $\bm{\eta}$ marginal $\int\tilde{\rho}\:\differential\bm{\xi} \propto \exp\left(-\bm{\eta}^{\top}\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\right)$.
+Following the steps as in \cite[Appendix E.1]{krichene2020global}, this is achieved at the stationary solution of \eqref{NewPDE} with $\bm{\xi}$ marginal $\int\tilde{\rho}\:\differential\bm{\eta} \propto \exp\left(-2U(\bm{\xi})\right)$.
+In other words, the stationary solution of \eqref{NewPDE} must be of the form
+$$\tilde{\rho}_{\infty}(\bm{\xi},\bm{\eta}) \propto \exp\left(-2U(\bm{\xi})+\bm{\eta}^{\top}\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\right),$$
+and is exactly where $\frac{\differential}{\differential t}\Phi = 0$.
+Therefore, $\frac{\differential}{\differential t}\Phi < 0$ along the transient solution of \eqref{NewPDE}, and $=0$ at the stationary solution of \eqref{NewPDE}.
 

### 2021-06-17 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

@@
 which is the sum of a potential energy (expected value of scaled $U$), a weighted kinetic energy (expected value of $F\left(\bm{\eta}\right)$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$).
-Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.
+Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov-like functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.
 
-However, it is not possible
+However, unlike the gradient drift case mentioned in Sec.
+\ref{subsec:GenGradDescent}, it is not possible to express the right hand side of \eqref{NewPDE} as the Wasserstein gradient of the functional \eqref{LyapButNotProx}.
 

### 2021-06-17 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

@@
 \ref{subsec:GenGradDescent}, it is not possible to express the right hand side of \eqref{NewPDE} as the Wasserstein gradient of the functional \eqref{LyapButNotProx}.
+The latter is defined as \cite[Ch.
+8]{ambrosio2008gradient}
+$$\nabla^{\text{Wasserstein}}\Phi := -\nabla\cdot\left(\rho\nabla\dfrac{\delta\Phi}{\delta\rho}\right),$$
+where $\nabla$ denotes the standard Euclidean gradient w.r.t.
+the vector $(\bm{\xi},\bm{\eta})^{\top}$, and $\frac{\delta}{\delta\rho}$ denotes the functional derivative.
+From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$.
+Thus, even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, we cannot interpret it as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}.
 
@@
 where $\nabla$ is the $2n\times 1$ Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and the vector field
-$$\bm{v} := \begin{pmatrix}
+\begin{align}
+\bm{v} := \begin{pmatrix}
  \bm{\eta}\\
  -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
- \end{pmatrix}
-.$$
+ \end{pmatrix}.
+\label{defvectorfieldv}	
+\end{align}
 
@@
  -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
- \end{pmatrix}\bigg\rangle\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta}	\nonumber\\
+ \end{pmatrix}\bigg\rangle\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
 &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

@@
 From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$.
-Thus, even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, we cannot interpret it as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}.
+Thus, even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, we cannot interpret the flow generated by \eqref{TransformedPDEivp} as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}.
+Thus, in \eqref{InfiniteDimProxOpDef}, we cannot construct $(\Phi,{\rm{dist}})$ by pairing \eqref{LyapButNotProx} and the Wasserstein metric.
+
+To set up a proximal recursion of the form \eqref{InfiniteDimProxOpDef} for \eqref{TransformedPDEivp}, we set $(\Phi,{\rm{dist}}) \equiv \left(\widetilde{\Phi},\widetilde{W}\right)$ where
+\begin{align}
+\widetilde{\Phi}(\tilde{\rho}) := \int_{\mathbb{T}^{n} \times \mathbb{R}^{n}}\left(F(\boldsymbol{\eta})+\frac{1}{2} \log \tilde{\rho}\right) \tilde{\rho}\: \mathrm{d} \boldsymbol{\xi} \mathrm{d} \boldsymbol{\eta},
+\label{Phitilde}	
+\end{align}
+and 
+\begin{align}
+\widetilde{W}^{2}\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right):=&\inf _{\pi \in \Pi\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right)}\!\int_{\mathbb{T}^{2 n} \times \mathbb{R}^{2 n}}\!\!\left\{\left\|\overline{\boldsymbol{\eta}}-\boldsymbol{\eta}+h \nabla_{\boldsymbol{\xi}} U(\boldsymbol{\xi})\right\|_{2}^{2}\right.\nonumber\\
+&\hspace*{-0.2in}\left.+12\left\|\frac{\boldsymbol{\xi}-\boldsymbol{\xi}}{h}-\frac{\overline{\boldsymbol{\eta}}+\boldsymbol{\eta}}{2}\right\|_{2}^{2}\right\} \mathrm{d} \pi(\boldsymbol{\xi}, \boldsymbol{\eta}, \overline{\boldsymbol{\xi}}, \overline{\boldsymbol{\eta}}).
+\label{Wtilde}	
+\end{align}
+In \eqref{Wtilde},
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

--- 2021-06-17 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$.
-Thus, even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, we cannot interpret it as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}.
+Thus, even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, we cannot interpret the flow generated by \eqref{TransformedPDEivp} as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}.
+Thus, in \eqref{InfiniteDimProxOpDef}, we cannot construct $(\Phi,{\rm{dist}})$ by pairing \eqref{LyapButNotProx} with the Wasserstein metric.
+
+To set up a variational recursion of the form \eqref{InfiniteDimProxOpDef} for \eqref{TransformedPDEivp}, we set $(\Phi,{\rm{dist}}) \equiv \left(\widetilde{\Phi},\widetilde{W}\right)$ where
+\begin{align}
+\widetilde{\Phi}(\tilde{\rho}) := \int_{\mathbb{T}^{n} \times \mathbb{R}^{n}}\left(F(\boldsymbol{\eta})+\frac{1}{2} \log \tilde{\rho}\right) \tilde{\rho}\: \mathrm{d} \boldsymbol{\xi} \mathrm{d} \boldsymbol{\eta},
+\label{Phitilde}	
+\end{align}
+and 
+\begin{align}
+\widetilde{W}^{2}\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right):=&\inf _{\pi \in \Pi\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right)}\!\int_{\mathbb{T}^{2 n} \times \mathbb{R}^{2 n}}\!\!\left\{\left\|\overline{\boldsymbol{\eta}}-\boldsymbol{\eta}+h \nabla_{\boldsymbol{\xi}} U(\boldsymbol{\xi})\right\|_{2}^{2}\right.\nonumber\\
+&\hspace*{-0.2in}\left.+12\left\|\frac{\overline{\boldsymbol{\xi}}-\boldsymbol{\xi}}{h}-\frac{\overline{\boldsymbol{\eta}}+\boldsymbol{\eta}}{2}\right\|_{2}^{2}\right\} \mathrm{d} \pi(\boldsymbol{\xi}, \boldsymbol{\eta}, \overline{\boldsymbol{\xi}}, \overline{\boldsymbol{\eta}}).
+\label{Wtilde}	
+\end{align}
+In \eqref{Wtilde},
+
+That the sequence of functions $\{\tilde{\varrho}_{k}\}$ for $k\in\mathbb{N}$, generated by the proximal recursion
+\begin{align}
+\tilde{\varrho}_{k} = {\text{prox}}^{{\rm{dist}}}_{h\widetilde{\Phi}}(\tilde{\varrho}_{k-1}), \quad k\in\mathbb{N}, \quad \varrho_{0}(\bm{x}) := \rho_{0}(\bm{x})
+\label{KineticProx}	
+\end{align}
+
+was established in \cite{duong2014conservative}.
 
@@
 
+% ==== M.H. Duong's prox for mixed conservative-dissipative =====
+
+\bibitem{duong2014conservative}
+M.H.~Duong, M.A.~Peletier, and J.~Zimmer, ``Conservative-dissipative Approximation Schemes for A Generalized Kramers Equation", \emph{Mathematical Methods in the Applied Sciences}, Vol.
+37, No.
+16, pp.
+2517--2540, 2014.
+
 \end{thebibliography}

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

@@
 The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively.
-The symbol $|\cdot|$ denotes the absolute value, $\otimes$ denotes the Kronecker product, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map.
+The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute value, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map.
 The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.
@@
  -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
- \end{pmatrix}\bigg\rangle\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
-&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
-&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
+ \end{pmatrix}\bigg\rangle\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
+&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
+&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
 \end{align}

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

@@
 \end{align}
-In \eqref{Wtilde},
+In \eqref{Wtilde}, $\Pi\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right)$ denotes the collection of all joint PDFs supported on $\mathbb{T}^{2 n} \times \mathbb{R}^{2 n}$ that have finite second raw moments, with the first marginal $\tilde{\varrho}$, and the second marginal $\tilde{\varrho}_{k-1}$.
 
@@
 \begin{align}
-\tilde{\varrho}_{k} = {\text{prox}}^{{\rm{dist}}}_{h\widetilde{\Phi}}(\tilde{\varrho}_{k-1}), \quad k\in\mathbb{N}, \quad \varrho_{0}(\bm{x}) := \rho_{0}(\bm{x})
+\tilde{\varrho}_{k} &= {\text{prox}}^{\widetilde{W}}_{h\widetilde{\Phi}}(\tilde{\varrho}_{k-1}) \nonumber\\
+&\equiv \underset{\tilde{\rho}\in\mathcal{P}_{2}}{\arg\inf}\:\frac{1}{2}\widetilde{W}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) + h\:\widetilde{\Phi}(\tilde{\rho}), \quad \tilde{\varrho}_{0} := \tilde{\rho}_{0},
 \label{KineticProx}	
 \end{align}
+converges to the flow generated by \eqref{TransformedPDEivp}, i.e., 
+$$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} \tilde{\varrho}(t=kh,\bm{\xi},\bm{\eta}),$$
+was established in \cite{duong2014conservative}.
 
-was established in \cite{duong2014conservative}.
+To compute \eqref{KineticProx}, we employ the probability-weighted scattered point cloud-based proximal algorithm proposed in \cite{caluya2019TAC}.
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

@@
 \begin{align}
-\widetilde{W}^{2}\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right):=&\inf _{\pi \in \Pi\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right)}\!\int_{\mathbb{T}^{2 n} \times \mathbb{R}^{2 n}}\!\!\left\{\left\|\overline{\boldsymbol{\eta}}-\boldsymbol{\eta}+h \nabla_{\boldsymbol{\xi}} U(\boldsymbol{\xi})\right\|_{2}^{2}\right.\nonumber\\
+\widetilde{W}^{2}\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right):=\!&\inf _{\pi \in \Pi\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right)}\!\int_{\mathbb{T}^{2 n} \times \mathbb{R}^{2 n}}\!\!\left\{\left\|\overline{\boldsymbol{\eta}}-\boldsymbol{\eta}+h \nabla_{\boldsymbol{\xi}} U(\boldsymbol{\xi})\right\|_{2}^{2}\right.\nonumber\\
 &\hspace*{-0.2in}\left.+12\left\|\frac{\overline{\boldsymbol{\xi}}-\boldsymbol{\xi}}{h}-\frac{\overline{\boldsymbol{\eta}}+\boldsymbol{\eta}}{2}\right\|_{2}^{2}\right\} \mathrm{d} \pi(\boldsymbol{\xi}, \boldsymbol{\eta}, \overline{\boldsymbol{\xi}}, \overline{\boldsymbol{\eta}}).
@@
 
-To compute \eqref{KineticProx}, we employ the probability-weighted scattered point cloud-based proximal algorithm proposed in \cite{caluya2019TAC}.
+To numerically compute \eqref{KineticProx}, we employ the probability-weighted scattered point cloud-based proximal algorithm proposed in \cite{caluya2019TAC}.
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

@@
 
-To numerically compute \eqref{KineticProx}, we employ the probability-weighted scattered point cloud-based proximal algorithm proposed in \cite{caluya2019TAC}.
+To numerically compute \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite $N$ samples.
+Specifically, the proximal algorithm recursively updates probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
 
@@
 \includegraphics[width=.8\linewidth]{BlockDiagm.pdf}
-\caption{\small{Schematic of the proposed algorithmic setup for propagating the joint state PDF as probability weighted scattered point cloud $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
-The location of the points $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ can be updated by Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\varrho_{k}^{i}\}_{i=1}^{N}$ can be updated via discrete version of the proximal recursion (\ref{ProxRecursionInfiniteDim}).}}
+\caption{\small{Schematic of the proposed algorithmic setup for propagating the joint state PDF as probability weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
+The location of the points $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ can be updated by Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ can be updated via discrete version of the proximal recursion (\ref{ProxRecursionInfiniteDim}).}}
 \label{fig:BlockDiagm}

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

@@
 To numerically compute \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite $N$ samples.
-Specifically, the proximal algorithm recursively updates probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
+Specifically, the proximal algorithm recursively updates probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where
+$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N.$$
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

--- 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 $$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} \tilde{\varrho}(t=kh,\bm{\xi},\bm{\eta}),$$
-was established in \cite{duong2014conservative}.
-
-To numerically compute \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite $N$ samples.
-Specifically, the proximal algorithm recursively updates probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
+was established in \cite{duong2014conservative}. To numerically compute \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite samples as explained next.
 
@@
 \section{Proximal Algorithm}\label{sec:ProxAlgo}
+The proximal algorithm recursively updates probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
+$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N.$$
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

--- 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 was established in \cite{duong2014conservative}.
-
-To numerically compute \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite $N$ samples.
-Specifically, the proximal algorithm recursively updates probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
+To numerically compute \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite number od samples, as explained next.
 
@@
 \section{Proximal Algorithm}\label{sec:ProxAlgo}
+Following \cite{caluya2019TAC}, our idea to numerically solve \eqref{KineticProx} is to recursively update probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where
+$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N.$$
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

--- 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 was established in \cite{duong2014conservative}.
-
-To numerically compute \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite $N$ samples.
-Specifically, the proximal algorithm recursively updates probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where
-$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N.$$
+To numerically perform the recursion \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite number of samples, as explained next.
 
@@
 \section{Proximal Algorithm}\label{sec:ProxAlgo}
+Following \cite{caluya2019TAC}, our idea to  \eqref{KineticProx} is to recursively update probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
+$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N.$$
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

@@
 converges to the flow generated by \eqref{TransformedPDEivp}, i.e., 
-$$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} \tilde{\varrho}(t=kh,\bm{\xi},\bm{\eta}),$$
+$$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} \tilde{\varrho}(t=kh,\bm{\xi},\bm{\eta})\;\text{in}\;L^{1}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right),$$
 was established in \cite{duong2014conservative}.

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

@@
 \section{Proximal Algorithm}\label{sec:ProxAlgo}
-Following \cite{caluya2019TAC}, our idea to  \eqref{KineticProx} is to recursively update probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
-$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N.$$
+Following \cite{caluya2019TAC}, we solve \eqref{KineticProx} by recursively updating probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
+$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N, \quad k\in\mathbb{N}.$$
+Thus, $\tilde{\varrho}_{k}^{i}$ is the joint PDF at the $i$th state sample
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

--- 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 \section{Proximal Algorithm}\label{sec:ProxAlgo}
-Following \cite{caluya2019TAC}, our idea to  \eqref{KineticProx} is to recursively update probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
-$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N.$$
+Following \cite{caluya2019TAC}, we solve \eqref{KineticProx} by recursively updating probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
+$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N, \quad k\in\mathbb{N}.$$
+Thus, $\tilde{\varrho}_{k}^{i}$ is the joint PDF value at $\bm{x}_{k}^{i}$, the $i$th state sample at the $k$th time step.
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix

--- 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 \section{Proximal Algorithm}\label{sec:ProxAlgo}
-Following \cite{caluya2019TAC}, our idea to  \eqref{KineticProx} is to recursively update probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
-$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N.$$
+Following \cite{caluya2019TAC}, we solve \eqref{KineticProx} by recursively updating probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
+$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N, \quad k\in\mathbb{N}.$$
+Thus, $\tilde{\varrho}_{k}^{i}$ is the joint PDF value obtained from \eqref{KineticProx} at $\bm{x}_{k}^{i}$, the $i$th state sample at the $k$th time step.
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating proof in the appendix
updating Section IV

@@
 
+\begin{figure}[htpb]
+\centering
+\includegraphics[width=.8\linewidth]{BlockDiagm.pdf}
+\caption{\small{Schematic of the proposed algorithmic setup for propagating the joint state PDF as probability weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
+The location of the points $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ can be updated by Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ can be updated via discrete version of the proximal recursion (\ref{ProxRecursionInfiniteDim}).}}
+\label{fig:BlockDiagm}
+\end{figure}
+
 \section{Proximal Algorithm}\label{sec:ProxAlgo}
@@
 Thus, $\tilde{\varrho}_{k}^{i}$ is the joint PDF value obtained from \eqref{KineticProx} at $\bm{x}_{k}^{i}$, the $i$th state sample at the $k$th time step.
+The schematic of the algorithm is shown in Fig.
+\ref{fig:BlockDiagm}.
+
+Finally, we transform the proximal updates back to the $\left(\bm{\theta},\bm{\omega}\right)$ state space via the pushforward $\bm{\Psi}^{-1}$ as
+\begin{align}
+\varrho_{k}\left(\bm{\theta},\bm{\omega}\right) &= \bm{\Psi}^{-1}_{\sharp} \tilde{\varrho}_{k}\left(\bm{\xi},\bm{\eta}\right) \nonumber\\
+&= \left(\prod_{i=1}^{n}m_{i}/\sigma_{i}\right)^{\!\!2}\tilde{\varrho}_{k}\left(\bm{M}\bm{\Sigma}^{-1}\bm{\theta},\bm{M}\bm{\Sigma}^{-1}\bm{\omega}\right).
+\label{BringBackOriginalStateSpace}	
+\end{align}
+The $\varrho_{k}\left(\bm{\theta},\bm{\omega}\right)$ from \eqref{BringBackOriginalStateSpace} approximates $\rho(t,\bm{\theta},\bm{\omega})$ (the transient solution of \eqref{KineticFPK}) in the sense \eqref{ConvergenceGuarantee}.
 
@@
 Therefore, $\frac{\differential}{\differential t}\Phi < 0$ along the transient solution of \eqref{NewPDE}, and $=0$ at the stationary solution of \eqref{NewPDE}.
-
-\begin{figure}
-\centering
-\includegraphics[width=.8\linewidth]{BlockDiagm.pdf}
-\caption{\small{Schematic of the proposed algorithmic setup for propagating the joint state PDF as probability weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
-The location of the points $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ can be updated by Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ can be updated via discrete version of the proximal recursion (\ref{ProxRecursionInfiniteDim}).}}
-\label{fig:BlockDiagm}
-\end{figure}
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section IV

@@
 
+\begin{figure}[t]
+\centering
+\includegraphics[width=.8\linewidth]{BlockDiagm.pdf}
+\caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
+The states $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}).}}
+\label{fig:BlockDiagm}
+\end{figure}
+
 \subsection{Proximal Update}\label{subsec:ProxUpdate}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[htpb]
-\centering
-\includegraphics[width=.8\linewidth]{BlockDiagm.pdf}
-\caption{\small{Schematic of the proposed algorithmic setup for propagating the joint state PDF as probability weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
-The location of the points $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ can be updated by Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ can be updated via discrete version of the proximal recursion (\ref{ProxRecursionInfiniteDim}).}}
-\label{fig:BlockDiagm}
-\end{figure}
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section IV

@@
 \caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
-The states $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}).}}
+The states $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}) as detailed in Sec.
+\ref{sec:ProxAlgo}.}}
 \label{fig:BlockDiagm}

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section IV

@@
 \section{Proximal Algorithm}\label{sec:ProxAlgo}
-Following \cite{caluya2019TAC}, we solve \eqref{KineticProx} by recursively updating probability-weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
+Following \cite{caluya2019TAC}, we solve \eqref{KineticProx} by recursively updating the probability-weighted scattered point clouds $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
 $$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N, \quad k\in\mathbb{N}.$$
@@
 \ref{fig:BlockDiagm}.
+
+In the numerical simulations reported next in Sec.
+\ref{sec:NumericalSimulations}, the states $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}).
+If one wishes so, the Euler-Maruyama scheme in Fig.
+\ref{fig:BlockDiagm} may be replaced by other SDE integrators, see e.g., \cite[Sec.
+III.B.2, Remark 1]{caluya2019TAC}.
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section IV

@@
 
-In the numerical simulations reported next in Sec.
+In the numerical simulations reported in Sec.
 \ref{sec:NumericalSimulations}, the states $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}).
@@
 III.B.2, Remark 1]{caluya2019TAC}.
+
+To numerically perform the proximal updates $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}\mapsto\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ for $k\in\mathbb{N}$, we implement an instance of the Algorithm 1 in \cite{caluya2019TAC}; see \cite[Sec.
+V.B]{caluya2019TAC}.
 

### 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section IV

--- 2021-06-18 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 III.B.2, Remark 1]{caluya2019TAC}.
+
+To numerically perform the proximal updates $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}\mapsto\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ for $k\in\mathbb{N}$, we implement an instance of the Algorithm 1 in \cite{caluya2019TAC}.
+This algorithm implements a dualization along
+with an entropic regularization of the variational update \eqref{KineticProx}, and then solves the same using a fixed point
+recursion that is provably contractive; see \cite[Sec. V.B]{caluya2019TAC}.
 

### 2021-06-19 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section IV

@@
 \end{pmatrix}\right)}{|\det\left(\bm{\Psi}\right)|} \nonumber\\
-&= \dfrac{\rho_{0}\left(\bm{\Sigma}\bm{M}^{-1}\bm{\xi},\bm{\Sigma}\bm{M}^{-1}\bm{\eta}\right)}{\left(\displaystyle\prod_{i=1}^{n}m_{i}/\sigma_{i}\right)^{\!2}},
+&= \dfrac{\rho_{0}\left(\bm{\Sigma}\bm{M}^{-1}\bm{\xi},\bm{\Sigma}\bm{M}^{-1}\bm{\eta}\right)}{\left(\displaystyle\prod_{j=1}^{n}m_{j}/\sigma_{j}\right)^{\!2}},
 \label{PushforwardOfIC}	
@@
 To numerically perform the proximal updates $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}\mapsto\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ for $k\in\mathbb{N}$, we implement an instance of the Algorithm 1 in \cite{caluya2019TAC}.
-This algorithm implements a dualization along
-with an entropic regularization of the variational update \eqref{KineticProx}, and then solves the same using a fixed point
-recursion that is provably contractive; see \cite[Sec. V.B]{caluya2019TAC}.
+This involves a dualization along
+with an entropic regularization of the variational update \eqref{KineticProx}, and then solving the same using a fixed point recursion that is provably contractive; see \cite[Sec.
+V.B]{caluya2019TAC}.
+This enables a nonparametric computation of $\tilde{\varrho}_{k}^{i} \equiv \tilde{\varrho}_{k}\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)$ for $i=1,\hdots,N$.
 
-Finally, we transform the proximal updates back to the $\left(\bm{\theta},\bm{\omega}\right)$ state space via the pushforward $\bm{\Psi}^{-1}$ as
+Finally, we transform the proximal updates back to the $\left(\bm{\theta},\bm{\omega}\right)^{\top}$ state space via the pushforward $\bm{\Psi}^{-1}$ as
 \begin{align}
-\varrho_{k}\left(\bm{\theta},\bm{\omega}\right) &= \bm{\Psi}^{-1}_{\sharp} \tilde{\varrho}_{k}\left(\bm{\xi},\bm{\eta}\right) \nonumber\\
-&= \left(\prod_{i=1}^{n}m_{i}/\sigma_{i}\right)^{\!\!2}\tilde{\varrho}_{k}\left(\bm{M}\bm{\Sigma}^{-1}\bm{\theta},\bm{M}\bm{\Sigma}^{-1}\bm{\omega}\right).
+\varrho_{k}\left(\bm{\theta}_{k}^{i},\bm{\omega}_{k}^{i}\right) &= \bm{\Psi}^{-1}_{\sharp} \tilde{\varrho}_{k}\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right) \nonumber\\
+&= \left(\prod_{j=1}^{n}m_{j}/\sigma_{j}\right)^{\!\!2}\tilde{\varrho}_{k}\left(\bm{M}\bm{\Sigma}^{-1}\bm{\theta}_{k}^{i},\bm{M}\bm{\Sigma}^{-1}\bm{\omega}_{k}^{i}\right),
 \label{BringBackOriginalStateSpace}	
 \end{align}
+for all $i=1,\hdots,N$.
 The $\varrho_{k}\left(\bm{\theta},\bm{\omega}\right)$ from \eqref{BringBackOriginalStateSpace} approximates $\rho(t,\bm{\theta},\bm{\omega})$ (the transient solution of \eqref{KineticFPK}) in the sense \eqref{ConvergenceGuarantee}.

### 2021-06-22 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section IV

@@
 \ref{subsubsec:technicalchallenges}.
-In the next Section, we discuss how a measure-valued variational recursion proposed in our recent works \cite{caluya2019ACC,caluya2019TAC,halder2020hopfield,caluya2021TAC} can be employed to address this challenge.
+In the next Section, we discuss how a measure-valued variational recursion proposed in our recent works \cite{caluya2019ACC,caluya2019TAC,halder2020hopfield,caluya2021TAC,caluya2021reflected} can be employed to address this challenge.
 
@@
 
+\bibitem{caluya2020reflected}
+K.F. Caluya, and A. Halder,``Reflected Schr$\backslash$" odinger Bridge: Density Control with Path Constraints", \emph{American Control Conference}, 2021.
+
 % === FPK for transient stability ====

### 2021-06-22 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section IV

@@
 To numerically perform the proximal updates $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}\mapsto\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ for $k\in\mathbb{N}$, we implement an instance of the Algorithm 1 in \cite{caluya2019TAC}.
-This involves a dualization along
+The algorithm involves a dualization along
 with an entropic regularization of the variational update \eqref{KineticProx}, and then solving the same using a fixed point recursion that is provably contractive; see \cite[Sec.

### 2021-06-23 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section IV
updating Section V

@@
 We use boldfaced small letters for vectors, and boldfaced capital letters for matrices.
-The set of natural numbers is denoted as $\mathbb{N}$.
+The set of natural numbers, real numbers and complex numbers are denoted as $\mathbb{N}$, $\mathbb{R}$, and $\mathbb{C}$, respectively.
 The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$.
@@
 \ref{subsubsec:technicalchallenges}.
-In the next Section, we discuss how a measure-valued variational recursion proposed in our recent works \cite{caluya2019ACC,caluya2019TAC,halder2020hopfield,caluya2021TAC,caluya2021reflected} can be employed to address this challenge.
+In the next Section, we discuss how a measure-valued variational recursion proposed in our recent works \cite{caluya2019ACC,caluya2019TAC,halder2020hopfield,caluya2021TAC,caluya2020reflected} can be employed to address this challenge.
 
@@
 \section{Numerical Simulations}\label{sec:NumericalSimulations}
+
+\subsection{IEEE Test System}\label{subsec:NumSimIEEEBus}
+
+\subsection{Synthetic Test System}\label{subsec:NumSimSynthetic}
 
@@
 \bibitem{caluya2020reflected}
-K.F. Caluya, and A. Halder,``Reflected Schr$\backslash$" odinger Bridge: Density Control with Path Constraints", \emph{American Control Conference}, 2021.
+K.F. Caluya, and A. Halder,``Reflected Schr\"{o}dinger Bridge: Density Control with Path Constraints", \emph{American Control Conference}, 2021.
 

### 2021-06-23 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \end{align}
-which is the sum of a potential energy (expected value of scaled $U$), a weighted kinetic energy (expected value of $F\left(\bm{\eta}\right)$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$).
+which is the sum of a potential energy (expected value of scaled $U$), a weighted kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$).
 Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov-like functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.

### 2021-06-23 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 
+\begin{table}[t]
+\centering
+\begin{tabular}{| l | l |} 
+\hline
+Description & Values\\
+\hline\hline
+&\\
+frequency & $f_{0} = 60$ Hz\\
+&\\
+inertia & $m_{i}\in{\rm{Unif}}\left([2,12]\right)/2\pi f_{0}$\\
+& \\
+damping coefficient & $\gamma_{i}\in{\rm{Unif}}\left([20,30]\right)/2\pi f_{0}$\\
+& \\
+diffusion coefficient & $\sigma_{i}\in{\rm{Unif}}\left([1,5]\right)$\\
+& \\
+\hline
+\end{tabular}
+\caption{Parameters used for the numerical simulation reported in Sec. \ref{subsec:NumSimSynthetic}.}
+\label{table:ParamUncertaintiesSynthetic}
+\end{table}
+
 \subsection{Synthetic Test System}\label{subsec:NumSimSynthetic}
+We consider a power network with $n=20$ generators.
 

### 2021-06-23 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

--- 2021-06-23 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively.
+The uniform probability distribution over an interval $[a,b]$ is denoted as ${\rm{Unif}}\left([a,b]\right)$.
 The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute value, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map.
@@
 
+\begin{table}[t]
+\centering
+\begin{tabular}{| l | l |} 
+\hline
+Description & Values\\
+\hline\hline
+&\\
+frequency & $f_{0} = 60$ Hz\\
+&\\
+inertia & $m_{i}\in{\rm{Unif}}\left([2,12]\right)/2\pi f_{0}$\\
+& \\
+damping coefficient & $\gamma_{i}\in{\rm{Unif}}\left([20,30]\right)/2\pi f_{0}$\\
+& \\
+diffusion coefficient & $\sigma_{i}\in{\rm{Unif}}\left([1,5]\right)$\\
+& \\
+tangent of phase shift & $\tan\varphi_{ij}\in{\rm{Unif}}\left([0,0.25]\right)$\\
+& \\
+mechanical power input & $P_{i}^{\text{mech}}\in{\rm{Unif}}\left([0,10]\right)$\\
+& \\
+coupling coefficient & $k_{ij}\in{\rm{Unif}}\left([0.7,1.2]\right)$\\
+& \\
+\hline
+\end{tabular}
+\caption{Parameters used for the numerical simulation reported in Sec. \ref{subsec:NumSimSynthetic}.}
+\label{table:ParamUncertaintiesSynthetic}
+\end{table}
+
 \subsection{Synthetic Test System}\label{subsec:NumSimSynthetic}
+We consider a power network with $n=20$ generators, and propagate the transient joint state PDFs supported over $\mathbb{T}^{20}\times\mathbb{R}^{20}$. We take the initial joint state PDF as
+\[\rho_{0}(t=0,\theta(0),\omega(0)) = {\rm{Unif}}\left([0,2\pi)^{n}\right) \times {\rm{Unif}}\left([-0.1,0.1]\right).\]
 

### 2021-06-23 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 The uniform probability distribution over an interval $[a,b]$ is denoted as ${\rm{Unif}}\left([a,b]\right)$.
+Likewise, the $n$ dimensional uniform probability distribution over $[a,b]^{n}$ is denoted as ${\rm{Unif}}\left([a,b]^{n}\right)$.
 The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute value, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map.
@@
 
-The rest of this paper is structured as follows. {\red{TBD}}
+The rest of this paper is structured as follows.
+Section \ref{sec:models} details the power system dynamics models at the microscopic or trajectory level, as well as at the macroscopic or ensemble level.
 
@@
 \subsection{Synthetic Test System}\label{subsec:NumSimSynthetic}
-We consider a power network with $n=20$ generators, and propagate the transient joint state PDFs supported over $\mathbb{T}^{20}\times\mathbb{R}^{20}$. We take the initial joint state PDF as
-\[\rho_{0}(t=0,\theta(0),\omega(0)) = {\rm{Unif}}\left([0,2\pi)^{n}\right) \times {\rm{Unif}}\left([-0.1,0.1]\right).\]
+We consider a power network with $n=20$ generators, and propagate the transient joint state PDFs supported over $\mathbb{T}^{20}\times\mathbb{R}^{20}$. We take the initial joint state PDF at $t=0$ as
+\begin{align*}
+\rho_{0}\equiv\rho(t=0,\theta(0),\omega(0)) = &\;{\rm{Unif}}\left([0,2\pi)^{n}\;\text{rad}\right)\\
+	&\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
+\end{align*}
+and the stochastic parameters as in Table \ref{table:ParamUncertaintiesSynthetic}.
 

### 2021-06-23 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 & \\
-mechanical power input & $P_{i}^{\text{mech}}\in{\rm{Unif}}\left([0,10]\right)$\\
+effective power input & $P_{i}\in{\rm{Unif}}\left([0,10]\right)$\\
 & \\
@@
 \end{tabular}
-\caption{Parameters used for the numerical simulation reported in Sec. \ref{subsec:NumSimSynthetic}.}
+\caption{Parameters used for the numerical simulation reported in Sec. \ref{subsec:NumSimSynthetic}. The indices $i,j\in\{1,\hdots,n\}$.}
 \label{table:ParamUncertaintiesSynthetic}

### 2021-06-23 code/Gen_20_prox_simulation.m

@@ -8,7 +8,7 @@ global num_Oscillator dim Gamma M Sigma K Pmech cc sigma m phi
 
 num_Oscillator = 20; dim  = 2*num_Oscillator; % dim = dimension of state space
 
-f0 = 50;
+f0 = 60;
 g_a = 20 ; g_b = 30; m_a = 2; m_b = 12 ; s_a = 1  ; s_b = 5;
 p_a = 0; p_b = 10 ;  k_a =.7 ; k_b = 1.2;
 
@@ -25,6 +25,7 @@ pmech = (p_b-p_a)*rand(num_Oscillator,1) + p_b;
 Gamma = diag(gamma);  M = diag(m); Sigma = diag(sigma); Pmech = diag(pmech);
 
 K = (k_b-k_a)*rand(num_Oscillator)+ k_a;
+K = K - diag(diag(K)); % zero the diagonal entries
 
 theta0_a = 0 ; theta0_b = 2*pi; omega0_a = -.1 ; omega0_b = .1;
 
@@ -58,9 +59,7 @@ invpsi =kron(eye(num_Oscillator),Sigma/M);
 
 
 for ii=1:num_Oscillator
-
-xi_0(:,ii) = wrapTo2PiMSigma(m(ii)/sigma(ii)*theta_0(:,ii),m(ii),sigma(ii));
-
+    xi_0(:,ii) = wrapTo2PiMSigma(m(ii)/sigma(ii)*theta_0(:,ii),m(ii),sigma(ii));
 end
 
 eta_0 = (psi_lower_diag*omega_0')';
@@ -97,12 +96,10 @@ mean_prox_omega(1,:) = sum(omega_upd(:,:,1).*rho_theta_omega_upd(:,1))/sum(rho_t
 mean_prox_theta(1,:) = weighted_angle_mean(theta_upd(:,:,1),rho_theta_omega_upd(:,1));
 
 tic;
-for j=1:numSteps
-   
+for j=1:numSteps 
     
    [drift_j,GradU] = PowerDrift(xi_eta_upd(:,1:num_Oscillator,j),xi_eta_upd(:,num_Oscillator+1:dim,j),nSample);
-  
-   
+    
     % SDE update for state
     xi_eta_upd(:,:,j+1) = PowerEulerMaruyama(h,xi_eta_upd(:,:,j),drift_j,nSample,num_Oscillator);
     

### 2021-06-23 code/PowerDrift.m

@@ -13,20 +13,14 @@ GradU_2 = -repmat((diag(Pmech/Sigma))',nSample,1);
 % end
 
 for l = 1:nSample
-    
-for i = 1:num_Oscillator
-    
+    for i = 1:num_Oscillator    
     sum = 0;
-    for j =1:num_Oscillator 
-       
+    for j =1:num_Oscillator        
         sum = sum +( K(i,j)/sigma(i) )*sin((sigma(i)/m(i))*xi(l,i)-(sigma(j)/m(j))*xi(l,j)+phi(i,j));
-    end
-    
+    end    
     GradU_1(l,i) = sum;
+    end
 end
-
-end
-
 GradU_xi = GradU_1 + GradU_2;
 
 driftk =[eta, -GradF_eta-GradU_xi];

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \hline
-Description & Values\\
+Parameter description & Values\\
 \hline\hline
@@
 & \\
-coupling coefficient & $k_{ij}\in{\rm{Unif}}\left([0.7,1.2]\right)$\\
+coupling coefficient & $k_{ij}\begin{cases}
+=0 & \text{for}\; i=j,\\
+\in{\rm{Unif}}\left([0.7,1.2]\right) & \text{for}\; i\neq j.	
+\end{cases}$\\
 & \\
@@
 \begin{align*}
-\rho_{0}\equiv\rho(t=0,\theta(0),\omega(0)) = &\;{\rm{Unif}}\left([0,2\pi)^{n}\;\text{rad}\right)\\
+\rho_{0}\equiv\rho(t=0,\theta(0),\omega(0)) = &\;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right)\\
 	&\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
 \end{align*}
-and the stochastic parameters as in Table \ref{table:ParamUncertaintiesSynthetic}.
+and following \cite[Sec. 5]{dorfler2012synchronization}, randomly generate the parameters as in Table \ref{table:ParamUncertaintiesSynthetic}. These parameter
+
+For $N=1000$ random samples
 

### 2021-06-24 code/Gen_20_prox_simulation.m

@@ -1,14 +1,12 @@
 close all; clear; clc;
 %% Parameters
-
 beta = 2; % inverse temperature
-
 % golbal parameters 
 global num_Oscillator dim Gamma M Sigma K Pmech cc sigma m phi
 
 num_Oscillator = 20; dim  = 2*num_Oscillator; % dim = dimension of state space
 
-f0 = 60;
+f0 = 50; % Hz
 g_a = 20 ; g_b = 30; m_a = 2; m_b = 12 ; s_a = 1  ; s_b = 5;
 p_a = 0; p_b = 10 ;  k_a =.7 ; k_b = 1.2;
 
@@ -36,7 +34,6 @@ h = 1e-3;                         % time step
 numSteps= 1000;                    % number of steps k, in discretization t=kh
 cc = 1e7;
 %% propagate joint PDF
-
 % samples from initial joint PDF 
 theta_0 = (theta0_b-theta0_a)*rand(nSample,num_Oscillator) + theta0_a;
 
@@ -56,7 +53,6 @@ psi_lower_diag = psi_upper_diag;
 psi = kron(eye(num_Oscillator),M/Sigma);
 
 invpsi =kron(eye(num_Oscillator),Sigma/M);
-
 
 for ii=1:num_Oscillator
     xi_0(:,ii) = wrapTo2PiMSigma(m(ii)/sigma(ii)*theta_0(:,ii),m(ii),sigma(ii));
@@ -115,8 +111,6 @@ for j=1:numSteps
    mean_prox_omega(j+1,:) = sum(omega_upd(:,:,j+1).*rho_theta_omega_upd(:,j+1))/sum(rho_theta_omega_upd(:,j+1));  
     
    mean_prox_theta(j+1,:) = weighted_angle_mean(theta_upd(:,:,j+1),rho_theta_omega_upd(:,j+1));
-   
-
 end
 
 toc
@@ -130,10 +124,7 @@ mean_mc_theta = squeeze(mean_mc_theta);
 mean_mc = [mean_mc_theta;mean_mc_omega];
 mean_prox = [mean_prox_theta';mean_prox_omega'];
 
-norm_diff_mean_mc_vs_prox = sqrt(sum((mean_mc - mean_prox).^2,1))/sqrt(sum(mean_mc.^2,1));
-
-
-
+norm_diff_mean_mc_vs_prox = sqrt(sum((mean_mc - mean_prox).^2,1))./sqrt(sum(mean_mc.^2,1));
 
 %% plots
[…]

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \end{align*}
-and following \cite[Sec. 5]{dorfler2012synchronization}, randomly generate the parameters as in Table \ref{table:ParamUncertaintiesSynthetic}. These parameter
+and following \cite[Sec.
+5]{dorfler2012synchronization}, randomly generate the parameters as in Table \ref{table:ParamUncertaintiesSynthetic}.
+These parameter ranges are consistent with the same found in \cite{sauerpai1998,andersonBook1977,kundurBook1994}.
 
@@
 
+% === Power system book
+
+\bibitem{andersonBook1977}
+P.M.~Anderson, and A.A.~Fouad, \emph{Power System Control and Stability}, Iowa State University Press, Ames, IA, 1977.
+
+\bibitem{kundurBook1994}
+P.~Kundur, \emph{Power System Stability and Control}, McGraw-Hill, New York, 1994.
+
 \end{thebibliography}
 
-\begin{IEEEbiography}{Michael Shell}
-Biography text here.
-\end{IEEEbiography}
+%\begin{IEEEbiography}{Michael Shell}
+%Biography text here.
+%\end{IEEEbiography}
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 The $\varrho_{k}\left(\bm{\theta},\bm{\omega}\right)$ from \eqref{BringBackOriginalStateSpace} approximates $\rho(t,\bm{\theta},\bm{\omega})$ (the transient solution of \eqref{KineticFPK}) in the sense \eqref{ConvergenceGuarantee}.
+For completeness, the overall algorithm is {\red{given}}.
 
@@
 & \\
-tangent of phase shift & $\tan\varphi_{ij}\in{\rm{Unif}}\left([0,0.25]\right)$\\
+tangent of phase shift & $\tan\varphi_{ij}\begin{cases}
+=0 & \text{for}\; i=j\\
+\in{\rm{Unif}}\left([0,0.25]\right)& \text{for}\; i\neq j
+\end{cases}$\\
 & \\
@@
 coupling coefficient & $k_{ij}\begin{cases}
-=0 & \text{for}\; i=j,\\
-\in{\rm{Unif}}\left([0.7,1.2]\right) & \text{for}\; i\neq j.	
+=0 & \text{for}\; i=j\\
+\in{\rm{Unif}}\left([0.7,1.2]\right) & \text{for}\; i\neq j	
 \end{cases}$\\
@@
 We consider a power network with $n=20$ generators, and propagate the transient joint state PDFs supported over $\mathbb{T}^{20}\times\mathbb{R}^{20}$. We take the initial joint state PDF at $t=0$ as
-\begin{align*}
+\begin{align}
 \rho_{0}\equiv\rho(t=0,\theta(0),\omega(0)) = &\;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right)\\
 	&\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
-\end{align*}
+\label{rho0synthetic}	
+\end{align}
 and following \cite[Sec.
@@
 
-For $N=1000$ random samples
+For $N=1000$ random samples from the initial joint PDF \eqref{rho0synthetic}, we use Algorithm
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \ref{subsec:GenGradDescent}, it is not possible to express the right hand side of \eqref{NewPDE} as the Wasserstein gradient of the functional \eqref{LyapButNotProx}.
-The latter is defined as \cite[Ch.
+To see this, recall that the Wasserstein gradient is defined as \cite[Ch.
 8]{ambrosio2008gradient}
@@
 \end{tabular}
-\caption{Parameters used for the numerical simulation reported in Sec. \ref{subsec:NumSimSynthetic}. The indices $i,j\in\{1,\hdots,n\}$.}
+\caption{Parameters used for the numerical simulation in Sec.
+\ref{subsec:NumSimSynthetic}.
+The indices $i,j\in\{1,\hdots,n\}$ where the number of generators $n=20$ in Sec.
+\ref{subsec:NumSimSynthetic}.}
 \label{table:ParamUncertaintiesSynthetic}

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \subsection{Generalized Gradient Descent}\label{subsec:GenGradDescent}
-Let $\mathcal{P}_{2}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right)$ denote the manifold of all joint PDFs supported over the state space $\mathbb{T}^{n} \times \mathbb{R}^{n}$, with finite second raw moments.
+Let $\mathcal{P}_{2}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right)$ denote the manifold of joint PDFs supported over the state space $\mathbb{T}^{n} \times \mathbb{R}^{n}$, with finite second raw moments.
 Symbolically,
@@
 The $\varrho_{k}\left(\bm{\theta},\bm{\omega}\right)$ from \eqref{BringBackOriginalStateSpace} approximates $\rho(t,\bm{\theta},\bm{\omega})$ (the transient solution of \eqref{KineticFPK}) in the sense \eqref{ConvergenceGuarantee}.
-For completeness, the overall algorithm is {\red{given}}.
+For completeness, the overall proximal algorithm is {\red{given}}.
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 Likewise, the $n$ dimensional uniform probability distribution over $[a,b]^{n}$ is denoted as ${\rm{Unif}}\left([a,b]^{n}\right)$.
-The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute value, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map.
+The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute value, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\odot$ and $\oslash$ respectively denote the elementwise (Hadamard) product and division, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map.
 The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.
@@
 
+\begin{algorithm}[t]
+\caption{Proposed proximal algorithm for $\bm{\varrho}_{k-1}\mapsto\bm{\varrho}_{k}$}
+\label{KineticProxAlgorithm}
+\begin{algorithmic}[1]
+\Procedure{ProxRecursion}{$\bm{\varrho}_{k-1}, h, \varepsilon, N, \delta, L$}
+\For{$i=1$ to $N$}
+\State $\bm{\zeta}_{k-1}(i) \gets \exp\left(-\left(\bm{\eta}_{k-1}^{i}\right)^{\top}\bm{\eta}_{k-1}^{i}-1\right)$
+\For{$j=1$ to $N$}
+\State $\bm{C}_{k}(i,j) \gets \parallel\bm{\eta}_{k-1}^{j} - \bm{\eta}_{k}^{i}+ h\nabla U(\bm{\xi}_{k}^{i})\parallel_{2}^{2}$
+\State \qquad\quad $+12\bigg\|\dfrac{\bm{\xi}_{k-1}^{j}-\bm{\xi}_{k}^{i}}{h} - \dfrac{\bm{\eta}_{k-1}^{j} + \bm{\eta}_{k}^{i}}{2}\bigg\|_{2}^{2}$
+\EndFor
+\EndFor
+\State $\bm{\Gamma}_{k} \gets \exp\left(-\bm{C}_{k}/2\varepsilon\right)$ \Comment{elementwise exponential}
+\State $\bm{z}_{0} \gets {\rm{rand}}_{N\times 1}$ \Comment{random vector of size $N\times 1$}
+\State $\bm{z} \gets \left[\bm{z}_{0}, \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
+\State $\bm{y} \gets \left[\bm{\varrho}_{k-1}\oslash\left(\bm{\Gamma}_{k}\bm{z}_{0}\right), \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
+\State $\ell = 1$
+\While{$\ell\leq L$}
+\State $\bm{z}(:,\ell+1) \gets \left(\bm{\zeta}_{k-1}\oslash\left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)\right)^{\frac{1}{1+2\varepsilon/h}}$
+\State $\bm{y}(:,\ell+1) \gets \bm{\varrho}_{k-1} \oslash \left(\bm{\Gamma}_{k}\bm{z}(:,\ell+1)\right)$
+\If{$\|\bm{y}(:,\ell+1)-\bm{y}(:,\ell)\|_{2}<\delta$ \& $\|\bm{z}(:,\ell+1)-\bm{z}(:,\ell)\|_{2}<\delta$} \Comment{error within tolerance}
+\State break
+\Else
+\State $\ell \gets \ell + 1$
+\EndIf
+\EndWhile\\
+\Return $\varrho_{k} \gets \bm{z}(:,\ell) \odot \left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)$ \Comment{proximal update}
+\EndProcedure	
+\end{algorithmic}
+\end{algorithm}
+
 Finally, we transform the proximal updates back to the $\left(\bm{\theta},\bm{\omega}\right)^{\top}$ state space via the pushforward $\bm{\Psi}^{-1}$ as
@@
 The $\varrho_{k}\left(\bm{\theta},\bm{\omega}\right)$ from \eqref{BringBackOriginalStateSpace} approximates $\rho(t,\bm{\theta},\bm{\omega})$ (the transient solution of \eqref{KineticFPK}) in the sense \eqref{ConvergenceGuarantee}.
-For completeness, the overall proximal algorithm is {\red{given}}.
+
+For completeness, the overall proximal algorithm is {\red{given}}. Its convergence guarantees can be found in \cite[Sec. III.C]{caluya2019TAC}.
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 Following \cite{caluya2019TAC}, we solve \eqref{KineticProx} by recursively updating the probability-weighted scattered point clouds $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
-$$\bm{x}_{k}^{i}\equiv\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N, \quad k\in\mathbb{N}.$$
+$$\tilde{\bm{x}}_{k}^{i}:=\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N, \quad k\in\mathbb{N}.$$
 Thus, $\tilde{\varrho}_{k}^{i}$ is the joint PDF value obtained from \eqref{KineticProx} at $\bm{x}_{k}^{i}$, the $i$th state sample at the $k$th time step.
@@
 In the numerical simulations reported in Sec.
-\ref{sec:NumericalSimulations}, the states $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}).
+\ref{sec:NumericalSimulations}, the states $\{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}).
 If one wishes so, the Euler-Maruyama scheme in Fig.
@@
 \begin{algorithm}[t]
-\caption{Proposed proximal algorithm for $\bm{\varrho}_{k-1}\mapsto\bm{\varrho}_{k}$}
-\label{KineticProxAlgorithm}
+\caption{Proposed proximal algorithm for $\tilde{\bm{\varrho}}_{k-1}\mapsto\tilde{\bm{\varrho}}_{k}$}
+\label{algo:KineticProx}
 \begin{algorithmic}[1]
-\Procedure{ProxRecursion}{$\bm{\varrho}_{k-1}, h, \varepsilon, N, \delta, L$}
+\Procedure{ProxRecursion}{$\bm{\varrho}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, L$}
 \For{$i=1$ to $N$}
@@
 \State $\bm{z} \gets \left[\bm{z}_{0}, \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
-\State $\bm{y} \gets \left[\bm{\varrho}_{k-1}\oslash\left(\bm{\Gamma}_{k}\bm{z}_{0}\right), \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
+\State $\bm{y} \gets \left[\tilde{\bm{\varrho}}_{k-1}\oslash\left(\bm{\Gamma}_{k}\bm{z}_{0}\right), \bm{0}_{N\times(L-1)}\right]$ \Comment{initialize}
 \State $\ell = 1$
@@
 \State $\bm{z}(:,\ell+1) \gets \left(\bm{\zeta}_{k-1}\oslash\left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)\right)^{\frac{1}{1+2\varepsilon/h}}$
-\State $\bm{y}(:,\ell+1) \gets \bm{\varrho}_{k-1} \oslash \left(\bm{\Gamma}_{k}\bm{z}(:,\ell+1)\right)$
+\State $\bm{y}(:,\ell+1) \gets \tilde{\bm{\varrho}}_{k-1} \oslash \left(\bm{\Gamma}_{k}\bm{z}(:,\ell+1)\right)$
 \If{$\|\bm{y}(:,\ell+1)-\bm{y}(:,\ell)\|_{2}<\delta$ \& $\|\bm{z}(:,\ell+1)-\bm{z}(:,\ell)\|_{2}<\delta$} \Comment{error within tolerance}
@@
 \EndWhile\\
-\Return $\varrho_{k} \gets \bm{z}(:,\ell) \odot \left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)$ \Comment{proximal update}
+\Return $\tilde{\bm{\varrho}}_{k} \gets \bm{z}(:,\ell) \odot \left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)$ \Comment{proximal update}
 \EndProcedure	
@@
 
-For completeness, the overall proximal algorithm is {\red{given}}. Its convergence guarantees can be found in \cite[Sec. III.C]{caluya2019TAC}.
+For completeness, the proximal algorithm for updating $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}\mapsto\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ is outlined in Algorithm \ref{algo:KineticProx}.
+Its convergence guarantees can be found in \cite[Sec.
+III.C]{caluya2019TAC}.
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \section{Proximal Algorithm}\label{sec:ProxAlgo}
-Following \cite{caluya2019TAC}, we solve \eqref{KineticProx} by recursively updating the probability-weighted scattered point clouds $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
+Following \cite{caluya2019TAC}, we solve \eqref{KineticProx} by recursively updating the probability-weighted scattered point clouds $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where
 $$\tilde{\bm{x}}_{k}^{i}:=\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N, \quad k\in\mathbb{N}.$$
-Thus, $\tilde{\varrho}_{k}^{i}$ is the joint PDF value obtained from \eqref{KineticProx} at $\bm{x}_{k}^{i}$, the $i$th state sample at the $k$th time step.
+Thus, $\tilde{\varrho}_{k}^{i}$ is the joint PDF value obtained from \eqref{KineticProx} at $\tilde{\bm{x}}_{k}^{i}$, the $i$th state sample at the $k$th time step.
 The schematic of the algorithm is shown in Fig.

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

--- 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 \includegraphics[width=.8\linewidth]{BlockDiagm.pdf}
-\caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
-The states $\{\bm{x}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}) as detailed in Sec.
+\caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
+The states $\{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}) as detailed in Sec.
 \ref{sec:ProxAlgo}.}}
@@
 \section{Proximal Algorithm}\label{sec:ProxAlgo}
-Following \cite{caluya2019TAC}, we solve \eqref{KineticProx} by recursively updating the probability-weighted scattered point clouds $\{\bm{x}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
+Following \cite{caluya2019TAC}, we solve \eqref{KineticProx} by recursively updating the probability-weighted scattered point clouds $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where
 $$\tilde{\bm{x}}_{k}^{i}:=\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N, \quad k\in\mathbb{N}.$$
-Thus, $\tilde{\varrho}_{k}^{i}$ is the joint PDF value obtained from \eqref{KineticProx} at $\bm{x}_{k}^{i}$, the $i$th state sample at the $k$th time step.
+Thus, $\tilde{\varrho}_{k}^{i}$ is the joint PDF value obtained from \eqref{KineticProx} at $\tilde{\bm{x}}_{k}^{i}$, the $i$th (transformed) state sample at the $k$th time step.
 The schematic of the algorithm is shown in Fig.

### 2021-06-24 code/Gen_20_prox_simulation.m
updating Section V

@@ -16,7 +16,9 @@ tp_b = 0.25;
 gamma = ((g_a-g_b)*rand(num_Oscillator,1) + g_a)/(2*pi*f0);
 m = ((m_b-m_a)*rand(num_Oscillator,1) + m_a)/(2*pi*f0);
 sigma = (s_b-s_a)*rand(num_Oscillator,1) + s_a;
+
 phi = atan((tp_b-tp_a)*rand(num_Oscillator) + tp_a);
+phi = phi - diag(diag(phi)); % zero the diagonal entries
 
 pmech = (p_b-p_a)*rand(num_Oscillator,1) + p_b;
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 For completeness, the proximal algorithm for updating $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}\mapsto\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ is outlined in Algorithm \ref{algo:KineticProx}.
-Its convergence guarantees can be found in \cite[Sec.
-III.C]{caluya2019TAC}.
+As shown in Fig.
+\ref{fig:BlockDiagm}, this algorithm takes the pre and post update state samples
+\[\{\tilde{\bm{x}}_{k-1}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k-1}^{i},\bm{\eta}_{k-1}^{i}\right)\}_{i=1}^{N}, \; \{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)\}_{i=1}^{N},\]
+and $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}$ as inputs. The Algorithm \ref{algo:KineticProx} has additional
+
+Its convergence guarantees can be found in \cite[Sec. III.C]{caluya2019TAC}.
 
@@
 2008, and Ph.D. from Texas A\&M University in 2014, all in Aerospace
-Engineering. His research interests are in stochastic systems, control
-and optimization with application focus on large scale cyber-physical
+Engineering. His research interests are in stochastic systems, control, learning and optimization with application focus on large scale cyber-physical
 systems.

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 As shown in Fig.
-\ref{fig:BlockDiagm}, this algorithm takes the pre and post update state samples
+\ref{fig:BlockDiagm}, at a conceptual level, this algorithm takes the pre and post update state samples
 \[\{\tilde{\bm{x}}_{k-1}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k-1}^{i},\bm{\eta}_{k-1}^{i}\right)\}_{i=1}^{N}, \; \{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)\}_{i=1}^{N},\]
-and $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}$ as inputs. The Algorithm \ref{algo:KineticProx} has additional
+and $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}$ as inputs.
+The Algorithm \ref{algo:KineticProx} needs additional inputs $h,\varepsilon,N,\delta,L$, which are parameters internal to the algorithm.
+Specifically, $h$ is the time-step size
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

--- 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 Thus, $\tilde{\varrho}_{k}^{i}$ is the joint PDF value obtained from \eqref{KineticProx} at $\tilde{\bm{x}}_{k}^{i}$, the $i$th (transformed) state sample at the $k$th time step.
-The schematic of the algorithm is shown in Fig.
+The high level schematic of the algorithm is shown in Fig.
 \ref{fig:BlockDiagm}.
@@
 As shown in Fig.
-\ref{fig:BlockDiagm}, this algorithm takes the pre and post update state samples
+\ref{fig:BlockDiagm}, this algorithm, at a conceptual level, takes the pre and post update state samples
 \[\{\tilde{\bm{x}}_{k-1}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k-1}^{i},\bm{\eta}_{k-1}^{i}\right)\}_{i=1}^{N}, \; \{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)\}_{i=1}^{N},\]
-and $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}$ as inputs. The Algorithm \ref{algo:KineticProx} has additional
+and $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}$ as inputs.
+The Algorithm \ref{algo:KineticProx} also needs input parameters $h,\varepsilon,N,\delta,L$.
+Specifically, $h$ is the time-step size, $\varepsilon$ is
 
@@
 These parameter ranges are consistent with the same found in \cite{sauerpai1998,andersonBook1977,kundurBook1994}.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\linewidth]{CompTimeSyntheticPowerSystem.png}
+\caption{\small{The computational time for propagating the transient joint PDFs over $\mathbb{T}^{20}\times\mathbb{R}^{20}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
+\label{fig:CompTimeSynthetic}
+\end{figure}
 

### 2021-06-24 code/Gen_20_prox_simulation.m

@@ -35,6 +35,7 @@ epsilon = 0.5;                      % regularizing coefficient
 h = 1e-3;                         % time step
 numSteps= 1000;                    % number of steps k, in discretization t=kh
 cc = 1e7;
+t_vec = h*(1:1:numSteps);
 %% propagate joint PDF
 % samples from initial joint PDF 
 theta_0 = (theta0_b-theta0_a)*rand(nSample,num_Oscillator) + theta0_a;
@@ -134,10 +135,15 @@ set(groot,'defaulttextinterpreter','latex');
 set(groot,'defaultLegendInterpreter','latex');
 
 figure(1)
-semilogy(comptime, 'LineWidth', 2)
+semilogy(t_vec, comptime, 'LineWidth', 2)
 set(gca,'FontSize',30)
-xlabel('Physical time $t=kh$','FontSize',30)
-ylabel('Computational time','FontSize',30)
+xlabel('Physical time $t=kh$ [s]','FontSize',30)
+ylabel('Computational time [s]','FontSize',30)
+ylim([1e-3 1.2e-2])
+YTick = [2e-3 8e-3 2e-2];
+YTickLabels = cellstr(num2str(round(log10(YTick(:))), '10^%d'));
+grid on
+
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@ -171,7 +177,7 @@ end
 legend('Mean MC','Mean Proximal')
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 figure(4)
-semilogy(norm_diff_mean_mc_vs_prox(5:end),'-k','Linewidth',2)
+semilogy(t_vec(4:end), norm_diff_mean_mc_vs_prox(5:end),'-k','Linewidth',2)
 set(gca,'FontSize',30)
-xlabel('Physical time $t=kh$','FontSize',30)
+xlabel('Physical time $t=kh$ [s]','FontSize',30)
 ylabel('Realtive error $\frac{\|\mu_{\rm{MC}}-\mu_{\rm{Prox}}\|_{2}}{\|\mu_{\rm{MC}}\|_{2}}$','FontSize',30,'interpreter','latex')

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \centering
-\includegraphics[width=0.9\linewidth]{CompTimeSyntheticPowerSystem.png}
-\caption{\small{The computational time for propagating the transient joint PDFs over $\mathbb{T}^{20}\times\mathbb{R}^{20}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
+\includegraphics[width=0.9\linewidth]{ComputationalTimeSynthetic.png}
+\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{20}\times\mathbb{R}^{20}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
 \label{fig:CompTimeSynthetic}

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py

@@ -1,0 +1,90 @@
+import numpy as np
+from matplotlib import rc
+import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+from mpl_toolkits.axes_grid1 import ImageGrid
+from itertools import cycle
+from pylab import *
+import glob
+from pylab import rcParams
+
+#====================================================
+# Make plots beautiful
+#====================================================
+
+pts_per_inch = 72.27
+# write "\the\textwidth" (or "\showthe\columnwidth" for a 2 collumn text)
+text_width_in_pts = 300.0
+# inside a figure environment in latex, the result will be on the
+# dvi/pdf next to the figure. See url above.
+text_width_in_inches = text_width_in_pts / pts_per_inch
+# figure.png or figure.eps will be intentionally larger, because it is prettier
+inverse_latex_scale = 2
+fig_proportion = (3.0 / 3.0)
+csize = inverse_latex_scale * fig_proportion * text_width_in_inches
+# always 1.0 on the first argument
+fig_size = (1.0 * csize, 0.85 * csize)
+# find out the fontsize of your latex text, and put it here
+text_size = inverse_latex_scale * 9
+label_size = inverse_latex_scale * 10
+tick_size = inverse_latex_scale * 8
+# learn how to configure:
+# http://matplotlib.sourceforge.net/users/customizing.html
+params = {'backend': 'ps',
+          'axes.labelsize': 16,
+          'legend.fontsize': tick_size,
+          'legend.handlelength': 2.5,
+          'legend.borderaxespad': 0,
+          'axes.labelsize': label_size,
+          'xtick.labelsize': tick_size,
+          'ytick.labelsize': tick_size,
+          'font.family': 'serif',
+          'font.size': text_size,
+          'font.serif': ['Computer Modern Roman'],
+          'ps.usedistiller': 'xpdf',
+          'text.usetex': True,
+          'figure.figsize': fig_size,
+          # include here any neede package for latex
+          'text.latex.preamble': [r'\usepackage{amsmath}'],
+          }
+plt.rcParams.update(params)
+fig = plt.figure(1, figsize=fig_size)  # figsize accepts only inches.
+# fig.subplots_adjust(left=0.04, right=0.98, top=0.93, bottom=0.15,
[…]

### 2021-06-24 code/Gen_20_prox_simulation.m

@@ -181,3 +181,14 @@ semilogy(t_vec(4:end), norm_diff_mean_mc_vs_prox(5:end),'-k','Linewidth',2)
 set(gca,'FontSize',30)
 xlabel('Physical time $t=kh$ [s]','FontSize',30)
 ylabel('Realtive error $\frac{\|\mu_{\rm{MC}}-\mu_{\rm{Prox}}\|_{2}}{\|\mu_{\rm{MC}}\|_{2}}$','FontSize',30,'interpreter','latex')
+
+
+%% save simulation data
+textfilename = 'TimeSynthetic.txt';
+dlmwrite(textfilename, t_vec,'delimiter','\t','precision','%.64f');
+
+textfilename = 'ComptimeSythetic.txt';
+dlmwrite(textfilename, comptime,'delimiter','\t','precision','%.64f');
+
+textfilename = 'RelErrMeanVectorMCvsProxSythetic.txt';
+dlmwrite(textfilename, norm_diff_mean_mc_vs_prox,'delimiter','\t','precision','%.64f');

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 
-\begin{figure}
+\begin{figure}[htpb]
 \centering
@@
 \caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{20}\times\mathbb{R}^{20}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
+\label{fig:CompTimeSynthetic}
+\end{figure}
+\begin{figure}[htpb]
+\centering
+\includegraphics[width=0.9\linewidth]{RelativeErrorMeanMCVersusProxSynthetic.png}
+\caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{\text{MC}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$, and the mean $\bm{\mu}_{\text{Prox}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec.
+\ref{subsec:NumSimSynthetic}.}}
 \label{fig:CompTimeSynthetic}

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py

@@ -53,7 +53,7 @@ fig = plt.figure(1, figsize=fig_size)  # figsize accepts only inches.
 #                     hspace=0.05, wspace=0.02)
 
 # =============================================================================
-# Computational time
+# Computational time plot 
 # =============================================================================
 Time_Synthetic = np.loadtxt("TimeSynthetic.txt")
 CompTime_Synthetic = np.loadtxt("ComptimeSythetic.txt")
@@ -82,9 +82,34 @@ fig.set_size_inches(10.2, 6.2)
 
 plt.savefig('ComputationalTimeSynthetic.png', dpi=400)
 
+# =============================================================================
+# Relative error in mean vector: MC versus Prox plot
+# =============================================================================
+RelErrMeanVectorMCvsProx_Synthetic = np.loadtxt("RelErrMeanVectorMCvsProxSythetic.txt")
+
+fig = plt.figure()
+ax = fig.add_subplot(1, 1, 1)
+
+ax.semilogy(Time_Synthetic[3:-1], RelErrMeanVectorMCvsProx_Synthetic[4:-1],'-', color='k', lw=1.5)
+
+# ax.set_ylim(1*10**-3, 1.2*10**-2)
+
+ax.tick_params(direction='in',which='both')
+ # axx.xaxis.tick_top()
+
+ax.grid(True,which="both",ls="-", color='0.75')
+# axx2.grid(True,which="both",ls="-", color='0.75')
+ax.tick_params(axis='both', labelsize=18)
 
 
+ax.set_ylabel(r"Realtive error $\frac{\|\boldsymbol{\mu}_{\rm{MC}}-\boldsymbol{\mu}_{\rm{Prox}}\|_{2}}{\|\boldsymbol{\mu}_{\rm{MC}}\|_{2}}$")
+ax.set_xlabel(r"Physical time $t=kh$ [s]")
+# # axx.yaxis.set_label_coords(-0.125,-0.05)
 
+# axx.legend(markerscale=1.5, numpoints=1,  ncol=1, bbox_to_anchor=(1.005, 1), frameon=False, prop={'size': 10.5})
+fig.set_size_inches(10.2, 6.2)
+
+plt.savefig('RelativeErrorMeanMCVersusProxSynthetic.png', dpi=400)
 
 
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \begin{algorithmic}[1]
-\Procedure{ProxRecursion}{$\bm{\varrho}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, L$}
+\Procedure{Prox}{$\bm{\varrho}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, L$}
 \For{$i=1$ to $N$}
@@
 
-For completeness, the proximal algorithm for updating $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}\mapsto\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ is outlined in Algorithm \ref{algo:KineticProx}.
+For completeness, the algorithm \textproc{Prox} for updating $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}\mapsto\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ is outlined in Algorithm \ref{algo:KineticProx}.
 As shown in Fig.
@@
 \[\{\tilde{\bm{x}}_{k-1}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k-1}^{i},\bm{\eta}_{k-1}^{i}\right)\}_{i=1}^{N}, \; \{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)\}_{i=1}^{N},\]
-and $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}$ as inputs.
+and $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}$ as inputs, and outputs the proximal updates $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
+For each $k\in\mathbb{N}$, the updated probability weighted point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are then brought back to the original state space via $\bm{\Psi}^{-1}$ as $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$, as explained earlier.
+
 The Algorithm \ref{algo:KineticProx} also needs input parameters $h,\varepsilon,N,\delta,L$.
-Specifically, $h$ is the time-step size, $\varepsilon$ is
-
-Its convergence guarantees can be found in \cite[Sec. III.C]{caluya2019TAC}.
+Specifically, $h$ is the time-step size, $\varepsilon$ is an entropic regularization weight internal to the \textproc{Prox} algorithm, and $N$ is the number of samples.
+The parameters $\delta$ and $L$ codify the numerical tolerance and maximum number of iterations, respectively, for the while loop in Algorithm \ref{algo:KineticProx}.
+Its convergence guarantees can be found in \cite[Sec.
+III.C]{caluya2019TAC}.
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \subsection{Synthetic Test System}\label{subsec:NumSimSynthetic}
-We consider a power network with $n=20$ generators, and propagate the transient joint state PDFs supported over $\mathbb{T}^{20}\times\mathbb{R}^{20}$. We take the initial joint state PDF at $t=0$ as
+We next consider a power network with $n=20$ generators, and propagate the transient joint state PDFs supported over $\mathbb{T}^{20}\times\mathbb{R}^{20}$.
+We take the initial joint state PDF at $t=0$ as
 \begin{align}

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 
-Finally, we transform the proximal updates back to the $\left(\bm{\theta},\bm{\omega}\right)^{\top}$ state space via the pushforward $\bm{\Psi}^{-1}$ as
+Finally, we transform the proximal updates back to the $\bm{x}\equiv\left(\bm{\theta},\bm{\omega}\right)^{\top}$ state space via the pushforward $\bm{\Psi}^{-1}$ as
 \begin{align}
@@
 \begin{align}
-\rho_{0}\equiv\rho(t=0,\theta(0),\omega(0)) = &\;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right)\\
+\rho_{0}\equiv\rho(t=0,\theta(0),\omega(0)) = &\;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right)\nonumber\\
 	&\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
@@
 \centering
+\vspace*{-0.1in}
 \includegraphics[width=0.9\linewidth]{ComputationalTimeSynthetic.png}
@@
 \centering
+\vspace*{-0.1in}
 \includegraphics[width=0.9\linewidth]{RelativeErrorMeanMCVersusProxSynthetic.png}
@@
 
-For $N=1000$ random samples from the initial joint PDF \eqref{rho0synthetic}, we use Algorithm
+With $N=1000$ random samples from the initial joint PDF \eqref{rho0synthetic}, we use the Euler-Maruyama scheme together with the Algorithm \ref{algo:KineticProx} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
+We used
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \ref{subsec:NumSimSynthetic}.}}
-\label{fig:CompTimeSynthetic}
+\label{fig:RelErrSynthetic}
 \end{figure}
 
-With $N=1000$ random samples from the initial joint PDF \eqref{rho0synthetic}, we use the Euler-Maruyama scheme together with the Algorithm \ref{algo:KineticProx} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
-We used
+With $N=1000$ random samples from the initial joint PDF \eqref{rho0synthetic}, and with the aforesaid randomly generated parameters, we employed the Euler-Maruyama scheme together with the Algorithm \ref{algo:KineticProx} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
+We used $h=10^{-3}$, $\varepsilon = 0.05$, $\delta = 10^{-3}$, $L = 300$.
+
+Fig. \ref{fig:CompTimeSynthetic} highlights that
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \includegraphics[width=0.9\linewidth]{RelativeErrorMeanMCVersusProxSynthetic.png}
-\caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{\text{MC}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$, and the mean $\bm{\mu}_{\text{Prox}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec.
+\caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$, and the mean $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec.
 \ref{subsec:NumSimSynthetic}.}}
@@
 
-With $N=1000$ random samples from the initial joint PDF \eqref{rho0synthetic}, and with the aforesaid randomly generated parameters, we employed the Euler-Maruyama scheme together with the Algorithm \ref{algo:KineticProx} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
+With $N=1000$ random samples from the initial joint PDF \eqref{rho0synthetic}, and with the aforesaid randomly generated parameters, we employed the Euler-Maruyama scheme together with Algorithm \ref{algo:KineticProx} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
 We used $h=10^{-3}$, $\varepsilon = 0.05$, $\delta = 10^{-3}$, $L = 300$.
 
-Fig. \ref{fig:CompTimeSynthetic} highlights that
+Fig.
+\ref{fig:CompTimeSynthetic} highlights that for computing high dimensional transient joint state PDFs as in here, the proposed variational framework enjoys remarkably fast computational time.
+To depict the numerical accuracy, Fig.
+\ref{fig:RelErrSynthetic} plots the (time-varying) relative error between the empirical (i.e., Monte Carlo estimate) mean vector $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$, and the ``proximal mean" vector $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$.
+The latter was computed using the proximal updates $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
 

### 2021-06-24 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py

@@ -102,7 +102,7 @@ ax.grid(True,which="both",ls="-", color='0.75')
 ax.tick_params(axis='both', labelsize=18)
 
 
-ax.set_ylabel(r"Realtive error $\frac{\|\boldsymbol{\mu}_{\rm{MC}}-\boldsymbol{\mu}_{\rm{Prox}}\|_{2}}{\|\boldsymbol{\mu}_{\rm{MC}}\|_{2}}$")
+ax.set_ylabel(r"Realtive error $\frac{\|\boldsymbol{\mu}_{k}^{\rm{MC}}-\boldsymbol{\mu}_{k}^{\rm{Prox}}\|_{2}}{\|\boldsymbol{\mu}_{k}^{\rm{MC}}\|_{2}}$")
 ax.set_xlabel(r"Physical time $t=kh$ [s]")
 # # axx.yaxis.set_label_coords(-0.125,-0.05)
 

### 2021-06-25 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 The rest of this paper is structured as follows.
-Section \ref{sec:models} details the power system dynamics models at the microscopic or trajectory level, as well as at the macroscopic or ensemble level.
+Section \ref{sec:models} details the power system dynamics models at the microscopic or trajectory level (Sec.
+\ref{subsec:samplepathdynamics}) as well as at the macroscopic or statistical ensemble level (Sec.
+\ref{subsec:macroscopic}).
+The proposed idea of realizing the flow of the joint state PDFs subject to the macroscopic power system dynamics via infinite dimensional proximal recursions, is explained in Section \ref{sec:prox}.
+Section \ref{sec:ProxAlgo} elucidates the corresponding proximal algorithm that enables the computation of the transient joint state PDFs via weighted point cloud evolution.
+Numerical simulations are reported in Section \ref{sec:NumericalSimulations}.
+Section \ref{sec:conclusion} concludes the paper.
 
@@
 and $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}$ as inputs, and outputs the proximal updates $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
-For each $k\in\mathbb{N}$, the updated probability weighted point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are then brought back to the original state space via $\bm{\Psi}^{-1}$ as $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$, as explained earlier.
+For each $k\in\mathbb{N}$, the updated probability weighted point clouds $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are then brought back to the original state space via $\bm{\Psi}^{-1}$ as $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$, as explained earlier.
 
@@
 \vspace*{-0.1in}
-\includegraphics[width=0.9\linewidth]{ComputationalTimeSynthetic.png}
+\includegraphics[width=0.95\linewidth]{ComputationalTimeSynthetic.png}
 \caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{20}\times\mathbb{R}^{20}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
@@
 \vspace*{-0.1in}
-\includegraphics[width=0.9\linewidth]{RelativeErrorMeanMCVersusProxSynthetic.png}
+\includegraphics[width=0.95\linewidth]{RelativeErrorMeanMCVersusProxSynthetic.png}
 \caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$, and the mean $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec.
@@
 \section{Conclusion}\label{sec:conclusion}
-The conclusion goes here.
+{\red{The conclusion goes here.}}
 

### 2021-06-25 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 We mention here that (\ref{ItoSDEvectorlevel}) has been used in \cite{ju2018analytical,ju2018stochastic} for uncertainty propagation via stochastic averaging approximation where the univariate energy PDF was proposed as a ``proxy" for the entire joint PDF.
-Most relevant to our approach in the power systems literature is the work in \cite{wang2013fokker}, which indeed voiced the need for computing the transient joint PDFs but only dealt with the single-machine-infinite-bus case -- simplest instance of (\ref{ItoSDEvectorlevel}).
+Most relevant to our approach in the power systems literature is the work in \cite{wang2013fokker}, which indeed voiced the need for computing the transient joint PDFs but only dealt with the single-machine-infinite-bus case -- simplest ($n=1$) instance of (\ref{ItoSDEvectorlevel}).
 The resulting bivariate Fokker-Planck PDE in \cite{wang2013fokker} was solved via finite element discretization, and revealed rich stochastic dynamics and nontrivial transient stability aspects even in this simple case.
@@
 &\\
-frequency & $f_{0} = 60$ Hz\\
+nominal frequency & $f_{0} = 60$ Hz\\
 &\\

### 2021-06-25 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Section V

@@
 \subsection{Synthetic Test System}\label{subsec:NumSimSynthetic}
-We next consider a power network with $n=20$ generators, and propagate the transient joint state PDFs supported over $\mathbb{T}^{20}\times\mathbb{R}^{20}$.
+We next consider a power network with $n=20$ generators, and propagate the transient joint state PDFs supported over the 40 dimensional state space $\mathbb{T}^{20}\times\mathbb{R}^{20}$.
 We take the initial joint state PDF at $t=0$ as

### 2021-06-25 code/IEEE14_prox_simulation.m

@@ -1,4 +1,4 @@
-close all; clear; clc;
+close all; clear all; clc;
 %% Parameters
 
 beta = 2; % inverse temperature
@@ -35,149 +35,149 @@ epsilon = 0.5;                      % regularizing coefficient
 h = 1e-3;                         % time step
 numSteps= 1e3;                    % number of steps k, in discretization t=kh
 cc = 1e7;
-%% popagate joint PDF
-
-% samples from initial joint PDF 
-theta_0 = (theta0_b-theta0_a)*rand(nSample,num_Oscillator) + theta0_a;
-
-rho_theta_0 = ones(nSample,1)*(1/(theta0_b-theta0_a))^num_Oscillator;
-
-omega_0 =(omega0_b-omega0_a)*rand(nSample,num_Oscillator) + omega0_a;
-
-% joint PDF values at the initial samples
-rho_omega_0 = ones(nSample,1)*(1/(omega0_b-omega0_a))^num_Oscillator;
-
-rho_theta_omega_0 = rho_omega_0.*rho_theta_0;
-
-psi_upper_diag = M/Sigma;
-
-psi_lower_diag = psi_upper_diag;
-
-psi = kron(eye(num_Oscillator),M/Sigma);
-
-invpsi =kron(eye(num_Oscillator),Sigma/M);
-
-sumcov_prox = zeros(dim,dim);
-    
-sumcov_mc = zeros(dim,dim);
-
-for ii=1:num_Oscillator
-
-xi_0(:,ii) = wrapTo2PiMSigma(m(ii)/sigma(ii)*theta_0(:,ii),m(ii),sigma(ii));
-
-end
-
-eta_0 = (psi_lower_diag*omega_0')';
-
-xi_eta_0 = [xi_0,eta_0];
-
-rho_xi_eta_0 = rho_theta_omega_0/(prod(m./sigma));
-
-% stores all the updated (states from the governing SDE
-xi_eta_upd = zeros(nSample,dim,numSteps+1);
-
-theta_upd = zeros(nSample,num_Oscillator,numSteps+1);
-
-omega_upd = zeros(nSample,num_Oscillator,numSteps+1);
-
-theta_omega_upd = zeros(nSample,dim,numSteps+1);
-
-% sets initial state
[…]

### 2021-06-25 code/phase_shift_matrix.m

@@ -1,21 +1,13 @@
 function phi = phase_shift_matrix(Y)
 
-m = size(Y);
-nc = m(2);
+nc = size(Y,2);
 
 for i = 1:nc
-    
-   
   for j = 1:nc
-   
      if i == j
-         
          phi(i,j) = 0;
-         
      else 
          phi(i,j) = -atan2(real(Y(i,j)),imag(Y(i,j)));
      end
   end
-    
-    
 end

### 2021-06-26 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 
-\author{Abhishek~Halder,~\IEEEmembership{Member,~IEEE,} Authors}
+\author{Abhishek~Halder,~\IEEEmembership{Member,~IEEE,} 
 %\author{Abhishek~Halder,~\IEEEmembership{Member,~IEEE,}
-%Xinbo~Geng,~\IEEEmembership{Member,~IEEE,}
-%        Kenneth~F. Caluya,
-%        and~Pegah~Ojaghi% <-this % stops a space
-%\thanks{A.
-Halder and K.F.
-Caluya are with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder,kcaluya\}@ucsc.edu}}.}% <-this % stops a space
-%\thanks{X.
-Halder is with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder\}@ucsc.edu}}.
-This research was partially supported by the NSF award 1923278.}}% <-this % stops a space
+Xinbo~Geng,~\IEEEmembership{Member,~IEEE,}
+        Kenneth~F. Caluya,
+        and~Pegah~Ojaghi% <-this % stops a space
+\thanks{A. Halder and K.F. Caluya are with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder,kcaluya\}@ucsc.edu}}.}% <-this % stops a space
+\thanks{X. Geng is with {\red{Edit next}} the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{ahalder@ucsc.edu}}.} %
+\thanks{P. Ojaghi is with the Department of Computer Science and Engineering, University of California, Santa Cruz, CA 95064, USA, {\texttt{pojaghi@ucsc.edu}}.}%
+\thanks{This research was partially supported by the NSF award 1923278.}}% <-this % stops a space
 %\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}
@@
 The proposed algorithm takes advantage of the exact nonlinearity structures in the trajectory-level dynamics of the networked power systems, and is nonparametric.
-Lifting the dynamics to the space of probability measures allows us to design a scalable algorithm that obviates gridding the underlying high dimensional state space or function approximation.
-The proximal recursion implements a generalized gradient flow, and evolves probability weighted scattered point clouds.
-We provide the theoretical details, convergence guarantees, and numerical examples on realistic test systems.
+Lifting the dynamics to the space of probability measures allows us to design a scalable algorithm that obviates gridding the underlying high dimensional state space which is computationally prohibitive.
+The proximal recursion implements a generalized infinite dimensional gradient flow, and evolves probability weighted scattered point clouds.
+We clarify the theoretical nuances and algorithmic details specific for the power system nonlinearities, and provide illustrative numerical examples.
 \end{abstract}
@@
 The potential (\ref{potential}) has a natural energy function interpretation and can also be motivated by a mechanical mass-spring-damper analogy \cite{dorfler2013synchronization,ishizaki2018}.
-The function $V$ is continuously differentiable, bounded below, and goes to $+\infty$ as $\|\bm{\theta}\|_{2}\rightarrow +\infty$.
+%The function $V$ is continuously differentiable, bounded below, and goes to $+\infty$ as $\|\bm{\theta}\|_{2}\rightarrow +\infty$.
 %In the development below, an important role will be played by the ``Hamiltonian-like" function
@@
 \ref{fig:CompTimeSynthetic} highlights that for computing high dimensional transient joint state PDFs as in here, the proposed variational framework enjoys remarkably fast computational time.
+Notice that constructing a histogram as an estimator of the joint state PDF by making a grid over the 40 dimensional state space, is computationally prohibitive.
+
 To depict the numerical accuracy, Fig.
@@
 \end{thebibliography}
-
-%\begin{IEEEbiography}{Michael Shell}
-%Biography text here.
-%\end{IEEEbiography}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{IEEEbiography}{Xinbo Geng}
+Biography text here.
+\end{IEEEbiography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{IEEEbiography}[{\includegraphics[height=1.25in]{kenneth.jpg}}]{Kenneth F. Caluya}
+is a Ph.D.
+student in the Department of Applied Mathematics at University of California, Santa Cruz.
+He obtained his B.S.
+degree in Mathematics at University of California, Riverside, in 2017.
+His research interests lie at the intersection of probability, convex optimization, partial differential equations, and dynamical systems.
+\end{IEEEbiography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{IEEEbiography}[{\includegraphics[height=1.25in]{parmita.png}}]{Pegah Ojaghi}
+is a Ph.D.
+student in the Department of Computer Science and Engineering at University of California, Santa Cruz.
+She obtained M.S.
+in Computer Engineering from the University of California, Santa Cruz in 2020, and M.S.
+in Control Engineering from the Imam Khomeini International University, Qazvin, Iran in 2015.
+She obtained B.S.
+in Control Engineering from the Hamedan University of Technology, Hamedan, Iran in 2012.
+\end{IEEEbiography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % You can push biographies down or up by placing

### 2021-06-26 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 \end{align}
-wherein $\mathcal{E}$ is the set of edges for the underlying graph modeling the power system network.
+wherein $\mathcal{E}$ is the set of edges for the underlying graph modeling the power system network topology.
 The potential (\ref{potential}) has a natural energy function interpretation and can also be motivated by a mechanical mass-spring-damper analogy \cite{dorfler2013synchronization,ishizaki2018}.
@@
 We now explain this idea in detail.
+
+\begin{figure}[htpb]
+\centering
+\includegraphics[width=0.9\linewidth]{ProxSchmatic.pdf}
+\caption{\small{A schematic of approximating the flow of the joint PDF trajectory $\rho(t,\bm{x})$ via the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.
+The curve $\rho(t,\bm{x})$ solves a kinetic Fokker-Planck PDE initial value problem.
+While the joint PDFs are supported over the finite dimensional base manifold $\mathbb{T}^{n}\times\mathbb{R}^{n}$, the proximal updates $\{\varrho_{k}\}_{k\in\mathbb{N}}$ evolve over the infinite dimensional manifold $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.}}
+\vspace*{-0.1in}
+\label{fig:ProxSchematic}
+\end{figure}
 
@@
 We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in therms of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous.
+Fig.
+\ref{fig:ProxSchematic} illustrates the idea of approximating the joint PDF trajectory $\rho(t,\bm{x})$ through the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ computed from a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.
 
@@
 \centering
-\vspace*{-0.1in}
 \includegraphics[width=0.95\linewidth]{ComputationalTimeSynthetic.png}
 \caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{20}\times\mathbb{R}^{20}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
+\vspace*{-0.1in}
 \label{fig:CompTimeSynthetic}
@@
 \centering
-\vspace*{-0.1in}
 \includegraphics[width=0.95\linewidth]{RelativeErrorMeanMCVersusProxSynthetic.png}
@@
 \ref{subsec:NumSimSynthetic}.}}
+\vspace*{-0.1in}
 \label{fig:RelErrSynthetic}

### 2021-06-27 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 \thanks{A. Halder and K.F. Caluya are with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder,kcaluya\}@ucsc.edu}}.}% <-this % stops a space
-\thanks{X. Geng is with {\red{Edit next}} the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{ahalder@ucsc.edu}}.} %
+\thanks{X. Geng is with the School of Electrical and Computer Engineering, Cornell University, Ithaca, NY 14850, USA, {\texttt{xg72@cornell.edu}}.} %
 \thanks{P. Ojaghi is with the Department of Computer Science and Engineering, University of California, Santa Cruz, CA 95064, USA, {\texttt{pojaghi@ucsc.edu}}.}%
@@
 The curve $\rho(t,\bm{x})$ solves a kinetic Fokker-Planck PDE initial value problem.
-While the joint PDFs are supported over the finite dimensional base manifold $\mathbb{T}^{n}\times\mathbb{R}^{n}$, the proximal updates $\{\varrho_{k}\}_{k\in\mathbb{N}}$ evolve over the infinite dimensional manifold $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.}}
+While the joint PDFs are supported over the finite dimensional base manifold $\mathbb{T}^{n}\times\mathbb{R}^{n}$, the proximal updates $\{\varrho_{k}\}_{k\in\mathbb{N}}$ (shown as circled markers with no face-color) evolve over the infinite dimensional manifold $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.}}
 \vspace*{-0.1in}
@@
 
-\begin{figure}[htpb]
+\begin{figure}[t]
 \centering
@@
 \end{figure}
-\begin{figure}[htpb]
+\begin{figure}[t]
 \centering
@@
 Halder}
-(S'10-M'14) is an Assistant Professor in the Department of Applied Mathematics,
+(S'10--M'14) is an Assistant Professor in the Department of Applied Mathematics,
 and is an affiliated faculty in the Department of Electrical and Computer Engineering at
@@
 
-\begin{IEEEbiography}{Xinbo Geng}
-Biography text here.
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{XinboGeng.jpg}}]{Xinbo Geng} (S'13--M'20)
+received his B.E.
+degree in Electrical Engineering from Tsinghua University, Beijing, China, in 2013.
+He received the M.S.
+and Ph.D.
+degrees from Department of Electrical and Computer Engineering, Texas A\&M University, College Station, TX, in 2015 and 2019, respectively.
+He is a postdoctoral researcher jointly affiliated with Cornell University and Texas A\&M University since 2020.
+He was a power system design and studies intern at National Renewable Energy Lab in 2016, a data analyst intern at ISO New England in 2017, a visiting student at UC Berkeley and MIT in 2018, and a visiting student at University of Wisconsin Madison in 2019.
+His main research interest is decision making under uncertainties.
 \end{IEEEbiography}

### 2021-06-27 code/Gen_20_prox_simulation.m

@@ -6,7 +6,7 @@ global num_Oscillator dim Gamma M Sigma K Pmech cc sigma m phi
 
 num_Oscillator = 20; dim  = 2*num_Oscillator; % dim = dimension of state space
 
-f0 = 50; % Hz
+f0 = 60; % Hz
 g_a = 20 ; g_b = 30; m_a = 2; m_b = 12 ; s_a = 1  ; s_b = 5;
 p_a = 0; p_b = 10 ;  k_a =.7 ; k_b = 1.2;
 

### 2021-06-27 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 The algorithm involves a dualization along
-with an entropic regularization of the variational update \eqref{KineticProx}, and then solving the same using a fixed point recursion that is provably contractive; see \cite[Sec.
-V.B]{caluya2019TAC}.
+with an entropic regularization of the variational update \eqref{KineticProx}, and then solving the same using a fixed point recursion that is provably contractive; we refer the interested readers to \cite[Sec.
+V.B]{caluya2019TAC} for details.
 This enables a nonparametric computation of $\tilde{\varrho}_{k}^{i} \equiv \tilde{\varrho}_{k}\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)$ for $i=1,\hdots,N$.

### 2021-06-27 code/EntropyRegularizedWasserstein.m

@@ -1,0 +1,16 @@
+function w_gam_squared = EntropyRegularizedWasserstein(X,Y,rho_x,rho_y,gam)
+
+% x,y each must be of size nsample x dimension
+
+%[X,Y] = meshgrid(x,y);
+
+C = (pdist2(X,Y)).^2;
+
+M = exp(-C/gam);
+
+OptimalCouplingMatrix_gam = SinkhornOMT(M, rho_x, rho_y);
+
+Elementwise_log = log(OptimalCouplingMatrix_gam);
+Elementwise_log(Elementwise_log == -Inf) = 0; % force log(0) = -Inf terms to zero
+
+w_gam_squared = trace(C'*OptimalCouplingMatrix_gam) - gam*sum(sum(OptimalCouplingMatrix_gam.*Elementwise_log));

### 2021-06-27 code/Gen_20_prox_simulation.m

@@ -115,9 +115,9 @@ for j=1:numSteps
     
    mean_prox_theta(j+1,:) = weighted_angle_mean(theta_upd(:,:,j+1),rho_theta_omega_upd(:,j+1));
 end
-
 toc
 
+%% MC and proximal mean vectors
 mean_mc_omega  = mean(squeeze(omega_upd));
 
 mean_mc_theta = weighted_angle_mean(theta_upd,ones(nSample,1)/nSample);
@@ -128,6 +128,16 @@ mean_mc = [mean_mc_theta;mean_mc_omega];
 mean_prox = [mean_prox_theta';mean_prox_omega'];
 
 norm_diff_mean_mc_vs_prox = sqrt(sum((mean_mc - mean_prox).^2,1))./sqrt(sum(mean_mc.^2,1));
+
+%% Wasserstein distance between MC and proximal joint PDFs
+
+W_gam_squared = zeros(numSteps,1); % pre-allocate
+gam = 0.05; % entropic regularization weight
+for j=1:numSteps
+    disp(['Now doing time-step ' num2str(j)])
+    %W_squared(j) = Wasserstein(theta_omega_upd(:,:,j+1),theta_omega_upd(:,:,j+1),rho_theta_omega_upd(:,j+1),ones(nSample,1)/nSample);
+    W_gam_squared(j) = EntropyRegularizedWasserstein(theta_omega_upd(:,:,j+1),theta_omega_upd(:,:,j+1),rho_theta_omega_upd(:,j+1),ones(nSample,1)/nSample,gam);
+end
 
 %% plots
 set(groot,'defaultAxesTickLabelInterpreter','latex');  
@@ -143,7 +153,6 @@ ylim([1e-3 1.2e-2])
 YTick = [2e-3 8e-3 2e-2];
 YTickLabels = cellstr(num2str(round(log10(YTick(:))), '10^%d'));
 grid on
-
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2021-06-27 code/SinkhornOMT.m

@@ -1,0 +1,24 @@
+function output_matrix = SinkhornOMT(input_matrix, rho_x, rho_y)
+
+v = ones(length(rho_x),1);
+
+max_iter = 1000; tol = 1e-5;
+
+iter = 1;
+
+while iter < max_iter
+    
+    u = rho_y ./ (input_matrix*v);
+    
+    v = rho_x ./ (input_matrix'*u);
+    
+    output_matrix = diag(v)*input_matrix*diag(u);
+    
+    if (norm(output_matrix'*ones(length(rho_x),1) - rho_y,inf) < tol) && (norm(output_matrix*ones(length(rho_y),1) - rho_x,inf) < tol)
+        break;
+    else
+        iter = iter + 1;
+    end
+    
+end
+

### 2021-06-27 code/Wasserstein.m

@@ -1,0 +1,21 @@
+function wsquared = Wasserstein(X,Y,rho_x,rho_y)
+
+% x,y each must be of size nsample x dimension
+
+m = size(X,1); n = size(Y,1);
+
+% Equality constraints
+Aeq = [kron(ones(1,n),speye(m));
+      kron(speye(n),ones(1,m))];
+
+beq = [rho_x/(sum(rho_x)); rho_y/(sum(rho_y))]; % stack marginal pmf vectors  
+% Cost matrix  
+%[X,Y]=meshgrid(x,y); C = (X-Y).^2; % only works for univariate PDFs
+C = (pdist2(X,Y)).^2;
+c = reshape(C, m*n, 1); % vectorize the cost for linprog
+clear C; % free some space
+
+% solve LP
+[pmf_opt,wsquared] = linprog(c,[],[],Aeq,beq,sparse(zeros(m*n,1)),Inf(m*n,1));
+
+%OptimalCouplingMatrix = reshape(pmf_opt,m,n);

### 2021-06-27 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 \section{Proximal Algorithm}\label{sec:ProxAlgo}
-Following \cite{caluya2019TAC}, we solve \eqref{KineticProx} by recursively updating the probability-weighted scattered point clouds $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where
+We solve \eqref{KineticProx} by recursively updating the probability-weighted scattered point clouds $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ where 
 $$\tilde{\bm{x}}_{k}^{i}:=\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)^{\!\top}, \quad i=1,\hdots,N, \quad k\in\mathbb{N}.$$

### 2021-06-28 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 \ref{fig:CompTimeSynthetic} highlights that for computing high dimensional transient joint state PDFs as in here, the proposed variational framework enjoys remarkably fast computational time.
-Notice that constructing a histogram as an estimator of the joint state PDF by making a grid over the 40 dimensional state space, is computationally prohibitive.
+Notice that constructing histograms or other conventional density estimators for the joint state PDF by making a grid over the 40 dimensional state space, is computationally prohibitive.
 

### 2021-06-28 code/Gen_20_prox_simulation.m

@@ -30,7 +30,7 @@ K = K - diag(diag(K)); % zero the diagonal entries
 theta0_a = 0 ; theta0_b = 2*pi; omega0_a = -.1 ; omega0_b = .1;
 
 % parameters for proximal recursion
-nSample = 1000;                      % number of samples                                                           
+nSample = 200;                      % number of samples                                                           
 epsilon = 0.5;                      % regularizing coefficient                                      
 h = 1e-3;                         % time step
 numSteps= 1000;                    % number of steps k, in discretization t=kh
@@ -131,12 +131,12 @@ norm_diff_mean_mc_vs_prox = sqrt(sum((mean_mc - mean_prox).^2,1))./sqrt(sum(mean
 
 %% Wasserstein distance between MC and proximal joint PDFs
 
-W_gam_squared = zeros(numSteps,1); % pre-allocate
-gam = 0.05; % entropic regularization weight
-for j=1:numSteps
+%W_squared = zeros(numSteps,1); % pre-allocate
+gam = 0.5; % entropic regularization weight
+for j=1:10:40
     disp(['Now doing time-step ' num2str(j)])
-    %W_squared(j) = Wasserstein(theta_omega_upd(:,:,j+1),theta_omega_upd(:,:,j+1),rho_theta_omega_upd(:,j+1),ones(nSample,1)/nSample);
-    W_gam_squared(j) = EntropyRegularizedWasserstein(theta_omega_upd(:,:,j+1),theta_omega_upd(:,:,j+1),rho_theta_omega_upd(:,j+1),ones(nSample,1)/nSample,gam);
+    W_squared(j) = Wasserstein(theta_omega_upd(:,:,j+1),theta_omega_upd(:,:,j+1),rho_theta_omega_upd(:,j+1),ones(nSample,1)/nSample);
+    % W_gam_squared(j) = EntropyRegularizedWasserstein(theta_omega_upd(:,:,j+1),theta_omega_upd(:,:,j+1),rho_theta_omega_upd(:,j+1),ones(nSample,1)/nSample,gam);
 end
 
 %% plots
[…]

### 2021-07-02 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 \end{subequations}
-where $P_{i}^{\text{mech}}$ is the mechanical power input, and $E_{i}$ is the internal voltage (magnitude) for generator $i$.
+where $P_{i}^{\text{mech}}$ is the mechanical power input, and $E_{i}$ is the internal voltage magnitude for generator $i$.
 
@@
 
-One can view (\ref{ItoSDEcomponentlevel}) as the noisy version of the second order nonuniform Kuramoto oscillator model \cite{dorfler2012synchronization,rodrigues2016kuramoto}, also known as the structure preserving power network model \cite{odun2012structure}, given by
+\begin{remark}
+We clarify here that the network-reduced model \eqref{ItoSDEcomponentlevel} is obtained from the so-called structure preserving power network model \cite{odun2012structure} with $n$ generators, after applying the Kron reduction, and therefore, has all-to-all connection topology.
+If the unreduced network has $n$ generators and $m$ buses, then the unreduced admittance matrix $\bm{Y}_{\rm{unreduced}} \in \mathbb{C}^{(n+m)\times(n+m)}$ satisfies
+\begin{align}
+\bm{Y}_{\rm{unreduced}} = .
+\label{unreducedY}	
+\end{align}
+The network reduced admittance matrix $\bm{Y}\in\mathbb{C}^{n\times n}$ is the Schur complement
+\[\bm{Y} = \bm{Y}_{\rm{unreduced}}.\]	
+\end{remark}
+
+One can view (\ref{ItoSDEcomponentlevel}) as the noisy version of the second order nonuniform Kuramoto oscillator model \cite{dorfler2012synchronization,rodrigues2016kuramoto}, given by
 \begin{align}
@@
 \begin{align}
-V(\bm{\theta}) := \displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{(i,j)\in\mathcal{E}}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right),
+V(\bm{\theta}) := \displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{i,j=1}^{n}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right).
 \label{potential}	
 \end{align}
-wherein $\mathcal{E}$ is the set of edges for the underlying graph modeling the power system network topology.
+%wherein $\mathcal{E}$ is the set of edges for the underlying graph modeling the power system network topology. 
 The potential (\ref{potential}) has a natural energy function interpretation and can also be motivated by a mechanical mass-spring-damper analogy \cite{dorfler2013synchronization,ishizaki2018}.

### 2021-07-04 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 
-\author{Abhishek~Halder,~\IEEEmembership{Member,~IEEE,} 
+\author{Abhishek~Halder,~\IEEEmembership{Senior Member,~IEEE,} 
 %\author{Abhishek~Halder,~\IEEEmembership{Member,~IEEE,}
@@
 The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively.
+We use the superscript $*$ to denote the complex conjugate.
 The uniform probability distribution over an interval $[a,b]$ is denoted as ${\rm{Unif}}\left([a,b]\right)$.
 Likewise, the $n$ dimensional uniform probability distribution over $[a,b]^{n}$ is denoted as ${\rm{Unif}}\left([a,b]^{n}\right)$.
-The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute value, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\odot$ and $\oslash$ respectively denote the elementwise (Hadamard) product and division, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map.
+The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute magnitude, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\odot$ and $\oslash$ respectively denote the elementwise (Hadamard) product and division, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map.
 The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.
@@
 \subsection{Sample Path Dynamics}\label{subsec:samplepathdynamics}
-We consider the coupled stochastic differential equations (SDEs) associated with the networked-reduced power systems model \cite[Ch.
+
+%\subsubsection{Deterministic dynamics and network reduced model}
+%Before considering stochastic uncertainties in the dynamics, let us briefly recall the network reduced power systems model \cite[Ch.
+7]{sauerpai1998} in the deterministic case.
+Suppose we denote the collection of generator buses in a networked power system as $\mathcal{G}$, and the collection of non-generator buses as $\mathcal{N}$.
+Assuming that the non-generator buses are modeled as constant current loads, with currents $I_i$, $i \in \mathcal{N}$, the power flow equations become
+%\begin{subequations}
+%\begin{eqnarray}
+%m_i \ddot{\theta} + \gamma_i \dot{\theta_i} = P_{mech,i} - P_{load,i} - \sum_{j} |V_i| |V_j| |Y_{ij}| \sin(\theta_i - \theta_j + \phi_{ij}) &~ i \in \mathcal{G}~\text{buses with generators} \\
+%0 = - \text{Real}(V_i I_i^\dagger) - \sum_{j} |V_i| |V_j| |Y_{ij}| \sin(\theta_i - \theta_j + \phi_{ij}) &~ i \in \mathcal{L}~\text{buses without generators} \\
+%\end{eqnarray}
+%\end{subequations}
+%Note: $P_{mech,i} \in \mathbb{R}$, $P_{load,i} \in \mathbb{R}$ and $I_i \in \mathbb{C}$ are all give parameters; load current is a complex number. 
+%
+%
+%\subsubsection{Stochastic dynamics}
+In this work, we consider the coupled stochastic differential equations (SDEs) associated with the networked-reduced power systems model \cite[Ch.
 7]{sauerpai1998}.
@@
 
+\begin{remark}
+We emphasize here that the network reduced model \eqref{ItoSDEcomponentlevel} is obtained from the so-called structure preserving power network model \cite{odun2012structure} after applying the Kron reduction \cite{dorfler2012kron}, and therefore, has all-to-all connection topology.
+\end{remark}
+
 With the $i$-th generator, we associate its inertia $m_{i}>0$ and damping coefficient $\gamma_{i}>0$.
@@
 \begin{align}
- P_{i} &= P_{i}^{\text{mech}} - E_{i}^{2} \Re\left(Y_{ii}\right), \\
+ P_{i} &= P_{i}^{\text{mech}} - P_{i}^{\text{load}} - |V_{i}|^{2} \Re\left(Y_{ii}\right) + \Re\left(V_i \cdot I_{i}^{*}\right),\\
  \varphi_{ij} &= \begin{cases} -\arctan\left(\dfrac{\Re\left(Y_{ij}\right)}{\Im\left(Y_{ij}\right)}\right), & \text{if}\;i\neq j,\\
@@
  \end{cases}\\
- k_{ij} &= \begin{cases} E_{i}E_{j} \vert Y_{ij}\vert, & \text{if}\;i\neq j,\\
+ k_{ij} &= \begin{cases} |V_{i}| |V_{j}| \vert Y_{ij}\vert, & \text{if}\;i\neq j,\\
  0, & \text{otherwise},
@@
 \end{subequations}
-where $P_{i}^{\text{mech}}$ is the mechanical power input, and $E_{i}$ is the internal voltage magnitude for generator $i$.
+where $P_{i}^{\text{mech}}$ is the mechanical power input, and $|V_{i}|$ is the voltage magnitude for generator $i$.
+{\red{{\textbf{Xinbo:}} One sentence saying any approximation made in (2) by using $V_{i}$ instead of the internal voltage $E_{i}$}}
 
@@
 \begin{remark}
-We clarify here that the network-reduced model \eqref{ItoSDEcomponentlevel} is obtained from the so-called structure preserving power network model \cite{odun2012structure} with $n$ generators, after applying the Kron reduction, and therefore, has all-to-all connection topology.
 If the unreduced network has $n$ generators and $m$ buses, then the unreduced admittance matrix $\bm{Y}_{\rm{unreduced}} \in \mathbb{C}^{(n+m)\times(n+m)}$ satisfies
@@
 subject to the initial condition \eqref{FPKInitCond} and the joint parameter PDF $\rho_{\text{param}}$.
+Of course, this subsumes special cases when either the initial condition or the parameter vector is deterministic.
 A direct numerical solution of this PDE initial value problem using conventional discretization (e.g., finite difference) or function approximation techniques will not be scalable in general, as explained in Sec.
@@
 
+% === Kron reduction
+
+\bibitem{dorfler2012kron}
+F.~Dorfler, and F.~Bullo, ``Kron reduction of graphs with applications to electrical networks", \emph{IEEE Transactions on Circuits and Systems I: Regular Papers}, Vol. 60, No. 1, pp. 150--163, 2012.
+
 \end{thebibliography}
@@
 Halder}
-(S'10--M'14) is an Assistant Professor in the Department of Applied Mathematics,
+(S'10-M'14-SM'21) is an Assistant Professor in the Department of Applied Mathematics,
 and is an affiliated faculty in the Department of Electrical and Computer Engineering at
@@
 
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{XinboGeng.jpg}}]{Xinbo Geng} (S'13--M'20)
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{XinboGeng.jpg}}]{Xinbo Geng} (S'13-M'20)
 received his B.E.

### 2021-07-04 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 \begin{remark}
-We emphasize here that the network reduced model \eqref{ItoSDEcomponentlevel} is obtained from the so-called structure preserving power network model \cite{odun2012structure} after applying the Kron reduction \cite{dorfler2012kron}, and therefore, has all-to-all connection topology.
+We emphasize here that the network-reduced model \eqref{ItoSDEcomponentlevel} is obtained from the so-called structure preserving power network model \cite{odun2012structure} after applying the Kron reduction \cite{dorfler2012kron}, and therefore, has all-to-all connection topology.
 \end{remark}
 
-With the $i$-th generator, we associate its inertia $m_{i}>0$ and damping coefficient $\gamma_{i}>0$.
+With the $i$th generator, we associate its inertia $m_{i}>0$ and damping coefficient $\gamma_{i}>0$.
 The other parameters: the \emph{effective power input} $P_{i}$, the \emph{phase shift} $\varphi_{ij}\in[0,\frac{\pi}{2})$, and the \emph{coupling coefficients} $k_{ij}\geq 0$, depend on the network reduced admittance matrix $\bm{Y}\equiv[Y_{ij}]_{i,j=1}^{n}\in\mathbb{C}^{n\times n}$.
@@
 \end{subequations}
-where $P_{i}^{\text{mech}}$ is the mechanical power input, and $|V_{i}|$ is the voltage magnitude for generator $i$.
+where $P_{i}^{\text{mech}}$ is the mechanical power input, $P_{i}^{\text{load}}$ is {\red{the real load (ignoring reactive load)}}, and $|V_{i}|$ is the voltage magnitude for generator $i$.
 {\red{{\textbf{Xinbo:}} One sentence saying any approximation made in (2) by using $V_{i}$ instead of the internal voltage $E_{i}$}}

### 2021-07-05 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 \subsubsection{Notation and organization}
-We use boldfaced small letters for vectors, and boldfaced capital letters for matrices.
+%We use boldfaced small letters for vectors, and boldfaced capital letters for matrices. 
 The set of natural numbers, real numbers and complex numbers are denoted as $\mathbb{N}$, $\mathbb{R}$, and $\mathbb{C}$, respectively.
@@
 where the state variables are the rotor angles $\theta_{i}\in[0,2\pi)$ and the rotor angular velocities $\omega_{i}\in\mathbb{R}$, for $i\in\{1,\hdots,n\}$.
-The stochastic forcing is modeled through the standard Wiener process $w_{i}(t)$, and the diffusion coefficient $\sigma_{i}>0$ denotes the intensity of stochastic forcing at the $i$-th generator.
+The stochastic forcing is modeled through the standard Wiener process $w_{i}(t)$, and the diffusion coefficient $\sigma_{i}>0$ denotes the intensity of stochastic forcing at the $i$th generator.
 
@@
 \begin{align}
- P_{i} &= P_{i}^{\text{mech}} - P_{i}^{\text{load}} - |V_{i}|^{2} \Re\left(Y_{ii}\right) + \Re\left(V_i \cdot I_{i}^{*}\right),\\
+ P_{i} &= P_{i}^{\text{mech}} - P_{i}^{\text{load}} - |E_{i}|^{2} \Re\left(Y_{ii}\right) + \Re\left(E_i \cdot I_{i}^{*}\right),\label{defPi}\\
  \varphi_{ij} &= \begin{cases} -\arctan\left(\dfrac{\Re\left(Y_{ij}\right)}{\Im\left(Y_{ij}\right)}\right), & \text{if}\;i\neq j,\\
@@
  \end{cases}\\
- k_{ij} &= \begin{cases} |V_{i}| |V_{j}| \vert Y_{ij}\vert, & \text{if}\;i\neq j,\\
+ k_{ij} &= \begin{cases} |E_{i}| |E_{j}| \vert Y_{ij}\vert, & \text{if}\;i\neq j,\\
  0, & \text{otherwise},
@@
 \end{align}
+\label{DefParam}
 \end{subequations}
-where $P_{i}^{\text{mech}}$ is the mechanical power input, $P_{i}^{\text{load}}$ is {\red{the real load (ignoring reactive load)}}, and $|V_{i}|$ is the voltage magnitude for generator $i$.
-{\red{{\textbf{Xinbo:}} One sentence saying any approximation made in (2) by using $V_{i}$ instead of the internal voltage $E_{i}$}}
+where $P_{i}^{\text{mech}}$ is the mechanical power input, $P_{i}^{\text{load}}$ is the real load, $E_{i}$ is the internal voltage, and $I_{i}\in\mathbb{C}$ is the current for generator $i$.
 
@@
 
-\begin{remark}
-If the unreduced network has $n$ generators and $m$ buses, then the unreduced admittance matrix $\bm{Y}_{\rm{unreduced}} \in \mathbb{C}^{(n+m)\times(n+m)}$ satisfies
-\begin{align}
-\bm{Y}_{\rm{unreduced}} = .
-\label{unreducedY}	
-\end{align}
-The network reduced admittance matrix $\bm{Y}\in\mathbb{C}^{n\times n}$ is the Schur complement
-\[\bm{Y} = \bm{Y}_{\rm{unreduced}}.\]	
-\end{remark}
+%Suppose we denote the collection of generator buses in an \emph{unreduced} power system as $\mathcal{G}$, and the collection of non-generator buses as $\mathcal{N}$. 
+Suppose that the \emph{unreduced} power network has $n$ generators and $m$ buses. Then the unreduced admittance matrix $\bm{Y}_{\rm{unreduced}}\in \mathbb{C}^{(n+m)\times(n+m)}$ can be partitioned as
+\begin{align}
+\bm{Y}_{\rm{unreduced}} = \begin{bmatrix} 
+\begin{array}{c|c}
+\bm{Y}_{\rm{bnd}} & \bm{Y}_{\rm{bnd{\text{-}}int}} \\
+\hline\vspace*{-0.12in}\\
+\bm{Y}_{\rm{bnd{\text{-}}int}}^{\top} & \bm{Y}_{\rm{int}}
+\end{array}	
+ \end{bmatrix},
+\label{PartionedAdmittanceMatrix}	
+\end{align} 
+where $\bm{Y}_{\rm{bnd}}\in\mathbb{C}^{n\times n}$, $\bm{Y}_{\rm{int}}\in\mathbb{C}^{m\times m}$, $\bm{Y}_{\rm{bnd{\text{-}}int}}\in\mathbb{C}^{n\times m}$.
+The matrix \eqref{PartionedAdmittanceMatrix} relates the unreduced current vector $I_{\rm{unreduced}}\in\mathbb{C}^{n+m}$ with the unreduced voltage vector $E_{\rm{unreduced}}\in\mathbb{C}^{n+m}$ via the Kirchhoff equations
+\begin{align}
+I_{\rm{unreduced}} = \bm{Y}_{\rm{unreduced}} E_{\rm{unreduced}},
+\label{DefUnreducedY}	
+\end{align}
+or equivalently, via its partitioned version associated with the interior and the boundary nodes:
+\begin{align}
+\begin{array}{c}
+\begin{bmatrix}
+I_{\rm{bnd}}\\
+I_{\rm{int}} 	
+\end{bmatrix}	
+\end{array} = \bm{Y}_{\rm{unreduced}} \begin{array}{c}
+\begin{bmatrix}
+E_{\rm{bnd}}\\
+E_{\rm{int}} 	
+\end{bmatrix}	
+\end{array}.
+\label{PartitionedKirchoffEquation}	
+\end{align}
+%The reduced admittance matrix
+%\begin{eqnarray}
+%\bm{Y} &=& Y_{\mathcal{G}\mathcal{G}} - Y_{\mathcal{G}\mathcal{L}} Y_{\mathcal{L}\mathcal{L}}^{-1} Y_{\mathcal{L}\mathcal{G}} \\
+%I_{red} &=& Y_{\mathcal{G}\mathcal{L}}Y_{\mathcal{L}\mathcal{L}}^{-1} I
+%\end{eqnarray}
+
+The network reduced admittance matrix $\bm{Y}\in\mathbb{C}^{n\times n}$ in \eqref{DefParam} is the Schur complement of \eqref{PartionedAdmittanceMatrix} w.r.t. the block $\bm{Y}_{\rm{int}}$, i.e.,
+\[\bm{Y} = \bm{Y}_{\rm{unreduced}} / \bm{Y}_{\rm{int}} := \bm{Y}_{\rm{bnd}} - \bm{Y}_{\rm{bnd{\text{-}}int}} \bm{Y}_{\rm{int}}^{-1}\bm{Y}_{\rm{bnd{\text{-}}int}}^{\top}.\]
+
+The network reduced current vector $I\in\mathbb{C}^{n}$ in \eqref{defPi} is obtained as
+\[I  = \bm{Y}E_{\rm{bnd}} - I_{\rm{bnd}}.\]
 
@@
 
-\begin{figure}[htpb]
+\begin{figure}[tb]
 \centering

### 2021-07-05 code/get_parameters_case14.m

@@ -1,0 +1,60 @@
+clear; clc;
+define_constants;
+
+
+mpc = loadcase('case14.m');
+results = runopf(mpc);
+
+z = mpc.branch(:,BR_R)+i*mpc.branch(:,BR_X);
+y = 1./z;
+k = abs(y);
+phi = angle(y);
+
+% line parameters
+lines = [mpc.branch(:,[F_BUS,T_BUS,BR_R, BR_X]), k, phi];
+writematrix(lines,'line-parameters.csv');
+
+% bus initial states (at equilibrium pt)
+buses = [mpc.bus(:,[BUS_I, VM]),deg2rad(mpc.bus(:,VA))];
+writematrix(buses,'bus-init.csv');
+
+% generator initial states (at equilibrium pt)
+w0 = 2*pi*60;
+H = [2.64; 2.61; 3.42; 2.45; 3.59];
+D = [4; 4; 4; 4; 4];
+m = 2*w0.*H;
+gamma = w0.*D; 
+gens = [results.gen(:,GEN_BUS),results.gen(:,PG)/mpc.baseMVA, H,D,m,gamma];
+writematrix(gens,'gen_parameters.csv');
+
+% Kron Reduction
+Y = makeYbus(mpc);
+
+nbus = size(mpc.bus,1);
+inner_bus_ind = find(mpc.bus(:,BUS_TYPE)==PQ);
+boundary_bus_ind = setdiff(1:nbus, inner_bus_ind);
+
+Y_red = Y(boundary_bus_ind ,boundary_bus_ind ) ...
+    - Y(boundary_bus_ind, inner_bus_ind)*inv(Y(inner_bus_ind,inner_bus_ind))*Y(inner_bus_ind,boundary_bus_ind);
+
+Y_red_real = real(Y_red);
+Y_red_imag = imag(Y_red);
+
+writematrix(Y_red_real,'Y_reduced_real_part.csv');
+writematrix(Y_red_imag,'Y_reduced_imag_part.csv');
+
+% load data
+% constant power load for those at generator buses
+load_at_genbus = [results.bus(boundary_bus_ind,BUS_I), results.bus(boundary_bus_ind,PD)/mpc.baseMVA];
+writematrix(load_at_genbus,'P_load_at_genbuses.csv');
+
+% constant current load for those at generator buses
+V_load = results.bus(inner_bus_ind, VM) .* exp(i*deg2rad(results.bus(inner_bus_ind, VA)) );
[…]

### 2021-07-05 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 \end{bmatrix}	
-\end{array}.
+\end{array},
 \label{PartitionedKirchoffEquation}	
 \end{align}
+where $I_{\rm{bnd}},E_{\rm{bnd}}\in\mathbb{C}^{n}$ and $I_{\rm{int}},E_{\rm{int}}\in\mathbb{C}^{m}$.
 %The reduced admittance matrix
@@
 \section{Numerical Simulations}\label{sec:NumericalSimulations}
-
-\subsection{IEEE Test System}\label{subsec:NumSimIEEEBus}
+To illustrate the proposed computational framework, we next provide two numerical simulation case studies.
+In Sec.
+\ref{subsec:NumSimIEEEBus}, we consider the prediction of stochastic states for the IEEE 14 bus system--our main objective being to highlight the .
+
+\subsection{IEEE 14 Bus System}\label{subsec:NumSimIEEEBus}
 

### 2021-07-05 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

@@
 \[\bm{Y} = \bm{Y}_{\rm{unreduced}} / \bm{Y}_{\rm{int}} := \bm{Y}_{\rm{bnd}} - \bm{Y}_{\rm{bnd{\text{-}}int}} \bm{Y}_{\rm{int}}^{-1}\bm{Y}_{\rm{bnd{\text{-}}int}}^{\top}.\]
-
 The network reduced current vector $I\in\mathbb{C}^{n}$ in \eqref{defPi} is obtained as

### 2021-07-05 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V.B with 50 generator simulation, 100 dim state space

@@
 subject to the initial condition \eqref{FPKInitCond} and the joint parameter PDF $\rho_{\text{param}}$.
-Of course, this subsumes special cases when either the initial condition or the parameter vector is deterministic.
+Of course, this subsumes special cases such as when either the initial condition or the parameter vector is deterministic.
 A direct numerical solution of this PDE initial value problem using conventional discretization (e.g., finite difference) or function approximation techniques will not be scalable in general, as explained in Sec.
@@
 In Sec.
-\ref{subsec:NumSimIEEEBus}, we consider the prediction of stochastic states for the IEEE 14 bus system--our main objective being to highlight the .
+\ref{subsec:NumSimIEEEBus}, we consider the prediction of stochastic states for the IEEE 14 bus system--{\red{our main objective being to highlight the interpretability and utility of the proposed method in power systems engineering}}.
+In Sec.
+\ref{subsec:NumSimSynthetic}, we propagate the joint PDFs over the {\red{100 dimensional state space}} of a synthetic power network with randomly generated parameters--our intent there is to highlight the scalability of the proposed method.
 
@@
 \ref{subsec:NumSimSynthetic}.
-The indices $i,j\in\{1,\hdots,n\}$ where the number of generators $n=20$ in Sec.
+The indices $i,j\in\{1,\hdots,n\}$ where the number of generators $n=50$ in Sec.
 \ref{subsec:NumSimSynthetic}.}
@@
 \subsection{Synthetic Test System}\label{subsec:NumSimSynthetic}
-We next consider a power network with $n=20$ generators, and propagate the transient joint state PDFs supported over the 40 dimensional state space $\mathbb{T}^{20}\times\mathbb{R}^{20}$.
+We next consider a power network with $n=50$ generators, and propagate the transient joint state PDFs supported over the 100 dimensional state space $\mathbb{T}^{50}\times\mathbb{R}^{50}$.
 We take the initial joint state PDF at $t=0$ as
@@
 \centering
-\includegraphics[width=0.95\linewidth]{ComputationalTimeSynthetic.png}
-\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{20}\times\mathbb{R}^{20}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
+\includegraphics[width=0.95\linewidth]{ComputationalTimeSynthetic50Gen.png}
+\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{50}\times\mathbb{R}^{50}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
 \vspace*{-0.1in}
@@
 \centering
-\includegraphics[width=0.95\linewidth]{RelativeErrorMeanMCVersusProxSynthetic.png}
-\caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$, and the mean $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec.
+\includegraphics[width=0.95\linewidth]{RelativeErrorMeanMCVersusProxSynthetic50Gen.png}
+\caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$, and the mean $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec.
 \ref{subsec:NumSimSynthetic}.}}
@@
 
-With $N=1000$ random samples from the initial joint PDF \eqref{rho0synthetic}, and with the aforesaid randomly generated parameters, we employed the Euler-Maruyama scheme together with Algorithm \ref{algo:KineticProx} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
+With $N=2000$ random samples from the initial joint PDF \eqref{rho0synthetic}, and with the aforesaid randomly generated parameters, we employed the Euler-Maruyama scheme together with Algorithm \ref{algo:KineticProx} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
 We used $h=10^{-3}$, $\varepsilon = 0.05$, $\delta = 10^{-3}$, $L = 300$.
@@
 To depict the numerical accuracy, Fig.
-\ref{fig:RelErrSynthetic} plots the (time-varying) relative error between the empirical (i.e., Monte Carlo estimate) mean vector $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$, and the ``proximal mean" vector $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{20}\times\mathbb{R}^{20}$.
+\ref{fig:RelErrSynthetic} plots the (time-varying) relative error between the empirical (i.e., Monte Carlo estimate) mean vector $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$, and the ``proximal mean" vector $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$.
 The latter was computed using the proximal updates $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
+Fig.
+\ref{fig:RelErrSynthetic} shows that the time varying statistics from these two computation match very well at all times.
 

### 2021-07-05 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py

@@ -55,15 +55,15 @@ fig = plt.figure(1, figsize=fig_size)  # figsize accepts only inches.
 # =============================================================================
 # Computational time plot 
 # =============================================================================
-Time_Synthetic = np.loadtxt("TimeSynthetic.txt")
-CompTime_Synthetic = np.loadtxt("ComptimeSythetic.txt")
+Time_Synthetic = np.loadtxt("TimeSynthetic50Gen.txt")
+CompTime_Synthetic = np.loadtxt("ComptimeSythetic50Gen.txt")
 
 fig = plt.figure()
 ax = fig.add_subplot(1, 1, 1)
 
 ax.semilogy(Time_Synthetic, CompTime_Synthetic,'-', color='b', lw=1.5)
 
-ax.set_ylim(1*10**-3, 1.2*10**-2)
+ax.set_ylim(5*10**-3, 5*10**-2)
 
 ax.tick_params(direction='in',which='both')
  # axx.xaxis.tick_top()
@@ -80,12 +80,12 @@ ax.set_xlabel(r"Physical time $t=kh$ [s]")
 # axx.legend(markerscale=1.5, numpoints=1,  ncol=1, bbox_to_anchor=(1.005, 1), frameon=False, prop={'size': 10.5})
 fig.set_size_inches(10.2, 6.2)
 
-plt.savefig('ComputationalTimeSynthetic.png', dpi=400)
+plt.savefig('ComputationalTimeSynthetic50Gen.png', dpi=400)
 
 # =============================================================================
 # Relative error in mean vector: MC versus Prox plot
 # =============================================================================
-RelErrMeanVectorMCvsProx_Synthetic = np.loadtxt("RelErrMeanVectorMCvsProxSythetic.txt")
+RelErrMeanVectorMCvsProx_Synthetic = np.loadtxt("RelErrMeanVectorMCvsProxSythetic50Gen.txt")
 
 fig = plt.figure()
 ax = fig.add_subplot(1, 1, 1)
@@ -109,7 +109,7 @@ ax.set_xlabel(r"Physical time $t=kh$ [s]")
 # axx.legend(markerscale=1.5, numpoints=1,  ncol=1, bbox_to_anchor=(1.005, 1), frameon=False, prop={'size': 10.5})
 fig.set_size_inches(10.2, 6.2)
 
-plt.savefig('RelativeErrorMeanMCVersusProxSynthetic.png', dpi=400)
[…]

### 2021-07-05 code/Gen_20_prox_simulation.m

@@ -30,7 +30,7 @@ K = K - diag(diag(K)); % zero the diagonal entries
 theta0_a = 0 ; theta0_b = 2*pi; omega0_a = -.1 ; omega0_b = .1;
 
 % parameters for proximal recursion
-nSample = 200;                      % number of samples                                                           
+nSample = 1000;                      % number of samples                                                           
 epsilon = 0.5;                      % regularizing coefficient                                      
 h = 1e-3;                         % time step
 numSteps= 1000;                    % number of steps k, in discretization t=kh
@@ -132,12 +132,12 @@ norm_diff_mean_mc_vs_prox = sqrt(sum((mean_mc - mean_prox).^2,1))./sqrt(sum(mean
 %% Wasserstein distance between MC and proximal joint PDFs
 
 %W_squared = zeros(numSteps,1); % pre-allocate
-gam = 0.5; % entropic regularization weight
-for j=1:10:40
-    disp(['Now doing time-step ' num2str(j)])
-    W_squared(j) = Wasserstein(theta_omega_upd(:,:,j+1),theta_omega_upd(:,:,j+1),rho_theta_omega_upd(:,j+1),ones(nSample,1)/nSample);
-    % W_gam_squared(j) = EntropyRegularizedWasserstein(theta_omega_upd(:,:,j+1),theta_omega_upd(:,:,j+1),rho_theta_omega_upd(:,j+1),ones(nSample,1)/nSample,gam);
-end
+% gam = 0.5; % entropic regularization weight
+% for j=1:10:40
+%     disp(['Now doing time-step ' num2str(j)])
+%     W_squared(j) = Wasserstein(theta_omega_upd(:,:,j+1),theta_omega_upd(:,:,j+1),rho_theta_omega_upd(:,j+1),ones(nSample,1)/nSample);
+%     % W_gam_squared(j) = EntropyRegularizedWasserstein(theta_omega_upd(:,:,j+1),theta_omega_upd(:,:,j+1),rho_theta_omega_upd(:,j+1),ones(nSample,1)/nSample,gam);
+% end
 
 %% plots
[…]

### 2021-07-05 code/Gen_50_prox_simulation.m

@@ -1,0 +1,208 @@
+close all; clear; clc;
+%% Parameters
+beta = 2; % inverse temperature
+% golbal parameters 
+global num_Oscillator dim Gamma M Sigma K Pmech cc sigma m phi
+
+num_Oscillator = 50; dim  = 2*num_Oscillator; % dim = dimension of state space
+
+f0 = 60; % Hz
+g_a = 20 ; g_b = 30; m_a = 2; m_b = 12 ; s_a = 1  ; s_b = 5;
+p_a = 0; p_b = 10 ;  k_a =.7 ; k_b = 1.2;
+
+tp_a = 0; 
+tp_b = 0.25;
+
+gamma = ((g_a-g_b)*rand(num_Oscillator,1) + g_a)/(2*pi*f0);
+m = ((m_b-m_a)*rand(num_Oscillator,1) + m_a)/(2*pi*f0);
+sigma = (s_b-s_a)*rand(num_Oscillator,1) + s_a;
+
+phi = atan((tp_b-tp_a)*rand(num_Oscillator) + tp_a);
+phi = phi - diag(diag(phi)); % zero the diagonal entries
+
+pmech = (p_b-p_a)*rand(num_Oscillator,1) + p_b;
+
+Gamma = diag(gamma);  M = diag(m); Sigma = diag(sigma); Pmech = diag(pmech);
+
+K = (k_b-k_a)*rand(num_Oscillator)+ k_a;
+K = K - diag(diag(K)); % zero the diagonal entries
+
+theta0_a = 0 ; theta0_b = 2*pi; omega0_a = -.1 ; omega0_b = .1;
+
+% parameters for proximal recursion
+nSample = 2000;                      % number of samples                                                           
+epsilon = 0.5;                      % regularizing coefficient                                      
+h = 1e-3;                         % time step
+numSteps= 1000;                    % number of steps k, in discretization t=kh
+cc = 1e7;
+t_vec = h*(1:1:numSteps);
+%% propagate joint PDF
+% samples from initial joint PDF 
+theta_0 = (theta0_b-theta0_a)*rand(nSample,num_Oscillator) + theta0_a;
+
[…]

### 2021-07-05 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V.B with 50 generator simulation, 100 dim state space

@@
 The resulting bivariate Fokker-Planck PDE in \cite{wang2013fokker} was solved via finite element discretization, and revealed rich stochastic dynamics and nontrivial transient stability aspects even in this simple case.
-However, it is unreasonable to expect that a finite element discretization, or in fact any spatial discretization scheme to solve (\ref{KineticFPK}) for moderately large $n$ in {\red{minutes}} of computational time, thereby limiting our current ability for realistic power systems simulation with stochastic variability.
+However, it is unreasonable to expect that a finite element discretization, or in fact any spatial discretization scheme to solve (\ref{KineticFPK}) for moderately large $n$ in seconds of computational time, thereby limiting our current ability for realistic power systems simulation with stochastic variability.
 This calls for fundamentally re-thinking what does it mean to solve the PDE (\ref{KineticFPK}) for dynamics (\ref{ItoSDEvectorlevel}).
@@
 In Sec.
-\ref{subsec:NumSimSynthetic}, we propagate the joint PDFs over the {\red{100 dimensional state space}} of a synthetic power network with randomly generated parameters--our intent there is to highlight the scalability of the proposed method.
+\ref{subsec:NumSimSynthetic}, we propagate the joint PDFs over the 100 dimensional state space of a synthetic power network with randomly generated parameters--our intent there is to highlight the scalability of the proposed method.
 
@@
 \ref{fig:CompTimeSynthetic} highlights that for computing high dimensional transient joint state PDFs as in here, the proposed variational framework enjoys remarkably fast computational time.
-Notice that constructing histograms or other conventional density estimators for the joint state PDF by making a grid over the 40 dimensional state space, is computationally prohibitive.
+Notice that constructing histograms or other conventional density estimators for the joint state PDF by making a grid over the 100 dimensional state space, is computationally prohibitive.
 

### 2021-07-07 code/IEEE14_prox_simulation.m
adding IEEE 14 bus sim results

@@ -1,31 +1,54 @@
 close all; clear all; clc;
 %% Parameters
-
 beta = 2; % inverse temperature
-
 % golbal parameters 
-global num_Oscillator dim Gamma M Sigma K Pmech cc sigma m phi
+global num_Oscillator dim Gamma M Sigma K P cc sigma m phi
 
 num_Oscillator = 5; dim  = 2*num_Oscillator; % dim = dimension of state space
 
-E = [1.06 1.045 1.01 1.07 1.09];
+f0 = 60; % nominal frequency (Hz)
 
-Y = Y_admittance_matrix;
-phi = phase_shift_matrix(Y);
-K = coupling_matrix(Y,E);
+% parameters for the IEEE 14 bus system
+Y_reduced_imag = readmatrix('Y_reduced_imag_part.csv');
+Y_reduced_real = readmatrix('Y_reduced_real_part.csv');
+% reduced admittance matrix
+Y_reduced = complex(Y_reduced_real,Y_reduced_imag);
 
-f0 = 60;
-g_a = 20 ; g_b = 30; m_a = 2; m_b = 12 ; s_a = 1  ; s_b = 5;
-p_a = 0; p_b = 10 ;  k_a =.7 ; k_b = 1.2;
+gen_param = readmatrix('gen_parameters.csv');
+gen_bus_idx = gen_param(:,1);
 
-gamma = ((g_a-g_b)*rand(num_Oscillator,1) + g_a)/(2*pi*f0);
-m = ((m_b-m_a)*rand(num_Oscillator,1) + m_a)/(2*pi*f0);
+bus_init = readmatrix('bus-init.csv');
+% generator bus voltage
+E_mag = bus_init(gen_bus_idx,2);
+E_phase = bus_init(gen_bus_idx,3);
+E = E_mag.*exp(1i*E_phase);
+% reduced current vector
+I_reduced_polar = readmatrix('I_red.csv');
+I_reduced_mag = I_reduced_polar(:,1);
+I_reduced_phase = I_reduced_polar(:,2);
+I_reduced = I_reduced_mag.*exp(1i*I_reduced_phase);
+
+P_mech = gen_param(:,2);
+P_load = readmatrix('P_load_at_genbuses.csv');
+P_load = P_load(:,2);
+% effective power input
[…]

### 2021-07-07 code/PowerDrift.m

@@ -1,10 +1,10 @@
 function [driftk,GradU] = PowerDrift(xi,eta,nSample)
 
-global num_Oscillator phi Gamma M Sigma Pmech K sigma  m
+global num_Oscillator phi Gamma M Sigma P K sigma  m
  
 GradF_eta = ((Gamma/M)*eta')';
 
-GradU_2 = -repmat((diag(Pmech/Sigma))',nSample,1);
+GradU_2 = -repmat((diag(P/Sigma))',nSample,1);
 % 
 % for l=1:nSample
 % 

### 2021-07-07 code/getMarginal2D.m

@@ -1,0 +1,26 @@
+function [XX1,XX2,marg] = getMarginal2D(X1,X2,PDF,nBins)
+
+Bins.x1 = linspace(min(X1),max(X1),nBins+1);
+Bins.x2 = linspace(min(X2),max(X2),nBins+1);
+
+dx1 = mean(diff(Bins.x1));
+dx2 = mean(diff(Bins.x2));
+
+A = dx1*dx2;
+
+[x1,x2] = meshgrid(Bins.x1,Bins.x2);
+
+
+for i=1:nBins
+    for j=1:nBins
+        ii = find(X1>= x1(i,j) & X1 < x1(i+1,j+1));
+        jj = find(X2(ii)>= x2(i,j) & X2(ii) < x2(i+1,j+1));
+        marg.MC(i,j) = length(jj);
+        marg.PF(i,j) = sum(PDF(ii(jj)));
+        XX1(i,j) = (x1(i,j) +  x1(i+1,j+1))/2.0;
+        XX2(i,j) = (x2(i,j) +  x2(i+1,j+1))/2.0;
+    end
+end
+
+marg.MC = marg.MC/sum(marg.MC(:))/A;
+marg.PF = marg.PF/sum(marg.PF(:))/A;

### 2021-07-07 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \ref{subsec:NumSimSynthetic}, we propagate the joint PDFs over the 100 dimensional state space of a synthetic power network with randomly generated parameters--our intent there is to highlight the scalability of the proposed method.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=0.95\linewidth]{IEEE14marginals.png}
+\caption{\small{Time evolution of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs, where $i=1,\hdots,5$, for the Kron-reduced IEEE 14 bus simulation setup described in Sec.
+\ref{subsec:NumSimIEEEBus}.
+The $5$ rows above correspond to the $5$ generator nodes--these are buses 1, 2, 3, 6 and 8 in Fig.
+\ref{fig:Motivating}.
+The columns correspond to the time snapshots.
+The colorbar at the bottom shows the marginal PDF values.}}
+\vspace*{-0.1in}
+\label{fig:IEEE14marginals}
+\end{figure*}
 

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \ref{subsec:NumSimIEEEBus}.
-The $5$ rows above correspond to the $5$ generator nodes--these are buses 1, 2, 3, 6 and 8 in Fig.
+The $5$ rows above correspond to the $5$ generator nodes; these are buses 1, 2, 3, 6 and 8 in Fig.
 \ref{fig:Motivating}.
@@
 
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.95\linewidth]{ComputationalTimeSyntheticIEEE14bus.png}
+\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ for the simulation set up in Sec. \ref{subsec:NumSimIEEEBus}.}}
+\vspace*{-0.1in}
+\label{fig:CompTimeSynthetic}
+\end{figure}
+
 \subsection{IEEE 14 Bus System}\label{subsec:NumSimIEEEBus}
+Following Sec.
+\ref{subsec:samplepathdynamics}, we consider the Kron-reduced dynamics \eqref{ItoSDEcomponentlevel} for the IEEE 14 bus system shown in Fig.
+\ref{fig:Motivating}.
+In this case, we have $n=5$ nodes which correspond to the buses 1, 2, 3, 6 and 8 in Fig.
+\ref{fig:Motivating}.
+{\red{The parameter values in \eqref{DefParam} use the network reduced admittance matrix computed from the CITE??
+The parameters $m_{i}, \gamma_{i}$ are obtained from ....}} We randomly generated $\sigma_{i}\in [1,5]$ as
+\[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]
+
+We take the initial joint PDF as
+\begin{align}
+\rho_{0}&\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = \vartheta_{0}\left(\bm{\theta}(0)\right)\Omega_{0}\left(\bm{\omega}(0)\right)\nonumber\\
+&=\underbrace{\left(\prod_{i=1}^{n=5}\dfrac{\exp\left(\kappa_{i}\cos\left(2\theta_{i}(0)-\mu_{\theta_{i}(0)}\right)\right)}{2\pi I_{0}(\kappa_{i})}\right)}_{=:\vartheta_{0}\left(\bm{\theta}(0)\right)}\nonumber\\
+	&\qquad\times \underbrace{{\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right)}_{=:\Omega_{0}\left(\bm{\omega}(0)\right)},
+\label{rho0IEEE14}		
+\end{align}
+whose $\bm{\theta}$ marginal $ \vartheta_{0}\left(\bm{\theta}(0)\right)$ is a product von Mises PDF \cite{mardia2008multivariate,mardia2014some} supported over $\mathbb{T}^{5}$ with mean angles (in rad)
+\begin{align}
+\left(\mu_{\theta_{1}}(0),\hdots,\mu_{\theta_{5}}(0)\right)=\left(0, 6.1963, 6.0612, 6.0350, 6.0500\right),
+\label{MeanAngleInitialIEEE14}	
+\end{align}
+and concentration parameter $\kappa_{i}\geq 0$ set as 
+\[(\kappa_{1},\hdots,\kappa_5) = \left(5,6,7,4,5\right).\]
+The mean angles \eqref{MeanAngleInitialIEEE14} were obtained as steady state power flow solutions {\red{for the generator phase angles with constant load??
+Please complete/edit this sentence}} Larger $\kappa_{i}$ entails a larger concentration around the mean angle $\mu_{\theta_{i}}(0)$ (setting $\kappa_i=0$ implies that $\theta_{i}(0)$ follows uniform distribution over $[0,2\pi)$).
+
+In \eqref{rho0IEEE14}, the $\bm{\omega}$ marginal $ \Omega_{0}\left(\bm{\omega}(0)\right)$ is a uniform PDF over $[-0.1,0.1]$ rad/s.
+Thus, the PDF $\rho_0$ in \eqref{rho0IEEE14} is supported over $\mathbb{T}^{5}\times [-0.1,0.1]$.
 
@@
 \begin{align}
-\rho_{0}\equiv\rho(t=0,\theta(0),\omega(0)) = &\;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right)\nonumber\\
+\rho_{0}\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = &\;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right)\nonumber\\
 	&\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
@@
 \bibitem{dorfler2012kron}
-F.~Dorfler, and F.~Bullo, ``Kron reduction of graphs with applications to electrical networks", \emph{IEEE Transactions on Circuits and Systems I: Regular Papers}, Vol. 60, No. 1, pp. 150--163, 2012.
+F.~Dorfler, and F.~Bullo, ``Kron Reduction of Graphs with Applications to Electrical Networks", \emph{IEEE Transactions on Circuits and Systems I: Regular Papers}, Vol. 60, No. 1, pp. 150--163, 2012.
+
+% === von Mises PDF on n torus: Tn
+\bibitem{mardia2008multivariate}
+K.V.~Mardia, G.~Hughes, C.C.~Taylor, H.~Singh, ``A Multivariate von Mises Distribution with Applications to Bioinformatics", \emph{Canadian Journal of Statistics}, Vol. 36, No. 1, pp. 99--109, 2008.
+
+\bibitem{mardia2014some}
+K.V.~Mardia, and J. Voss, ``Some Fundamental Properties of a
+Multivariate von Mises Distribution", \emph{Communications in Statistics - Theory and Methods}, Vol. 43, No. 6, pp. 1132--1144, 2014.
 

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py

@@ -55,15 +55,18 @@ fig = plt.figure(1, figsize=fig_size)  # figsize accepts only inches.
 # =============================================================================
 # Computational time plot 
 # =============================================================================
-Time_Synthetic = np.loadtxt("TimeSynthetic50Gen.txt")
-CompTime_Synthetic = np.loadtxt("ComptimeSythetic50Gen.txt")
+# Time_Synthetic = np.loadtxt("TimeSynthetic50Gen.txt")
+# CompTime_Synthetic = np.loadtxt("ComptimeSythetic50Gen.txt")
+
+Time_Synthetic = np.loadtxt("TimeSyntheticIEEE14bus.txt")
+CompTime_Synthetic = np.loadtxt("ComptimeSytheticIEEE14bus.txt")
 
 fig = plt.figure()
 ax = fig.add_subplot(1, 1, 1)
 
 ax.semilogy(Time_Synthetic, CompTime_Synthetic,'-', color='b', lw=1.5)
 
-ax.set_ylim(5*10**-3, 5*10**-2)
+ax.set_ylim(10**-3, 6*10**-2)
 
 ax.tick_params(direction='in',which='both')
  # axx.xaxis.tick_top()
@@ -80,36 +83,37 @@ ax.set_xlabel(r"Physical time $t=kh$ [s]")
 # axx.legend(markerscale=1.5, numpoints=1,  ncol=1, bbox_to_anchor=(1.005, 1), frameon=False, prop={'size': 10.5})
 fig.set_size_inches(10.2, 6.2)
 
-plt.savefig('ComputationalTimeSynthetic50Gen.png', dpi=400)
+# plt.savefig('ComputationalTimeSynthetic50Gen.png', dpi=400)
+plt.savefig('ComputationalTimeSyntheticIEEE14bus.png', dpi=400)
 
 # =============================================================================
 # Relative error in mean vector: MC versus Prox plot
 # =============================================================================
-RelErrMeanVectorMCvsProx_Synthetic = np.loadtxt("RelErrMeanVectorMCvsProxSythetic50Gen.txt")
[…]

### 2021-07-08 code/IEEE14_prox_simulation.m

@@ -261,3 +261,10 @@ for j=1:num_Oscillator
         dlmwrite(textfilename_marg2D, marg2D_theta_omega_save{tt,j}.PF,'delimiter','\t','precision','%f');
     end    
 end
+
+%% save simulation time data
+textfilename = 'TimeSyntheticIEEE14bus.txt';
+dlmwrite(textfilename, t_vec,'delimiter','\t','precision','%.64f');
+
+textfilename = 'ComptimeSytheticIEEE14bus.txt';
+dlmwrite(textfilename, comptime,'delimiter','\t','precision','%.64f');

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \vspace*{-0.1in}
-\label{fig:CompTimeSynthetic}
+\label{fig:CompTimeIEEE14bus}
 \end{figure}
@@
 The mean angles \eqref{MeanAngleInitialIEEE14} were obtained as steady state power flow solutions {\red{for the generator phase angles with constant load??
-Please complete/edit this sentence}} Larger $\kappa_{i}$ entails a larger concentration around the mean angle $\mu_{\theta_{i}}(0)$ (setting $\kappa_i=0$ implies that $\theta_{i}(0)$ follows uniform distribution over $[0,2\pi)$).
+Please complete/edit this sentence}} Larger $\kappa_{i}$ entails a higher concentration around the mean angle $\mu_{\theta_{i}}(0)$ (setting $\kappa_i=0$ implies that $\theta_{i}(0)$ follows uniform distribution over $[0,2\pi)$).
 
@@
 Thus, the PDF $\rho_0$ in \eqref{rho0IEEE14} is supported over $\mathbb{T}^{5}\times [-0.1,0.1]$.
+
+With step size $h=10^{-3}$, we discretized time $t\in[0,1]$.
+Using Algorithm \ref{algo:KineticProx}, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
+We used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $L=300$.
+
+Fig.
+\ref{fig:IEEE14marginals} shows the time snapshots of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs for generator $i=1,\hdots, n=5$, computed using the weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
+Notice that in Fig.
+\ref{fig:IEEE14marginals}, the horizontal axes of the subfigures have periodic boundary conditions, i.e., $(\theta_{i},\omega_{i})\in\mathbb{T}\times\mathbb{R}$ (cylinder).
+Fig.
+\ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via Algorithm \ref{algo:KineticProx}.
 
@@
 With $N=2000$ random samples from the initial joint PDF \eqref{rho0synthetic}, and with the aforesaid randomly generated parameters, we employed the Euler-Maruyama scheme together with Algorithm \ref{algo:KineticProx} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
-We used $h=10^{-3}$, $\varepsilon = 0.05$, $\delta = 10^{-3}$, $L = 300$.
+As in Sec.
+\ref{subsec:NumSimIEEEBus}, we used $h=10^{-3}$, $\varepsilon = 0.05$, $\delta = 10^{-3}$, $L = 300$.
 
@@
 
-\begin{IEEEbiography}[{\includegraphics[height=1.25in]{Abhishek-Halder.jpg}}]{Abhishek
-Halder}
-(S'10-M'14-SM'21) is an Assistant Professor in the Department of Applied Mathematics,
-and is an affiliated faculty in the Department of Electrical and Computer Engineering at
-University of California, Santa Cruz. Before that he held postdoctoral
-positions in the Department of Mechanical and Aerospace Engineering at
-University of California, Irvine, and in the Department of Electrical
-and Computer Engineering at Texas A\&M University. He obtained his
-Bachelors and Masters from Indian Institute of Technology Kharagpur in
-2008, and Ph.D. from Texas A\&M University in 2014, all in Aerospace
-Engineering. His research interests are in stochastic systems, control, learning and optimization with application focus on large scale cyber-physical
-systems.
-\end{IEEEbiography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{XinboGeng.jpg}}]{Xinbo Geng} (S'13-M'20)
-received his B.E.
+%\begin{IEEEbiography}[{\includegraphics[height=1.25in]{Abhishek-Halder.jpg}}]{Abhishek
+%Halder}
+%(S'10-M'14-SM'21) is an Assistant Professor in the Department of Applied Mathematics,
+%and is an affiliated faculty in the Department of Electrical and Computer Engineering at
+%University of California, Santa Cruz. Before that he held postdoctoral
+%positions in the Department of Mechanical and Aerospace Engineering at
+%University of California, Irvine, and in the Department of Electrical
+%and Computer Engineering at Texas A\&M University. He obtained his
+%Bachelors and Masters from Indian Institute of Technology Kharagpur in
+%2008, and Ph.D. from Texas A\&M University in 2014, all in Aerospace
+%Engineering. His research interests are in stochastic systems, control, learning and optimization with application focus on large scale cyber-physical
+%systems.
+%\end{IEEEbiography}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{XinboGeng.jpg}}]{Xinbo Geng} (S'13-M'20)
+%received his B.E.
 degree in Electrical Engineering from Tsinghua University, Beijing, China, in 2013.
@@
 degrees from Department of Electrical and Computer Engineering, Texas A\&M University, College Station, TX, in 2015 and 2019, respectively.
-He is a postdoctoral researcher jointly affiliated with Cornell University and Texas A\&M University since 2020.
-He was a power system design and studies intern at National Renewable Energy Lab in 2016, a data analyst intern at ISO New England in 2017, a visiting student at UC Berkeley and MIT in 2018, and a visiting student at University of Wisconsin Madison in 2019.
-His main research interest is decision making under uncertainties.
-\end{IEEEbiography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{IEEEbiography}[{\includegraphics[height=1.25in]{kenneth.jpg}}]{Kenneth F. Caluya}
-is a Ph.D.
+%He is a postdoctoral researcher jointly affiliated with Cornell University and Texas A\&M University since 2020.
+%He was a power system design and studies intern at National Renewable Energy Lab in 2016, a data analyst intern at ISO New England in 2017, a visiting student at UC Berkeley and MIT in 2018, and a visiting student at University of Wisconsin Madison in 2019.
+%His main research interest is decision making under uncertainties.
+%\end{IEEEbiography}
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%\begin{IEEEbiography}[{\includegraphics[height=1.25in]{kenneth.jpg}}]{Kenneth F. Caluya}
+%is a Ph.D.
 student in the Department of Applied Mathematics at University of California, Santa Cruz.
@@
 His research interests lie at the intersection of probability, convex optimization, partial differential equations, and dynamical systems.
-\end{IEEEbiography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{IEEEbiography}[{\includegraphics[height=1.25in]{parmita.png}}]{Pegah Ojaghi}
-is a Ph.D.
+%\end{IEEEbiography}
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%\begin{IEEEbiography}[{\includegraphics[height=1.25in]{parmita.png}}]{Pegah Ojaghi}
+%is a Ph.D.
 student in the Department of Computer Science and Engineering at University of California, Santa Cruz.
@@
 in Control Engineering from the Hamedan University of Technology, Hamedan, Iran in 2012.
-\end{IEEEbiography}
+%\end{IEEEbiography}
 

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 
+All simulations were performed in MATLAB R2019b on an iMac with 3.4 GHz Quad-Core Intel Core i5 processor and 8 GB memory.
+
 \begin{figure*}[t]
@@
 \end{align}
-whose $\bm{\theta}$ marginal $ \vartheta_{0}\left(\bm{\theta}(0)\right)$ is a product von Mises PDF \cite{mardia2008multivariate,mardia2014some} supported over $\mathbb{T}^{5}$ with mean angles (in rad)
+whose $\bm{\theta}$ marginal $ \vartheta_{0}\left(\bm{\theta}(0)\right)$ is a product von Mises PDF\footnote{the factor $2$ inside the cosine ensures the range $[0,2\pi)$; it can be dispensed if we instead use the range $[-\pi,\pi)$ as common in directional statistics literature.} \cite{mardia2008multivariate,mardia2014some} supported over $\mathbb{T}^{5}$ with mean angles (in rad)
 \begin{align}
@@
 \[(\kappa_{1},\hdots,\kappa_5) = \left(5,6,7,4,5\right).\]
+In \eqref{rho0IEEE14}, $I_{0}(\cdot)$ denotes the modified Bessel function of the first kind with order zero, and can be evaluated via MATLAB command \texttt{besseli(0,$\cdot$)}.
+
 The mean angles \eqref{MeanAngleInitialIEEE14} were obtained as steady state power flow solutions {\red{for the generator phase angles with constant load??

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \[(\kappa_{1},\hdots,\kappa_5) = \left(5,6,7,4,5\right).\]
-In \eqref{rho0IEEE14}, $I_{0}(\cdot)$ denotes the modified Bessel function of the first kind with order zero, and can be evaluated via MATLAB command \texttt{besseli(0,$\cdot$)}.
+In \eqref{rho0IEEE14}, $I_{0}(\cdot)$ denotes the modified Bessel function of the first kind with order zero, given by $I_{0}(\kappa) = \sum_{r=0}^{\infty} \frac{\left(\kappa^{2}/4\right)^{r}}{(r!)^{2}}$, and can be evaluated via MATLAB command \texttt{besseli(0,$\cdot$)}.
 

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 With step size $h=10^{-3}$, we discretized time $t\in[0,1]$.
-Using Algorithm \ref{algo:KineticProx}, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
+Using the Euler-Maruyama scheme and the Algorithm \ref{algo:KineticProx}, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
 We used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $L=300$.
@@
 Fig.
-\ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via Algorithm \ref{algo:KineticProx}.
+\ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the Euler-Maruyama and the Algorithm \ref{algo:KineticProx}.
 

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \end{align}
-whose $\bm{\theta}$ marginal $ \vartheta_{0}\left(\bm{\theta}(0)\right)$ is a product von Mises PDF\footnote{the factor $2$ inside the cosine ensures the range $[0,2\pi)$; it can be dispensed if we instead use the range $[-\pi,\pi)$ as common in directional statistics literature.} \cite{mardia2008multivariate,mardia2014some} supported over $\mathbb{T}^{5}$ with mean angles (in rad)
+whose $\bm{\theta}$ marginal $ \vartheta_{0}\left(\bm{\theta}(0)\right)$ is a product von Mises PDF\footnote{the factor $2$ inside the cosine ensures the range $[0,2\pi)$; it can be dispensed if we instead use the range $[-\pi,\pi)$ for the angular variables, as common in the directional statistics literature \cite{mardia2009directional}.} \cite{mardia2008multivariate,mardia2014some} supported over $\mathbb{T}^{5}$ with mean angles (in rad)
 \begin{align}
@@
 Multivariate von Mises Distribution", \emph{Communications in Statistics - Theory and Methods}, Vol. 43, No. 6, pp. 1132--1144, 2014.
+
+\bibitem{mardia2009directional}
+K.V.~Mardia, and P.E.~Jupp, \emph{Directional Statistics}, John Wiley \& Sons, 2009.
 

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \subsubsection{Contributions of this paper}
-Our contributions are twofold. {\red{TBD}}
-
-Our technical approach juxtaposes with the existing literature discussed above, in that we approximate neither the dynamical nonlinearity nor the statistics.
+Our main contribution is to demonstrate that by harnessing recent developments in generalized gradient flows \cite{ambrosio2008gradient} and proximal algorithms \cite{parikh2014proximal}, it is possible to perform \emph{nonparametric} propagation of the joint state PDFs subject to the stochastic nonlinear dynamics of a networked power system, and that the computation can be implemented in a gridless and scalable manner.
+
+Our technical approach juxtaposes with the existing power systems literature discussed above, in that we approximate neither the dynamical nonlinearity nor the statistics.
 Instead of treating the exact nonlinearity as bane, we exploit the geometry induced by the power system's dynamical nonlinearity over the manifold of time-varying joint state PDFs, thereby enabling a new proximal algorithm to compute the transient joint PDFs.
@@
 \centering
-\includegraphics[width=0.95\linewidth]{IEEE14marginals.png}
+\includegraphics[width=0.9\linewidth]{IEEE14marginals.png}
 \caption{\small{Time evolution of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs, where $i=1,\hdots,5$, for the Kron-reduced IEEE 14 bus simulation setup described in Sec.
@@
 
+% Proximal algorithm
+% =====================================
+\bibitem{parikh2014proximal}
+N.~Parikh, and S.~Boyd, ``Proximal Algorithms", \emph{Foundations and Trends in Optimization}, Vol. 1, No. 3, pp. 127--239, 2014.
+
 % == Kuramoto second order ===
@@
 \bibitem{mardia2009directional}
-K.V.~Mardia, and P.E.~Jupp, \emph{Directional Statistics}, John Wiley \& Sons, 2009.
+K.V.~Mardia, and P.E.~Jupp, \emph{Directional Statistics}, Wiley, 1st edition, 1999.
 

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \subsubsection{Contributions of this paper}
-Our main contribution is to demonstrate that by harnessing recent developments in generalized gradient flows \cite{ambrosio2008gradient} and proximal algorithms \cite{parikh2014proximal}, it is possible to perform \emph{nonparametric} propagation of the joint state PDFs subject to the stochastic nonlinear dynamics of a networked power system, and that the computation can be implemented in a gridless and scalable manner.
+Our main contribution is to demonstrate that by harnessing recent developments in generalized gradient flows \cite{ambrosio2008gradient} and proximal algorithms \cite{parikh2014proximal}, it is possible to perform \emph{nonparametric} propagation of the joint state PDFs subject to the stochastic nonlinear dynamics of a networked power system, and that the computation can be performed in a gridless and scalable manner.
 

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \begin{IEEEkeywords}
-Uncertainty propagation, power system dynamics, optimal transport, proximal operator.
+Uncertainty propagation, power system dynamics, kinetic Fokker-Planck equation, proximal operator.
 \end{IEEEkeywords}
@@
 Section \ref{sec:ProxAlgo} elucidates the corresponding proximal algorithm that enables the computation of the transient joint state PDFs via weighted point cloud evolution.
-Numerical simulations are reported in Section \ref{sec:NumericalSimulations}.
+Numerical simulations illustrating the proposed framework are reported in Section \ref{sec:NumericalSimulations}.
 Section \ref{sec:conclusion} concludes the paper.

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \begin{align}
-U(\bm{\xi}) &:= \!\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\!\!
-\displaystyle\sum_{(i,j)\in\mathcal{E}}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} + \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right), \label{defU}\\
+U(\bm{\xi}) &:= \!\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\displaystyle\sum_{i,j=1}^{n}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} + \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right), \label{defU}\\
 F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
@@
 With step size $h=10^{-3}$, we discretized time $t\in[0,1]$.
-Using the Euler-Maruyama scheme and the Algorithm \ref{algo:KineticProx}, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
-We used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $L=300$.
+Using the algorithm detailed in Sec.
+\ref{sec:ProxAlgo}, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
+In Algorithm \ref{algo:KineticProx}, we used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $L=300$.
 
@@
 
-With $N=2000$ random samples from the initial joint PDF \eqref{rho0synthetic}, and with the aforesaid randomly generated parameters, we employed the Euler-Maruyama scheme together with Algorithm \ref{algo:KineticProx} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
+With $N=2000$ random samples from the initial joint PDF \eqref{rho0synthetic}, and with the aforesaid randomly generated parameters, we employed the procedure detailed in Sec.
+\ref{sec:ProxAlgo} for propagating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
 As in Sec.
-\ref{subsec:NumSimIEEEBus}, we used $h=10^{-3}$, $\varepsilon = 0.05$, $\delta = 10^{-3}$, $L = 300$.
+\ref{subsec:NumSimIEEEBus}, we used $h=10^{-3}$, $\varepsilon = 0.05$, $\delta = 10^{-3}$, $L = 300$ in Algorithm \ref{algo:KineticProx}.
 

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \includegraphics[width=0.9\linewidth]{IEEE14marginals.png}
-\caption{\small{Time evolution of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs, where $i=1,\hdots,5$, for the Kron-reduced IEEE 14 bus simulation setup described in Sec.
+\caption{\small{Time evolution of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs of the generator $i=1,\hdots,5$, for the Kron-reduced IEEE 14 bus simulation setup described in Sec.
 \ref{subsec:NumSimIEEEBus}.

### 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 Our main contribution is to demonstrate that by harnessing recent developments in generalized gradient flows \cite{ambrosio2008gradient} and proximal algorithms \cite{parikh2014proximal}, it is possible to perform \emph{nonparametric} propagation of the joint state PDFs subject to the stochastic nonlinear dynamics of a networked power system, and that the computation can be performed in a gridless and scalable manner.
+In doing so, the paper also makes theoretical contribution by introducing a change of coordinates that transforms the joint PDF evolution equations in way that is amenable for the aforesaid generalized proximal recursions.
 
@@
 Fig.
-\ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the Euler-Maruyama and the Algorithm \ref{algo:KineticProx}.
+\ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec.
+\ref{sec:ProxAlgo}.
 

### 2021-07-09 code/IEEE14_prox_simulation.m
adding IEEE 14 bus sim results

@@ -235,6 +235,8 @@ for j=1:num_Oscillator
         if j==1
             title(['$t=$' num2str(t_now_vec(tt))],'interpreter','latex')
         end
+        set(gca,'XTick',0:pi/2:2*pi, 'XTickLabel',{'$0$','$\frac{\pi}{2}$','$\pi$','$\frac{3\pi}{2}$','$2\pi$'})
+        set(get(gca,'XAxis'),'TickDir','out')
         axis tight
         hold on
     end    

### 2021-07-09 Data/create_network_reduced_files.m
creating 3 cases for simulation, generated data included

@@ -1,0 +1,95 @@
+function create_network_reduced_files(mpc,  second_series, gen_series, PD_series, QD_series, VM_series, VA_series, note)
+
+define_constants;
+nbus = size(mpc.bus,1);
+
+
+nt = size(second_series,1);
+ngen = size(gen_series,1);
+
+assert(nt==size(gen_series,2) ); assert(ngen==size(gen_series,1));
+assert(nt==size(PD_series,2) ); assert(nbus==size(PD_series,1) );
+assert(nt==size(QD_series,2) ); assert(nbus==size(QD_series,1) );
+assert(nt==size(VM_series,2) ); assert(nbus==size(VM_series,1) );
+assert(nt==size(VA_series,2) ); assert(nbus==size(VA_series,1) );
+
+z = mpc.branch(:,BR_R)+i*mpc.branch(:,BR_X);
+y = 1./z;
+k = abs(y);
+phi = angle(y);
+
+% line parameters
+lines = [mpc.branch(:,[F_BUS,T_BUS,BR_R, BR_X]), k, phi];
+line_header = {'from-bus','no-bus','resistance-R','reactance-X','admittance-y-magnitude','admittance-y-angle'};
+write_csv_with_header([note,'_line_parameters.csv'],line_header,lines);
+
+% bus initial states (at equilibrium pt)
+buses = [mpc.bus(:,[BUS_I, VM]),deg2rad(mpc.bus(:,VA))];
+writematrix(buses,'bus-init.csv');
+
+% generator initial states (at equilibrium pt)
+w0 = 2*pi*60;
+H = [2.64; 2.61; 3.42; 2.45; 3.59];
+D = [4; 4; 4; 4; 4];
+m = 2*w0.*H;
+gamma = w0.*D; 
+
+gens = [mpc.gen(:,GEN_BUS), H,D,m,gamma];
+gen_header = {'BusNum(in 14 bus)','H','D','m','gamma'};
+write_csv_with_header([note,'_gen_parameters.csv'],gen_header,gens);
+
+% Kron Reduction
+Y = makeYbus(mpc);
+
+
+inner_bus_ind = find(mpc.bus(:,BUS_TYPE)==PQ);
+boundary_bus_ind = setdiff(1:nbus, inner_bus_ind);
+
+Y_red = Y(boundary_bus_ind ,boundary_bus_ind ) ...
[…]

### 2021-07-09 Data/report.tex

\documentclass[a4paper,11pt]{article}

\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{color}
\usepackage[top=1in,bottom=1in,margin=1in]{geometry}

\title{Structure Preserving Power Systems Model}
% sort your names alphabetically by last name
\author{
  Abhishek Halder
}
\date{} % change this accordingly

\begin{document}

\maketitle

\subsection*{Dynamics}
We consider a power network with $n>1$ generators, where the $i$th generator's dynamics is modeled as
\begin{subequations}
\begin{align}
   \dot{\theta}_{i} &= \omega_{i},\\
   m_{i}\dot{\omega}_{i} &= P_{i} - \gamma_{i}\omega_{i} - \displaystyle\sum_{j=1}^{n}k_{ij}\sin(\theta_{i}-\theta_{j} - \varphi_{ij}).
\end{align}
\label{GenDyn}
\end{subequations}
The state variables are the \textbf{rotor angles} $\theta_{i}\in\mathbb{S}^{1}\equiv [0,2\pi)$, and the \textbf{rotor angular velocities} $\omega_{i}\in\mathbb{R}$ for $i=1,...,n$. Recall that the $n$-torus $\mathbb{T}^{n} = \underbrace{\mathbb{S}^{1} \times \hdots \times \mathbb{S}^{1}}_{n \; \text{times}}$. Thus the state space for (\ref{GenDyn}) is $\mathbb{T}^{n}\times\mathbb{R}^{n}$.

\subsection*{Parameters}
The parameters $m_{i}>0,\gamma_{i}>0$ respectively denote the inertia and damping coefficient for the $i$th generator.

The other three parameters: the \textbf{effective power input} $P_{i}$, the \textbf{phase shift} $\varphi_{ij}\in[0,\frac{\pi}{2})$, and the \textbf{coupling} $k_{ij}\geq 0$, depend on the network reduced admittance matrix $Y\in\mathbb{C}^{n\times n}$.

Specifically,
\begin{subequations}
\begin{align}
 P_{i} &= P_{i}^{\text{mech}} - E_{i}^{2} \Re\left(Y_{ii}\right), \\
 \varphi_{ij} &= \begin{cases} -\arctan\left(\dfrac{\Re\left(Y_{ij}\right)}{\Im\left(Y_{ij}\right)}\right), & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}\\
 k_{ij} &= \begin{cases} E_{i}E_{j} \vert Y_{ij}\vert, & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}
\end{align}
\end{subequations}
where $P_{i}^{\text{mech}}$ is the mechanical power input, and $E_{i}$ is the internal voltage (magnitude) for generator $i$.

\subsection*{What we need}
The IEEE 14 bus has 5 generators. We want $m_i, \gamma_i, P_{i}^{\text{mech}}, E_{i}$ and the $5\times 5$ admittance matrix $Y$ for the IEEE 14 bus test system.

\section{Generator Parameters} % (fold)
\label{sec:generator_parameters}
Important notes:
\begin{itemize}
\item \textbf{angles $\theta$ and rotor speeds $\omega$ must be in rad (NOT deg).}
\item \textbf{everything is in per unit (p.u)}. For example, $P_{\text{elec}} = 0.8$p.u. with $\text{baseMVA}=100$ means $P_{\text{elec}} = 80$MVA.
\end{itemize}
\textcolor{red}{Question: where is the voltage magnitude $|V_i|$ of bus $i$? \eqref{GenDyn} only has angles.}

GENCLS Model:
\begin{subequations}
\begin{align}
& \dot{\theta}  = \omega_0 \omega \\
& \dot{\omega} = \frac{1}{2H}(P_{\text{mech}} - D \omega - P_{\text{elec}})
\end{align}
\end{subequations}
Comparing GENCLS model with \eqref{GenDyn}:
\begin{subequations}
\begin{align}
& m = 2\omega_0 H \\
& \gamma = \omega_0 D
\end{align}
\end{subequations}
\begin{table}[tb]
	\caption{Generator Parameters for GENCLS Model ($\omega_0 = 2\pi\times 60\text{Hz}$)}
	\label{tab:GENCLS_Parameters}
	\centering

	\begin{tabular}{c|c|cccc}
	\hline

	\hline
	\textbf{Generator} & \textbf{Bus} & $H$ & $D$ & $m = 2\omega_0 H$ & $\gamma = \omega_0 D$ \\
	\hline
	1	 & 1 & 2.64 & 4.00 & $1.9905 \times 10^3$ & $1.5080 \times 10^3$ \\
	2	 & 2 & 2.61 & 4.00 & $1.9679 \times 10^3$ & $1.5080 \times 10^3$ \\
	3	 & 3 & 3.42 & 4.00 & $2.5786 \times 10^3$ & $1.5080 \times 10^3$ \\
	4	 & 6 & 2.45 & 4.00 & $1.8473 \times 10^3$ & $1.5080 \times 10^3$ \\
	5	 & 8 & 3.59 & 4.00 & $2.7068 \times 10^3$ & $1.5080 \times 10^3$ \\
	\hline

	\hline
	\end{tabular}
\end{table}
Other parameters are in CSV files
\begin{description}
\item[case14.m] Matpower casefile,
\item[get\_parameters\_case14.m] Matlab script to get all scripts and write into csv files
\item[gen\_parameters.csv] everything in Table \ref{tab:GENCLS_Parameters}
\item[bus-init.csv] about initial values of $\theta_i$
\item[line-parameters.csv] everything about lines, $i,j,k_{ij},\phi_{ij}$
\end{description}
When simulating line failures, you can remove any one of 20 lines of the 14-bus system (any line in line-parameters.csv).

\section{Load Modeling} % (fold)
\label{sec:load_modeling}
The most commonly used load modeling is ZIP load:
\begin{itemize}
\item Z: constant impedance Z
\item I: constant current I
\item P: constant power P
\end{itemize}
So given the voltage at bus $i$, the real load is calculated as
\begin{equation}
P^{\text{load}}_i = P_i + \text{Real}(V_i I_i^\dagger) + \text{Real}(V_i * (V_i/Z_i)^\dagger)
\end{equation}
Note we are ignoring reactive load (imaginary parts) in the swing equation.

\subsection{Load Models in Kron Reduction} % (fold)
In Florian's slides, they are using constant current and impedance loads (see Figure on page 8/41).
\begin{itemize}
\item the constant impedance part is modeled as self-loops in Kron reduction (see Page 33/41), then integrated in the $Y_{red}$ matrix
\item the constant current load can be considered in the Kron reduction, more details below
\item I believe Kron reduction has trouble dealing with constant power load, so Florian (and us) don't consider it (for load buses).
\end{itemize}

% subsection constant_current_load (end)

\subsection{Constant Current Load} % (fold)
\label{sub:constant_current_load}
Assume all load at non-generator buses are modeled as constant current load, with current $I_i$ $i \in \mathcal{L}$.

Original equations with constant current load
\begin{subequations}
\begin{eqnarray}
M_i \ddot{\theta} + D_i \dot{\theta_i} = P_{mech,i} - P_{load,i} - \sum_{j} |V_i| |V_j| |Y_{ij}| \sin(\theta_i - \theta_j + \phi_{ij}) &~ i \in \mathcal{G}~\text{buses with generators} \\
0 = - \text{Real}(V_i I_i^\dagger) - \sum_{j} |V_i| |V_j| |Y_{ij}| \sin(\theta_i - \theta_j + \phi_{ij}) &~ i \in \mathcal{L}~\text{buses without generators} \\
\end{eqnarray}
\end{subequations}
Note: $P_{mech,i} \in \mathbb{R}$, $P_{load,i} \in \mathbb{R}$ and $I_i \in \mathbb{C}$ are all give parameters; load current is a complex number.

After Kron reduction
\begin{eqnarray}
Y_{red} &=& Y_{\mathcal{G}\mathcal{G}} - Y_{\mathcal{G}\mathcal{L}} Y_{\mathcal{L}\mathcal{L}}^{-1} Y_{\mathcal{L}\mathcal{G}} \\
I_{red} &=& Y_{\mathcal{G}\mathcal{L}}Y_{\mathcal{L}\mathcal{L}}^{-1} I
\end{eqnarray}

\begin{eqnarray}
M_i \ddot{\theta} + D_i \dot{\theta_i} = P_i - \sum_{j=1, j\ne i}^n |V_i| |V_j| |Y_{red,ij}| \sin(\theta_i - \theta_j + \phi_{red,ij}) - Y_{},\\
~i \in \mathcal{G}~\text{buses with generators}
\end{eqnarray}
where 
\begin{eqnarray}
P_i = P_{mech,i} - P_{load,i} - |V_i|^2 \text{Real}(Y_{red,ii}) + |V_i| |I_{red,i}| \cos(\theta_i - \alpha_i)
\end{eqnarray}

These files/parameters are given
\begin{description}
\item[gen\_parameters.csv] everything in Table \ref{tab:GENCLS_Parameters}
\item[bus-init.csv] about initial values of $\theta_i$
\item[line-parameters.csv] everything about lines, $i,j,k_{ij},\phi_{ij}$
\end{description}
Additional:
\begin{description}
\item[case14.m] Matpower casefile,
\item[get\_parameters\_case14.m] Matlab script to get all scripts and write into csv files
\end{description}
% subsection constant_current_load (end)

% section load_modeling (end)

\end{document}

### 2021-07-09 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results
add references, minor changes in the Numerical simulation section

@@
 \label{fig:Motivating}
+\vspace*{-0.1in}
 \end{figure}
@@
 \label{fig:UncertaintyPropagation}
+\vspace*{-0.1in}
 \end{figure}
@@
 \centering
-\includegraphics[width=.8\linewidth]{BlockDiagm.pdf}
+\includegraphics[width=.7\linewidth]{BlockDiagm.pdf}
 \caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
@@
 \label{fig:BlockDiagm}
+\vspace*{-0.1in}
 \end{figure}
@@
 \ref{fig:Motivating}.
-{\red{The parameter values in \eqref{DefParam} use the network reduced admittance matrix computed from the CITE??
-The parameters $m_{i}, \gamma_{i}$ are obtained from ....}} We randomly generated $\sigma_{i}\in [1,5]$ as
+{\red{The original parameters of the 14-bus system are from Matpower \cite{zimmerman2010matpower}.
+The calculation of network reduced admittance matrices in \eqref{DefParam} is described in \ref{subsec:samplepathdynamics} and \cite{dorfler2012kron}.
+The parameters $m_{i}, \gamma_{i}$ are obtained from ANDES \cite{cui2021andes}.}}
+We randomly generated $\sigma_{i}\in [1,5]$ as
 \[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]
@@
 
-The mean angles \eqref{MeanAngleInitialIEEE14} were obtained as steady state power flow solutions {\red{for the generator phase angles with constant load??
-Please complete/edit this sentence}} Larger $\kappa_{i}$ entails a higher concentration around the mean angle $\mu_{\theta_{i}}(0)$ (setting $\kappa_i=0$ implies that $\theta_{i}(0)$ follows uniform distribution over $[0,2\pi)$).
+The mean angles \eqref{MeanAngleInitialIEEE14} were obtained from steady state AC power flow solutions.
+Larger $\kappa_{i}$ entails a higher concentration around the mean angle $\mu_{\theta_{i}}(0)$ (setting $\kappa_i=0$ implies that $\theta_{i}(0)$ follows uniform distribution over $[0,2\pi)$).
 
@@
 \label{table:ParamUncertaintiesSynthetic}
+\vspace*{-0.1in}
 \end{table}
@@
 K.V.~Mardia, and P.E.~Jupp, \emph{Directional Statistics}, Wiley, 1st edition, 1999.
+
+\bibitem{zimmerman2010matpower}
+R.D.~Zimmerman, C.E.~Murillo-S{\'a}nchez, R.J.
+Thomas, ``MATPOWER: Steady-state operations, planning, and analysis tools for power systems research and education'', \emph{IEEE Transactions on power systems},  Vol.
+26, No.
+1, pp.
+12--19, 2010.
+
+\bibitem{cui2021andes}
+H.~Cui, F.~Li and K.~Tomsovic, ``Hybrid Symbolic-Numeric Framework for Power System Modeling and Analysis,'' \emph{IEEE Transactions on Power Systems}, vol. 36, no. 2, pp. 1373-1384, 2021.
 

### 2021-07-09 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

--- 2021-07-08 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 %\author{Abhishek~Halder,~\IEEEmembership{Member,~IEEE,}
-Xinbo~Geng,~\IEEEmembership{Member,~IEEE,}
         Kenneth~F. Caluya,
-        and~Pegah~Ojaghi% <-this % stops a space
+        Pegah~Ojaghi,
+        and~Xinbo~Geng,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
 \thanks{A. Halder and K.F. Caluya are with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder,kcaluya\}@ucsc.edu}}.}% <-this % stops a space
@@
 The columns correspond to the time snapshots.
-The colorbar at the bottom shows the marginal PDF values.}}
+The colorbar at the bottom shows the marginal PDF values.
+The $\theta_i$ are in rad, the $\omega_i$ are in rad/s.}}
 \vspace*{-0.1in}

### 2021-07-09 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 
+%\begin{document}
+%
+%\title{\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions}
+%
+%\author{Abhishek~Halder,~\IEEEmembership{Senior Member,~IEEE,} 
+%%\author{Abhishek~Halder,~\IEEEmembership{Member,~IEEE,}
+%        Kenneth~F. Caluya,
+%        Pegah~Ojaghi,
+%        and~Xinbo~Geng,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
+%\thanks{A.
+Halder and K.F.
+Caluya are with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder,kcaluya\}@ucsc.edu}}.}% <-this % stops a space
+%=======
+
 \begin{document}
@@
 \author{Abhishek~Halder,~\IEEEmembership{Senior Member,~IEEE,} 
-%\author{Abhishek~Halder,~\IEEEmembership{Member,~IEEE,}
         Kenneth~F. Caluya,
@@
 \thanks{A. Halder and K.F. Caluya are with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder,kcaluya\}@ucsc.edu}}.}% <-this % stops a space
+\thanks{P. Ojaghi is with the Department of Computer Science and Engineering, University of California, Santa Cruz, CA 95064, USA, {\texttt{pojaghi@ucsc.edu}}.}%
 \thanks{X. Geng is with the School of Electrical and Computer Engineering, Cornell University, Ithaca, NY 14850, USA, {\texttt{xg72@cornell.edu}}.} %
-\thanks{P. Ojaghi is with the Department of Computer Science and Engineering, University of California, Santa Cruz, CA 95064, USA, {\texttt{pojaghi@ucsc.edu}}.}%
 \thanks{This research was partially supported by the NSF award 1923278.}}% <-this % stops a space
@@
 \subsection{IEEE 14 Bus System}\label{subsec:NumSimIEEEBus}
-Following Sec.
-\ref{subsec:samplepathdynamics}, we consider the Kron-reduced dynamics \eqref{ItoSDEcomponentlevel} for the IEEE 14 bus system shown in Fig.
+We consider the Kron-reduced dynamics \eqref{ItoSDEcomponentlevel} for the IEEE 14 bus system shown in Fig.
 \ref{fig:Motivating}.
@@
 \ref{fig:Motivating}.
-{\red{The parameter values in \eqref{DefParam} use the network reduced admittance matrix computed from the CITE??
-The parameters $m_{i}, \gamma_{i}$ are obtained from ....}} We randomly generated $\sigma_{i}\in [1,5]$ as
+We obtained the parameters of the 14 bus system from MATPOWER \cite{zimmerman2010matpower}.
+The calculation of Kron-reduced admittance matrix and current vector in \eqref{DefParam} followed Sec.
+\ref{subsec:samplepathdynamics}; see also \cite{dorfler2012kron}.
+The parameters $m_{i}, \gamma_{i}$ are obtained from the open-source Python-based
+library ANDES \cite{cui2021andes}.
+
+We randomly generated the noise coefficients $\sigma_{i}\in [1,5]$ as
 \[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]
@@
 K.V.~Mardia, and P.E.~Jupp, \emph{Directional Statistics}, Wiley, 1st edition, 1999.
+
+\bibitem{zimmerman2010matpower}
+R.D.~Zimmerman, C.E.~Murillo-S{\'a}nchez, and R.J.
+Thomas, ``MATPOWER: Steady-state Operations, Planning, and Analysis Tools for Power Systems Research and Education'', \emph{IEEE Transactions on Power Systems},  Vol.
+26, No.
+1, pp.
+12--19, 2010.
+
+\bibitem{cui2021andes}
+H.~Cui, F.~Li, and K.~Tomsovic, ``Hybrid Symbolic-Numeric Framework for Power System Modeling and Analysis,'' \emph{IEEE Transactions on Power Systems}, vol. 36, no. 2, pp. 1373-1384, 2021.
 

### 2021-07-09 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 The curve $\rho(t,\bm{x})$ solves a kinetic Fokker-Planck PDE initial value problem.
+Points on this curve (shown as three filled circled markers at three specific instances $t_0,t_1,t_2$) are joint PDFs.
 While the joint PDFs are supported over the finite dimensional base manifold $\mathbb{T}^{n}\times\mathbb{R}^{n}$, the proximal updates $\{\varrho_{k}\}_{k\in\mathbb{N}}$ (shown as circled markers with no face-color) evolve over the infinite dimensional manifold $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.}}
@@
 \ref{subsec:samplepathdynamics}; see also \cite{dorfler2012kron}.
-The parameters $m_{i}, \gamma_{i}$ are obtained from the open-source Python-based
+The parameters $m_{i}, \gamma_{i}$ were obtained from the open-source Python-based
 library ANDES \cite{cui2021andes}.

### 2021-07-09 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Short.tex
adding IEEE 14 bus sim results

--- 2021-07-09 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 In Sec.
-\ref{subsec:NumSimIEEEBus}, we consider the prediction of stochastic states for the IEEE 14 bus system--{\red{our main objective being to highlight the interpretability and utility of the proposed method in power systems engineering}}.
+\ref{subsec:NumSimIEEEBus}, we consider the prediction of stochastic states for the IEEE 14 bus system--our main objective being to highlight the interpretability and utility of the proposed method in power systems engineering.
 In Sec.
@@
 \section{Conclusion}\label{sec:conclusion}
-{\red{The conclusion goes here.}}
+We proposed a nonparametric algorithm to propagate the joint state PDFs subject to networked power system dynamics with stochastic initial conditions, parameters and process noise.
+Our development is built on recent advances in generalized gradient flows on the space of PDFs, and new analytical results specific to the power system's stochastic dynamics that are presented in this paper.
+The novelty of the proposed proximal algorithm is that it does not approximate the dynamical or statistical nonlinearities.
+Instead, the algorithm allows gridless computation by actually exploiting the structure of the PDE governing the evolution of the joint state PDF that is induced by the underlying nonlinear sample path dynamics of a networked power system.
+Numerical case studies reveal that the proposed computational framework can be a scalable way to propagate stochastic uncertainties in realistic power networks.
+
+Possible future directions include generalizing the proposed method for more complex generator dynamics, and to use this framework to actively steer the uncertainties via optimal control on the space of joint PDFs.
 

### 2021-07-11 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 
-The corresponding macroscopic dynamics governing the flow of the joint state PDF $\rho(t,\bm{\theta},\bm{\omega})$ is given by a kinetic Fokker-Planck partial differential equation (PDE)
+The corresponding macroscopic dynamics governing the flow of the joint state PDF $\rho(t,\bm{\theta},\bm{\omega})$ is given by a kinetic Fokker-Planck \cite{villani2006hypocoercive} partial differential equation (PDE)
 \begin{align}
@@
 T. Ishizaki, A. Chakrabortty, and J-I. Imura. ``Graph-Theoretic Analysis of Power Systems", \emph{Proceedings of the IEEE}, Vol. 106, No. 5, pp. 931--952, 2018.
+
+\bibitem{villani2006hypocoercive}
+C.~Villani, ``Hypocoercive Diffusion Operators", \emph{in International Congress
+of Mathematicians}, Vol. 3, pp. 473--498, 2006.
 

### 2021-07-31 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results
minor changes in the manuscript, major changes sent over email

@@
         Pegah~Ojaghi,
-        and~Xinbo~Geng,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
+        and~Xinbo~Geng% <-this % stops a space
 \thanks{A. Halder and K.F. Caluya are with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder,kcaluya\}@ucsc.edu}}.}% <-this % stops a space
@@
 We emphasize here that the network-reduced model \eqref{ItoSDEcomponentlevel} is obtained from the so-called structure preserving power network model \cite{odun2012structure} after applying the Kron reduction \cite{dorfler2012kron}, and therefore, has all-to-all connection topology.
+The derivation of the Kron-reduced parameters is outlined below.
 \end{remark}
@@
 Finally, we transform the proximal updates back to the $\bm{x}\equiv\left(\bm{\theta},\bm{\omega}\right)^{\top}$ state space via the pushforward $\bm{\Psi}^{-1}$ as
-\begin{align}
-\varrho_{k}\left(\bm{\theta}_{k}^{i},\bm{\omega}_{k}^{i}\right) &= \bm{\Psi}^{-1}_{\sharp} \tilde{\varrho}_{k}\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right) \nonumber\\
-&= \left(\prod_{j=1}^{n}m_{j}/\sigma_{j}\right)^{\!\!2}\tilde{\varrho}_{k}\left(\bm{M}\bm{\Sigma}^{-1}\bm{\theta}_{k}^{i},\bm{M}\bm{\Sigma}^{-1}\bm{\omega}_{k}^{i}\right),
+% \begin{align}
+% \varrho_{k}\left(\bm{\theta}_{k}^{i},\bm{\omega}_{k}^{i}\right) &= \bm{\Psi}^{-1}_{\sharp} \tilde{\varrho}_{k}\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right) \nonumber\\
+% &= \left(\prod_{j=1}^{n}m_{j}/\sigma_{j}\right)^{\!\!2}\tilde{\varrho}_{k}\left(\bm{M}\bm{\Sigma}^{-1}\bm{\theta}_{k}^{i},\bm{M}\bm{\Sigma}^{-1}\bm{\omega}_{k}^{i}\right),
+% \label{BringBackOriginalStateSpace}	
+% \end{align}
+\begin{multline}
+\varrho_{k}\left(\bm{\theta}_{k}^{i},\bm{\omega}_{k}^{i}\right) = \bm{\Psi}^{-1}_{\sharp} \tilde{\varrho}_{k}\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right) \\
+= \left(\prod_{j=1}^{n}m_{j}/\sigma_{j}\right)^{\!\!2}\tilde{\varrho}_{k}\left(\bm{M}\bm{\Sigma}^{-1}\bm{\theta}_{k}^{i},\bm{M}\bm{\Sigma}^{-1}\bm{\omega}_{k}^{i}\right),
 \label{BringBackOriginalStateSpace}	
-\end{align}
+\end{multline}
 for all $i=1,\hdots,N$.
@@
 \ref{fig:Motivating}.
-We obtained the parameters of the 14 bus system from MATPOWER \cite{zimmerman2010matpower}.
+We obtained the parameters of the 14-bus system from MATPOWER \cite{zimmerman2010matpower}.
 The calculation of Kron-reduced admittance matrix and current vector in \eqref{DefParam} followed Sec.
@@
 We take the initial joint state PDF at $t=0$ as
-\begin{align}
-\rho_{0}\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = &\;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right)\nonumber\\
-	&\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
+% \begin{align}
+% \rho_{0}\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = &\;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right)\nonumber\\
+% 	&\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
+% \label{rho0synthetic}	
+% \end{align}
+\begin{multline}
+\rho_{0}\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = \;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right) \\
+	\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
 \label{rho0synthetic}	
-\end{align}
+\end{multline}
 and following \cite[Sec.

### 2021-08-03 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

--- 2021-07-11 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 
-Notice that the proximal recursions given by \eqref{InfiniteDimProxOpDef}-\eqref{ProxRecursionInfiniteDim} define an infinite dimensional gradient descent of the functional $h\Phi$ over $\mathcal{P}_{2}$ w.r.t.
+We can interpret the proximal operator \eqref{InfiniteDimProxOpDef} as a generalized gradient step for the functional $\Phi$ in the manifold $\mathcal{P}_{2}$.
+The proximal recursions \eqref{ProxRecursionInfiniteDim} define an infinite dimensional gradient descent of the functional $h\Phi$ over $\mathcal{P}_{2}$ w.r.t.
 the distance ${\rm{dist}}$.

### 2021-08-03 code/IsoContour.m

@@ -1,0 +1,207 @@
+function [Q,iv,VT,H]=IsoContour(TR,F,iv,vis)
+% Retrieve line segments comprising level-sets of a scalar field defined at
+% the vertices of triangular surface mesh.
+%
+% INPUT:
+%   - TR    : input surface mesh represented as an object of 'TriRep' 
+%             class, 'triangulation' class, or a cell such that TR={Tri,X},
+%             where Tri is an M-by-3 array of faces and X is an N-by-3 
+%             array of vertex coordinates. 
+%   - F     : N-by-1 array specifiying values of the scalar field at the 
+%             mesh vertices. 
+%   - iv    : 1-by-K vector specifying either the values OR number 
+%             level-sets. In case of the latter, level-sets will be 
+%             distributed uniformly between the extremal values of F.
+%             To get a level-set at one specific value of F (e.g., Fs), 
+%             specify iv as Fs*[1 1];
+%   - vis   : axes handle (or logical value) indicating where (or whether)
+%             the computed level-sets should be plotted. 
+%
+% OUTPUT: 
+%   - Q     : 1-by-K cell containing coordinates of unorderdered line 
+%             segments comprising the level-sets.
+%   - iv    : 1-by-K array of level-set values.
+%   - VT    : 1-by-K cell containg linear interpolation coefficients along 
+%             the with indices of mesh vertices that bound the edges 
+%             intersected by the computed level-sets.
+%
+% AUTHOR: Anton Semechko (a.semechko@gmail.com)
+%
+
+
+if nargin<2 || isempty(TR) || isempty(F)
+   error('Insufficient number of input arguments') 
+end
+
+[Tri,X,fmt]=GetMeshData(TR);
+
+F=F(:);
+if ~isnumeric(F) || ~ismatrix(F) || numel(F)~=size(X,1) || sum(isnan(F) | isinf(F))>0
+    edit('Invalid entry for 2nd input argument (F)')
+end
+
[…]

### 2021-08-03 code/ThreeCases_IEEE14_prox_simulation.m

@@ -1,14 +1,15 @@
 close all; clear all; clc;
+set(groot,'defaultAxesTickLabelInterpreter','latex');  
+set(groot,'defaulttextinterpreter','latex');
+set(groot,'defaultLegendInterpreter','latex');
 %% Parameters
 beta = 2; % inverse temperature
 % golbal parameters 
 global num_Oscillator dim Gamma M Sigma K P cc sigma m phi
 
-num_Oscillator = 5; dim  = 2*num_Oscillator; % dim = dimension of state space
-
 f0 = 60; % nominal frequency (Hz)
 
-bus_init = readmatrix('Users\abhishekhaldermac\NSF-1923278-Stochastic-Uncertainty-Management\Data\bus-init.csv');
+bus_init = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/bus-init.csv');
 
 % prefixes
 % case0_norminal_
@@ -16,13 +17,15 @@ bus_init = readmatrix('Users\abhishekhaldermac\NSF-1923278-Stochastic-Uncertaint
 % case2_time_series_
 
 % parameters for the IEEE 14 bus system
-Y_reduced_imag = readmatrix('case1_line_13_failure_Y_reduced_imag_part.csv');
-Y_reduced_real = readmatrix('case1_line_13_failure_Y_reduced_real_part.csv');
+Y_reduced_imag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_Y_reduced_imag_part.csv');
+Y_reduced_real = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_Y_reduced_real_part.csv');
 % reduced admittance matrix
 Y_reduced = complex(Y_reduced_real,Y_reduced_imag);
 
-gen_param = readmatrix('case1_line_13_failure_gen_parameters.csv');
+gen_param = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_gen_parameters.csv');
 gen_bus_idx = gen_param(:,1);
+
[…]

### 2021-08-03 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \centering
-\includegraphics[width=0.9\linewidth]{IEEE14marginals.png}
-\caption{\small{Time evolution of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs of the generator $i=1,\hdots,5$, for the Kron-reduced IEEE 14 bus simulation setup described in Sec.
+\includegraphics[width=0.9\linewidth]{IEEE14marginalsCase0.png}
+\caption{\small{Time evolution of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs of the generator $i=1,\hdots,5$, for the Kron-reduced IEEE 14 bus simulation setup \textbf{Case I} described in Sec.
 \ref{subsec:NumSimIEEEBus}.
@@
 \vspace*{-0.1in}
-\label{fig:IEEE14marginals}
+\label{fig:IEEE14marginalsCase0}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=0.9\linewidth]{IEEE14marginalsCase1.png}
+\caption{\small{Time evolution of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs of the generator $i=1,\hdots,5$, for the Kron-reduced IEEE 14 bus simulation setup \textbf{Case II} described in Sec.
+\ref{subsec:NumSimIEEEBus}.
+The $5$ rows above correspond to the $5$ generator nodes; these are buses 1, 2, 3, 6 and 8 in Fig.
+\ref{fig:Motivating}.
+The columns correspond to the time snapshots.
+The colorbar at the bottom shows the marginal PDF values.
+The $\theta_i$ are in rad, the $\omega_i$ are in rad/s.}}
+\vspace*{-0.1in}
+\label{fig:IEEE14marginalsCase1}
 \end{figure*}
@@
 Fig.
-\ref{fig:IEEE14marginals} shows the time snapshots of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs for generator $i=1,\hdots, n=5$, computed using the weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
-Notice that in Fig.
+\ref{fig:IEEE14marginalsCase0} and Fig.
+\ref{fig:IEEE14marginalsCase1} show the time snapshots of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs for \textbf{Case I} and \textbf{Case II} respectively, associated with the generator $i=1,\hdots, n=5$.
+These marginals are computed using the corresponding weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
+%Notice that in Fig.
 \ref{fig:IEEE14marginals}, the horizontal axes of the subfigures have periodic boundary conditions, i.e., $(\theta_{i},\omega_{i})\in\mathbb{T}\times\mathbb{R}$ (cylinder).
-Fig.
-\ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec.
-\ref{sec:ProxAlgo}.
+Fig. \ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec.  \ref{sec:ProxAlgo}.
 

### 2021-08-03 code/Case2_IEEE14_prox_simulation.m

@@ -1,0 +1,327 @@
+close all; clear all; clc;
+set(groot,'defaultAxesTickLabelInterpreter','latex');  
+set(groot,'defaulttextinterpreter','latex');
+set(groot,'defaultLegendInterpreter','latex');
+%% Parameters
+beta = 2; % inverse temperature
+% golbal parameters 
+global num_Oscillator dim Gamma M Sigma K P cc sigma m phi
+
+f0 = 60; % nominal frequency (Hz)
+
+bus_init = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/bus-init.csv');
+
+% prefixes
+% case0_norminal_
+% case1_line_13_failure_
+% case2_time_series_
+
+% parameters for the IEEE 14 bus system
+Y_reduced_imag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case2_time_series_Y_reduced_imag_part.csv');
+Y_reduced_real = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case2_time_series_Y_reduced_real_part.csv');
+% reduced admittance matrix
+Y_reduced = complex(Y_reduced_real,Y_reduced_imag);
+
+gen_param = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case2_time_series_gen_parameters.csv');
+gen_bus_idx = gen_param(:,1);
+
+num_Oscillator = numel(gen_bus_idx); dim  = 2*num_Oscillator;
+
+% generator bus voltage
+E_mag = bus_init(gen_bus_idx,2);
+E_phase = bus_init(gen_bus_idx,3);
+E = E_mag.*exp(1i*E_phase);
+% reduced current vector
+%I_reduced_polar = readmatrix('I_red.csv');
+%I_reduced_mag = I_reduced_polar(:,1);
+%I_reduced_phase = I_reduced_polar(:,2);
+I_reduced_mag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case2_time_series_I_red_magnitude.csv');
[…]

### 2021-08-03 code/ThreeCases_IEEE14_prox_simulation.m

@@ -17,12 +17,12 @@ bus_init = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertain
 % case2_time_series_
 
 % parameters for the IEEE 14 bus system
-Y_reduced_imag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_Y_reduced_imag_part.csv');
-Y_reduced_real = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_Y_reduced_real_part.csv');
+Y_reduced_imag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_Y_reduced_imag_part.csv');
+Y_reduced_real = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_Y_reduced_real_part.csv');
 % reduced admittance matrix
 Y_reduced = complex(Y_reduced_real,Y_reduced_imag);
 
-gen_param = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_gen_parameters.csv');
+gen_param = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_gen_parameters.csv');
 gen_bus_idx = gen_param(:,1);
 
 num_Oscillator = numel(gen_bus_idx); dim  = 2*num_Oscillator;
@@ -35,17 +35,17 @@ E = E_mag.*exp(1i*E_phase);
 %I_reduced_polar = readmatrix('I_red.csv');
 %I_reduced_mag = I_reduced_polar(:,1);
 %I_reduced_phase = I_reduced_polar(:,2);
-I_reduced_mag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_I_red_magnitude.csv');
+I_reduced_mag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_I_red_magnitude.csv');
 I_reduced_mag = (I_reduced_mag(2,2:end))';
[…]

### 2021-08-04 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results

@@
 \subsection{Synthetic Test System}\label{subsec:NumSimSynthetic}
-We next consider a power network with $n=50$ generators, and propagate the transient joint state PDFs supported over the 100 dimensional state space $\mathbb{T}^{50}\times\mathbb{R}^{50}$.
+To highlight the scalability of the proposed method, we next consider a power network with $n=50$ generators, and propagate the transient joint state PDFs supported over the 100 dimensional state space $\mathbb{T}^{50}\times\mathbb{R}^{50}$.
 We take the initial joint state PDF at $t=0$ as
@@
 \ref{fig:CompTimeSynthetic} highlights that for computing high dimensional transient joint state PDFs as in here, the proposed variational framework enjoys remarkably fast computational time.
-Notice that constructing histograms or other conventional density estimators for the joint state PDF by making a grid over the 100 dimensional state space, is computationally prohibitive.
+Notice that constructing histograms or other conventional density estimators for the joint state PDF by making a grid over the 100 dimensional state space, is computationally prohibitive in a general purpose computer.
 

### 2021-08-04 code/Case2_IEEE14_prox_simulation.m

@@ -43,12 +43,31 @@ I_reduced_phase = (I_reduced_phase(2,2:end))';
 I_reduced = I_reduced_mag.*exp(1i*I_reduced_phase);
 
 %P_mech = gen_param(:,2);
-P_mech = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case2_time_series_P_mech_in_per_unit.csv');
+P_mech_coarse = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case2_time_series_P_mech_in_per_unit.csv');
 %P_mech = (P_mech(2:end))';
-P_load = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case2_time_series_P_load_at_genbuses.csv');
+P_load_coarse = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case2_time_series_P_load_at_genbuses.csv');
 %P_load = (P_load(2:end))';
 
-t_vec_coarse = P_mech(:,1);
+h = 1e-3;                         % time step
+t_vec_coarse = P_mech_coarse(:,1);
+t_vec = t_vec_coarse(1):h:t_vec_coarse(end);
+
+%initialize
+P_mech = zeros(numel(t_vec),num_Oscillator);
+P_load = zeros(numel(t_vec),num_Oscillator);
+% interpolate the time varying powers to a finer grid
+figure
+for j=2:size(P_mech_coarse,2)-1    
+    P_mech(:,j) = interp1(t_vec_coarse,P_mech_coarse(:,j),t_vec,'spline');
+    P_load(:,j) = interp1(t_vec_coarse,P_load_coarse(:,j),t_vec,'spline');
+
+    plot(t_vec,P_mech_coarse(:,j),'LineWidth',2)    
+    xlabel('$t$','FontSize', 30,'Interpreter','latex')
+end
+set(gca,'FontSize',30)
+legend('$P_{1}^{\text{mech}}$','$P_{2}^{\text{mech}}$','$P_{3}^{\text{mech}}$','$P_{4}^{\text{mech}}$','$P_{5}^{\text{mech}}$','FontSize', 30,'Interpreter','latex');
+axis tight
+
 
 % effective power input
[…]

### 2021-08-05 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
minor changes

@@
 In doing so, the paper also makes theoretical contribution by introducing a change of coordinates that transforms the joint PDF evolution equations in way that is amenable for the aforesaid generalized proximal recursions.
+
+From a methodological perspective, while proximal algorithms are well-studied \cite{parikh2014proximal,rockafellar1976monotone,combettes2011proximal} in finite dimensional optimization context, the utility of generalized proximal recursions to compute the transient solutions of partial differential equations (PDEs) is relatively less known.
+This paper highlights this connection by solving the PDE induced by the stochastic power system dynamics.
 
@@
 
-The corresponding macroscopic dynamics governing the flow of the joint state PDF $\rho(t,\bm{\theta},\bm{\omega})$ is given by a kinetic Fokker-Planck \cite{villani2006hypocoercive} partial differential equation (PDE)
+The corresponding macroscopic dynamics governing the flow of the joint state PDF $\rho(t,\bm{\theta},\bm{\omega})$ is given by a kinetic Fokker-Planck \cite{villani2006hypocoercive} PDE
 \begin{align}
@@
 N.~Parikh, and S.~Boyd, ``Proximal Algorithms", \emph{Foundations and Trends in Optimization}, Vol. 1, No. 3, pp. 127--239, 2014.
+
+\bibitem{rockafellar1976monotone}
+R.T.~Rockafellar, ``Monotone Operators and the Proximal Point Algorithm", \emph{SIAM Journal on Control and Optimization}, Vol. 14, No. 5, pp. 877--898, 1976.
+
+\bibitem{combettes2011proximal}
+P.L.~Combettes, and J-C.~Pesquet, ``Proximal Splitting Methods in Signal Processing", in \emph{Fixed-point algorithms for inverse problems in science and engineering}, pp. 185--212, 2011.
 

### 2021-08-08 code/ThreeCases_IEEE14_prox_simulation.m
adding IEEE 14 bus sim results

@@ -178,21 +178,21 @@ norm_diff_mean_mc_vs_prox = sqrt(sum((mean_mc - mean_prox).^2,1))./sqrt(sum(mean
 
 %% plots
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-figure(1)
-semilogy(t_vec, comptime, 'LineWidth', 2)
-set(gca,'FontSize',30)
-xlabel('Physical time $t=kh$ [s]','FontSize',30)
-ylabel('Computational time [s]','FontSize',30)
-% ylim([1e-3 1.2e-2])
-% YTick = [2e-3 8e-3 2e-2];
-% YTickLabels = cellstr(num2str(round(log10(YTick(:))), '10^%d'));
-grid on
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-figure(2)
-semilogy(t_vec(4:end), norm_diff_mean_mc_vs_prox(5:end),'-k','Linewidth',2)
-set(gca,'FontSize',30)
-xlabel('Physical time $t=kh$ [s]','FontSize',30)
-ylabel('Realtive error $\frac{\|\mu_{\rm{MC}}-\mu_{\rm{Prox}}\|_{2}}{\|\mu_{\rm{MC}}\|_{2}}$','FontSize',30,'interpreter','latex')
+% figure(1)
+% semilogy(t_vec, comptime, 'LineWidth', 2)
+% set(gca,'FontSize',30)
+% xlabel('Physical time $t=kh$ [s]','FontSize',30)
+% ylabel('Computational time [s]','FontSize',30)
+% % ylim([1e-3 1.2e-2])
+% % YTick = [2e-3 8e-3 2e-2];
+% % YTickLabels = cellstr(num2str(round(log10(YTick(:))), '10^%d'));
+% grid on
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% figure(2)
+% semilogy(t_vec(4:end), norm_diff_mean_mc_vs_prox(5:end),'-k','Linewidth',2)
+% set(gca,'FontSize',30)
[…]

### 2021-08-08 code/ThreeCases_IEEE14_prox_simulation.m
adding IEEE 14 bus sim results

@@ -240,13 +240,16 @@ xt = [1 0 -1 0 -0.5]; yt = [0 1 -0.64 -1 -1.25];
 figure('color','w')
 for j=1:num_Oscillator
     for tt=1:length(t_now_idx_vec)
-        
+        % univariate omega marginals
+        [omega_save1D{tt,j},marg1D_omega_save{tt,j}] = getMarginal1D(omega_upd(:,j,tt),rho_theta_omega_upd(:,tt),nBins);
+
+        % bivariate marginals
         [theta_temp{tt,j},omega_save{tt,j},marg2D_theta_omega_temp{tt,j}] = getMarginal2D(theta_upd(:,j,tt),omega_upd(:,j,tt),rho_theta_omega_upd(:,tt),nBins);
-             
+        % make new bivariate grid     
         [theta_save{tt,j},omega_save{tt,j}] = meshgrid(linspace(0,2*pi,nBins),linspace(min(omega_save{tt,j}(:)),max(omega_save{tt,j}(:)),nBins));
-        
+        % interpolate bivariate marginal on new bivariarte grid
         marg2D_theta_omega_save{tt,j} = interp2(theta_temp{tt,j},omega_save{tt,j},marg2D_theta_omega_temp{tt,j}.PF,theta_save{tt,j},omega_save{tt,j},'spline');
-        
+        % zero the spurious negative values of the bivariate marginal resulting from interpolation
         marg2D_theta_omega_save{tt,j}(marg2D_theta_omega_save{tt,j}<0) = 0;
         
         maxMargValMatrix(tt,j) = max(max(marg2D_theta_omega_save{tt,j}));
@@ -303,9 +306,33 @@ h = colorbar('peer', cbax, 'southoutside', ...
   'position', [left1 cb_bottom cb_width cb_height],...
   'FontSize',30,'TickLabelInterpreter','latex');
 
-% save 2D marginal data as .txt file
+% plot 1D omega marginals
+figure;
 for j=1:num_Oscillator
     for tt=1:length(t_now_idx_vec)
+        subplot(num_Oscillator,length(t_now_idx_vec),tt+(j-1)*length(t_now_idx_vec))
+        plot(omega_save1D{tt,j},marg1D_omega_save{tt,j}.MC,'--ro','LineWidth',2)
+        hold on
+        plot(omega_save1D{tt,j},marg1D_omega_save{tt,j}.PF1,'-bs','LineWidth',2)
+        hold on
[…]

### 2021-08-08 code/getMarginal1D.m

@@ -1,0 +1,27 @@
+function [X1,marg] = getMarginal1D(X,PDF,nBins)
+
+Grid = linspace(min(X),max(X),nBins+1);
+
+dx = mean(diff(Grid));
+
+ 
+
+for i=1:nBins
+
+    ii = find(X>=Grid(i) & X < Grid(i+1));
+
+    marg.MC(i,1) = length(ii); % MC marginal via frequentist counting 
+
+    marg.PF1(i,1) = sum(PDF(ii)); % PF marginal via arithmatic mean
+
+    marg.PF2(i,1) = median(PDF(ii)); % PF marginal via median
+
+    X1(i) = (Grid(i)+Grid(i+1))/2.0;
+
+end
+
+marg.MC = marg.MC/sum(marg.MC)/dx;
+
+marg.PF1 = marg.PF1/sum(marg.PF1)/dx;
+
+marg.PF2 = marg.PF2/sum(marg.PF2)/dx;

### 2021-08-10 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py
adding IEEE 14 bus sim results

@@ -1,17 +1,21 @@
 import numpy as np
-from matplotlib import rc
 import matplotlib.pyplot as plt
+import matplotlib.patches as mpatches
+from matplotlib import rc
 from mpl_toolkits.mplot3d import Axes3D
 from mpl_toolkits.axes_grid1 import ImageGrid
 from itertools import cycle
 from pylab import *
 import glob
 from pylab import rcParams
+from matplotlib.patches import FancyArrowPatch
+from mpl_toolkits.mplot3d import proj3d
+from scipy.interpolate import interp1d
 
 #====================================================
 # Make plots beautiful
 #====================================================
-
+myalphavalue=0.6
 pts_per_inch = 72.27
 # write "\the\textwidth" (or "\showthe\columnwidth" for a 2 collumn text)
 text_width_in_pts = 300.0
@@ -58,33 +62,33 @@ fig = plt.figure(1, figsize=fig_size)  # figsize accepts only inches.
 # Time_Synthetic = np.loadtxt("TimeSynthetic50Gen.txt")
 # CompTime_Synthetic = np.loadtxt("ComptimeSythetic50Gen.txt")
 
-Time_Synthetic = np.loadtxt("TimeSyntheticIEEE14bus.txt")
-CompTime_Synthetic = np.loadtxt("ComptimeSytheticIEEE14bus.txt")
+# Time_Synthetic = np.loadtxt("TimeSyntheticIEEE14bus.txt")
+# CompTime_Synthetic = np.loadtxt("ComptimeSytheticIEEE14bus.txt")
 
-fig = plt.figure()
-ax = fig.add_subplot(1, 1, 1)
+# fig = plt.figure()
+# ax = fig.add_subplot(1, 1, 1)
 
-ax.semilogy(Time_Synthetic, CompTime_Synthetic,'-', color='b', lw=1.5)
+# ax.semilogy(Time_Synthetic, CompTime_Synthetic,'-', color='b', lw=1.5)
 
-ax.set_ylim(10**-3, 6*10**-2)
+# ax.set_ylim(10**-3, 6*10**-2)
 
-ax.tick_params(direction='in',which='both')
- # axx.xaxis.tick_top()
+# ax.tick_params(direction='in',which='both')
+#  # axx.xaxis.tick_top()
 
[…]

### 2021-08-10 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py
adding IEEE 14 bus sim results

@@ -15,7 +15,7 @@ from scipy.interpolate import interp1d
 #====================================================
 # Make plots beautiful
 #====================================================
-myalphavalue=0.6
+myalphavalue=0.3
 pts_per_inch = 72.27
 # write "\the\textwidth" (or "\showthe\columnwidth" for a 2 collumn text)
 text_width_in_pts = 300.0
@@ -120,50 +120,217 @@ fig = plt.figure(1, figsize=fig_size)  # figsize accepts only inches.
 # plt.savefig('RelativeErrorMeanMCVersusProxSynthetic50Gen.png', dpi=400)
 
 # =============================================================================
-# Univariate omega marginal plot: Case 1 
+# Univariate omega marginal plot: Case 0 
 # =============================================================================
-# Time_Synthetic = np.loadtxt("case1_line_13_failure_TimeSyntheticIEEE14bus.txt")
 
-t_vec = [0.2, 0.4, 0.6, 0.8, 1.0]
+case0omega1t1 = np.loadtxt("/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/code/case0_norminal_IEEE14BusGenIdx1omega1Dt0.2.txt")
+case0omega1t2 = np.loadtxt("/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/code/case0_norminal_IEEE14BusGenIdx1omega1Dt0.4.txt")
+case0omega1t3 = np.loadtxt("/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/code/case0_norminal_IEEE14BusGenIdx1omega1Dt0.6.txt")
+case0omega1t4 = np.loadtxt("/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/code/case0_norminal_IEEE14BusGenIdx1omega1Dt0.8.txt")
+case0omega1t5 = np.loadtxt("/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/code/case0_norminal_IEEE14BusGenIdx1omega1Dt1.txt")
 
[…]

### 2021-08-10 code/ThreeCases_IEEE14_prox_simulation.m

@@ -17,12 +17,12 @@ bus_init = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertain
 % case2_time_series_
 
 % parameters for the IEEE 14 bus system
-Y_reduced_imag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_Y_reduced_imag_part.csv');
-Y_reduced_real = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_Y_reduced_real_part.csv');
+Y_reduced_imag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_Y_reduced_imag_part.csv');
+Y_reduced_real = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_Y_reduced_real_part.csv');
 % reduced admittance matrix
 Y_reduced = complex(Y_reduced_real,Y_reduced_imag);
 
-gen_param = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_gen_parameters.csv');
+gen_param = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_gen_parameters.csv');
 gen_bus_idx = gen_param(:,1);
 
 num_Oscillator = numel(gen_bus_idx); dim  = 2*num_Oscillator;
@@ -35,17 +35,17 @@ E = E_mag.*exp(1i*E_phase);
 %I_reduced_polar = readmatrix('I_red.csv');
 %I_reduced_mag = I_reduced_polar(:,1);
 %I_reduced_phase = I_reduced_polar(:,2);
-I_reduced_mag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_I_red_magnitude.csv');
+I_reduced_mag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_I_red_magnitude.csv');
 I_reduced_mag = (I_reduced_mag(2,2:end))';
[…]

### 2021-08-25 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
adding IEEE 14 bus sim results
updated the signs in the potentials
updates

@@
 \usepackage{amsmath,graphicx,amsfonts,amssymb,amsthm,epsfig,mathrsfs,balance}
-\usepackage{subcaption,tikz,color,multirow,algorithm,algpseudocode,algorithmicx}
+\usepackage{subfigure,tikz,color,multirow,algorithm,algpseudocode,algorithmicx}
 \usepackage{cite,url,framed,bm,dsfont,lipsum}
@@
         Pegah~Ojaghi,
-        and~Xinbo~Geng% <-this % stops a space
+        and~Xinbo~Geng,~\IEEEmembership{Member,~IEEE}% <-this % stops a space
 \thanks{A. Halder and K.F. Caluya are with the Department of Applied Mathematics, University of California, Santa Cruz, CA 95064, USA, {\texttt{\{ahalder,kcaluya\}@ucsc.edu}}.}% <-this % stops a space
@@
 \begin{align}
-V(\bm{\theta}) := \displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{i,j=1}^{n}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right).
+V(\bm{\theta}) := -\displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{i,j=1}^{n}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right).
 \label{potential}	
@@
 \begin{align}
-U(\bm{\xi}) &:= \!\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\displaystyle\sum_{i,j=1}^{n}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} + \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right), \label{defU}\\
+U(\bm{\xi}) &:= \!-\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\displaystyle\sum_{i,j=1}^{n}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right), \label{defU}\\
 F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
@@
 
+\begin{figure*}[ht!]
+   \subfigure[]{\includegraphics[width=0.322\textwidth]{Case0_OmegaMarginals.png}}
+   \hspace{\fill}
+   \subfigure[]{\includegraphics[width=0.322\textwidth]{Case1_OmegaMarginals.png}}
+   \hspace{\fill}
+   \subfigure[]{\includegraphics[width=0.309\textwidth]{Case0Case1Combined_Gen4_OmegaMarginals.png}}
+\caption{\small{For $i=1,\hdots,5$, the $\omega_{i}$ marginals (a) for Case I, and (b) for Case II, resulting from the IEEE 14 bus simulation described in Sec.
+\ref{subsec:NumSimIEEEBus}.
+(c) Comparison of the $\omega_{4}$ marginals for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios.}}
+\vspace*{-0.1in}
+\label{fig:IEEE14omegamarginalsCase12}
+\end{figure*}
+
 \begin{figure}[t]
@@
 
-We randomly generated the noise coefficients $\sigma_{i}\in [1,5]$ as
+Our numerical simulations for the IEEE 14 bus system consider the following three cases:
+
+\noindent\textbf{Case I.} Nominal case where ...
+
+\noindent\textbf{Case II.} post-contingency simulation after the line 13 fails.
+
+\noindent\textbf{Case III.}	load data are time-varying ...
+
+For all the three cases, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
 \[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]
@@
 Using the algorithm detailed in Sec.
-\ref{sec:ProxAlgo}, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
+\ref{sec:ProxAlgo}, for all the three cases, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
 In Algorithm \ref{algo:KineticProx}, we used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $L=300$.
@@
 These marginals are computed using the corresponding weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
-%Notice that in Fig.
-\ref{fig:IEEE14marginals}, the horizontal axes of the subfigures have periodic boundary conditions, i.e., $(\theta_{i},\omega_{i})\in\mathbb{T}\times\mathbb{R}$ (cylinder).
+
+ %Notice that in Fig. \ref{fig:IEEE14marginals}, the horizontal axes of the subfigures have periodic boundary conditions, i.e., $(\theta_{i},\omega_{i})\in\mathbb{T}\times\mathbb{R}$ (cylinder).
+
+Fig.
+\ref{fig:IEEE14omegamarginalsCase12} shows the corresponding univariate $\omega_{i}$-marginal PDFs.
+In particular, Fig.
+\ref{fig:IEEE14omegamarginalsCase12}(c) helps compare the $\omega_{4}$ marginal evolution for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios.
+Since Case II considers the failure of line 13, we expect from Fig.
+\ref{fig:Motivating} that the transient effects of this fault will be prominent for bus 6.
+Indeed, Fig.
+\ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both Case I and II, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger dispersion.
+
 Fig. \ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec.  \ref{sec:ProxAlgo}.

### 2021-08-25 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updates

@@
 \ref{fig:IEEE14omegamarginalsCase12}(c) helps compare the $\omega_{4}$ marginal evolution for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios.
-Since Case II considers the failure of line 13, we expect from Fig.
-\ref{fig:Motivating} that the transient effects of this fault will be prominent for bus 6.
+Since Case II considers the failure of line 13, from Fig.
+\ref{fig:Motivating} we intuitively expect that the transient effects of this fault will be prominent for bus 6 (generator 4).
 Indeed, Fig.
-\ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both Case I and II, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger dispersion.
+\ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both Case I and II, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger spread or dispersion of the probability mass, as expected.
+{\red{That the generator 4 states in Case II have larger uncertainties than in Case I, is also visible in the bivariate marginal plots if we compare the fourth row in Fig.
+\ref{fig:IEEE14marginalsCase0} with the same in Fig.
+\ref{fig:IEEE14marginalsCase1}.}}
 

### 2021-08-26 code/Case3_IEEE14_prox_simulation.m
updates

@@ -1,0 +1,43 @@
+close all; clear all; clc;
+set(groot,'defaultAxesTickLabelInterpreter','latex');  
+set(groot,'defaulttextinterpreter','latex');
+set(groot,'defaultLegendInterpreter','latex');
+%% Parameters
+beta = 2; % inverse temperature
+% golbal parameters 
+global num_Oscillator dim Gamma M Sigma K P cc sigma m phi
+
+f0 = 60; % nominal frequency (Hz)
+
+bus_init = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/bus-init.csv');
+
+% parameters for the IEEE 14 bus system
+Y_reduced_imag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case3_realworld_time_series_Y_reduced_imag_part.csv');
+Y_reduced_real = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case3_realworld_time_series_Y_reduced_real_part.csv');
+% reduced admittance matrix
+Y_reduced = complex(Y_reduced_real,Y_reduced_imag);
+
+gen_param = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case3_realworld_time_series_gen_parameters.csv');
+gen_bus_idx = gen_param(:,1);
+
+num_Oscillator = numel(gen_bus_idx); dim  = 2*num_Oscillator;
+
+% generator bus voltage
+E_mag = bus_init(gen_bus_idx,2);
+E_phase = bus_init(gen_bus_idx,3);
+E = E_mag.*exp(1i*E_phase);
+
+I_reduced_mag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case3_realworld_time_series_I_red_magnitude.csv');
+I_reduced_mag = (I_reduced_mag(2,2:end))';
+I_reduced_phase = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case3_realworld_time_series_I_red_angle.csv');
+I_reduced_phase = (I_reduced_phase(2,2:end))';
+
+I_reduced = I_reduced_mag.*exp(1i*I_reduced_phase);
+
[…]

### 2021-08-26 code/Case3_IEEE14_prox_simulation.m
updates

@@ -34,10 +34,128 @@ I_reduced_phase = (I_reduced_phase(2,2:end))';
 
 I_reduced = I_reduced_mag.*exp(1i*I_reduced_phase);
 
-%P_mech = gen_param(:,2);
+% select the time window from one day time-varying power data 
+% at 1 sec resolution
+initial_sec = 50000; final_sec = 51000;
+% get the power trajectories within that time window
 P_mech_coarse = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case3_realworld_time_series_P_mech_in_per_unit.csv');
-%P_mech = (P_mech(2:end))';
+P_mech_coarse = (P_mech_coarse(initial_sec:1:final_sec,2:end));
 P_load_coarse = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case3_realworld_time_series_P_load_at_genbuses.csv');
-%P_load = (P_load(2:end))';
+P_load_coarse = (P_load_coarse(initial_sec:1:final_sec,2:end));
 
+h = 1e-3; % time step
+t_vec_coarse = initial_sec:1:final_sec; % time vector with 1 sec step-size
+t_vec = initial_sec:h:final_sec; % time vector with h sec step-size
 
+%initialize
+P_mech = zeros(numel(t_vec),num_Oscillator);
+P_load = zeros(numel(t_vec),num_Oscillator);
+% interpolate the time varying powers to a finer time grid
+figure
+color_cell = {'r','m','b','c','g'};
+for j=1:size(P_mech_coarse,2)    
+    P_mech(:,j) = interp1(t_vec_coarse,P_mech_coarse(:,j),t_vec,'spline');
+    P_load(:,j) = interp1(t_vec_coarse,P_load_coarse(:,j),t_vec,'spline');
+
+    plot(t_vec,P_mech(:,j),'-','color',color_cell{j},'LineWidth',2)    
+    xlabel('$t$ [s]','FontSize', 30,'Interpreter','latex')
+    hold on
+    plot(t_vec,P_load(:,j),'--','color',color_cell{j},'LineWidth',2)    
+    xlabel('$t$ [s]','FontSize', 30,'Interpreter','latex')
+    hold on
+end
[…]

### 2021-08-26 code/Gen_50_prox_simulation.m

@@ -1,4 +1,4 @@
-close all; clear; clc;
+cclose all; clear; clc;
 %% Parameters
 beta = 2; % inverse temperature
 % golbal parameters 

### 2021-08-27 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

--- 2021-08-25 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
@@
 In Sec.
-\ref{subsec:NumSimIEEEBus}, we consider the prediction of stochastic states for the IEEE 14 bus system--{\red{our main objective being to highlight the interpretability and utility of the proposed method in power systems engineering}}.
+\ref{subsec:NumSimIEEEBus}, we consider the prediction of transient stochastic states for the IEEE 14 bus system for the nominal case as well as for the case when a line failure occurs.
 In Sec.
@@
 \centering
-\includegraphics[width=0.95\linewidth]{ComputationalTimeSyntheticIEEE14bus.png}
+\includegraphics[width=0.95\linewidth]{ComputationalTimeSyntheticIEEE14busCase1and2combined.png}
 \caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ for the simulation set up in Sec. \ref{subsec:NumSimIEEEBus}.}}
@@
 
-Fig. \ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec.  \ref{sec:ProxAlgo}.
+Fig. \ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec. \ref{sec:ProxAlgo}, for both Cases I and II.
 

### 2021-08-27 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py

@@ -57,7 +57,39 @@ fig = plt.figure(1, figsize=fig_size)  # figsize accepts only inches.
 #                     hspace=0.05, wspace=0.02)
 
 # =============================================================================
-# Computational time plot 
+# Computational time plot [IEEE 14 bus 5 gen, Case I and Case II]
+# =============================================================================
+Time_Synthetic = np.loadtxt("TimeSyntheticIEEE14bus.txt")
+CompTime_Synthetic_Case1 = np.loadtxt("case0_norminal_ComptimeSytheticIEEE14bus.txt")
+CompTime_Synthetic_Case2 = np.loadtxt("case1_line_13_failure_ComptimeSytheticIEEE14bus.txt")
+
+fig = plt.figure()
+ax = fig.add_subplot(1, 1, 1)
+
+ax.semilogy(Time_Synthetic, CompTime_Synthetic_Case1,'-', color='b', lw=1.5, alpha=0.7,label=r"Case I")
+ax.semilogy(Time_Synthetic, CompTime_Synthetic_Case2,':', color='r', lw=1.5, alpha=0.7,label=r"Case II")
+
+# ax.set_ylim(10**-3, 6*10**-2)
+
+ax.tick_params(direction='in',which='both')
+#  # axx.xaxis.tick_top()
+ax.grid(True,which="both",ls="-", color='0.75')
+# # axx2.grid(True,which="both",ls="-", color='0.75')
+ax.tick_params(axis='both', labelsize=18)
+
+
+ax.set_ylabel(r"Computational time [s]")
+ax.set_xlabel(r"Physical time $t=kh$ [s]")
+# # # axx.yaxis.set_label_coords(-0.125,-0.05)
+
+ax.legend(markerscale=1.5, numpoints=1,  ncol=1, bbox_to_anchor=(1.005, 1), frameon=False, prop={'size': 10.5})
+fig.set_size_inches(10.2, 6.2)
+
+#plt.savefig('ComputationalTimeSynthetic50Gen.png', dpi=400)
+plt.savefig('ComputationalTimeSyntheticIEEE14busCase1and2combined.png', dpi=400)
+
+# =============================================================================
[…]

### 2021-08-27 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updates

@@
 
-Our numerical simulations for the IEEE 14 bus system consider the following three cases:
-
-\noindent\textbf{Case I.} Nominal case where ...
-
-\noindent\textbf{Case II.} post-contingency simulation after the line 13 fails.
-
-\noindent\textbf{Case III.}	load data are time-varying ...
+Our numerical simulations for the IEEE 14 bus system consider the following two cases:
+
+\noindent\textbf{Case I.} {\red{Nominal case where ...}}
+
+\noindent\textbf{Case II.} {\red{post-contingency simulation after the line 13 fails.}}
+
+%\noindent\textbf{Case III.}	load data are time-varying ...
 

### 2021-08-27 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

@@
 
+\begin{remark}\label{remark:timevaryingEffectivePowerInput}
+The proposed computational framework can be applied as is when $P_{i}^{\rm{mech}}, P_{i}^{\rm{load}}$, and thus $P_{i}$ in \eqref{defPi}, are bounded time-varying functions.
+{\red{This can be the case, for instance, when ...}}
+\end{remark}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \ref{subsec:NumSimSynthetic}, we propagate the joint PDFs over the 100 dimensional state space of a synthetic power network with randomly generated parameters--our intent there is to highlight the scalability of the proposed method.
-
 All simulations were performed in MATLAB R2019b on an iMac with 3.4 GHz Quad-Core Intel Core i5 processor and 8 GB memory.
@@
 
-Our numerical simulations for the IEEE 14 bus system consider the following three cases:
-
-\noindent\textbf{Case I.} Nominal case where ...
-
-\noindent\textbf{Case II.} post-contingency simulation after the line 13 fails.
-
-\noindent\textbf{Case III.}	load data are time-varying ...
-
-For all the three cases, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
+Our numerical simulations for the IEEE 14 bus system consider the following two cases:
+
+\noindent\textbf{Case I.} {\red{Nominal case where ...}}
+
+\noindent\textbf{Case II.} {\red{post-contingency simulation after the failure of line 13 (see Fig. \ref{fig:Motivating}).}}
+
+%\noindent\textbf{Case III.}	load data are time-varying ...
+
+For all the cases mentioned above, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
 \[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]
 
-We take the initial joint PDF as
+To account the pertinent geometry of $\mathbb{T}^{n}\times\mathbb{R}^{n}$, we take the initial joint PDF as
 \begin{align}

### 2021-08-27 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

@@
 
-\begin{figure*}[ht!]
-   \subfigure[]{\includegraphics[width=0.322\textwidth]{Case0_OmegaMarginals.png}}
-   \hspace{\fill}
-   \subfigure[]{\includegraphics[width=0.322\textwidth]{Case1_OmegaMarginals.png}}
-   \hspace{\fill}
-   \subfigure[]{\includegraphics[width=0.309\textwidth]{Case0Case1Combined_Gen4_OmegaMarginals.png}}
-\caption{\small{For $i=1,\hdots,5$, the $\omega_{i}$ marginals (a) for Case I, and (b) for Case II, resulting from the IEEE 14 bus simulation described in Sec.
+%\begin{figure*}[ht!]
+%   \subfigure[]{\includegraphics[width=0.322\textwidth]{Case0_OmegaMarginals.png}}
+%   \hspace{\fill}
+%   \subfigure[]{\includegraphics[width=0.322\textwidth]{Case1_OmegaMarginals.png}}
+%   \hspace{\fill}
+%   \subfigure[]{\includegraphics[width=0.309\textwidth]{Case0Case1Combined_Gen4_OmegaMarginals.png}}
+%\caption{\small{For $i=1,\hdots,5$, the $\omega_{i}$ marginals (a) for Case I, and (b) for Case II, resulting from the IEEE 14 bus simulation described in Sec.
 \ref{subsec:NumSimIEEEBus}.
 (c) Comparison of the $\omega_{4}$ marginals for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios.}}
+%\vspace*{-0.1in}
+%\label{fig:IEEE14omegamarginalsCase12}
+%\end{figure*}
+
+\begin{figure}[ht!]
+\includegraphics[width=0.95\linewidth]{Case0Case1Combined_Gen4_OmegaMarginals.png}
+\caption{\small{Comparison of the $\omega_{4}$ marginals for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios in the IEEE 14 bus simulation described in Sec.
+\ref{subsec:NumSimIEEEBus}.}}
 \vspace*{-0.1in}
 \label{fig:IEEE14omegamarginalsCase12}
-\end{figure*}
+\end{figure}
 
@@
 \centering
-\includegraphics[width=0.95\linewidth]{ComputationalTimeSyntheticIEEE14busCase1and2combined.png}
+\includegraphics[width=0.9\linewidth]{ComputationalTimeSyntheticIEEE14busCase1and2combined.png}
 \caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ for the simulation set up in Sec. \ref{subsec:NumSimIEEEBus}.}}

### 2021-08-28 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updates

@@
 \label{fig:BlockDiagm}
-\vspace*{-0.1in}
+\vspace*{-0.2in}
 \end{figure}
@@
 
+\begin{remark}\label{remark:timevaryingEffectivePowerInput}
+The proposed computational framework is also applicable when $P_{i}^{\rm{mech}}, P_{i}^{\rm{load}}$, and thus $P_{i}$ in \eqref{defPi}, are bounded time-varying functions.
+This can be the case, for instance, with rapid fluctuations from renewables in the short term.
+\end{remark}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{figure}[t]
+\begin{figure}[htpb]
 \centering
 \includegraphics[width=0.95\linewidth]{ComputationalTimeSyntheticIEEE14busCase1and2combined.png}
-\caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ for the simulation set up in Sec. \ref{subsec:NumSimIEEEBus}.}}
+\caption{\small{The computational times for propagating the transient joint state PDFs over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ for the simulation set up in Sec. \ref{subsec:NumSimIEEEBus}.}}
 \vspace*{-0.1in}
@@
 
-\noindent\textbf{Case I.} {\red{Nominal case where ...}}
-
-\noindent\textbf{Case II.} {\red{post-contingency simulation after the line 13 fails.}}
+\noindent\textbf{Case I.} \textbf{Nominal case} where the $P_{i}^{\text{mech}}$ and $P_{i}^{\text{load}}$ are from the steady state power flow solutions without the process noise, and assuming to remain constant over the simulation time horizon.
+The transient stochastic simulations (i.e., with process noise) are then performed using the proposed framework.
+
+\noindent\textbf{Case II.} \textbf{Post-contingency case} where the line 13 (see Fig.
+\ref{fig:Motivating}) fails at time $t=0$, and starting with the parameters and initial conditions of \textbf{Case I}, we simulate the post-contingency stochastic state evolution with process noise.
+%In other words, \textbf{Case II} starts  and tracks the stochastic states towards a post-contingency equilibrium.
 
@@
 
-For all the three cases, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
+For both the cases mentioned above, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
 \[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]
 
-We take the initial joint PDF as
+To account the pertinent geometry of $\mathbb{T}^{n}\times\mathbb{R}^{n}$, we take the initial joint PDF as
 \begin{align}
@@
 \ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both Case I and II, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger spread or dispersion of the probability mass, as expected.
-{\red{That the generator 4 states in Case II have larger uncertainties than in Case I, is also visible in the bivariate marginal plots if we compare the fourth row in Fig.
+That the generator 4 states in Case II have larger uncertainties than in Case I, is also visible in the bivariate marginal plots if we compare the fourth row in Fig.
 \ref{fig:IEEE14marginalsCase0} with the same in Fig.
-\ref{fig:IEEE14marginalsCase1}.}}
+\ref{fig:IEEE14marginalsCase1}.
 
@@
 \label{table:ParamUncertaintiesSynthetic}
-\vspace*{-0.1in}
+\vspace*{-0.35in}
 \end{table}
@@
 \ref{subsec:NumSimSynthetic}.}}
-\vspace*{-0.1in}
+\vspace*{-0.2in}
 \label{fig:RelErrSynthetic}
@@
 Instead, the algorithm allows gridless computation by actually exploiting the structure of the PDE governing the evolution of the joint state PDF that is induced by the underlying nonlinear sample path dynamics of a networked power system.
-Numerical case studies reveal that the proposed computational framework can be a scalable way to propagate stochastic uncertainties in realistic power networks.
+Numerical case studies reveal that the computational framework can be a scalable way to propagate stochastic uncertainties in realistic power networks.
 

### 2021-08-28 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

@@
 \centering
-\includegraphics[width=0.9\linewidth]{UncPropBlockDiagm.pdf}
+\includegraphics[width=0.8\linewidth]{UncPropBlockDiagm.pdf}
 \caption{\small{Block diagram for joint state PDF propagation.}}
 \label{fig:UncertaintyPropagation}
-\vspace*{-0.1in}
+\vspace*{-0.15in}
 \end{figure}
@@
 \centering
-\includegraphics[width=0.9\linewidth]{ProxSchmatic.pdf}
+\includegraphics[width=0.7\linewidth]{ProxSchmatic.pdf}
 \caption{\small{A schematic of approximating the flow of the joint PDF trajectory $\rho(t,\bm{x})$ via the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.
@@
 \centering
-\includegraphics[width=.7\linewidth]{BlockDiagm.pdf}
+\includegraphics[width=.6\linewidth]{BlockDiagm.pdf}
 \caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
@@
 \begin{remark}\label{remark:timevaryingEffectivePowerInput}
-The proposed computational framework can be applied as is when $P_{i}^{\rm{mech}}, P_{i}^{\rm{load}}$, and thus $P_{i}$ in \eqref{defPi}, are bounded time-varying functions.
-{\red{This can be the case, for instance, when ...}}
+The proposed computational framework is also applicable when $P_{i}^{\rm{mech}}, P_{i}^{\rm{load}}$, and thus $P_{i}$ in \eqref{defPi}, are bounded time-varying functions.
+This can be the case, for instance, with rapid fluctuations from renewables in the short term.
 \end{remark}
@@
 \centering
-\includegraphics[width=0.9\linewidth]{IEEE14marginalsCase1.png}
-\caption{\small{Time evolution of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs of the generator $i=1,\hdots,5$, for the Kron-reduced IEEE 14 bus simulation setup \textbf{Case II} described in Sec.
-\ref{subsec:NumSimIEEEBus}.
-The $5$ rows above correspond to the $5$ generator nodes; these are buses 1, 2, 3, 6 and 8 in Fig.
-\ref{fig:Motivating}.
-The columns correspond to the time snapshots.
-The colorbar at the bottom shows the marginal PDF values.
-The $\theta_i$ are in rad, the $\omega_i$ are in rad/s.}}
+\includegraphics[width=0.9\linewidth]{IEEE14BivariateMarginalComparison.pdf}
+\caption{\small{Comparison of the $(\theta_{4},\omega_{4})$ marginals for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios in the IEEE 14 bus simulation described in Sec.
+\ref{subsec:NumSimIEEEBus}.}}
 \vspace*{-0.1in}
-\label{fig:IEEE14marginalsCase1}
+\label{fig:IEEE14marginalsBivariateComparison}
 \end{figure*}
@@
 
-\noindent\textbf{Case I.} {\red{Nominal case where ...}}
-
-\noindent\textbf{Case II.} {\red{post-contingency simulation after the failure of line 13 (see Fig. \ref{fig:Motivating}).}}
+\noindent\textbf{Case I.} \textbf{Nominal case} where the $P_{i}^{\text{mech}}$ and $P_{i}^{\text{load}}$ are from the steady state power flow solutions without the process noise, and assuming to remain constant over the simulation time horizon.
+The transient stochastic simulations (i.e., with process noise) are then performed using the proposed framework.
+
+\noindent\textbf{Case II.} \textbf{Post-contingency case} where the line 13 (see Fig.
+\ref{fig:Motivating}) fails at time $t=0$, and starting with the parameters and initial conditions of \textbf{Case I}, we simulate the post-contingency stochastic state evolution with process noise.
 
@@
 
-For all the cases mentioned above, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
+For both the cases mentioned above, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
 \[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]
@@
 Fig.
-\ref{fig:IEEE14marginalsCase0} and Fig.
-\ref{fig:IEEE14marginalsCase1} show the time snapshots of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs for \textbf{Case I} and \textbf{Case II} respectively, associated with the generator $i=1,\hdots, n=5$.
+\ref{fig:IEEE14marginalsCase0} shows the time snapshots of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs for \textbf{Case I} associated with the generator $i=1,\hdots, n=5$.
 These marginals are computed using the corresponding weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
 
- %Notice that in Fig. \ref{fig:IEEE14marginals}, the horizontal axes of the subfigures have periodic boundary conditions, i.e., $(\theta_{i},\omega_{i})\in\mathbb{T}\times\mathbb{R}$ (cylinder).
-
-Fig.
-\ref{fig:IEEE14omegamarginalsCase12} shows the corresponding univariate $\omega_{i}$-marginal PDFs.
-In particular, Fig.
-\ref{fig:IEEE14omegamarginalsCase12}(c) helps compare the $\omega_{4}$ marginal evolution for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios.
-Since Case II considers the failure of line 13, from Fig.
+Since \textbf{Case II} considers the failure of line 13, from Fig.
 \ref{fig:Motivating} we intuitively expect that the transient effects of this fault will be prominent for bus 6 (generator 4).
-Indeed, Fig.
-\ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both Case I and II, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger spread or dispersion of the probability mass, as expected.
-{\red{That the generator 4 states in Case II have larger uncertainties than in Case I, is also visible in the bivariate marginal plots if we compare the fourth row in Fig.
+The comparison of the $(\theta_4,\omega_4)$ bivariate marginals between \textbf{Cases I} and \textbf{II} shown in Fig.
+\ref{fig:IEEE14marginalsBivariateComparison} indeed confirms that the post contingency transient stochastic states for bus 6 have higher dispersion or spread of the probability mass.
+
+Likewise, Fig.
+\ref{fig:IEEE14omegamarginalsCase12} compares the $\omega_{4}$ marginal evolution for bus 6 in the nominal (\textbf{Case I}) and post-fault (\textbf{Case II}) scenarios.
+As expected, Fig.
+\ref{fig:IEEE14omegamarginalsCase12} highlights that although the $\omega_{4}$ marginals in both \textbf{Cases I} and \textbf{II}, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger dispersion.
+%{\red{That the generator 4 states in Case II have larger uncertainties than in Case I, is also visible in the bivariate marginal plots if we compare the fourth row in Fig.
 \ref{fig:IEEE14marginalsCase0} with the same in Fig.
@@
 H.~Cui, F.~Li, and K.~Tomsovic, ``Hybrid Symbolic-Numeric Framework for Power System Modeling and Analysis,'' \emph{IEEE Transactions on Power Systems}, vol. 36, no. 2, pp. 1373-1384, 2021.
+
+% expanded version of this paper (exceeds 10 pages)
+
+\bibitem{halder2021stochasticExpanded}
+A.~Halder, K.F.~Caluya, P.~Ojaghi, and X.~Geng, ``Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions", \emph{extended preprint}, \\
+doi: \url{http://dx.doi.org/10.13140/RG.2.2.30956.28805}, 2021.
 

### 2021-08-28 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

@@
 \label{fig:UncertaintyPropagation}
-\vspace*{-0.15in}
+\vspace*{-0.2in}
 \end{figure}
@@
 While the joint PDFs are supported over the finite dimensional base manifold $\mathbb{T}^{n}\times\mathbb{R}^{n}$, the proximal updates $\{\varrho_{k}\}_{k\in\mathbb{N}}$ (shown as circled markers with no face-color) evolve over the infinite dimensional manifold $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.}}
-\vspace*{-0.1in}
+\vspace*{-0.2in}
 \label{fig:ProxSchematic}
@@
 which is the sum of a potential energy (expected value of scaled $U$), a weighted kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$).
-Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov-like functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.
+Indeed, it can be shown \cite[Appendix A]{halder2021stochasticExpanded} that the functional \eqref{LyapButNotProx} is a Lyapunov-like functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.
 
@@
 the vector $(\bm{\xi},\bm{\eta})^{\top}$, and $\frac{\delta}{\delta\rho}$ denotes the functional derivative.
-From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$.
-Thus, even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, we cannot interpret the flow generated by \eqref{TransformedPDEivp} as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}.
+%From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$.
+Thus,
+From \cite[Appendix A (see eq.
+(31), (33) and (34a))]{halder2021stochasticExpanded} it follows that even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, we cannot interpret the flow generated by \eqref{TransformedPDEivp} as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}.
 Thus, in \eqref{InfiniteDimProxOpDef}, we cannot construct $(\Phi,{\rm{dist}})$ by pairing \eqref{LyapButNotProx} with the Wasserstein metric.
@@
 This enables a nonparametric computation of $\tilde{\varrho}_{k}^{i} \equiv \tilde{\varrho}_{k}\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)$ for $i=1,\hdots,N$.
-
-\begin{algorithm}[t]
+{\small{\begin{algorithm}[t]
 \caption{Proposed proximal algorithm for $\tilde{\bm{\varrho}}_{k-1}\mapsto\tilde{\bm{\varrho}}_{k}$}
@@
 \end{algorithmic}
-\end{algorithm}
+\end{algorithm}}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\vspace*{-0.2in}
 \section{Numerical Simulations}\label{sec:NumericalSimulations}
@@
 \centering
-\includegraphics[width=0.9\linewidth]{IEEE14marginalsCase0.png}
+\includegraphics[width=0.75\linewidth]{IEEE14marginalsCase0.png}
 \caption{\small{Time evolution of the $(\theta_{i},\omega_{i})$ bivariate marginal PDFs of the generator $i=1,\hdots,5$, for the Kron-reduced IEEE 14 bus simulation setup \textbf{Case I} described in Sec.
@@
 \centering
-\includegraphics[width=0.9\linewidth]{IEEE14BivariateMarginalComparison.pdf}
+\includegraphics[width=0.8\linewidth]{IEEE14BivariateMarginalComparison.pdf}
 \caption{\small{Comparison of the $(\theta_{4},\omega_{4})$ marginals for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios in the IEEE 14 bus simulation described in Sec.
 \ref{subsec:NumSimIEEEBus}.}}
-\vspace*{-0.1in}
+\vspace*{-0.2in}
 \label{fig:IEEE14marginalsBivariateComparison}
@@
 \begin{figure}[ht!]
-\includegraphics[width=0.95\linewidth]{Case0Case1Combined_Gen4_OmegaMarginals.png}
+\includegraphics[width=0.85\linewidth]{Case0Case1Combined_Gen4_OmegaMarginals.png}
 \caption{\small{Comparison of the $\omega_{4}$ marginals for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios in the IEEE 14 bus simulation described in Sec.
@@
 \centering
-\includegraphics[width=0.9\linewidth]{ComputationalTimeSyntheticIEEE14busCase1and2combined.png}
+\includegraphics[width=0.8\linewidth]{ComputationalTimeSyntheticIEEE14busCase1and2combined.png}
 \caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ for the simulation set up in Sec. \ref{subsec:NumSimIEEEBus}.}}
-\vspace*{-0.1in}
+\vspace*{-0.2in}
 \label{fig:CompTimeIEEE14bus}
@@
 
+\vspace*{-0.2in}
 \subsection{IEEE 14 Bus System}\label{subsec:NumSimIEEEBus}
@@
 For both the cases mentioned above, we used the randomly generated noise coefficients $\sigma_{i}\in [1,5]$ as
-\[(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).\]
+\vspace*{-0.1in}
+$$(\sigma_{1},\hdots,\sigma_5) = \left(2.4628, 4.9266, 4.8724, 1.4215, 3.8681\right).$$
 
@@
 
-Likewise, Fig.
+Fig.
 \ref{fig:IEEE14omegamarginalsCase12} compares the $\omega_{4}$ marginal evolution for bus 6 in the nominal (\textbf{Case I}) and post-fault (\textbf{Case II}) scenarios.
@@
 \ref{fig:IEEE14omegamarginalsCase12} highlights that although the $\omega_{4}$ marginals in both \textbf{Cases I} and \textbf{II}, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger dispersion.
+For more simulation results, we refer the readers to the extended version of this manuscript \cite[see Fig.
+6 and 7(a),(b)]{halder2021stochasticExpanded}.
+
 %{\red{That the generator 4 states in Case II have larger uncertainties than in Case I, is also visible in the bivariate marginal plots if we compare the fourth row in Fig.
@@
 
-Fig. \ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec. \ref{sec:ProxAlgo}, for both Cases I and II.
+Fig. \ref{fig:CompTimeIEEE14bus} shows the computational times for updating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec. \ref{sec:ProxAlgo} for Cases I and II.
 
@@
 \end{tabular}
-\caption{Parameters used for the numerical simulation in Sec.
-\ref{subsec:NumSimSynthetic}.
-The indices $i,j\in\{1,\hdots,n\}$ where the number of generators $n=50$ in Sec.
-\ref{subsec:NumSimSynthetic}.}
+\caption{Parameters used in Sec. \ref{subsec:NumSimSynthetic}. The indices $i,j\in\{1,\hdots,n\}$ where the number of generators $n=50$ in Sec. \ref{subsec:NumSimSynthetic}.}
 \label{table:ParamUncertaintiesSynthetic}
-\vspace*{-0.1in}
+\vspace*{-0.2in}
 \end{table}
@@
 \centering
-\includegraphics[width=0.95\linewidth]{ComputationalTimeSynthetic50Gen.png}
+\includegraphics[width=0.85\linewidth]{ComputationalTimeSynthetic50Gen.png}
 \caption{\small{The computational time for propagating the transient joint state PDFs over $\mathbb{T}^{50}\times\mathbb{R}^{50}$ for the simulation set up in Sec. \ref{subsec:NumSimSynthetic}.}}
-\vspace*{-0.1in}
+\vspace*{-0.2in}
 \label{fig:CompTimeSynthetic}
@@
 \centering
-\includegraphics[width=0.95\linewidth]{RelativeErrorMeanMCVersusProxSynthetic50Gen.png}
+\includegraphics[width=0.85\linewidth]{RelativeErrorMeanMCVersusProxSynthetic50Gen.png}
 \caption{\small{The relative error between the empirical mean (i.e., Monte Carlo estimate) $\bm{\mu}_{k}^{\text{MC}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$, and the mean $\bm{\mu}_{k}^{\text{Prox}}\in\mathbb{T}^{50}\times\mathbb{R}^{50}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec.
 \ref{subsec:NumSimSynthetic}.}}
-\vspace*{-0.1in}
+\vspace*{-0.2in}
 \label{fig:RelErrSynthetic}
@@
 
+\vspace*{-0.1in}
 \section{Conclusion}\label{sec:conclusion}
-We proposed a nonparametric algorithm to propagate the joint state PDFs subject to networked power system dynamics with stochastic initial conditions, parameters and process noise.
+We propose a nonparametric algorithm to propagate the joint state PDFs subject to networked power system dynamics with stochastic initial conditions, parameters and process noise.
 Our development is built on recent advances in generalized gradient flows on the space of PDFs, and new analytical results specific to the power system's stochastic dynamics that are presented in this paper.
@@
 
-Possible future directions include generalizing the proposed method for more complex generator dynamics, and to use this framework to actively steer the uncertainties via optimal control on the space of joint PDFs.
-
-\appendices
-\section{Showing $\Phi(\tilde{\rho})$ is Decreasing along the Flow Generated by \eqref{TransformedPDEivp}}\label{AppLyap}
-Let $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ be the transient solution of \eqref{TransformedPDEivp}. Define
-$$\bm{\ell}_{\bm{\xi}} := \frac{1}{2}\nabla_{\bm{\xi}}\log\tilde{\rho}, \quad \bm{\ell}_{\bm{\eta}} := \frac{1}{2}\nabla_{\bm{\eta}}\log\tilde{\rho},$$
-and rewrite \eqref{NewPDE} as
-$$\dfrac{\partial\tilde{\rho}}{\partial t} = - \nabla\cdot\left(\tilde{\rho}\bm{v}\right)$$
-where $\nabla$ is the $2n\times 1$ Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and the vector field
-\begin{align}
-\bm{v} := \begin{pmatrix}
- \bm{\eta}\\
- -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
- \end{pmatrix}.
-\label{defvectorfieldv}	
-\end{align}
-
-Now consider the time derivative of the functional \eqref{LyapButNotProx} along \eqref{NewPDE}. Following \cite[Proposition 1]{krichene2020global}, 
-\begin{align}
-\frac{\differential}{\differential t}\Phi = \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}} \bigg\langle\nabla\dfrac{\delta\Phi}{\delta\tilde{\rho}}, \tilde{\rho}\bm{v}\bigg\rangle\:\differential\bm{\xi}\differential\bm{\eta},
-\label{dPhidt}	
-\end{align}
-wherein the functional derivative
-\begin{align}
-\dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
-\label{FunctionalDerivative}
-\end{align}
-Recalling the explicit form of $F(\bm{\eta})$ from \eqref{defF}, and using \eqref{FunctionalDerivative}, we simplify \eqref{dPhidt} as
-\begin{strip}
-\hrule
-\begin{subequations}
-\begin{align}
-\frac{\differential}{\differential t}\Phi &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\langle\begin{pmatrix}
-\bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
-\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}	
-\end{pmatrix}, \begin{pmatrix}
- \bm{\eta}\\
- -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
- \end{pmatrix}\bigg\rangle\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
-&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
-&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
-\end{align}
-\label{ExplicitdPhidt}	
-\end{subequations}
-\hrule	
-\end{strip}
-where \eqref{lastequality} follows from the fact that the integrals involving $\langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle$ and $\langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle$ in \eqref{lastbutoneequality} are zero, see \cite[Appendix D.3]{krichene2020global}.
-
-From \eqref{ExplicitdPhidt}, it follows that $\frac{\differential}{\differential t}\Phi \leq 0$, i.e., $\Phi$ is non-increasing along the flow generated by \eqref{TransformedPDEivp}.
+%Possible future directions include generalizing the proposed method for more complex generator dynamics, and to use this framework to actively steer the uncertainties via optimal control on the space of joint PDFs.
+
+%\appendices
+%\section{Showing $\Phi(\tilde{\rho})$ is Decreasing along the Flow Generated by \eqref{TransformedPDEivp}}\label{AppLyap}
+%Let $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ be the transient solution of \eqref{TransformedPDEivp}. Define
+%$$\bm{\ell}_{\bm{\xi}} := \frac{1}{2}\nabla_{\bm{\xi}}\log\tilde{\rho}, \quad \bm{\ell}_{\bm{\eta}} := \frac{1}{2}\nabla_{\bm{\eta}}\log\tilde{\rho},$$
+%and rewrite \eqref{NewPDE} as
+%$$\dfrac{\partial\tilde{\rho}}{\partial t} = - \nabla\cdot\left(\tilde{\rho}\bm{v}\right)$$
+%where $\nabla$ is the $2n\times 1$ Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and the vector field
+%\begin{align}
+%\bm{v} := \begin{pmatrix}
+% \bm{\eta}\\
+% -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
+% \end{pmatrix}.
+%\label{defvectorfieldv}	
+%\end{align}
+%
+%
+%Now consider the time derivative of the functional \eqref{LyapButNotProx} along \eqref{NewPDE}. Following \cite[Proposition 1]{krichene2020global}, 
+%\begin{align}
+%\frac{\differential}{\differential t}\Phi = \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}} \bigg\langle\nabla\dfrac{\delta\Phi}{\delta\tilde{\rho}}, \tilde{\rho}\bm{v}\bigg\rangle\:\differential\bm{\xi}\differential\bm{\eta},
+%\label{dPhidt}	
+%\end{align}
+%wherein the functional derivative
+%\begin{align}
+%\dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
+%\label{FunctionalDerivative}
+%\end{align}
+%Recalling the explicit form of $F(\bm{\eta})$ from \eqref{defF}, and using \eqref{FunctionalDerivative}, we simplify \eqref{dPhidt} as
+%\begin{strip}
+%\hrule
+%\begin{subequations}
+%\begin{align}
+%\frac{\differential}{\differential t}\Phi &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\langle\begin{pmatrix}
+%\bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
+%\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}	
+%\end{pmatrix}, \begin{pmatrix}
+% \bm{\eta}\\
+% -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
+% \end{pmatrix}\bigg\rangle\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
+%&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
+%&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
+%\end{align}
+%\label{ExplicitdPhidt}	
+%\end{subequations}
+%\hrule	
+%\end{strip}
+%where \eqref{lastequality} follows from the fact that the integrals involving $\langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle$ and $\langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle$ in \eqref{lastbutoneequality} are zero, see \cite[Appendix D.3]{krichene2020global}.
+%
+%From \eqref{ExplicitdPhidt}, it follows that $\frac{\differential}{\differential t}\Phi \leq 0$, i.e., $\Phi$ is non-increasing along the flow generated by \eqref{TransformedPDEivp}.
 Furthermore, \eqref{lastequality} tells us that $\frac{\differential}{\differential t}\Phi = 0$ if and only if
-$$\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}} = 0,$$
-i.e., $\Phi$ is stationary if and only if $\tilde{\rho}$ has $\bm{\eta}$ marginal $\int\tilde{\rho}\:\differential\bm{\xi} \propto \exp\left(-\bm{\eta}^{\top}\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\right)$.
+%$$\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}} = 0,$$
+%i.e., $\Phi$ is stationary if and only if $\tilde{\rho}$ has $\bm{\eta}$ marginal $\int\tilde{\rho}\:\differential\bm{\xi} \propto \exp\left(-\bm{\eta}^{\top}\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\right)$.
 Following the steps as in \cite[Appendix E.1]{krichene2020global}, this is achieved at the stationary solution of \eqref{NewPDE} with $\bm{\xi}$ marginal $\int\tilde{\rho}\:\differential\bm{\eta} \propto \exp\left(-2U(\bm{\xi})\right)$.
 In other words, the stationary solution of \eqref{NewPDE} must be of the form
-$$\tilde{\rho}_{\infty}(\bm{\xi},\bm{\eta}) \propto \exp\left(-2U(\bm{\xi})+\bm{\eta}^{\top}\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\right),$$
-and is exactly where $\frac{\differential}{\differential t}\Phi = 0$.
+%$$\tilde{\rho}_{\infty}(\bm{\xi},\bm{\eta}) \propto \exp\left(-2U(\bm{\xi})+\bm{\eta}^{\top}\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\right),$$
+%and is exactly where $\frac{\differential}{\differential t}\Phi = 0$.
 Therefore, $\frac{\differential}{\differential t}\Phi < 0$ along the transient solution of \eqref{NewPDE}, and $=0$ at the stationary solution of \eqref{NewPDE}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\vspace*{-0.1in}
 \begin{thebibliography}{99}
@@
 A.~Halder, K.F.~Caluya, P.~Ojaghi, and X.~Geng, ``Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions", \emph{extended preprint}, \\
-doi: \url{http://dx.doi.org/10.13140/RG.2.2.30956.28805}, 2021.
+URL: \url{http://dx.doi.org/10.13140/RG.2.2.30956.28805}, 2021.
 

### 2021-08-28 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

@@
 In particular, the functional $\Phi$ serves as a Lyapunov functional in the sense $\frac{\differential}{\differential t}\Phi < 0$ along the transient solution of the Fokker-Planck PDE initial value problem.
-This idea has since been generalized to many other types of PDE initial value problems, see e.g., \cite{ambrosio2008gradient,santambrogio2017euclidean}.
+This idea has since been generalized to many other types of PDE initial value problems, see e.g., \cite{ambrosio2008gradient}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace*{-0.2in}
 \section{Numerical Simulations}\label{sec:NumericalSimulations}
@@
 
-\vspace*{-0.2in}
 \subsection{IEEE 14 Bus System}\label{subsec:NumSimIEEEBus}
@@
 The comparison of the $(\theta_4,\omega_4)$ bivariate marginals between \textbf{Cases I} and \textbf{II} shown in Fig.
-\ref{fig:IEEE14marginalsBivariateComparison} indeed confirms that the post contingency transient stochastic states for bus 6 have higher dispersion or spread of the probability mass.
+\ref{fig:IEEE14marginalsBivariateComparison} indeed confirms that the post-contingency transient stochastic states for bus 6 have higher dispersion or spread of the probability mass.
 
@@
 
-\bibitem{santambrogio2017euclidean}
-F.~Santambrogio, ``$\{$Euclidean, Metric, and Wasserstein$\}$ Gradient Flows: An Overview", \emph{Bulletin of Mathematical Sciences}, Vol. 7, No. 1, pp. 87--154, 2017.
+%\bibitem{santambrogio2017euclidean}
+%F.~Santambrogio, ``$\{$Euclidean, Metric, and Wasserstein$\}$ Gradient Flows: An Overview", \emph{Bulletin of Mathematical Sciences}, Vol. 7, No. 1, pp. 87--154, 2017.
 
@@
 
-\bibitem{krichene2020global}
-W. Krichene, K.F.~Caluya, and A.~Halder, ``Global Convergence of Second-order Dynamics in Two-layer Neural Networks", \emph{arXiv preprint arXiv:2007.06852}, 2020.
+%\bibitem{krichene2020global}
+%W. Krichene, K.F.~Caluya, and A.~Halder, ``Global Convergence of Second-order Dynamics in Two-layer Neural Networks", \emph{arXiv preprint arXiv:2007.06852}, 2020.
 

### 2021-08-28 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updates

@@
 &=\underbrace{\left(\prod_{i=1}^{n=5}\dfrac{\exp\left(\kappa_{i}\cos\left(2\theta_{i}(0)-\mu_{\theta_{i}(0)}\right)\right)}{2\pi I_{0}(\kappa_{i})}\right)}_{=:\vartheta_{0}\left(\bm{\theta}(0)\right)}\nonumber\\
-	&\qquad\times \underbrace{{\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right)}_{=:\Omega_{0}\left(\bm{\omega}(0)\right)},
+	&\qquad\times \underbrace{{\rm{Unif}}\left(\left([-0.1,0.1]\;\text{rad/s}\right)^{5}\right)}_{=:\Omega_{0}\left(\bm{\omega}(0)\right)},
 \label{rho0IEEE14}		
@@
 
-In \eqref{rho0IEEE14}, the $\bm{\omega}$ marginal $ \Omega_{0}\left(\bm{\omega}(0)\right)$ is a uniform PDF over $[-0.1,0.1]$ rad/s.
-Thus, the PDF $\rho_0$ in \eqref{rho0IEEE14} is supported over $\mathbb{T}^{5}\times [-0.1,0.1]$.
+In \eqref{rho0IEEE14}, the $\bm{\omega}$ marginal $ \Omega_{0}\left(\bm{\omega}(0)\right)$ is a uniform PDF over the hypercube $\left([-0.1,0.1]\;\text{rad/s}\right)^{5}$.
+Thus, the PDF $\rho_0$ in \eqref{rho0IEEE14} is supported over $\mathbb{T}^{5}\times [-0.1,0.1]^{5}$.
 

### 2021-08-28 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

@@
 \rho_{0}\equiv\rho(t=0,\bm{\theta}(0),\bm{\omega}(0)) = \;{\rm{Unif}}\left(\left([0,2\pi)\;\text{rad}\right)^{n}\right) \\
-	\qquad\times {\rm{Unif}}\left([-0.1,0.1]\;\text{rad/s}\right),
+	\qquad\times {\rm{Unif}}\left(\left([-0.1,0.1]\;\text{rad/s}\right)^{n}\right),
 \label{rho0synthetic}	
@@
 A.~Halder, K.F.~Caluya, P.~Ojaghi, and X.~Geng, ``Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions", \emph{extended preprint}, \\
-URL: \url{http://dx.doi.org/10.13140/RG.2.2.30956.28805}, 2021.
+URL: \url{http://dx.doi.org/10.13140/RG.2.2.30956.28805/1}, 2021.
 

### 2021-08-30 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updates

@@
 Our main contribution is to demonstrate that by harnessing recent developments in generalized gradient flows \cite{ambrosio2008gradient} and proximal algorithms \cite{parikh2014proximal}, it is possible to perform \emph{nonparametric} propagation of the joint state PDFs subject to the stochastic nonlinear dynamics of a networked power system, and that the computation can be performed in a gridless and scalable manner.
-In doing so, the paper also makes theoretical contribution by introducing a change of coordinates that transforms the joint PDF evolution equations in way that is amenable for the aforesaid generalized proximal recursions.
+In doing so, the paper also makes theoretical contribution by introducing a change of coordinates that transforms the joint PDF evolution equations in a way that is amenable for the aforesaid generalized proximal recursions.
 
@@
 Given the PDE \eqref{KineticFPK}, we would like to design the functional pair $\left(\Phi,{\rm{dist}}\right)$ such that the sequence of functions $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by the proximal recursion \eqref{ProxRecursionInfiniteDim}, in the small time step limit, converges to the flow $\rho(t=kh,\bm{\theta},\bm{\omega})$ generated by the PDE initial value problem of interest.
-In particular,
+In particular\footnote{$L^{1}$ denotes the Lebesgue space of absolutely integrable functions.},
 \begin{align}
@@
 \begin{algorithmic}[1]
-\Procedure{Prox}{$\bm{\varrho}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, L$}
+\Procedure{Prox}{$\bm{\varrho}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, \ell_{\max}$}
 \For{$i=1$ to $N$}
@@
 \State $\ell = 1$
-\While{$\ell\leq L$}
+\While{$\ell\leq \ell_{\max}$}
 \State $\bm{z}(:,\ell+1) \gets \left(\bm{\zeta}_{k-1}\oslash\left(\bm{\Gamma}_{k}^{\top}\bm{y}(:,\ell)\right)\right)^{\frac{1}{1+2\varepsilon/h}}$
@@
 Specifically, $h$ is the time-step size, $\varepsilon$ is an entropic regularization weight internal to the \textproc{Prox} algorithm, and $N$ is the number of samples.
-The parameters $\delta$ and $L$ codify the numerical tolerance and maximum number of iterations, respectively, for the while loop in Algorithm \ref{algo:KineticProx}.
+The parameters $\delta$ and $\ell_{\max}$ codify the numerical tolerance and maximum number of iterations, respectively, for the while loop in Algorithm \ref{algo:KineticProx}.
 Its convergence guarantees can be found in \cite[Sec.
@@
 \ref{sec:ProxAlgo}, for all the three cases, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
-In Algorithm \ref{algo:KineticProx}, we used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $L=300$.
+In Algorithm \ref{algo:KineticProx}, we used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $\ell_{\max}=300$.
 
@@
 \end{tabular}
+\vspace*{0.1in}
 \caption{Parameters used for the numerical simulation in Sec.
@@
 As in Sec.
-\ref{subsec:NumSimIEEEBus}, we used $h=10^{-3}$, $\varepsilon = 0.05$, $\delta = 10^{-3}$, $L = 300$ in Algorithm \ref{algo:KineticProx}.
+\ref{subsec:NumSimIEEEBus}, we used $h=10^{-3}$, $\varepsilon = 0.05$, $\delta = 10^{-3}$, $\ell_{\max} = 300$ in Algorithm \ref{algo:KineticProx}.
 

### 2021-08-30 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

@@
 \end{tabular}
+\vspace*{0.05in}
 \caption{Parameters used in Sec. \ref{subsec:NumSimSynthetic}. The indices $i,j\in\{1,\hdots,n\}$ where the number of generators $n=50$ in Sec. \ref{subsec:NumSimSynthetic}.}
@@
 \end{figure}
-\begin{figure}[t]
+\begin{figure}[h]
 \centering

### 2021-08-30 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updates

@@
    \subfigure[]{\includegraphics[width=0.309\textwidth]{Case0Case1Combined_Gen4_OmegaMarginals.png}}
-\caption{\small{For $i=1,\hdots,5$, the $\omega_{i}$ marginals (a) for Case I, and (b) for Case II, resulting from the IEEE 14 bus simulation described in Sec.
+\caption{\small{For $i=1,\hdots,5$, the $\omega_{i}$ marginals (a) for \textbf{Case I}, and (b) for \textbf{Case II}, resulting from the IEEE 14 bus simulation described in Sec.
 \ref{subsec:NumSimIEEEBus}.
-(c) Comparison of the $\omega_{4}$ marginals for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios.}}
+(c) Comparison of the $\omega_{4}$ marginals for bus 6 in the nominal (\textbf{Case I}) and post-fault (\textbf{Case II}) scenarios.}}
 \vspace*{-0.1in}
@@
 Using the algorithm detailed in Sec.
-\ref{sec:ProxAlgo}, for all the three cases, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
+\ref{sec:ProxAlgo}, for both cases, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
 In Algorithm \ref{algo:KineticProx}, we used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $\ell_{\max}=300$.
@@
 In particular, Fig.
-\ref{fig:IEEE14omegamarginalsCase12}(c) helps compare the $\omega_{4}$ marginal evolution for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios.
-Since Case II considers the failure of line 13, from Fig.
+\ref{fig:IEEE14omegamarginalsCase12}(c) helps compare the $\omega_{4}$ marginal evolution for bus 6 in the nominal (\textbf{Case I}) and post-fault (\textbf{Case II}) scenarios.
+Since \textbf{Case II} considers the failure of line 13, from Fig.
 \ref{fig:Motivating} we intuitively expect that the transient effects of this fault will be prominent for bus 6 (generator 4).
 Indeed, Fig.
-\ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both Case I and II, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger spread or dispersion of the probability mass, as expected.
-That the generator 4 states in Case II have larger uncertainties than in Case I, is also visible in the bivariate marginal plots if we compare the fourth row in Fig.
+\ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both \textbf{Case I} and \textbf{II}, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger spread or dispersion of the probability mass, as expected.
+That the generator 4 states in \textbf{Case II} have larger uncertainties than in \textbf{Case I}, is also visible in the bivariate marginal plots if we compare the fourth row in Fig.
 \ref{fig:IEEE14marginalsCase0} with the same in Fig.
@@
 
-Fig. \ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec. \ref{sec:ProxAlgo}, for both Cases I and II.
+Fig.
+\ref{fig:CompTimeIEEE14bus} shows the computational times needed to update $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec.
+\ref{sec:ProxAlgo}, for both \textbf{Cases I} and \textbf{II}.
 
@@
 \end{tabular}
-\vspace*{0.1in}
+\vspace*{0.05in}
 \caption{Parameters used for the numerical simulation in Sec.

### 2021-08-30 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

@@
 \includegraphics[width=0.8\linewidth]{IEEE14BivariateMarginalComparison.pdf}
-\caption{\small{Comparison of the $(\theta_{4},\omega_{4})$ marginals for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios in the IEEE 14 bus simulation described in Sec.
+\caption{\small{Comparison of the $(\theta_{4},\omega_{4})$ marginals for bus 6 in the nominal (\textbf{Case I}) and post-fault (\textbf{Case II}) scenarios in the IEEE 14 bus simulation described in Sec.
 \ref{subsec:NumSimIEEEBus}.}}
@@
 \includegraphics[width=0.85\linewidth]{Case0Case1Combined_Gen4_OmegaMarginals.png}
-\caption{\small{Comparison of the $\omega_{4}$ marginals for bus 6 in the nominal (Case I) and post-fault (Case II) scenarios in the IEEE 14 bus simulation described in Sec.
+\caption{\small{Comparison of the $\omega_{4}$ marginals for bus 6 in the nominal (\textbf{Case I}) and post-fault (\textbf{Case II}) scenarios in the IEEE 14 bus simulation described in Sec.
 \ref{subsec:NumSimIEEEBus}.}}
@@
 
-Fig. \ref{fig:CompTimeIEEE14bus} shows the computational times for updating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec. \ref{sec:ProxAlgo} for Cases I and II.
+Fig.
+\ref{fig:CompTimeIEEE14bus} shows the computational times for updating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec.
+\ref{sec:ProxAlgo} for \textbf{Cases I} and \textbf{II}.
 

### 2021-08-30 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
updates

@@
 Fig.
-\ref{fig:CompTimeIEEE14bus} shows the computational times for updating $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec.
+\ref{fig:CompTimeIEEE14bus} shows the computational times for updating the point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ via the algorithm in Sec.
 \ref{sec:ProxAlgo} for \textbf{Cases I} and \textbf{II}.

### 2021-09-07 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_10pg.tex
fixed indexing issues with equations (7) and (8)

@@
 The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively.
-We use the superscript $*$ to denote the complex conjugate.
+We use the superscript $*$ to denote the complex conjugate, and the superscript $^{\top}$ to denote the matrix transpose.
 The uniform probability distribution over an interval $[a,b]$ is denoted as ${\rm{Unif}}\left([a,b]\right)$.
@@
 With the $i$th generator, we associate its inertia $m_{i}>0$ and damping coefficient $\gamma_{i}>0$.
-The other parameters: the \emph{effective power input} $P_{i}$, the \emph{phase shift} $\varphi_{ij}\in[0,\frac{\pi}{2})$, and the \emph{coupling coefficients} $k_{ij}\geq 0$, depend on the network reduced admittance matrix $\bm{Y}\equiv[Y_{ij}]_{i,j=1}^{n}\in\mathbb{C}^{n\times n}$.
+The other parameters: the \emph{effective power input} $P_{i}$, the \emph{phase shift} $\varphi_{ij}\in[0,\frac{\pi}{2})$, and the \emph{coupling coefficients} $k_{ij}\geq 0$, depend on the network reduced admittance matrix $\bm{Y}\equiv[Y_{ij}]_{i,j=1}^{n}\in\mathbb{C}^{n\times n}$, $\bm{Y}=\bm{Y}^{\top}$.
 Specifically,
@@
 \begin{align}
-V(\bm{\theta}) := -\displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{i,j=1}^{n}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right).
+V(\bm{\theta}) := -\displaystyle\sum_{i=1}^{n} P_{i}\theta_{i} + \!\displaystyle\sum_{\substack{i,j=1\\i<j}}^{n}\!k_{ij}\left(1 - \cos(\theta_{i}-\theta_{j}-\varphi_{ij})\right).
 \label{potential}	

### 2021-12-04 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

\documentclass[12pt,onecolumn]{IEEEtran}
\usepackage{amsmath ,amssymb,euscript ,yfonts,psfrag,latexsym,dsfont,graphicx,bbm,color,amstext,wasysym,subfig,flushend,parskip}
\usepackage{bm,cite,soul,url}
\usepackage{verbatim}
\usepackage[normalem]{ulem}
\usepackage{amsthm}
%\usepackage{refcheck}
%\usepackage{showkeys}
%\usepackage{epstopdf,hyperref,pdfsync,url}
\graphicspath{{./},{./figures/}}
\newcommand{\mR}{{\mathbb R}}
\newcommand{\D}{{\mathbb D}}
\newcommand{\E}{{\mathbb E}}
\newcommand{\mcN}{{\mathcal N}}
\newcommand{\mcR}{{\mathcal R}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\tr}{\operatorname{trace}}
\newcommand{\ignore}[1]{}
\newcommand{\mike}{\color{magenta}}

\newcommand{\blue}{\color{blue}}
\newcommand{\red}{\textcolor{red}}

\newcommand{\bQ}{{\bm{Q}}}
\newcommand{\bPi}{\bm{\Pi}}
\newcommand{\bv}{{\bm{v}}}
\newcommand{\balpha}{{\bm{\alpha}}}
\newcommand{\bbeta}{{\bm{\beta}}}
\newcommand{\bq}{{\bm{q}}}
\newcommand{\bp}{{\bm{p}}}
\newcommand{\bbf}{{\bm{f}}}
\newcommand{\differential}{{\mathrm{d}}}

\newtheorem{claim}{Claim}

\def\spacingset#1{\def\baselinestretch{#1}\small\normalsize}
\setlength{\parskip}{10pt}
\setlength{\parindent}{20pt}
\spacingset{1}

\definecolor{grey}{rgb}{0.6,0.6,0.6}
\definecolor{lg}{rgb}{0.6,.6,0.6}
\newcommand{\nib}{\noindent  {\bf Response:} }

\begin{document}
%\newcommand{\bp}{{\bm{p}}}
\noindent
{\large {\bf Revision of
Manuscript Number}: IEEE-TPWRS-01391-2021, version 1\hfill \today\\[.1in]
{\bf Journal}: IEEE Transactions on Power Systems \\[.1in]
{\bf Author}: Abhishek Halder, Kenneth F. Caluya, Pegah Ojaghi, Xinbo Geng}\\

\section*{\large \bf Response to decision:}

{\noindent\blue
The perceptive input from all reviewers and the AE are much appreciated.

Please find below itemized responses to the individual comments addressing the concerns and detailing the revisions. All the edits in the revised manuscript are marked in blue.\\

\noindent
Sincerely,\\
The authors
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\newpage
\spacingset{1}

\section*{\large \bf Response to AE:}

\noindent
{\bf AE.1. Comments by AE}:\\
{\em The authors are requested to revise and resubmit their paper after addressing reviewer feedback. Of particular importance is that reviewers point out the model leveraged for analysis is overly simplistic.}

{\nib {\blue Thanks for this comment.
}}

\noindent
{\bf AE.2. Comments by AE}:\\
{\em Authors are also encouraged to revisit their overview of prior art.}

{\nib {\blue Please see }}

\noindent
{\bf AE.3. Comments by AE}:\\
{\em Presentation of technical points and results can also be improved.}

{\nib {\blue Please see }}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\spacingset{1}

\section*{\large \bf Response to Reviewer \#1:}

\noindent
{\bf R1.1. Comment by reviewer}:\\
{\em The original contribution of the article is not clear.  Indeed, what is original new from this article, not the summary from existing techniques?}

{\nib {\blue This is. 
}}

\noindent
{\bf R1.2. Comments by reviewer}:\\
{\em The structure of the article should be highly improved. A large portion of the contexts is simply repeating the well-known power system classical model, kroon reduction, and the FPF equations.  Only from section IV, page 6, I can see some of the proposed weighted points methods.}

{\nib{ \blue We have.}}

\noindent
{\bf R1.3. Comments by reviewer}:\\
{\em I have major doubts about the model used in the article for the following reasons.\\
(1)     The paper claim the proposed method can propagate stochastic uncertainties in realistic power networks. However, only a classical model is used, which is not realistic at all.\\
(2)     Why do active power and reactive power from load have no randomness at all. This is too unrealistic.\\
(3)     The diffusion term of this article is only considered in the frequency. The process noise in the rotor angle is fully ignored. This is also unreasonable.\\
To improve your work, I highly suggested the authors follow [R1], which is missing in your literature but can help the authors better know Power system SDE.  Also, it is suggested to introduce the OU process to model the randomness in the load, which cannot be ignored in practice.\\
R1. A systematic method to model power systems as stochastic differential-algebraic equations, F. Milano, R. Zarate-Minano, IEEE Transactions on Power Systems, 28(4), 4537-4544.}

{\nib{ \blue TBD.}}

\noindent
{\bf R1.4. Comments by reviewer}:\\
{\em There is no comparison of the proposed method compared with other state-of-the-art methods. Therefore, it is hard for the readers to judge the importance of this article.}

{\nib{ \blue TBD.}}

\noindent
{\bf R1.5. Comments by reviewer}:\\
{\em Contributions are over-claimed in this article:\\
(1) The scalability of the proposed method cannot convince the readers since you are using a too simple model and ignored many important random dimensions from the loads.\\
(2) The proposed method claims to be nonparametric, however, all the inputs are assumed to be either wiener process or uniform distribution, which are all parametric distributions.\\
(3) If the ``nonparametric" is only referred to the output PDF, I would not think of it as a major contribution since there are a lot of existing techniques that can depict nonparametric PDF.}

{\nib{ \blue TBD.}}

\noindent
{\bf R1.6. Comments by reviewer}:\\
{\em The simulation time is too short to convince the readers. Only 1 second is considered here. Note that the errors can be quickly accumulated in the iteration.}

{\nib{ \blue TBD.}}

\noindent
{\bf R1.7. Comments by reviewer}:\\
{\em It is suggested to count the total computing time, not on each step.}

{\nib{ \blue TBD.}}

\noindent
{\bf R1.8. Comments by reviewer}:\\
{\em If the proposed method depends on the Kronecker product, how can it be scaled to a higher dimension?}

{\nib{ \blue TBD.}}

\noindent
{\bf R1.9. Comments by reviewer}:\\
{\em beta is undefined in equation (15a).}

{\nib{ \blue TBD.}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section*{\large \bf Response to Reviewer \#2:}

\noindent
{\bf R2.1. Comments by reviewer}:\\
{\em Overall, the paper is well-written and theoretically sound.}

{\nib \blue{Thank you for the kind and encouraging remarks.}}

\noindent
{\bf R2.2. Comments by reviewer}:\\
{\em The paper's language and presentation make it seem like it is more oriented towards the applied mathematics and control communities. It is highly suggested that the authors revise the exposition of ideas and concepts, putting more emphasis on the power system so that it can be well-understood by the power system experts.}

{\nib \blue{TBD.}}

\noindent
{\bf R2.3. Comments by reviewer}:\\
{\em How will the proposed approach be used for the cases in which dependencies among uncertainty sources exist? An explanation and illustration would be helpful.}

{\nib \blue{TBD.}}

\noindent
{\bf R2.4. Comments by reviewer}:\\
{\em Incorporation of power system dynamics into the proposed uncertainty propagation framework is a bit vague. The proposed study seems not to have utilized any tool capable of simulating power system dynamics. Could the authors explain the part pertaining to the dynamic simulation in more detail?}

{\nib \blue{TBD.}}

\noindent
{\bf R2.5. Comments by reviewer}:\\
{\em With regard to the point above, this reviewer would expect to see uncertainty bounds for the system dynamic response such as the trajectory of the rotor angle. Without such information, the method's usability is difficult to gauge.}

{\nib \blue{TBD.}}

\noindent
{\bf R2.6. Comments by reviewer}:\\
{\em Instead of using a purely synthetic system, the authors are highly encouraged to choose available test cases such as the IEEE 118-bus system and a larger one, if possible.}

{\nib \blue{TBD.}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section*{\large \bf Response to Reviewer \#3:}

\noindent
{\bf R3.1. Comments by reviewer}:\\
{\em My main concern with this work is that all its derivations are deeply model-dependent and, unfortunately, the model that the author consider is not adequate to represent modern power systems. Moreover, second order model of synchronous machines are adequate only in a time scale of few seconds, so, even if the system were effectively including only conventional power plants, the model would be acceptable only for very short-term transients, i.e., transients for which the propagation of the noise is immaterial. In other words, the proposed model is useless.}

{\nib \blue{TBD.}}

\noindent
{\bf R3.2. Comments by reviewer}:\\
{\em The literature is incomplete, many recent works on SDAEs for power system applications are not mentioned.}

{\nib \blue{TBD.}}

\noindent
{\bf R3.3. Comments by reviewer}:\\
{\em There are several typos in the test (especially in the references).}

{\nib \blue{TBD.}}

\noindent
{\bf R3.4. Comments by reviewer}:\\
{\em Figures 5 and 6 are unreadable.}

{\nib \blue{TBD.}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section*{\large \bf Response to Reviewer \#4:}

\noindent
{\bf R4.1. Comments by reviewer}:\\
{\em The main contribution of this paper lies in applying results on the governing partial differential equation (PDE) for the joint PDF to avoid discretization of the state space, thereby circumventing computational challenges. The reviewer appreciates that this paper is rigorous in development and thinks it deals with an important problem.}

{\nib \blue{We appreciate the positive remark.}}

\noindent
{\bf R4.2. Comments by reviewer}:\\
{\em The key contribution of this paper is that the proposed method can reduce the computation burden when solving for the transient PDF. However, no comparison is provided with existing methods. The authors should provide comparisons to the exiting approaches, for example, based on finite element discretization or other numerical methods. The test example could be performed using the 5-generator system or larger systems. Besides, referring to Figs. 8 and 9, the total computation time should be reported for a longer horizon.}

{\nib \blue{TBD.}}

\noindent
{\bf R4.3. Comments by reviewer}:\\
{\em Regarding the presentation, the reviewer suggests making the paper more self-contained. Some discussion in this paper highly relies on the authors' previous work, which can be improved by either adding a summarization or trimming out the materials.

Following the same vein, there are some unnecessary information and technical points provided in this paper that may be distracting.
For example, on page 3, why would the authors mention the Kuramoto oscillator model?
On page 4, is the detailed discussion regarding Wasserstein-2 metric necessary? Similarly, on page 5, the discussion about the natural energy functional and how the gradient drift case does not apply also seems too detailed.}

{\nib \blue{TBD.}}

\noindent
{\bf R4.4. Comments by reviewer}:\\
{\em Along with [25], there are recent developments that over-approximates the reachable set for nonlinear systems under uncertain initial conditions, parameters, and external disturbances:
``Reach-Set Estimation for DAE Systems under Uncertainty and Disturbances Using Trajectory Sensitivity and Logarithmic Norm".}

{\nib \blue{TBD.}}

\noindent
{\bf R4.5. Comments by reviewer}:\\
{\em On page 7, the authors stated about ``To account the pertinent geometry of $\mathbb{T}^n \times \mathbb{R}^n$". What exactly does this mean? Do the authors indicate that in (28) the two groups of state values are independent? Or do the authors mean the specific choices of the product von Mises PDF and the uniform distribution? If so, what are the connections?}

{\nib \blue{TBD.}}

\noindent
{\bf R4.6. Comments by reviewer}:\\
{\em On page 5, is there a mistake in ``random samples from (28), and evaluated at (28)"?}

{\nib \blue{TBD.}}

\noindent
{\bf R4.7. Comments by reviewer}:\\
{\em As a validation, it would be good to also provide information of higher-order moments along with the mean given in Fig. 10.}

{\nib \blue{TBD.}}

\noindent
{\bf R4.8. Comments by reviewer}:\\
{\em There are some typos in this paper:\\
Page 4: in therms of\\
Page 5: in the right-hand side of the arrow in the expression following ``converges to the flow generated by (22)'', it should be plain $\rho$}

{\nib \blue{TBD.}}

\end{document}

### 2021-12-16 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 {\nib {\blue Thanks for this comment.
+
+In addition to the itemized responses to the reviewers' comments and the corresponding revisions marked in blue in the manuscript, we have made the \ul{following additional fixes and improvements} (also marked in blue in the revised manuscript):\\
+(i) In equation (8), we corrected a typo in its right hand side: the second summand runs over pairs $(i,j)$ such that $i<j$.\\
+(ii)  
 }}
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{That was not a typo.
+We meant that (28) was \ul{used in two ways}: it was used to generate initial state samples, and then the initial joint PDF values at those samples were evaluated via the functional form of (28), i.e., as in function evaluation.
+This is needed in our case because unlike standard Monte Carlo, we \ul{co-evolve} the state samples and joint PDF values at those state samples.
+\ul{To clarify this further, in the revised manuscript, we have slightly re-phrased that sentence}.}}
 
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{Corrected both. Thanks for pointing them out.}}
 

### 2021-12-16 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
initiated TPS rebuttal version 1

@@
 \newcommand{\red}{\color{red}}
+\newcommand{\blue}{\color{blue}}
 
@@
 \end{align}
-We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in therms of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous.
+We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in {\blue{terms}} of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous.
 Fig.
@@
 converges to the flow generated by \eqref{TransformedPDEivp}, i.e., 
-$$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} \tilde{\varrho}(t=kh,\bm{\xi},\bm{\eta})\;\text{in}\;L^{1}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right),$$
+$$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} {\blue{\tilde{\rho}}}(t=kh,\bm{\xi},\bm{\eta})\;\text{in}\;L^{1}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right),$$
 was established in \cite{duong2014conservative}.
@@
 The proposed computational framework is also applicable when $P_{i}^{\rm{mech}}, P_{i}^{\rm{load}}$, and thus $P_{i}$ in \eqref{defPi}, are bounded time-varying functions.
-This can be the case, for instance, with rapid fluctuations from renewables in the short term.
+This can be the case, for instance, with rapid fluctuations from renewables {\blue{or from loads}} in the short term.
 \end{remark}
@@
 Using the algorithm detailed in Sec.
-\ref{sec:ProxAlgo}, for both cases, we propagated $N=1000$ random samples from \eqref{rho0IEEE14}, and evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
+\ref{sec:ProxAlgo}, for both cases, we propagated $N=1000$ random {\blue{state}} samples from \eqref{rho0IEEE14}, and {\blue{corresponding joint PDF values}} evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
 In Algorithm \ref{algo:KineticProx}, we used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $\ell_{\max}=300$.

### 2021-12-16 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{Thanks for providing that interesting recent reference. We have cited the same in the revised manuscript.}}
 
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{We simply mean that the joint PDF should be supported on $\mathbb{T}^n \times \mathbb{R}^n$, or a submanifold thereof.
+For instance, one cannot assume the initial joint PDF to be a $2n$ dimensional multivariate normal since that is supported over $\mathbb{R}^{2n}$, not over $\mathbb{T}^n \times \mathbb{R}^n$.
+It is, however, perfectly fine if some or all of the $2n$ components of the initial state vector are statistically correlated.
+As long as the support is $\mathbb{T}^n \times \mathbb{R}^n$, there is no issue if the angular variables are correlated with the Euclidean (i.e., angular velocity) variables.}}
 
@@
 This is needed in our case because unlike standard Monte Carlo, we \ul{co-evolve} the state samples and joint PDF values at those state samples.
-\ul{To clarify this further, in the revised manuscript, we have slightly re-phrased that sentence}.}}
+\ul{To clarify this further, in the revised manuscript, we have slightly expanded that sentence}.}}
 

### 2021-12-16 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
initiated TPS rebuttal version 1

@@
 
-In a different vein, \emph{deterministic bounded uncertainty models} for power flow simulations have been used \cite{dimitrovski2004boundary,hiskens2006sensitivity,chen2012method,althoff2014formal} for set-valued analysis.
+In a different vein, \emph{deterministic bounded uncertainty models} for power flow simulations have been used \cite{dimitrovski2004boundary,hiskens2006sensitivity,chen2012method,althoff2014formal,geng2020reach} for set-valued analysis.
 These, however, require approximating the underlying nonlinear differential algebraic equations (DAEs) appearing in power system dynamic simulation, and thus lead to conservative analysis.
@@
 M. Althoff, ``Formal and Compositional Analysis of Power Systems using Reachable Sets", \emph{IEEE Transactions on Power Systems}, Vol. 29, No. 5, pp. 2270--2280, 2014.
+
+{\blue{\bibitem{geng2020reach}
+S.~Geng, and I.A.~Hiskens, ``Reach-Set Estimation for DAE Systems under Uncertainty and Disturbances Using Trajectory Sensitivity and Logarithmic Norm", \emph{21st IFAC World Congress}, Vol.
+53, No.
+2, pp.
+1955--1961, 2020.}}
 

### 2021-12-16 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 
-{\nib{ \blue TBD.}}
+{\nib{ \blue The proposed method \ul{does not entail computation of a Kronecker product}.
+In the manuscript, the Kronecker product with the $2\times 2$ identity matrix appeared in the definition of the change-of-variable (17).
+\ul{Its only purpose is to define} that change-of-variable, and \ul{thereby derive} the PDE in the new variables.
+The proximal recursion (for the PDE thus derived) detailed in Algorithm 1 \ul{does not have any Kronecker product computation}.}}
 
@@
 
-{\nib{ \blue TBD.}}
+{\nib{ \blue Thanks for the comment.
+The $\beta$ in (15a) is the same $\beta$ in (16), as in part of the same sentence.
+To avoid any confusion, in the revised manuscript, we have moved the phrase ``for some $\beta>0$" in (15a).}}
 

### 2021-12-25 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 
-{\nib{ \blue TBD.}}
+{\nib{ \blue While the depiction of the computational times in Figs.
+8 and 9 in the revised manuscript are somewhat unconventional, there is a reason why we choose to plot that way.
+At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1.
+This proximal update involves execution of a contractive fixed point recursion.
+The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update.
+To claim that the proposed method is }}
 
@@
 The $\beta$ in (15a) is the same $\beta$ in (16), as in part of the same sentence.
-To avoid any confusion, in the revised manuscript, we have moved the phrase ``for some $\beta>0$" in (15a).}}
+\ul{To avoid any confusion, in the revised manuscript, we have moved the phrase ``for some $\beta>0$" in (15a)}.}}
 
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{TBD FOR FIRST PARA.
+
+We appreciate the reviewer's point that the sentence containing equation (6) and the mention of Kuramoto model can be removed without affecting the rest of the development.
+However, we prefer to keep it since it adds didactic value, and many engineers may feel comfortable by the explicit mention that (1) is simply a rewriting of (6).
+Furthermore, an astute reader, encountering that sentence will realize that the scope of the technical contribution of this paper is broad because the algorithmic framework developed here will be applicable to several other non-power system applications (mentioned in the two cited references in that sentence) where the second order Kuramoto model is used.
+Since this is just a matter of one sentence, we will prefer to keep it.
+We hope you approve.
+
+\ul{On the other hand, we believe the detailed discussion regarding Wasserstein gradient flow and why that standard development cannot be applied in this case, is in fact, the heart of this paper}.
+Please note that \ul{ours is not a recipe paper} where a numerical algorithm is proposed backed by some theory.
+The key contribution of this paper, from the authors' viewpoint, is to help distill out the core technical difficulty that hinders simply importing standard techniques from the optimal transport literature.
+We believe this clarification of what really is the technical crux of the matter--versus what are peripheral issues--should be illuminating to not just power systems researchers, but also to applied mathematicians underscoring how the standard theory and algorithms for measure-valued gradient flows need to be adapted in engineering applications.}}
 
@@
 
-{\nib \blue{Thanks for providing that interesting recent reference. We have cited the same in the revised manuscript.}}
+{\nib \blue{Thanks for providing that interesting recent reference. \ul{We have cited the same in the revised manuscript}.}}
 
@@
 It is, however, perfectly fine if some or all of the $2n$ components of the initial state vector are statistically correlated.
-As long as the support is $\mathbb{T}^n \times \mathbb{R}^n$, there is no issue if the angular variables are correlated with the Euclidean (i.e., angular velocity) variables.}}
+As long as the support is $\mathbb{T}^n \times \mathbb{R}^n$, there is no issue if the angular variables are correlated with the Euclidean (i.e., angular velocity) variables.
+
+}}
 
@@
 {\nib \blue{That was not a typo.
-We meant that (28) was \ul{used in two ways}: it was used to generate initial state samples, and then the initial joint PDF values at those samples were evaluated via the functional form of (28), i.e., as in function evaluation.
+Indeed, equation (28) was \ul{used in two ways}: it was used to generate initial state samples, and then the initial joint PDF values at those samples were evaluated via the functional form of (28), i.e., as in function evaluation.
 This is needed in our case because unlike standard Monte Carlo, we \ul{co-evolve} the state samples and joint PDF values at those state samples.

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 
-{\nib{ \blue While the depiction of the computational times in Figs.
-8 and 9 in the revised manuscript are somewhat unconventional, there is a reason why we choose to plot that way.
+{\nib{ \blue While the depiction of the computational times in Figs. 8 and 9 in the revised manuscript are somewhat unconventional, we explain why we choose to plot that way.
+
 At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1.
@@
 The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update.
-To claim that the proposed method is }}
+To claim that the proposed method is practical, we therefore, need to demonstrate that the computational timescale is faster than the physical timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time.
+On the other hand, theoretical consistency (see the unnumbered equation after (26), before Sec.
+IV) requires that $h$ itself should be small.
+Thus, even if one has a provably convergent algorithm that can perform the one step proximal update, it is not clear if the computational time needed to do so can be much faster than $h$.
+If not, then the method would be statistically consistent but practically not useful.
+The Figs.
+8 and 9 in the revised manuscript show that the computational time needed to perform the proximal updates, for all $t_{k}$, are indeed orders-of-magnitude faster than $h$.}}
 
@@
 Please note that \ul{ours is not a recipe paper} where a numerical algorithm is proposed backed by some theory.
-The key contribution of this paper, from the authors' viewpoint, is to help distill out the core technical difficulty that hinders simply importing standard techniques from the optimal transport literature.
+\ul{The key contribution of this paper, from the authors' viewpoint, is to help distill out the core technical difficulty that hinders simply importing standard techniques from the optimal transport literature}.
 We believe this clarification of what really is the technical crux of the matter--versus what are peripheral issues--should be illuminating to not just power systems researchers, but also to applied mathematicians underscoring how the standard theory and algorithms for measure-valued gradient flows need to be adapted in engineering applications.}}

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update.
-To claim that the proposed method is practical, we therefore, need to demonstrate that the computational timescale is faster than the physical timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time.
+\ul{To claim that the proposed method is practical, we therefore, need to demonstrate that the computational timescale is faster than the physical timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time}.
 On the other hand, theoretical consistency (see the unnumbered equation after (26), before Sec.
 IV) requires that $h$ itself should be small.
-Thus, even if one has a provably convergent algorithm that can perform the one step proximal update, it is not clear if the computational time needed to do so can be much faster than $h$.
+Thus, even if one has a provably convergent algorithm that can perform the one step proximal update, it is not clear \emph{a priori} if the computational time needed to do so can be much faster than $h$.
 If not, then the method would be statistically consistent but practically not useful.
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{In our paper, we do not make any assumption on statistical independence of the uncertainties.
+\ul{In fact, all stochastic uncertainties are considered to be ``joint", i.e., statistically correlated}.}}
 

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 {\nib \blue{In our paper, we do not make any assumption on statistical independence of the uncertainties.
-\ul{In fact, all stochastic uncertainties are considered to be ``joint", i.e., statistically correlated}.}}
+\ul{In fact, all stochastic uncertainties are considered to be ``joint", i.e., statistically correlated}.
+Please note that the initial joint PDFs (28) and (30) are taken to be in that form only for the sake of concrete example simulations.
+Their only purpose, in the simulations reported in Sec IV, is to generate joint samples, and to evaluate the initial joint PDF values at the joint samples thus generated.
+If this initial scattered point cloud data is supplied to us only as a numerical table, without specifying any functional form whatsoever, nothing changes in the proposed algorithm.
+Similarly, the parameters in Table I could be correlated random vectors, and all we need are the \ul{joint} parametric weighted point cloud data.
+In fact, this is how we expect the proposed computational framework to be applied in practice.}}
 

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 \ul{In fact, all stochastic uncertainties are considered to be ``joint", i.e., statistically correlated}.
-Please note that the initial joint PDFs (28) and (30) are taken to be in that form only for the sake of concrete example simulations.
-Their only purpose, in the simulations reported in Sec IV, is to generate joint samples, and to evaluate the initial joint PDF values at the joint samples thus generated.
+Please note that the initial joint PDFs (28) and (30) are taken to be in that form for the sake of concrete example simulations.
+Their only purpose, in the simulations reported in Sec V, is to generate joint samples, and to evaluate the initial joint PDF values at the joint samples thus generated.
 If this initial scattered point cloud data is supplied to us only as a numerical table, without specifying any functional form whatsoever, nothing changes in the proposed algorithm.
-Similarly, the parameters in Table I could be correlated random vectors, and all we need are the \ul{joint} parametric weighted point cloud data.
+Similarly, the parameters in Table I could instead be correlated random vectors, and all we need are the \ul{joint} parametric weighted point cloud data.
 In fact, this is how we expect the proposed computational framework to be applied in practice.}}
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{In this paper, we are interested in propagating the joint PDF values of the states of a power system dynamic model subject to joint stochastic uncertainties.
+Our focus is on explicitly computing these transient joint state PDFs at the time-varying state vector samples, which amounts to co-evolving the states and the joint PDF values over time.
+\ul{This is different from Monte Carlo or the traditional density estimation techniques since they approximate the joint PDFs via post-processing, not as an online computation}.
+To the best of our knowledge, there does not exist a power system simulation tool which can do this online computation while making neither statistical nor dynamical approximation.
+
+Please note that in Sec. V, we have used MATPOWER and the Python-library ANDES (refs. [59] and [60]) for obtaining some parameters, as explained in the first paragraph of V-A. }}
 

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 
-Please note that in Sec. V, we have used MATPOWER and the Python-library ANDES (refs. [59] and [60]) for obtaining some parameters, as explained in the first paragraph of V-A. }}
+Please note that in Sec.
+V, we have used MATPOWER and the Python-library ANDES (with inline citation of refs.
+[59] and [60] therein) for obtaining some simulation parameters, as explained in the first paragraph of V-A.}}
 

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 V, we have used MATPOWER and the Python-library ANDES (with inline citation of refs.
-[59] and [60] therein) for obtaining some simulation parameters, as explained in the first paragraph of V-A.}}
+[61] and [62] therein) for obtaining some simulation parameters, as explained in the first paragraph of V-A.}}
 
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{The reviewer's point is well taken. TBD.}}
 
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{We appreciate the comment.
+As per the reviewer's suggestion, in the revised manuscript, we have included a sentence (marked blue) in Sec.
+I.1, second paragraph (last sentence) mentioning the usage of SDAEs in power system dynamic simulation.
+That sentence in the revised manuscript contains two newly added references: [22] and [23], both on simulating SDAE sample paths.}}
 
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{TBD. We corrected the two typos pointed out in comment \textbf{R4.8} in this rebuttal document.}}
 

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
initiated TPS rebuttal version 1

@@
 However, this leads to moment closure problems since the dimensions of the time-varying sufficient statistics associated with the corresponding transient joint state PDFs are not known in general.
+{\blue{We also note the usage of \emph{stochastic} differential algebraic equations (DAEs) for power system dynamic simulation \cite{wang2011numerical,milano2013systematic}.
+These studies, however, simulate the sample paths and does not directly propagate the joint state PDFs.}}
 
 In a different vein, \emph{deterministic bounded uncertainty models} for power flow simulations have been used \cite{dimitrovski2004boundary,hiskens2006sensitivity,chen2012method,althoff2014formal,geng2020reach} for set-valued analysis.
-These, however, require approximating the underlying nonlinear differential algebraic equations (DAEs) appearing in power system dynamic simulation, and thus lead to conservative analysis.
+These, however, require approximating the underlying nonlinear (deterministic) DAEs appearing in power system dynamic simulation, and thus lead to conservative analysis.
 For example, the method in \cite{althoff2014formal} requires converting the nonlinear DAEs to linear DAEs in such a way that guarantees set-valued over-approximation of the reachable sets.
@@
 \begin{align}
-\rho_{\infty}\left(\bm{\theta},\bm{\omega}\right) &= \dfrac{1}{Z}\exp\left(-\beta H\right), \\
+\rho_{\infty}\left(\bm{\theta},\bm{\omega}\right) &= \dfrac{1}{Z}\exp\left(-\beta H\right), \quad\text{for some}\;\beta > 0,\\
 H(\bm{\theta},\bm{\omega}) &:= V(\bm{\theta}) + \frac{1}{2}\langle\bm{\omega},\bm{M\omega}\rangle,	
@@
 \begin{align}
-\bm{\Sigma}\bm{\Sigma}^{\top} = \beta^{-1}\left(\bm{\Gamma} + \bm{\Gamma}^{\top}\right) \quad\text{for some}\;\beta > 0.
+\bm{\Sigma}\bm{\Sigma}^{\top} = \beta^{-1}\left(\bm{\Gamma} + \bm{\Gamma}^{\top}\right).
 \label{EinsteinRelation}	
@@
 \begin{algorithmic}[1]
-\Procedure{Prox}{$\bm{\varrho}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, \ell_{\max}$}
+\Procedure{Prox}{$\tilde{\bm{\varrho}}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, \ell_{\max}$}
 \For{$i=1$ to $N$}
@@
 doi: 10.1109/PESGM.2018.8586023
+
+% === stochastic DAEs ======
+
+\bibitem{wang2011numerical}
+K.~Wang, and M.L.~Crow, ``Numerical simulation of stochastic differential algebraic equations for power system transient stability with random loads", \emph{2011 IEEE Power \& Energy Society General Meeting (PESGM)}, pp.
+1--8, 2011.
+doi: 10.1109/PES.2011.6039188
+
+\bibitem{milano2013systematic}
+F.~Milano, and R.~Z{\'a}rate-Mi{\~n}ano, ``A systematic method to model power systems as stochastic differential algebraic equations", \emph{IEEE Transactions on Power Systems}, Vol.
+28, No.
+4, pp.
+4537--4544, 2013.
 

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 {\nib \blue{We appreciate the comment.
-As per the reviewer's suggestion, in the revised manuscript, we have included a sentence (marked blue) in Sec.
-I.1, second paragraph (last sentence) mentioning the usage of SDAEs in power system dynamic simulation.
-That sentence in the revised manuscript contains two newly added references: [22] and [23], both on simulating SDAE sample paths.}}
+As per the reviewer's suggestion, in the revised manuscript, we have included two sentences (marked blue) in Sec.
+I.1, toward the end of second paragraph, mentioning the usage of SDAEs in power system dynamic simulation.
+Therein, the first sentence contains two newly added references: [22] and [23], both on simulating SDAE sample paths.}}
 

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 
-{\nib {\blue Please see }}
+{\nib {\blue As suggested in the comments, the revised manuscript has expanded the overview of prior art.
+Specifically, we have included \ul{two new references [22], [23] on stochastic differential algebraic equations (SDAEs) and cited them inline in Sec.
+I.1, second paragraph, as detailed in our response to the comment \textbf{R3.2} in this document}.
+This includes the reference suggested in comment \textbf{R1.3}.
+We have also included the reference suggested in comment \textbf{R4.4}; \ul{this is reference [28] in the revised manuscript.
+Its inline citation is in Sec.
+I.1, last paragraph, first sentence}.}}
 
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{We appreciate the comment.
+In the revised manuscript, we have made some changes toward this direction.
+This includes slightly expanded overview of the related power system literature (see the two sentences added in Sec.
+I.1, second paragraph with two new references [22], [23]), citing recent reference [28] in Sec.
+I.1, third paragraph.
+Please note that the literature review in Sec.
+I.1 is completely based on the power system literature.
+
+Having said that, we believe our technical contribution is a fundamentally interdisciplinary in the sense it is not mere translational research.
+In other words, we are not simply importing ideas from a different domain and applying them in a power system problem.
+Instead, we make novel contributions at the (possibly unanticipated) intersection of two areas: measure-valued gradient flows and power system dynamics.
+The presentation of the paper reflects the same, perhaps rightfully so.
+Notice in particular that our presentation is bottom-up: we start from power system literature review and mention that the problem addressed in our paper was already recognized as a technical challenge in that literature.
+We move onto the theoretical ideas and tools much later in the manuscript, as the build up demands.
+
+We expect a significant share of the readers of this work to be power system engineers, but not exclusively so since this contribution is also of significant interest to the researchers in control, optimization and applied mathematics.
+We do not believe this issue is exceptional for our paper since several high impact papers in the Trans.
+Power Systems were indeed at the intersection of these areas, and are not only well-read by the researchers from either side but these works have positively contributed toward more interdisciplinary research collaboration.}}
 

### 2021-12-26 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
initiated TPS rebuttal version 1

@@
 
-Having said that, we believe our technical contribution is a fundamentally interdisciplinary in the sense it is not mere translational research.
+Having said that, we believe our technical contribution is fundamentally interdisciplinary in the sense it is not mere translational research.
 In other words, we are not simply importing ideas from a different domain and applying them in a power system problem.
@@
 
-We expect a significant share of the readers of this work to be power system engineers, but not exclusively so since this contribution is also of significant interest to the researchers in control, optimization and applied mathematics.
+We expect a significant share of the readers of this work to be power system engineers, but not exclusively so since this contribution will also be of interest to the researchers in control, optimization and applied mathematics.
 We do not believe this issue is exceptional for our paper since several high impact papers in the Trans.

### 2021-12-29 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TCST rebuttal round 2

@@
 This includes the reference suggested in comment \textbf{R1.3}.
-We have also included the reference suggested in comment \textbf{R4.4}; \ul{this is reference [28] in the revised manuscript.
+\ul{We have also included the reference suggested in comment \textbf{R4.4}; this is reference [28] in the revised manuscript.
 Its inline citation is in Sec.
@@
 {\nib \blue{We appreciate the comment.
-In the revised manuscript, we have made some changes toward this direction.
+\ul{In the revised manuscript, we have made some changes toward this direction.
 This includes slightly expanded overview of the related power system literature (see the two sentences added in Sec.
@@
 Please note that the literature review in Sec.
-I.1 is completely based on the power system literature.
+I.1 is completely based on the power system literature}.
 
@@
 We do not believe this issue is exceptional for our paper since several high impact papers in the Trans.
-Power Systems were indeed at the intersection of these areas, and are not only well-read by the researchers from either side but these works have positively contributed toward more interdisciplinary research collaboration.}}
+Power Systems were indeed at the intersection of these areas, and are not only well-read by the researchers from either side but these works have positively contributed toward increasing interdisciplinary research collaboration.}}
 
@@
 {\nib \blue{We appreciate the comment.
-As per the reviewer's suggestion, in the revised manuscript, we have included two sentences (marked blue) in Sec.
-I.1, toward the end of second paragraph, mentioning the usage of SDAEs in power system dynamic simulation.
+\ul{As per the reviewer's suggestion, in the revised manuscript, we have included two sentences (marked blue) in Sec.
+I.1, toward the end of second paragraph, mentioning the usage of SDAEs in power system dynamic simulation}.
 Therein, the first sentence contains two newly added references: [22] and [23], both on simulating SDAE sample paths.}}

### 2021-12-29 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
initiated TPS rebuttal version 1
TCST rebuttal round 2

@@
 {\blue{We also note the usage of \emph{stochastic} differential algebraic equations (DAEs) for power system dynamic simulation \cite{wang2011numerical,milano2013systematic}.
-These studies, however, simulate the sample paths and does not directly propagate the joint state PDFs.}}
+These studies, however, simulate the sample paths and do not directly propagate the joint state PDFs.}}
 
@@
 
-\bibitem{wang2011numerical}
+{\blue{\bibitem{wang2011numerical}
 K.~Wang, and M.L.~Crow, ``Numerical simulation of stochastic differential algebraic equations for power system transient stability with random loads", \emph{2011 IEEE Power \& Energy Society General Meeting (PESGM)}, pp.
@@
 4, pp.
-4537--4544, 2013.
+4537--4544, 2013.}}
 

### 2021-12-30 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 
-{\nib \blue{TBD FOR FIRST PARA.
+{\nib \blue{We have followed the reviewer's suggestion to improve the presentation.
+\ul{Specifically, in the revised manuscript's Sec.
+III-A, last two paragraphs summarize the technical relation of this paper w.r.t.
+the authors' previous work.
+In particular, we have added three more sentences with a new reference and broken what was a single paragraph in the previous version into two paragraphs.}
 
@@
 \ul{On the other hand, we believe the detailed discussion regarding Wasserstein gradient flow and why that standard development cannot be applied in this case, is in fact, the heart of this paper}.
-Please note that \ul{ours is not a recipe paper} where a numerical algorithm is proposed backed by some theory.
+Please note that \ul{ours is not a recipe paper} where a numerical algorithm is proposed backed by some theory and simulations.
 \ul{The key contribution of this paper, from the authors' viewpoint, is to help distill out the core technical difficulty that hinders simply importing standard techniques from the optimal transport literature}.

### 2021-12-30 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
TPS  rebuttal round 1

@@
 A point cloud-based proximal algorithm was proposed in \cite{caluya2019ACC,caluya2019TAC} which was reported to have very fast runtime.
+{\blue{The main idea in these references was to \emph{co-evolve} the time-varying state samples (via numerical SDE solver) as well as the joint PDF values (via \eqref{ProxRecursionInfiniteDim}) evaluated at those samples.
+The resulting computation is an \emph{online} propagation of the joint PDFs as opposed to the computation in Monte Carlo or density estimation methods \cite{silverman1998density}.
+The latter methods only propagate the state samples and then approximate the joint PDFs as \emph{offline} post-processing.}}
+
 Notice that even though the drift in \eqref{ItoSDEvectorlevel} is \emph{not} a gradient vector field, the algorithm in \cite[Sec.
@@
 \bibitem{villani2003topics}
-C. Villani, \emph{Topics in optimal transportation}, No. 58, American Mathematical Society, 2003.
+C. Villani, \emph{Topics in Optimal Transportation}, No. 58, American Mathematical Society, 2003.
 
@@
 F.~Santambrogio, ``$\{$Euclidean, Metric, and Wasserstein$\}$ Gradient Flows: An Overview", \emph{Bulletin of Mathematical Sciences}, Vol. 7, No. 1, pp. 87--154, 2017.
+
+% === density estimation ====
+
+\bibitem{silverman1998density}
+{\blue{B.W.~Silverman, \emph{Density Estimation for Statistics and Data Analysis}, 1st ed., Chapman \& Hall/CRC, 1998.}}
 

### 2021-12-31 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
TPS  rebuttal round 1

--- 2021-12-29 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
@@
 A point cloud-based proximal algorithm was proposed in \cite{caluya2019ACC,caluya2019TAC} which was reported to have very fast runtime.
+{\blue{The main idea in these references was to \emph{co-evolve} the time-varying state samples (via SDE discretization schemes such as the Euler-Maruyama method) as well as the joint PDF values (via \eqref{ProxRecursionInfiniteDim}) evaluated at those samples.
+The resulting computation is an \emph{online} propagation of the joint PDFs as opposed to the \emph{offline} computation in Monte Carlo or density estimation methods \cite{silverman1998density}.
+The latter methods only propagate the state samples and then approximate the joint PDFs as post-processing.}}
+
 Notice that even though the drift in \eqref{ItoSDEvectorlevel} is \emph{not} a gradient vector field, the algorithm in \cite[Sec.

### 2021-12-31 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
TPS  rebuttal round 1

@@
 A point cloud-based proximal algorithm was proposed in \cite{caluya2019ACC,caluya2019TAC} which was reported to have very fast runtime.
-{\blue{The main idea in these references was to \emph{co-evolve} the time-varying state samples (via SDE discretization schemes such as the Euler-Maruyama method) as well as the joint PDF values (via \eqref{ProxRecursionInfiniteDim}) evaluated at those samples.
+{\blue{The main idea in these references was to \emph{co-evolve} the time-varying state samples (via SDE discretization schemes such as the Euler-Maruyama scheme) as well as the joint PDF values (via \eqref{ProxRecursionInfiniteDim}) evaluated at those samples.
 The resulting computation is an \emph{online} propagation of the joint PDFs as opposed to the \emph{offline} computation in Monte Carlo or density estimation methods \cite{silverman1998density}.

### 2022-01-01 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 
-{\nib {\blue Please see }}
+{\nib {\blue \ul{Following the comment \textbf{R4.3} to improve the presentation, in the revised manuscript, we have expanded the last two paragraphs of Sec.
+III-A with a new reference [50]}.
+Please see our response to that comment below for details.} {\red{TBD on figure clarity}}
+}
 
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{In the previous version, the strict 10 page limit enforced small figure sizes.
+In the revised manuscript, Figures 5 and 6 are made larger.
+Also, the Fig.
+6 in the revised manuscript shows the transient $(\theta_i,\omega_i)$ bivariate marginals for \ul{all the five generators in Case II}; the previous version showed the same only for the generator 4 (bus 6).}}
 

### 2022-01-01 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

--- 2021-12-30 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{Thanks for this comment.
+In the previous version, the strict 10 page limit enforced small figure sizes.
+In the revised manuscript, \ul{Figures 5 and 6 are made larger}.
+Also, the Fig.
+6 in the revised manuscript shows the transient $(\theta_i,\omega_i)$ bivariate marginals for \ul{all the five generators in Case II}; the previous version showed the same only for the generator 4 (bus 6).
+We believe these changes have improved clarity.
+
+Regarding the style of plots, we had considered plotting the bivariate marginals as planar contour plots (as in Fig.
+1 in this rebuttal document) \ul{but decided against it because of the following two reasons}.
+
+\emph{First}, \ul{in our opinion, this kind of plot is scientifically misleading as they hide the fact that these bivariate marginal PDFs are not really supported on the plane, or the subsets thereof, but instead have cylindrical supports}.
+Our message that the proposed method explicitly accounts this manifold geometry in the computation, would be missed had we chosen to make the plots as per populist cognitive bias on behalf of the readers.
+That the planar plots evoke a sense of familiarity cannot be a scientific argument in favor of it.
+
+\emph{Second}, \ul{the planar contour marginal plots require additional mental efforts from the readers to correctly parse the statistical information conveyed}.
+For instance, consider the fourth column, second row subplot in Fig.
+1 in this rebuttal document.
+It is upon the reader to parse that the two high probability regions on the left and right of that subplot are actually referring to the same angular high probability region around $\theta_{2}=0$, thanks to the periodic boundary condition on the angular variable.
+Our philosophy of technical communication is that the onus is on the authors, not on the readers, to present the scientific information as correctly as possible.
+And yes, we are aware that many papers in the power system literature make plots in the rectangular domain ignoring the manifold nature of the support, which in our experience, continue to add pain to the graduate students and beginning researchers trying to understand the statistical meaning conveyed in those plots.
+We prefer ours  not be another paper contributing to the accrual of such.}}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.99\linewidth]{IEEE14marginals.png}
+\caption{\small{Example $(\theta_i,\omega_i)$ bivariate marginals, for the IEEE 14 bus system's generators $i=1,\hdots, 5$, shown as planar contour plots.
+The columns correspond to time snapshots; the rows correspond to the five generators.
+The colorbar corresponds to the value of the bivariate marginals.
+The simulation set up is described in Sec.
+V-A of the manuscript.}}
+\vspace*{-0.1in}
+\label{fig:PlanarContourBivariateMarginals}
+\end{figure}
 

### 2022-01-01 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 Having said that, we believe our technical contribution is fundamentally interdisciplinary in the sense it is not mere translational research.
-In other words, we are not simply importing ideas from a different domain and applying them in a power system problem.
+In other words, we are not simply importing ideas from a different domain and applying them to a power system problem.
 Instead, we make novel contributions at the (possibly unanticipated) intersection of two areas: measure-valued gradient flows and power system dynamics.
@@
 And yes, we are aware that many papers in the power system literature make plots in the rectangular domain ignoring the manifold nature of the support, which in our experience, continue to add pain to the graduate students and beginning researchers trying to understand the statistical meaning conveyed in those plots.
-We prefer ours  not be another paper contributing to the accrual of such.}}
+We prefer ours  not be contributing to the accrual of such pain, and instead make plots that depict the variables as they are.}}
 

### 2022-01-01 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 III-A with a new reference [50]}.
-Please see our response to that comment below for details.} {\red{TBD on figure clarity}}
-}
+Please see our response to that comment below for details.
+
+\ul{As detailed in the response to comment \textbf{R3.4}, in the revised manuscript, we have enlarged Figures 5 and 6, and expanded the plot in Fig.
+6}.
+There we also explained why those plot styles are used.
+
+\ul{Furthermore, Fig.
+7 in the revised manuscript has also been expanded to include the univariate angular velocity transient marginal plots for all the five generators of the IEEE 14 bus system; previous version had this plot only for the generator 4 (buss 6)}.
+
+{\red{New plots}}
+}}
 
@@
 It is upon the reader to parse that the two high probability regions on the left and right of that subplot are actually referring to the same angular high probability region around $\theta_{2}=0$, thanks to the periodic boundary condition on the angular variable.
-Our philosophy of technical communication is that the onus is on the authors, not on the readers, to present the scientific information as correctly as possible.
+Our philosophy of technical communication is that the onus is on the authors, not on the readers, to clarify the scientific information as correctly as possible.
 And yes, we are aware that many papers in the power system literature make plots in the rectangular domain ignoring the manifold nature of the support, which in our experience, continue to add pain to the graduate students and beginning researchers trying to understand the statistical meaning conveyed in those plots.

### 2022-01-02 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 
-{\nib \blue{The reviewer's point is well taken. TBD.}}
+{\nib \blue{The reviewer's point is well taken in that the second order model may not be adequate to capture all the complexities of modern power systems, and }}
 
@@
 1 in this rebuttal document.
-It is upon the reader to parse that the two high probability regions on the left and right of that subplot are actually referring to the same angular high probability region around $\theta_{2}=0$, thanks to the periodic boundary condition on the angular variable.
+It is upon the reader to parse that the two high probability regions on the left and right of that subplot are actually referring to the same angular high probability region around $\theta_{2}=0$ rad, thanks to the periodic boundary condition on the angular variable.
 Our philosophy of technical communication is that the onus is on the authors, not on the readers, to clarify the scientific information as correctly as possible.
-And yes, we are aware that many papers in the power system literature make plots in the rectangular domain ignoring the manifold nature of the support, which in our experience, continue to add pain to the graduate students and beginning researchers trying to understand the statistical meaning conveyed in those plots.
+And yes, we are aware that many papers in the power system literature make plots in the rectangular domain ignoring the manifold nature of the support, which in our experience, continue to add pain to the graduate students and beginning researchers trying to decipher the statistical meaning conveyed in those plots.
 We prefer ours  not be contributing to the accrual of such pain, and instead make plots that depict the variables as they are.}}
@@
 the authors' previous work.
-In particular, we have added three more sentences with a new reference and broken what was a single paragraph in the previous version into two paragraphs.}
+In particular, we have added three more sentences with a new reference.
+To improve readability, have also broken what was a single paragraph in the previous version, into two separate paragraphs.}
 
@@
 
-\ul{On the other hand, we believe the detailed discussion regarding Wasserstein gradient flow and why that standard development cannot be applied in this case, is in fact, the heart of this paper}.
+\ul{On the other hand, the detailed discussion regarding Wasserstein gradient flow and why that standard development cannot be applied in this case, is in fact, the heart of this paper}.
 Please note that \ul{ours is not a recipe paper} where a numerical algorithm is proposed backed by some theory and simulations.

### 2022-01-02 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 
-{\nib \blue{The reviewer's point is well taken in that the second order model may not be adequate to capture all the complexities of modern power systems, and }}
+{\nib \blue{The reviewer's point is well taken in the sense the second order model may not be adequate to capture all the complexities of modern power systems, and further research is needed to design/extend nonparametric joint PDF propagation methods such as the one we propose with higher order generator dynamics.
+\ul{Please note that the last sentence of our Conclusion section explicitly stated so}.
+
+At the same time, it is important to understand how exactly this paper is making a significant stride w.r.t. the current state-of-the-art even with the .}}
 

### 2022-01-02 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 
-{\nib {\blue \ul{Following the comment \textbf{R4.3} to improve the presentation, in the revised manuscript, we have expanded the last two paragraphs of Sec.
+{\nib {\blue We have made several updates to clarify the technical issues and to improve the presentation.
+
+\ul{Following the comment \textbf{R4.3} to improve the presentation, in the revised manuscript, we have expanded the last two paragraphs of Sec.
 III-A with a new reference [50]}.
@@
 
-At the same time, it is important to understand how exactly this paper is making a significant stride w.r.t. the current state-of-the-art even with the .}}
+\ul{At the same time, it is important to understand how exactly this paper is making a significant stride w.r.t.
+the current state-of-the-art even with the second order dynamics}.
+The last paragraph of our Sec.
+II-B clearly asserts that the state-of-the-art in directly computing the joint PDF remains the ref.
+[45] published in this journal in 2013, which considered the second order dynamics for the single machine infinite bus case via finite element discretization.
+In the same paragraph, we also cited two 2018 references [16, 17] published again in this journal, which did not directly propagate the state PDFs and instead computed some proxy for the same.
+\ul{So the key  question becomes: if the power system literature has already acknowledged the importance of solving this problem, why is there no progress since [45] for even moderate $n$ and even assuming seemingly simple second order models?
+What exactly is the fundamental impediment standing in the way of nonparametric computation of the joint state PDF subject to the power system dynamics}?
+Our Sec.
+I.2 clearly lists the challenges for the same.
+The crux is that addressing this challenge is not a matter of innovating on a better discretization scheme or innovating on computing architecture.
+\ul{See specifically the last two sentences of our Sec.
+II-B}.
+\ul{This paper is breaking the stalemate by fundamentally and rigorously innovating on what it really means to solve the underlying PDE, albeit subject to the second order dynamics}.
+Of course, we expect follow up research to extend the proposed approach for higher order models but we cannot have progress with a complex model unless we first demonstrate how to tackle the simpler second order case -- a case that the existing literature cannot tackle for nonparametric online PDF computation.
+
+\ul{Last but not the least, we find the reviewer's assertion ``In other words, the proposed model is useless" vacuous for the following reasons}.
+
+(i) We \ul{do not propose} the second order model.
+We use the second order SDE that albeit simple, is widely used in power system dynamic simulation, as the sample path dynamics.
+In this paper, we are not interested in simulating sample paths.
+Our interest is in nonparametrically computing the transient joint PDFs subject to the given sample path dynamics in the form of It\^{o} SDE.
+Our usage of this particular sample path dynamics does not preempt the possibility of designing a variant of the proposed proximal recursion for performing computation of the same nature when the given It\^{o} SDE is of higher order.
+Indeed, this is what we alluded to in the last sentence of the Conclusions section.
+
+(ii) \ul{What we propose is not a model}.
+Transcribing the Fokker-Planck operator into a generalized gradient flow, and thereby deriving the corresponding proximal recursion, is a purely mathematical construct.
+There are no modeling assumptions made in that derivation.
+Our contribution in this paper is to initiate and showcase this new way of solving the transient Fokker-Planck PDE subject to a simple but popular power system dynamic model.
+Extending this framework to higher order sample path dynamics is mentioned as future work, and is beyond the scope of this particular paper.
+
+}}
 

### 2022-01-03 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 II-B}.
-\ul{This paper is breaking the stalemate by fundamentally and rigorously innovating on what it really means to solve the underlying PDE, albeit subject to the second order dynamics}.
-Of course, we expect follow up research to extend the proposed approach for higher order models but we cannot have progress with a complex model unless we first demonstrate how to tackle the simpler second order case -- a case that the existing literature cannot tackle for nonparametric online PDF computation.
+\ul{The present paper is breaking the stalemate by fundamentally and rigorously innovating on what it really means to solve the underlying PDE, albeit subject to the second order dynamics}.
+As we have stated up front, we expect follow up research to extend the proposed approach for higher order models but we cannot have progress with a complex model unless we first demonstrate how to tackle the simpler second order case -- a case that the existing literature cannot tackle for nonparametric online PDF computation.
 
@@
 Our interest is in nonparametrically computing the transient joint PDFs subject to the given sample path dynamics in the form of It\^{o} SDE.
-Our usage of this particular sample path dynamics does not preempt the possibility of designing a variant of the proposed proximal recursion for performing computation of the same nature when the given It\^{o} SDE is of higher order.
-Indeed, this is what we alluded to in the last sentence of the Conclusions section.
+\ul{Our usage of this particular sample path dynamics does not preempt the possibility of designing a variant of the proposed proximal recursion for performing computation of the same nature when the given It\^{o} SDE is of higher order.
+Indeed, this is what we alluded to in the last sentence of the Conclusions section}.
 
@@
 There are no modeling assumptions made in that derivation.
-Our contribution in this paper is to initiate and showcase this new way of solving the transient Fokker-Planck PDE subject to a simple but popular power system dynamic model.
+\ul{Our contribution in this paper is to initiate and showcase this new way of solving the transient Fokker-Planck PDE subject to a simple but popular power system dynamic model}.
 Extending this framework to higher order sample path dynamics is mentioned as future work, and is beyond the scope of this particular paper.

### 2022-01-07 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 
-{\nib {\blue Thanks for this comment.
+{\nib {\blue We are indebted to the AE and the reviewers for their help to improve the manuscript.
 
@@
 (ii)
+
+{\red{Regarding the usage of simple model, please see our response to comment \textbf{R3.1} and ??
+In summary, the current state-of-the-art in power system literature for this problem is ref.
+[45] with second order dynamics + single machine infinite bus ($n=1$ in our manuscript's notation).
+Given this context, our manuscript enables a new kind of computation that can handle moderately large $n$ with second order sample path dynamics.
+This computation is online and approximates neither the statistics nor the dynamics.
+\ul{We clarify two important points here}:\\(i) The proposed proximal recursion framework on the space of probability measures is not conceptually limited by the second order dynamics, and we expect follow up research to design custom proximal recursions for higher order synchronous machine models.
+In this sense, the present paper should be seen as the stepping stone laying the foundation for handling more complex dynamical models.
+The authors have been completely up front by explicitly mentioning this in the Conclusions section.
+We have also reiterated the same in this rebuttal.\\(ii) Deriving proximal recursion requires understanding the generalized gradient flow structure over probability measures.
+Our research disposition is that one cannot bypass understanding the generalized gradient flow structure for the simpler second order sample path dynamics before attempting the same for more complex sample path models.
+In fact, a significant contribution of our article, in the authors' opinion, is that it distills out what exactly is technically challenging and what are the peripheral issues.}}
 }}
@@
 
-{\nib {\blue This is. 
+{\nib {\blue The original contribution of this article is to design an algorithm for \emph{online nonparametric} computation of the joint state PDFs subject to stochastic uncertainties (initial conditions, parameters, process noise and combinations thereof) in power system dynamics.
+
+The qualifier ``online" means that it is not a variant of Monte Carlo where samples are first propagated \ul{and then} a function approximation is performed offline to estimate the joint state PDF at a snapshot.
+The qualifier ``nonparametric" means that neither the statistics nor the dynamics are approximated \emph{a priori}.
+This is novel because to the best of the author's knowledge, there does not exist a power systems paper which can accomplish the same.
+The state-of-the-art in power system literature is discussed in Sec I.1.
+More importantly, Sec.
+I.2 details why the state-of-the-art is what it is.
+\ul{We specifically explain the contribution of this article in Sec.
+I.3 titled ``Contributions of this paper".}
 }}
@@
 
-{\nib{ \blue We have.}}
+{\nib{ \blue We thank the reviewer for this comment.
+This paper is \emph{not} about novel implementation of an existing idea; the idea itself is new and consequently, needs to be contextualized and developed in the paper.
+\ul{The discussion toward the end of Sec.
+II and that in Sec III are, in our opinion, more fundamental contribution than the implementation in Sec.
+IV.
+Those parts of the paper really mark the innovation since they explain why someone, even if conversant in infinite dimensional gradient flow literature, cannot directly apply existing techniques to this problem.
+Then Sec.
+III-C and III-D explain how to circumvent the specific difficulties arising in this problem.
+A reader jumping to Sec.
+IV assuming that everything before that is known, is mistaken}.
+Notice that we clearly laid out the navigational chart of this paper in the paragraph right before Sec.
+II starts.}}
 
@@
 V, we have used MATPOWER and the Python-library ANDES (with inline citation of refs.
-[61] and [62] therein) for obtaining some simulation parameters, as explained in the first paragraph of V-A.}}
+[62] and [63] therein) for obtaining some simulation parameters, as explained in the first paragraph of V-A.}}
 
@@
 
-\ul{Last but not the least, we find the reviewer's assertion ``In other words, the proposed model is useless" vacuous for the following reasons}.
+\ul{Last but not the least, we find the reviewer's assertion ``In other words, the proposed model is useless" moot for the following reasons}.
 

### 2022-01-07 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 {\nib{ \blue We thank the reviewer for this comment.
-This paper is \emph{not} about novel implementation of an existing idea; the idea itself is new and consequently, needs to be contextualized and developed in the paper.
+This paper is \emph{not} about novel implementation of an existing idea; the idea itself is new and consequently needs to be contextualized and developed in the paper.
 \ul{The discussion toward the end of Sec.

### 2022-01-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 {\nib{ \blue We thank the reviewer for this comment.
-This paper is \emph{not} about novel implementation of an existing idea; the idea itself is new and consequently needs to be contextualized and developed in the paper.
+This paper is \emph{not} about novel implementation of an existing idea.
+The idea itself is new and consequently needs to be contextualized and developed in the paper.
 \ul{The discussion toward the end of Sec.
@@
 IV assuming that everything before that is known, is mistaken}.
-Notice that we clearly laid out the navigational chart of this paper in the paragraph right before Sec.
+Notice that we clearly laid out the structure of this paper (which section does what) in the paragraph right before Sec.
 II starts.}}
@@
 
-{\nib{ \blue TBD.}}
+{\nib{ \blue Our responses to the three numbered sub-comments below.
+
+(1) The second order sample path dynamics model we used is standard in power systems literature, and specifically as we pointed}}
 
@@
 
-{\nib \blue{We appreciate the comment.
-\ul{As per the reviewer's suggestion, in the revised manuscript, we have included two sentences (marked blue) in Sec.
+{\nib \blue{\ul{As per the reviewer's suggestion, in the revised manuscript, we have included two sentences (marked blue) in Sec.
 I.1, toward the end of second paragraph, mentioning the usage of SDAEs in power system dynamic simulation}.

### 2022-01-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 
-{\red{Regarding the usage of simple model, please see our response to comment \textbf{R3.1} and ??
+{\red{Regarding the usage of simple model, please see our response to comment \textbf{R3.1} and \textbf{1.3, sub-comment (1)}.
 In summary, the current state-of-the-art in power system literature for this problem is ref.
@@
 
-{\nib{ \blue Our responses to the three numbered sub-comments below.
+{\nib{ \blue Our responses to the \ul{three numbered sub-comments} below.
 
-(1) The second order sample path dynamics model we used is standard in power systems literature, and specifically as we pointed}}
+(1) As pointed out in the last paragraph of Sec.
+II-B, the current state-of-the-art in power system literature for the problem we are addressing is ref.
+[45] with second order dynamics + single machine infinite bus ($n=1$ in our manuscript's notation).
+Given this context, our manuscript enables a new kind of computation that can handle moderately large $n$ with second order sample path dynamics.
+This computation is online and approximates neither the statistics nor the dynamics.
+\ul{We clarify two important points here}:\\ \textbf{[1a]} The proposed proximal recursion framework on the space of probability measures is not conceptually limited by the second order dynamics, and we expect follow up research to design custom proximal recursions for \ul{higher order synchronous machine models}.
+In this sense, the present paper should be seen as the stepping stone laying the foundation for handling more complex dynamical models.
+The authors have been completely up front by explicitly mentioning this in the Conclusions section.
+\\ \textbf{[1b]} Deriving proximal recursion requires understanding the generalized gradient flow structure over probability measures.
+\ul{Our research disposition is that one cannot bypass understanding the generalized gradient flow structure for the simpler second order sample path dynamics before attempting the same for more complex sample path models}.
+In fact, a significant contribution of our article, in the authors' opinion, is that it distills out what exactly is technically challenging and what are the peripheral issues.
+
+(2) The }}
 

### 2022-01-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 
-{\nib {\blue As suggested in the comments, the revised manuscript has expanded the overview of prior art.
+{\nib {\blue Following the suggestions, the revised manuscript has expanded the overview of prior art.
 Specifically, we have included \ul{two new references [22], [23] on stochastic differential algebraic equations (SDAEs) and cited them inline in Sec.
@@
 
-{\nib{ \blue Our responses to the \ul{three numbered sub-comments} below.
+{\nib{ \blue \ul{Our responses to the sub-comments below}.
 
-(1) As pointed out in the last paragraph of Sec.
+(1) \textbf{Regarding usage of classical model:} Whether a model is relevant depends on the problem context.
+As pointed out in the last paragraph of Sec.
 II-B, the current state-of-the-art in power system literature for the problem we are addressing is ref.
@@
 In this sense, the present paper should be seen as the stepping stone laying the foundation for handling more complex dynamical models.
-The authors have been completely up front by explicitly mentioning this in the Conclusions section.
+\ul{The authors have been completely up front by explicitly mentioning this in the Conclusions section}.
 \\ \textbf{[1b]} Deriving proximal recursion requires understanding the generalized gradient flow structure over probability measures.
@@
 
-(2) The }}
+(2) \textbf{Regarding randomness in active and reactive power:} {\red{TBD}}
+
+(3) \textbf{Regarding the diffusion term:} The reason why the diffusion term is present in the RHS of (1b) but not in the RHS of (1a) is because (1) is the state space representation of the second order Langevin dynamics (6).
+For the Langevin dynamics (6), its corresponding It\^{o} SDE \ul{has to be} (1).
+This is simply re-writing of a second order differential equation as an equivalent system of first order differential equations.
+There is no extra assumption here.
+\ul{Indeed, it is mathematically incorrect to add a noise term in the RHS of (1a) while claiming consistency with (6)}.
+
+\ul{Notice that the absence of additive noise in the RHS of (1a) does \emph{not} imply that the rotor angles are deterministic trajectories}.
+The reason why the rotor angles are stochastic processes is because the process noise in (1b) affects the rotor angle sample paths via dynamical nonlinear coupling with the angular velocities, i.e., these are second order stochastic effects.
+\ul{In short, the noise in rotor angles is \emph{not} ignored; in fact the rotor angles are nonlinearly colored (i.e., non-white) noisy sample paths}.
+
+This reviewer may also have \ul{mathematical misconception} in thinking that a model that has additive noise in the RHS of \emph{all} SDE components is somehow more general than the case where the additive noise terms are absent from some but not all components of an SDE RHS.
+\ul{In reality, the complexity is in the other way around}.
+The SDEs of the form (1) are the so-called ``degenerate It\^{o} SDEs" which are much more difficult to analyze than the non-degenerate versions (noise in the RHS of all SDE components).
+For instance, this degeneracy makes the joint PDF evolution equation (10) a \emph{kinetic} Fokker-Planck PDE, which is different in structure than the standard Fokker-Planck a.k.a.
+Kolmogorov's forward PDE.
+\ul{The gradient flow structure and the associated measure-valued proximal recursions for the kinetic Fokker-Planck PDE are less understood than the same for the standard Fokker-Planck PDE.
+This is exactly why the methodological contribution of this paper is important}.
+
+\textbf{Regarding the reference [R1]:} \ul{In the revised manuscript, we have included the ref.
+[R1] suggested by the reviewer with inline citation in Sec.
+I.1, toward the end of second paragraph.
+Please note that the topic of [R1] is different from the topic of our paper: how to perform online nonparametric computation of transient joint state PDFs}.
+
+\textbf{Regarding OU process in load:} \ul{It appears that the reviewer have missed what is actually going on in our paper}.
+The Kron reduced second order stochastic model implies that in the unreduced sense, the load dynamics was first order noisy Kuramoto, in addition to the synchronous machine dynamics being second order noisy Kuramoto.
+Therefore, we do have stochastic load dynamics, which is in fact more realistic than OU process model.
+While OU process assumes linear drift with constant diffusion, our setting considers the load SDEs having drifts with trigonometric nonlinearity and constant diffusion.
+
+}}
 

### 2022-01-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 A reader jumping to Sec.
-IV assuming that everything before that is known, is mistaken}.
+IV assuming that everything before that is known, is completely mistaken}.
 Notice that we clearly laid out the structure of this paper (which section does what) in the paragraph right before Sec.
@@
 I.1, toward the end of second paragraph.
-Please note that the topic of [R1] is different from the topic of our paper: how to perform online nonparametric computation of transient joint state PDFs}.
+Please note that the topic of [R1] is different from the topic of our paper.
+We are interested in how to perform online nonparametric computation of transient joint state PDFs}.
+This is not Monte Carlo.
 
-\textbf{Regarding OU process in load:} \ul{It appears that the reviewer have missed what is actually going on in our paper}.
+\textbf{Regarding OU process in load:} \ul{It appears that the reviewer might have missed what is actually going on in our paper}.
 The Kron reduced second order stochastic model implies that in the unreduced sense, the load dynamics was first order noisy Kuramoto, in addition to the synchronous machine dynamics being second order noisy Kuramoto.

### 2022-01-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 We are interested in how to perform online nonparametric computation of transient joint state PDFs}.
-This is not Monte Carlo.
+We are not doing Monte Carlo.
 
@@
 
-{\nib{ \blue While the depiction of the computational times in Figs. 8 and 9 in the revised manuscript are somewhat unconventional, we explain why we choose to plot that way.
+{\nib{ \blue While the depiction of the computational times in Figs.
+8 and 9 in the revised manuscript are somewhat unconventional, we explain why we choose to plot that way.
+\ul{To aid clarity, in the revised manuscript, we have added an excerpt of the following at the end of Sec.
+V-A.}
 
@@
 And yes, we are aware that many papers in the power system literature make plots in the rectangular domain ignoring the manifold nature of the support, which in our experience, continue to add pain to the graduate students and beginning researchers trying to decipher the statistical meaning conveyed in those plots.
-We prefer ours  not be contributing to the accrual of such pain, and instead make plots that depict the variables as they are.}}
+We thus prefer to make plots that depict the variables as they are.}}
 

### 2022-01-09 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
TPS  rebuttal round 1

@@
 \ref{sec:ProxAlgo}, for both \textbf{Cases I} and \textbf{II}.
+{\blue{While the depiction of the computational times in Fig.
+\ref{fig:CompTimeIEEE14bus} appears unconventional, we next explain why that is meaningful in our framework.
+
+At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1.
+This proximal update has no analytical solution and instead requires execution of a contractive fixed point recursion.
+The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update.
+To claim that the proposed method is practical, we therefore, need to demonstrate that the computational timescale is faster than the physical timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time.
+On the other hand, theoretical consistency (see the last paragraph of Sec.
+\ref{subsec:ProxUpdate}) requires that $h$ itself should be small.
+Thus, even if one has a provably convergent algorithm that can perform the one step proximal update, it is not clear \emph{a priori} if the computational time needed to do so can be much faster than $h$.
+If not, then the method would be statistically consistent but practically not useful.
+The Fig \ref{fig:CompTimeIEEE14bus} shows that the computational time needed to perform the proximal updates, for all $t_{k}$, are indeed orders-of-magnitude faster than $h$.
+}}
 

### 2022-01-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 
-{\nib{ \blue While the depiction of the computational times in Figs.
+{\nib{ \blue While the depiction of the per-step computational times in Figs.
 8 and 9 in the revised manuscript are somewhat unconventional, we explain why we choose to plot that way.
@@
 At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1.
-This proximal update involves execution of a contractive fixed point recursion.
+This proximal update has no analytical solution and instead requires execution of a contractive fixed point recursion.
 The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update.

### 2022-01-10 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 \ul{The gradient flow structure and the associated measure-valued proximal recursions for the kinetic Fokker-Planck PDE are less understood than the same for the standard Fokker-Planck PDE.
-This is exactly why the methodological contribution of this paper is important}.
+This is exactly why the methodological contribution of this paper is important, even if one completely ignores the algorithmic contribution}.
 
@@
 
-{\nib{ \blue TBD.}}
+{\nib{ \blue \ul{Our responses to the sub-comments below}.
+
+(1) \textbf{Regarding simple model and ignoring the randomness in loads}: Please refer to our response to \textbf{R1.3, sub-comment (1)} where we explained why this model is relevant for this particular problem, and why this is not limiting to our methodological contribution.
+In response to \textbf{R1.3, last unnumbered sub-comment in bold text: ``Regarding OU process in load"}, we explained that we are not ignoring randomness from loads.
+
+(2) \textbf{Regarding nonparametric}: \ul{We are \emph{not} making any parametric assumptions}.
+Our usage of the word ``nonparametric" is in the standard sense used in mathematical statistics literature.
+That the additive noise process is Wiener does not invite any loss of generality since if it was colored, we could always re-cast the same in our SDE form via a linear map.
+\ul{Furthermore, we have zero assumptions on the initial joint state PDF $\rho_{0}$ and on the joint PDF of the parameters appearing in the model.
+As a matter of fact, we don't even need to know the functional form of these joint PDFs.
+For instance, our framework remains unchanged if we are given the initial joint state and parameter distributional data as numerical table containing samples from these joints \emph{as well as} the joint PDF values evaluated at those samples}.
+The latter might be coming from some measurement-based sensor fusion/Bayesian algorithm.
+The specific product von Mises and uniform PDFs in Sec.
+V-A and V-B are only used for generating data to perform our numerical example simulations.
+\ul{This again underscores how a reader jumping straight to Sec.
+V without contemplative reading and understanding of the earlier Sections will make prematurely incorrect claims}.
+
+(3) \textbf{Regarding depict nonparametric PDF}: \ul{This paper is not about nonparametric \emph{depiction} but about nonparametric \emph{computation} of the transient joint state PDFs}.
+For example, Monte Carlo followed by function approximation via Kernel density estimates and its variants, or other function approximants (e.g., RBF, Gaussian mixture, neural networks) are all parametric methods.
+We enable computing the value of joint state PDF at a time-varying sample location \emph{online (not as offline post-processing)} without assuming any functional representation of the transient joint which the computed values correspond to.
+Notice that we are doing this joint computation without gridding the state space, which would be impossible had we followed the traditional scientific computing route: finite element/volume methods, tensor decomposition techniques etc.
+\ul{We believe the proposed computational framework is novel, highly non-trivial and is specifically useful in this power system uncertainty propagation context}.
+}}
 

### 2022-01-10 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
TPS  rebuttal round 1

@@
 (2) \textbf{Regarding nonparametric}: \ul{We are \emph{not} making any parametric assumptions}.
-Our usage of the word ``nonparametric" is in the standard sense used in mathematical statistics literature.
+Our usage of the word ``nonparametric" is standard in mathematical statistics literature.
 That the additive noise process is Wiener does not invite any loss of generality since if it was colored, we could always re-cast the same in our SDE form via a linear map.
@@
 The specific product von Mises and uniform PDFs in Sec.
-V-A and V-B are only used for generating data to perform our numerical example simulations.
+V-A and V-B are only used for generating example data to perform our illustrative numerical simulations.
 \ul{This again underscores how a reader jumping straight to Sec.
-V without contemplative reading and understanding of the earlier Sections will make prematurely incorrect claims}.
+V without contemplative reading and understanding of the earlier Sections may make prematurely incorrect claims}.
 

### 2022-02-19 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
polishing rebuttal round 1

@@
 The qualifier ``nonparametric" means that neither the statistics nor the dynamics are approximated \emph{a priori}.
-This is novel because to the best of the author's knowledge, there does not exist a power systems paper which can accomplish the same.
+This is novel because the existing power systems literature makes at least one of the aforesaid approximations.
 The state-of-the-art in power system literature is discussed in Sec I.1.
@@
 
-This reviewer may also have \ul{mathematical misconception} in thinking that a model that has additive noise in the RHS of \emph{all} SDE components is somehow more general than the case where the additive noise terms are absent from some but not all components of an SDE RHS.
+This reviewer may also have misunderstood in thinking that a model that has additive noise in the RHS of \emph{all} SDE components is somehow more general than the case where the additive noise terms are absent from some but not all components of an SDE RHS.
 \ul{In reality, the complexity is in the other way around}.
@@
 We are interested in how to perform online nonparametric computation of transient joint state PDFs}.
-We are not doing Monte Carlo.
+Our computation is different from Monte Carlo.
 
-\textbf{Regarding OU process in load:} \ul{It appears that the reviewer might have missed what is actually going on in our paper}.
+\textbf{Regarding OU process in load:} \ul{The reviewer might have missed the following key point of our paper}.
 The Kron reduced second order stochastic model implies that in the unreduced sense, the load dynamics was first order noisy Kuramoto, in addition to the synchronous machine dynamics being second order noisy Kuramoto.
 Therefore, we do have stochastic load dynamics, which is in fact more realistic than OU process model.
-While OU process assumes linear drift with constant diffusion, our setting considers the load SDEs having drifts with trigonometric nonlinearity and constant diffusion.
+While OU process assumes linear drift with constant diffusion, whereas our setting considers the load SDEs having drifts with trigonometric nonlinearity and constant diffusion.
 
@@
 
-{\nib{ \blue TBD.}}
+{\nib{ \blue {\red{TBD}}.}}
 
@@
 (2) \textbf{Regarding nonparametric}: \ul{We are \emph{not} making any parametric assumptions}.
-Our usage of the word ``nonparametric" is standard in mathematical statistics literature.
+Our usage of the word ``nonparametric" is standard as in mathematical statistics literature.
 That the additive noise process is Wiener does not invite any loss of generality since if it was colored, we could always re-cast the same in our SDE form via a linear map.
@@
 V-A and V-B are only used for generating example data to perform our illustrative numerical simulations.
-\ul{This again underscores how a reader jumping straight to Sec.
-V without contemplative reading and understanding of the earlier Sections may make prematurely incorrect claims}.
+\ul{This again underscores how jumping straight to Sec.
+V without contemplative reading and understanding of the earlier Sections may lead to misleading conclusions}.
 
@@
 
-{\nib{ \blue TBD.}}
+{\nib{ \blue Please note that the convergence guarantee given by equation (14) in the revised manuscript is pointwise in time, i.e., for each $k\in\mathbb{N}$.
+Unlike traditional numerical methods, there is no concept of {\red{TBD.}}}}
 

### 2022-02-19 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
polishing rebuttal round 1

@@
 
-{\nib{ \blue Please note that the convergence guarantee given by equation (14) in the revised manuscript is pointwise in time, i.e., for each $k\in\mathbb{N}$.
-Unlike traditional numerical methods, there is no concept of {\red{TBD.}}}}
+{\nib{ \blue {\red{TBD.}}}}
 
@@
 
-Having said that, we believe our technical contribution is fundamentally interdisciplinary in the sense it is not mere translational research.
+Having said that, we believe our technical contribution is indeed interdisciplinary in the sense it is not mere translational research.
 In other words, we are not simply importing ideas from a different domain and applying them to a power system problem.
@@
 We expect a significant share of the readers of this work to be power system engineers, but not exclusively so since this contribution will also be of interest to the researchers in control, optimization and applied mathematics.
-We do not believe this issue is exceptional for our paper since several high impact papers in the Trans.
+We do not believe this issue is exceptional for our paper since several high impact papers in the IEEE Trans.
 Power Systems were indeed at the intersection of these areas, and are not only well-read by the researchers from either side but these works have positively contributed toward increasing interdisciplinary research collaboration.}}
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{ {\red{TBD.}}}}
 
@@
 
-{\nib \blue{TBD.}}
+{\nib \blue{Please note that the synthetic system in Sec.
+V-B is actually simulating a higher dimensional system than the IEEE 118-bus system where the 19 generators would have resulted in the transient joint PDFs with 38 dimensional supports.
+In comparison, the synthetic system is evolving joint state PDFs having 100 dimensional supports, i.e., in $\mathbb{T}^{50}\times\mathbb{R}^{50}$.
+Our only intention behind simulating the synthetic system was to show that the computation can handle large dimensional joint state PDF supports.
+We also point out that the parameter ranges reported in Table I are consistent with the existing literature: refs.
+[25, Sec.
+5], [20], [56], [57].
+We hope the reviewer agrees that this simulation is adequate for the intended purpose.}}
 
@@
 II-B}.
-\ul{The present paper is breaking the stalemate by fundamentally and rigorously innovating on what it really means to solve the underlying PDE, albeit subject to the second order dynamics}.
+\ul{The present paper is breaking the stalemate by mathematically innovating on what it really means to solve the underlying PDE, albeit subject to the second order dynamics}.
 As we have stated up front, we expect follow up research to extend the proposed approach for higher order models but we cannot have progress with a complex model unless we first demonstrate how to tackle the simpler second order case -- a case that the existing literature cannot tackle for nonparametric online PDF computation.
@@
 We use the second order SDE that albeit simple, is widely used in power system dynamic simulation, as the sample path dynamics.
-In this paper, we are not interested in simulating sample paths.
+In this paper, we are not interested in simulating sample paths via variants of Monte Carlo.
 Our interest is in nonparametrically computing the transient joint PDFs subject to the given sample path dynamics in the form of It\^{o} SDE.
@@
 (ii) \ul{What we propose is not a model}.
-Transcribing the Fokker-Planck operator into a generalized gradient flow, and thereby deriving the corresponding proximal recursion, is a purely mathematical construct.
+Transcribing the kinetic Fokker-Planck operator into a generalized gradient flow, and thereby deriving the corresponding proximal recursion, is a purely mathematical construct.
 There are no modeling assumptions made in that derivation.
-\ul{Our contribution in this paper is to initiate and showcase this new way of solving the transient Fokker-Planck PDE subject to a simple but popular power system dynamic model}.
+\ul{Our contribution in this paper is to initiate and showcase this new way of computing the transient solutions of the kinetic Fokker-Planck PDE arising from a simple but popular power system dynamic model}.
 Extending this framework to higher order sample path dynamics is mentioned as future work, and is beyond the scope of this particular paper.
@@
 \ul{On the other hand, the detailed discussion regarding Wasserstein gradient flow and why that standard development cannot be applied in this case, is in fact, the heart of this paper}.
-Please note that \ul{ours is not a recipe paper} where a numerical algorithm is proposed backed by some theory and simulations.
-\ul{The key contribution of this paper, from the authors' viewpoint, is to help distill out the core technical difficulty that hinders simply importing standard techniques from the optimal transport literature}.
+Please note that this is far from a recipe paper where a numerical algorithm is proposed backed by some theory and simulations.
+\ul{The key contribution of this paper, instead, is to help distill out the core technical difficulty that hinders simply importing standard techniques from the optimal transport literature}.
 We believe this clarification of what really is the technical crux of the matter--versus what are peripheral issues--should be illuminating to not just power systems researchers, but also to applied mathematicians underscoring how the standard theory and algorithms for measure-valued gradient flows need to be adapted in engineering applications.}}

### 2022-02-19 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
polishing rebuttal round 1

@@
 Our research disposition is that one cannot bypass understanding the generalized gradient flow structure for the simpler second order sample path dynamics before attempting the same for more complex sample path models.
-In fact, a significant contribution of our article, in the authors' opinion, is that it distills out what exactly is technically challenging and what are the peripheral issues.}}
+In fact, a significant contribution of our article, in the authors' opinion, is that it distills out what exactly is mathematically challenging.}}
 }}
@@
 A reader jumping to Sec.
-IV assuming that everything before that is known, is completely mistaken}.
+IV assuming that everything before that is known, would be completely mistaken}.
 Notice that we clearly laid out the structure of this paper (which section does what) in the paragraph right before Sec.
@@
 V-A and V-B are only used for generating example data to perform our illustrative numerical simulations.
-\ul{This again underscores how jumping straight to Sec.
-V without contemplative reading and understanding of the earlier Sections may lead to misleading conclusions}.
+\ul{This again underscores how looking into numerical simulations without contemplative reading and understanding of the earlier Sections may lead to misleading conclusions}.
 
@@
 
-{\nib \blue{TBD. We corrected the two typos pointed out in comment \textbf{R4.8} in this rebuttal document.}}
+{\nib \blue{In Sec. V-A, we corrected ``assuming" $\rightarrow$ ``are assumed". We corrected the two typos pointed out in comment \textbf{R4.8} in this rebuttal document.}}
 
@@
 \emph{First}, \ul{in our opinion, this kind of plot is scientifically misleading as they hide the fact that these bivariate marginal PDFs are not really supported on the plane, or the subsets thereof, but instead have cylindrical supports}.
-Our message that the proposed method explicitly accounts this manifold geometry in the computation, would be missed had we chosen to make the plots as per populist cognitive bias on behalf of the readers.
+Our message that the proposed method explicitly accounts this manifold geometry in the computation, would be missed had we chosen to make the plots the other way.%as per populist cognitive bias on behalf of the readers.
 That the planar plots evoke a sense of familiarity cannot be a scientific argument in favor of it.
@@
 Our philosophy of technical communication is that the onus is on the authors, not on the readers, to clarify the scientific information as correctly as possible.
-And yes, we are aware that many papers in the power system literature make plots in the rectangular domain ignoring the manifold nature of the support, which in our experience, continue to add pain to the graduate students and beginning researchers trying to decipher the statistical meaning conveyed in those plots.
+%We are aware that many papers in the power system literature make plots in the rectangular domain ignoring the manifold nature of the support, which in our experience, continue to add pain to the graduate students and beginning researchers trying to decipher the statistical meaning conveyed in those plots.
 We thus prefer to make plots that depict the variables as they are.}}

### 2022-02-19 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
polishing rebuttal round 1

@@
 Lifting the dynamics to the space of probability measures allows us to design a scalable algorithm that obviates gridding the underlying high dimensional state space which is computationally prohibitive.
-The proximal recursion implements a generalized infinite dimensional gradient flow, and evolves probability weighted scattered point clouds.
-We clarify the theoretical nuances and algorithmic details specific for the power system nonlinearities, and provide illustrative numerical examples.
+The proximal recursion implements a generalized infinite dimensional gradient flow, and evolves probability-weighted scattered point clouds.
+We clarify the theoretical nuances and algorithmic details specific to the power system nonlinearities, and provide illustrative numerical examples.
 \end{abstract}
@@
 Several studies \cite{timko1983monte,nwankpa1992stochastic,odun2012structure,ghanavati2016identifying,apostolopoulou2016assessment} have reported that even small stochastic effects can significantly alter the assessment of transient stability, or the performance of automatic generation control.
-However, the lack of a rigorous yet scalable stochastic computational framework continues to impede \cite{schwalbe2015mathematical} our ability to perform transient analysis involving time varying joint probability density functions (PDFs) over the states of a large power system network.
+However, the lack of a rigorous yet scalable stochastic computational framework continues to impede \cite{schwalbe2015mathematical} our ability to perform transient analysis involving time-varying joint probability density functions (PDFs) over the states of a large power system network.
 In this paper, we present a new algorithm to address this computational need.
@@
 As an alternative, \emph{probabilistic small signal analysis} \cite{nwankpa1992stochastic,rueda2009assessment,huang2013quasi,dhople2013analysis} have appeared in the power systems literature, albeit at the expense of the additional assumption that the random perturbations remain ``small".
-\emph{Polynomial chaos and related stochastic collocation methods} \cite{hockenberry2004evaluation,xu2019propagating} can do away with the ``small stochastic perturbation" assumption but due to the finite-dimensional approximation of the probability space, computational performance degrades if the long-term statistics are desired.
+\emph{Polynomial chaos and related stochastic collocation methods} \cite{hockenberry2004evaluation,xu2019propagating} can do away with the ``small stochastic perturbation" assumption but due to the finite dimensional approximation of the probability space, computational performance degrades if the long-term statistics are desired.
 Furthermore, to cope with the stochasticity, these techniques require simulating a higher-dimensional system than the dimension of the physical state space, which further limits the scalability for nonlinear simulation.
@@
 For example, the method in \cite{althoff2014formal} requires converting the nonlinear DAEs to linear DAEs in such a way that guarantees set-valued over-approximation of the reachable sets.
-Likewise, the convex optimization-based bounded uncertainty propagation methods in \cite{choi2017propagating,choi2018propagating} require second order approximation of the power flow state variables as function of the uncertainties.
+Likewise, the convex optimization-based bounded uncertainty propagation methods in \cite{choi2017propagating,choi2018propagating} require second order approximation of the power flow state variables as a function of the uncertainties.
 
@@
 Our technical approach juxtaposes with the existing power systems literature discussed above, in that we approximate neither the dynamical nonlinearity nor the statistics.
-Instead of treating the exact nonlinearity as bane, we exploit the geometry induced by the power system's dynamical nonlinearity over the manifold of time-varying joint state PDFs, thereby enabling a new proximal algorithm to compute the transient joint PDFs.
+Instead of treating the exact nonlinearity as a bane, we exploit the geometry induced by the power system's dynamical nonlinearity over the manifold of time-varying joint state PDFs, thereby enabling a new proximal algorithm to compute the transient joint PDFs.
 
@@
 %We use boldfaced small letters for vectors, and boldfaced capital letters for matrices. 
-The set of natural numbers, real numbers and complex numbers are denoted as $\mathbb{N}$, $\mathbb{R}$, and $\mathbb{C}$, respectively.
+The set of natural numbers, real numbers, and complex numbers are denoted as $\mathbb{N}$, $\mathbb{R}$, and $\mathbb{C}$, respectively.
 The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$.
@@
 \includegraphics[width=.7\linewidth]{BlockDiagm.pdf}
-\caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability weighted scattered point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
+\caption{\small{Schematic of the proposed proximal algorithm for propagating the joint state PDF as probability-weighted scattered point cloud $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
 The states $\{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}$ are updated by the Euler-Maruyama scheme applied to (\ref{ItoSDEvectorlevel}); the corresponding probability weights $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are updated via discrete version of the proximal recursion (\ref{KineticProx}) as detailed in Sec.
@@
 As shown in Fig.
-\ref{fig:BlockDiagm}, this algorithm, at a conceptual level, takes the pre and post update state samples
+\ref{fig:BlockDiagm}, this algorithm, at a conceptual level, takes the pre and post-update state samples
 \[\{\tilde{\bm{x}}_{k-1}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k-1}^{i},\bm{\eta}_{k-1}^{i}\right)\}_{i=1}^{N}, \; \{\tilde{\bm{x}}_{k}^{i}\}_{i=1}^{N}\equiv\{\left(\bm{\xi}_{k}^{i},\bm{\eta}_{k}^{i}\right)\}_{i=1}^{N},\]
 and $\{\tilde{\varrho}_{k-1}^{i}\}_{i=1}^{N}$ as inputs, and outputs the proximal updates $\{\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$.
-For each $k\in\mathbb{N}$, the updated probability weighted point clouds $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are then brought back to the original state space via $\bm{\Psi}^{-1}$ as $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$, as explained earlier.
-
-The Algorithm \ref{algo:KineticProx} also needs input parameters $h,\varepsilon,N,\delta,L$.
+For each $k\in\mathbb{N}$, the updated probability-weighted point clouds $\{\tilde{\bm{x}}_{k}^{i},\tilde{\varrho}_{k}^{i}\}_{i=1}^{N}$ are then brought back to the original state space via $\bm{\Psi}^{-1}$ as $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$, as explained earlier.
+
+Algorithm \ref{algo:KineticProx} also needs input parameters $h,\varepsilon,N,\delta,L$.
 Specifically, $h$ is the time-step size, $\varepsilon$ is an entropic regularization weight internal to the \textproc{Prox} algorithm, and $N$ is the number of samples.
@@
 
-\noindent\textbf{Case I.} \textbf{Nominal case} where the $P_{i}^{\text{mech}}$ and $P_{i}^{\text{load}}$ are from the steady state power flow solutions without the process noise, and assuming to remain constant over the simulation time horizon.
+\noindent\textbf{Case I.} \textbf{Nominal case} where the $P_{i}^{\text{mech}}$ and $P_{i}^{\text{load}}$ are from the steady state power flow solutions without the process noise, and {\blue{are assumed}} to remain constant over the simulation time horizon.
 The transient stochastic simulations (i.e., with process noise) are then performed using the proposed framework.

### 2022-02-19 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
polishing rebuttal round 1

@@
 \ref{subsec:NumSimIEEEBus}.
-The $5$ rows above correspond to the $5$ generator nodes; these are buses 1, 2, 3, 6 and 8 in Fig.
+The five rows above correspond to the five generator nodes; these are buses 1, 2, 3, 6 and 8 in Fig.
 \ref{fig:Motivating}.
@@
 Indeed, Fig.
-\ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both \textbf{Case I} and \textbf{II}, at all times have highest probability around $0$ rad/s, the post-fault marginals show larger spread or dispersion of the probability mass, as expected.
+\ref{fig:IEEE14omegamarginalsCase12}(c) shows that although the $\omega_{4}$ marginals in both \textbf{Case I} and \textbf{II}, at all times have the highest probability around $0$ rad/s, the post-fault marginals show larger spread or dispersion of the probability mass, as expected.
 That the generator 4 states in \textbf{Case II} have larger uncertainties than in \textbf{Case I}, is also visible in the bivariate marginal plots if we compare the fourth row in Fig.
@@
 Fig.
-\ref{fig:RelErrSynthetic} shows that the time varying statistics from these two computation match very well at all times.
+\ref{fig:RelErrSynthetic} shows that the time-varying statistics from these two computation match very well at all times.
 

### 2022-02-20 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
polishing rebuttal round 1

@@
 \end{pmatrix}
-\! = \!\begin{pmatrix}
+\! = \!\begin{pmatrix}\!
 \bm{\eta}\\
--\nabla_{\bm{\xi}}U(\bm{\xi}) -\nabla_{\bm{\eta}}F\left(\bm{\eta}\right)  	
-\end{pmatrix}{\mathrm{d}}t + \!\begin{pmatrix}
+-\bm{\Upsilon}\nabla_{\bm{\xi}}U(\bm{\xi}) -\nabla_{\bm{\eta}}F\left(\bm{\eta}\right)  	
+\!\end{pmatrix}\!{\mathrm{d}}t + \!\begin{pmatrix}\!
  \bm{0}_{n\times n}\\
  \bm{I}_{n}	
- \end{pmatrix}{\mathrm{d}}\bm{w},
+ \!\end{pmatrix}\!{\mathrm{d}}\bm{w},
 \label{XiEtaVectorSDE}	
 \end{align}
-where the potentials
+where {\blue{the diagonal matrix $\bm{\Upsilon}:= \left(\prod_{i=1}^{n}\sigma_{i}^{2}m_{i}^{-2}\right)\bm{M\Sigma}^{-2}$ has positive entries along the main diagonal, and}} the potentials
 {\small{\begin{subequations}
 \begin{align}
-U(\bm{\xi}) &:= \!-\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\displaystyle\sum_{i,j=1}^{n}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right), \label{defU}\\
-F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
+U(\bm{\xi}) &:= \left(\displaystyle\sum_{i<j}k_{ij}\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right) \right.\nonumber\\
+&\left.\qquad\qquad-\displaystyle\sum_{i=1}^{n}\!\frac{\sigma_{i}}{m_{i}}P_{i}\xi_{i}\right)\left(\prod_{i=1}^{n}\frac{m_{i}^{2}}{\sigma_{i}^{2}}\right), \label{defU}
+%\!-\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\displaystyle\sum_{i,j=1}^{n}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right)
+\\
+F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{\Sigma}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
 \end{align}
@@
 \end{pmatrix}\right)}{|\det\left(\bm{\Psi}\right)|} \nonumber\\
-&= \dfrac{\rho_{0}\left(\bm{\Sigma}\bm{M}^{-1}\bm{\xi},\bm{\Sigma}\bm{M}^{-1}\bm{\eta}\right)}{\left(\displaystyle\prod_{j=1}^{n}m_{j}/\sigma_{j}\right)^{\!2}},
+&= \dfrac{\rho_{0}\left(\bm{\Sigma}\bm{M}^{-1}\bm{\xi},\bm{\Sigma}\bm{M}^{-1}\bm{\eta}\right)}{\left(\displaystyle\prod_{i=1}^{n}m_{i}/\sigma_{i}\right)^{\!2}},
 \label{PushforwardOfIC}	
@@
 \begin{align}
-&\dfrac{\partial\tilde{\rho}}{\partial t} = -\langle\bm{\eta},\nabla_{\bm{\xi}}\tilde{\rho}\rangle + \nabla_{\bm{\eta}}\cdot\left(\tilde{\rho}\left(\nabla_{\bm{
-\xi}}U(\bm{\xi})+\nabla_{\bm{\eta}}F(\bm{\eta})\right)\right) + \dfrac{1}{2}\Delta_{\bm{\eta}}\tilde{\rho}, \label{NewPDE}\\
+&\dfrac{\partial\tilde{\rho}}{\partial t} = -\langle\bm{\eta},\nabla_{\bm{\xi}}\tilde{\rho}\rangle + \nabla_{\bm{\eta}}\cdot\left(\tilde{\rho}\left(\bm{\Upsilon}\nabla_{\bm{
+\xi}}U(\bm{\xi})+\nabla_{\bm{\eta}}F(\bm{\eta})\right)\right) \nonumber\\
+&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad+ \dfrac{1}{2}\Delta_{\bm{\eta}}\tilde{\rho}, \label{NewPDE}\\
 &\tilde{\rho}(t=t_{0},\bm{\xi},\bm{\eta}) = \underbrace{\tilde{\rho}_{0}(\bm{\xi},\bm{\eta})}_{\text{from}\;\eqref{PushforwardOfIC}}. \label{NewIC}	
@@
 \begin{align}
-\Phi(\tilde{\rho}) := \!\!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\!\!\left(\!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\!\right)\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta},
+\Phi(\tilde{\rho}) &:= \!\!\displaystyle\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)}}\times\mathbb{R}^{n}}\!\!\!\left(\!U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\!\right)\nonumber\\
+&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta},
 \label{LyapButNotProx}	
 \end{align}
-which is the sum of a potential energy (expected value of scaled $U$), a weighted kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$).
-Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov-like functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.
-
+which is the sum of a potential energy (expected value of $U$), a weighted kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$).
+%Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov-like functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.
 However, unlike the gradient drift case mentioned in Sec.
 \ref{subsec:GenGradDescent}, it is not possible to express the right hand side of \eqref{NewPDE} as the Wasserstein gradient of the functional \eqref{LyapButNotProx}.
-To see this, recall that the Wasserstein gradient is defined as \cite[Ch.
+This can be verified via direct computation by recalling that the Wasserstein gradient is defined as \cite[Ch.
 8]{ambrosio2008gradient}
@@
 the vector $(\bm{\xi},\bm{\eta})^{\top}$, and $\frac{\delta}{\delta\rho}$ denotes the functional derivative.
-From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$.
-Thus, even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, we cannot interpret the flow generated by \eqref{TransformedPDEivp} as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}.
-Thus, in \eqref{InfiniteDimProxOpDef}, we cannot construct $(\Phi,{\rm{dist}})$ by pairing \eqref{LyapButNotProx} with the Wasserstein metric.
-
-To set up a variational recursion of the form \eqref{InfiniteDimProxOpDef} for \eqref{TransformedPDEivp}, we set $(\Phi,{\rm{dist}}) \equiv \left(\widetilde{\Phi},\widetilde{W}\right)$ where
-\begin{align}
-\widetilde{\Phi}(\tilde{\rho}) := \int_{\mathbb{T}^{n} \times \mathbb{R}^{n}}\left(F(\boldsymbol{\eta})+\frac{1}{2} \log \tilde{\rho}\right) \tilde{\rho}\: \mathrm{d} \boldsymbol{\xi} \mathrm{d} \boldsymbol{\eta},
+%From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$.
+Thus, 
+%even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, 
+we cannot interpret the flow generated by \eqref{TransformedPDEivp} as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}.
+Consequently, in \eqref{InfiniteDimProxOpDef}, we cannot construct $(\Phi,{\rm{dist}})$ by pairing \eqref{LyapButNotProx} with the Wasserstein metric.
+
+To set up a variational recursion of the form \eqref{InfiniteDimProxOpDef} for \eqref{TransformedPDEivp}, we set $(\Phi,{\rm{dist}}) \equiv \left(\widetilde{\Phi},\widetilde{W}_{h,\bm{\Upsilon}}\right)$ where
+\begin{align}
+\widetilde{\Phi}(\tilde{\rho}) := \!\!\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)}} \times \mathbb{R}^{n}}\!\!\left(\!F(\boldsymbol{\eta})+\frac{1}{2} \log \tilde{\rho}\!\right) \tilde{\rho}\: \mathrm{d} \boldsymbol{\xi} \mathrm{d} \boldsymbol{\eta},
 \label{Phitilde}	
@@
 \begin{align}
-\widetilde{W}^{2}\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right):=\!&\inf _{\pi \in \Pi\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right)}\!\int_{\mathbb{T}^{2 n} \times \mathbb{R}^{2 n}}\!\!\left\{\left\|\overline{\boldsymbol{\eta}}-\boldsymbol{\eta}+h \nabla_{\boldsymbol{\xi}} U(\boldsymbol{\xi})\right\|_{2}^{2}\right.\nonumber\\
-&\hspace*{-0.2in}\left.+12\left\|\frac{\overline{\boldsymbol{\xi}}-\boldsymbol{\xi}}{h}-\frac{\overline{\boldsymbol{\eta}}+\boldsymbol{\eta}}{2}\right\|_{2}^{2}\right\} \mathrm{d} \pi(\boldsymbol{\xi}, \boldsymbol{\eta}, \overline{\boldsymbol{\xi}}, \overline{\boldsymbol{\eta}}).
+&\widetilde{W}_{h,\bm{\Upsilon}}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) := \!\!\underset{\pi\in\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)}{\inf}\!\displaystyle\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)}} \times \mathbb{R}^{2n}}\!\!s_{h,\bm{\Upsilon}}\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right)\nonumber\\
+&\differential\pi\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right),
 \label{Wtilde}    
 \end{align}
-In \eqref{Wtilde}, $\Pi\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right)$ denotes the collection of all joint PDFs supported on $\mathbb{T}^{2 n} \times \mathbb{R}^{2 n}$ that have finite second raw moments, with the first marginal $\tilde{\varrho}$, and the second marginal $\tilde{\varrho}_{k-1}$.
+where $h>0$ is the step-size in proximal recursion, and $\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)$ denotes the set of joint probability measures over the product space $\mathbb{T}^{2n} \times \mathbb{R}^{2n}$ that have finite second moments  with the first marginal $\tilde{\varrho}$, and the second marginal $\tilde{\varrho}_{k-1}$.
+The ``ground cost" $s_{h,\bm{\Upsilon}}$ in \eqref{Wtilde} is given by
+ \begin{align}
+     &s_{h,\bm{\Upsilon}}\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right):=\nonumber\\
+    &\bigg\langle\!\!\left(\bar{\bm{\eta}}-\bm{\eta}+h\bm{\Upsilon} \nabla_{\bm{\xi}} \bm{U}(\bm{\xi})\right),\bm{\Upsilon}^{-1}\!\left(\bar{\bm{\eta}}-\bm{\eta}+h\bm{\Upsilon} \nabla_{\bm{\xi}}\bm{U}(\bm{\xi})\right)\!\!\bigg\rangle\nonumber\\
+    & + 12 \bigg\langle\!\left(\!\frac{\bar{\bm{\xi}}-\bm{\xi}}{h}-\frac{\bar{\bm{\eta}}-\bm{\eta}}{h}\!\right),\bm{\Upsilon}^{-1}\!\left(\!\frac{\bar{\bm{\xi}}-\bm{\xi}}{h}-\frac{\bar{\bm{\eta}}-\bm{\eta}}{h}\!\right)\!\bigg\rangle.
+    \end{align}
+
+%$\Pi\left(\tilde{\varrho}, \tilde{\varrho}_{k-1}\right)$ denotes the collection of all joint PDFs supported on $\mathbb{T}^{2 n} \times \mathbb{R}^{2 n}$ that have finite second raw moments,
 
@@
 \tilde{\varrho}_{k} &= {\text{prox}}^{\widetilde{W}}_{h\widetilde{\Phi}}(\tilde{\varrho}_{k-1}) \nonumber\\
-&\equiv \underset{\tilde{\rho}\in\mathcal{P}_{2}}{\arg\inf}\:\frac{1}{2}\widetilde{W}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) + h\:\widetilde{\Phi}(\tilde{\rho}), \quad \tilde{\varrho}_{0} := \tilde{\rho}_{0},
+&\equiv \underset{\tilde{\varrho}\in\mathcal{P}_{2}}{\arg\inf}\:\frac{1}{2}\widetilde{W}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) + h\:\widetilde{\Phi}(\tilde{\varrho}), \quad \tilde{\varrho}_{0} := \tilde{\rho}_{0},
 \label{KineticProx}	
@@
 
-\appendices
-\section{Showing $\Phi(\tilde{\rho})$ is Decreasing along the Flow Generated by \eqref{TransformedPDEivp}}\label{AppLyap}
-Let $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ be the transient solution of \eqref{TransformedPDEivp}. Define
-$$\bm{\ell}_{\bm{\xi}} := \frac{1}{2}\nabla_{\bm{\xi}}\log\tilde{\rho}, \quad \bm{\ell}_{\bm{\eta}} := \frac{1}{2}\nabla_{\bm{\eta}}\log\tilde{\rho},$$
-and rewrite \eqref{NewPDE} as
-$$\dfrac{\partial\tilde{\rho}}{\partial t} = - \nabla\cdot\left(\tilde{\rho}\bm{v}\right)$$
-where $\nabla$ is the $2n\times 1$ Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and the vector field
-\begin{align}
-\bm{v} := \begin{pmatrix}
- \bm{\eta}\\
- -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
- \end{pmatrix}.
-\label{defvectorfieldv}	
-\end{align}
-
-Now consider the time derivative of the functional \eqref{LyapButNotProx} along \eqref{NewPDE}. Following \cite[Proposition 1]{krichene2020global}, 
-\begin{align}
-\frac{\differential}{\differential t}\Phi = \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}} \bigg\langle\nabla\dfrac{\delta\Phi}{\delta\tilde{\rho}}, \tilde{\rho}\bm{v}\bigg\rangle\:\differential\bm{\xi}\differential\bm{\eta},
-\label{dPhidt}	
-\end{align}
-wherein the functional derivative
-\begin{align}
-\dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
-\label{FunctionalDerivative}
-\end{align}
-Recalling the explicit form of $F(\bm{\eta})$ from \eqref{defF}, and using \eqref{FunctionalDerivative}, we simplify \eqref{dPhidt} as
-\begin{strip}
-\hrule
-\begin{subequations}
-\begin{align}
-\frac{\differential}{\differential t}\Phi &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\langle\begin{pmatrix}
-\bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
-\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}	
-\end{pmatrix}, \begin{pmatrix}
- \bm{\eta}\\
- -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
- \end{pmatrix}\bigg\rangle\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
-&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
-&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
-\end{align}
-\label{ExplicitdPhidt}	
-\end{subequations}
-\hrule	
-\end{strip}
-where \eqref{lastequality} follows from the fact that the integrals involving $\langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle$ and $\langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle$ in \eqref{lastbutoneequality} are zero, see \cite[Appendix D.3]{krichene2020global}.
-
-From \eqref{ExplicitdPhidt}, it follows that $\frac{\differential}{\differential t}\Phi \leq 0$, i.e., $\Phi$ is non-increasing along the flow generated by \eqref{TransformedPDEivp}.
+%\appendices
+%\section{Showing $\Phi(\tilde{\rho})$ is Decreasing along the Flow Generated by \eqref{TransformedPDEivp}}\label{AppLyap}
+%Let $\tilde{\rho}(t,\bm{\xi},\bm{\eta})$ be the transient solution of \eqref{TransformedPDEivp}. Define
+%$$\bm{\ell}_{\bm{\xi}} := \frac{1}{2}\nabla_{\bm{\xi}}\log\tilde{\rho}, \quad \bm{\ell}_{\bm{\eta}} := \frac{1}{2}\nabla_{\bm{\eta}}\log\tilde{\rho},$$
+%and rewrite \eqref{NewPDE} as
+%$$\dfrac{\partial\tilde{\rho}}{\partial t} = - \nabla\cdot\left(\tilde{\rho}\bm{v}\right)$$
+%where $\nabla$ is the $2n\times 1$ Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and the vector field
+%\begin{align}
+%\bm{v} := \begin{pmatrix}
+% \bm{\eta}\\
+% -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
+% \end{pmatrix}.
+%\label{defvectorfieldv}	
+%\end{align}
+%
+%
+%Now consider the time derivative of the functional \eqref{LyapButNotProx} along \eqref{NewPDE}. Following \cite[Proposition 1]{krichene2020global}, 
+%\begin{align}
+%\frac{\differential}{\differential t}\Phi = \displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}} \bigg\langle\nabla\dfrac{\delta\Phi}{\delta\tilde{\rho}}, \tilde{\rho}\bm{v}\bigg\rangle\:\differential\bm{\xi}\differential\bm{\eta},
+%\label{dPhidt}	
+%\end{align}
+%wherein the functional derivative
+%\begin{align}
+%\dfrac{\delta\Phi}{\delta\tilde{\rho}}	= \!\bm{M}^{-1}\bm{\Gamma}U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2} + \frac{1}{2}\log\tilde{\rho}.
+%\label{FunctionalDerivative}
+%\end{align}
+%Recalling the explicit form of $F(\bm{\eta})$ from \eqref{defF}, and using \eqref{FunctionalDerivative}, we simplify \eqref{dPhidt} as
+%\begin{strip}
+%\hrule
+%\begin{subequations}
+%\begin{align}
+%\frac{\differential}{\differential t}\Phi &= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\langle\begin{pmatrix}
+%\bm{M}^{-1}\bm{\Gamma}\nabla_{\bm{\xi}}U(\bm{\xi}) + \bm{\ell}_{\bm{\xi}}\\
+%\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}	
+%\end{pmatrix}, \begin{pmatrix}
+% \bm{\eta}\\
+% -\nabla_{\bm{\xi}}U(\bm{\xi}) - \bm{M}^{-1}\bm{\Gamma}\bm{\eta} - \bm{\ell}_{\bm{\eta}}	
+% \end{pmatrix}\bigg\rangle\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta}	\label{firstequality}\\
+%&= \!\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\!\bigg\{-\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\|_{2}^{2} - \|\bm{\ell}_{\bm{\eta}}\|_{2}^{2} - 2\langle \bm{M}^{-1}\bm{\Gamma}\bm{\eta},\bm{\ell}_{\bm{\eta}}\rangle + \langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle - \langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle\bigg\}\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi}\:\differential\bm{\eta} \label{lastbutoneequality}\\
+%&= -\displaystyle\int_{\mathbb{T}^{n}\times\mathbb{R}^{n}}\|\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}}\|_{2}^{2}\:\tilde{\rho}(t,\bm{\xi},\bm{\eta})\:\differential\bm{\xi} \:\differential\bm{\eta}, \label{lastequality}
+%\end{align}
+%\label{ExplicitdPhidt}	
+%\end{subequations}
+%\hrule	
+%\end{strip}
+%where \eqref{lastequality} follows from the fact that the integrals involving $\langle \bm{\ell}_{\bm{\xi}},\bm{\eta}\rangle$ and $\langle\bm{\ell}_{\bm{\eta}},\nabla_{\bm{\xi}}U(\bm{\xi})\rangle$ in \eqref{lastbutoneequality} are zero, see \cite[Appendix D.3]{krichene2020global}.
+%
+%From \eqref{ExplicitdPhidt}, it follows that $\frac{\differential}{\differential t}\Phi \leq 0$, i.e., $\Phi$ is non-increasing along the flow generated by \eqref{TransformedPDEivp}.
 Furthermore, \eqref{lastequality} tells us that $\frac{\differential}{\differential t}\Phi = 0$ if and only if
-$$\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}} = 0,$$
-i.e., $\Phi$ is stationary if and only if $\tilde{\rho}$ has $\bm{\eta}$ marginal $\int\tilde{\rho}\:\differential\bm{\xi} \propto \exp\left(-\bm{\eta}^{\top}\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\right)$.
+%$$\bm{M}^{-1}\bm{\Gamma}\bm{\eta} + \bm{\ell}_{\bm{\eta}} = 0,$$
+%i.e., $\Phi$ is stationary if and only if $\tilde{\rho}$ has $\bm{\eta}$ marginal $\int\tilde{\rho}\:\differential\bm{\xi} \propto \exp\left(-\bm{\eta}^{\top}\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\right)$.
 Following the steps as in \cite[Appendix E.1]{krichene2020global}, this is achieved at the stationary solution of \eqref{NewPDE} with $\bm{\xi}$ marginal $\int\tilde{\rho}\:\differential\bm{\eta} \propto \exp\left(-2U(\bm{\xi})\right)$.
 In other words, the stationary solution of \eqref{NewPDE} must be of the form
-$$\tilde{\rho}_{\infty}(\bm{\xi},\bm{\eta}) \propto \exp\left(-2U(\bm{\xi})+\bm{\eta}^{\top}\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\right),$$
-and is exactly where $\frac{\differential}{\differential t}\Phi = 0$.
+%$$\tilde{\rho}_{\infty}(\bm{\xi},\bm{\eta}) \propto \exp\left(-2U(\bm{\xi})+\bm{\eta}^{\top}\bm{M}^{-1}\bm{\Gamma}\bm{\eta}\right),$$
+%and is exactly where $\frac{\differential}{\differential t}\Phi = 0$.
 Therefore, $\frac{\differential}{\differential t}\Phi < 0$ along the transient solution of \eqref{NewPDE}, and $=0$ at the stationary solution of \eqref{NewPDE}.

### 2022-02-20 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
polishing rebuttal round 1

@@
 \begin{align}
-&\widetilde{W}_{h,\bm{\Upsilon}}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) := \!\!\underset{\pi\in\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)}{\inf}\!\displaystyle\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)}} \times \mathbb{R}^{2n}}\!\!s_{h,\bm{\Upsilon}}\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right)\nonumber\\
-&\differential\pi\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right),
+&\widetilde{W}_{h,\bm{\Upsilon}}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) := \!\underset{\pi\in\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)}{\inf}\!\displaystyle\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)^{2}}} \times \mathbb{R}^{2n}}\nonumber\\
+&\qquad\qquad\qquad\qquad s_{h,\bm{\Upsilon}}\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right)\differential\pi\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right),
 \label{Wtilde}    
 \end{align}
-where $h>0$ is the step-size in proximal recursion, and $\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)$ denotes the set of joint probability measures over the product space $\mathbb{T}^{2n} \times \mathbb{R}^{2n}$ that have finite second moments  with the first marginal $\tilde{\varrho}$, and the second marginal $\tilde{\varrho}_{k-1}$.
+wherein $h>0$ denotes the step-size in proximal recursion, and $\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)$ denotes the set of joint probability measures over the product space ${\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)^{2}}} \times \mathbb{R}^{2n}$ that have finite second moments  with the first marginal $\tilde{\varrho}$, and the second marginal $\tilde{\varrho}_{k-1}$.
 The ``ground cost" $s_{h,\bm{\Upsilon}}$ in \eqref{Wtilde} is given by
@@
 $$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} {\blue{\tilde{\rho}}}(t=kh,\bm{\xi},\bm{\eta})\;\text{in}\;L^{1}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right),$$
-was established in \cite{duong2014conservative}.
+can be established following the arguments in \cite{duong2014conservative}.
 To numerically perform the recursion \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite number of samples, as explained next.
@@
 \begin{algorithmic}[1]
-\Procedure{Prox}{$\tilde{\bm{\varrho}}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, h, \varepsilon, N, \delta, \ell_{\max}$}
+\Procedure{Prox}{$\tilde{\bm{\varrho}}_{k-1}, \tilde{\bm{x}}_{k-1}, \tilde{\bm{x}}_{k}, \bm{\Upsilon}, h, \varepsilon, N, \delta, \ell_{\max}$}
 \For{$i=1$ to $N$}
-\State $\bm{\zeta}_{k-1}(i) \gets \exp\left(-\left(\bm{\eta}_{k-1}^{i}\right)^{\top}\bm{\eta}_{k-1}^{i}-1\right)$
+\State $\bm{\zeta}_{k-1}(i) \gets \exp\left(-F\left(\bm{\eta}_{k-1}^{i}\right)-1\right)$
 \For{$j=1$ to $N$}
-\State $\bm{C}_{k}(i,j) \gets \parallel\bm{\eta}_{k-1}^{j} - \bm{\eta}_{k}^{i}+ h\nabla U(\bm{\xi}_{k}^{i})\parallel_{2}^{2}$
-\State \qquad\quad $+12\bigg\|\dfrac{\bm{\xi}_{k-1}^{j}-\bm{\xi}_{k}^{i}}{h} - \dfrac{\bm{\eta}_{k-1}^{j} + \bm{\eta}_{k}^{i}}{2}\bigg\|_{2}^{2}$
+\State $\bm{C}_{k}(i,j) \gets s_{h,\bm{\Upsilon}}\left(\bm{\xi}_{k-1}^{i},\bm{\eta}_{k-1}^{i},\bm{\xi}_{k}^{j},\bm{\eta}_{k}^{j}\right)$
 \EndFor

### 2022-02-27 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
polishing rebuttal round 1

@@
 (i) In equation (8), we corrected a typo in its right hand side: the second summand runs over pairs $(i,j)$ such that $i<j$.\\
-(ii)
+(ii) To make the presentation tighter, the Appendix appearing in the previous version has been eschewed.
+We have accordingly adjusted and condensed the material in Sec.
+III-D of the revised manuscript.\\
+(iii) Some mathematical typos have been fixed in the revised manuscript's Sec. III-C and III-D.
 
@@
 
-{\nib \blue{In Sec. V-A, we corrected ``assuming" $\rightarrow$ ``are assumed". We corrected the two typos pointed out in comment \textbf{R4.8} in this rebuttal document.}}
+{\nib \blue{In Sec.
+V-A, we corrected ``assuming" $\rightarrow$ ``are assumed".
+We corrected the two typos pointed out in comment \textbf{R4.8} in this rebuttal document.
+We also corrected minor mathematical typos in (19), (20a), (22a), as well the transformed domains in Sec.
+III-D.}}
 
@@
 \ul{On the other hand, the detailed discussion regarding Wasserstein gradient flow and why that standard development cannot be applied in this case, is in fact, the heart of this paper}.
-Please note that this is far from a recipe paper where a numerical algorithm is proposed backed by some theory and simulations.
-\ul{The key contribution of this paper, instead, is to help distill out the core technical difficulty that hinders simply importing standard techniques from the optimal transport literature}.
-We believe this clarification of what really is the technical crux of the matter--versus what are peripheral issues--should be illuminating to not just power systems researchers, but also to applied mathematicians underscoring how the standard theory and algorithms for measure-valued gradient flows need to be adapted in engineering applications.}}
+\ul{A key contribution of this paper is to help distill out and tackle the core mathematical difficulty that hinders simply importing standard techniques from the optimal transport literature.
+To make the presentation tighter, in the revised manuscript, we have eschewed the Appendix and accordingly adjusted the material in Sec.
+III-D.}}}
 

### 2022-02-28 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
polishing rebuttal round 1

@@
 We do not believe this issue is exceptional for our paper since several high impact papers in the IEEE Trans.
-Power Systems were indeed at the intersection of these areas, and are not only well-read by the researchers from either side but these works have positively contributed toward increasing interdisciplinary research collaboration.}}
+Power Systems were indeed at the intersection of these areas, and have positively contributed toward increasing interdisciplinary research collaboration.}}
 
@@
 
-{\nib \blue{ {\red{TBD.}}}}
+{\nib \blue{Thanks for this comment.
+\ul{The revised manuscript's newly incorporated Fig.
+8 shows the statistical uncertainty bounds for the rotor angle sample paths for the simulation in Sec.
+V-A}.
+The inline text citing this figure in Sec.
+V-A is highlighted in blue.}}
 
@@
 
-{\nib \blue{ {\red{TBD.}}}}
+{\nib \blue{\ul{Regarding the comparison with existing methods}, we have provided comparison with Monte Carlo in the revised manuscript's Fig.
+11 {\red{and another??
+(in response to comment \textbf{R4.7})}}.
+The new Fig.
+8 in the revised manuscript also shows the consistency for the rotor angle stochastic predictions w.r.t.
+Monte Carlo for the simulation reported in Sec.
+V-A, as requested by comment \textbf{R2.5}.
+Please note that performing finite element simulation for 5-generator system is a very tall order since the state space then is 10 dimensional, and even if we do a very coarse 10 discretizations per dimension, then such a simulation would require us to evolve a joint PDF over $10^{10}$ grid points.
+This requires dedicated high-performance computing architecture and an undertaking well beyond the scope of this paper.
+This also explains why the state-of-the-art for direct computation of transient joint PDF in power systems literature remains single-machine-infinite-bus $(n=1)$ as in the 2013 ref.
+[45] which indeed performed the finite element simulation with a two dimensional discretization, and mentioned the difficulty of scaling this computational approach for even moderately large sized systems.
+
+\ul{Regarding the computational time for longer horizon}, {\red{TBD}}}}
 

### 2022-02-28 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
polishing rebuttal round 1

@@
 
+\begin{figure*}[ht!]
+   \subfigure[Case I]{\includegraphics[width=0.49\textwidth]{IEEE14_Case0_RotorAngles.png}}
+   \hspace{\fill}
+   \subfigure[Case II]{\includegraphics[width=0.49\textwidth]{IEEE14_Case1_RotorAngles.png}}
+\caption{\small{Boxplots for the rotor angles $\theta_{i}(t)$, $i=1,\hdots,5$, for the IEEE 14 bus system simulation in Sec.
+\ref{subsec:NumSimIEEEBus} in the (a) nominal \textbf{Case I}, and in the post-fault \textbf{Case II} scenarios, respectively.
+As usual, the boxes show the respective interquartile ranges, and the whiskers show the respective minimum and maximum values.
+In each box plot, we have superimposed  the respective \emph{mean} rotor angle sample path.}}
+\vspace*{-0.1in}
+\label{fig:IEEE14theta}
+\end{figure*}
+
 \begin{figure}[htpb]
@@
 \ref{fig:IEEE14marginalsCase1}.
+{\blue{The statistical uncertainties in the rotor angles are shown in Fig.
+\ref{fig:IEEE14theta} for both \textbf{Case I} and \textbf{Case II}.}}
 

### 2022-02-28 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py

@@ -11,6 +11,7 @@ from pylab import rcParams
 from matplotlib.patches import FancyArrowPatch
 from mpl_toolkits.mplot3d import proj3d
 from scipy.interpolate import interp1d
+from fractions import Fraction as frac
 
 #====================================================
 # Make plots beautiful
@@ -60,33 +61,33 @@ fig = plt.figure(1, figsize=fig_size)  # figsize accepts only inches.
 # Computational time plot [IEEE 14 bus 5 gen, Case I and Case II]
 # =============================================================================
 Time_Synthetic = np.loadtxt("TimeSyntheticIEEE14bus.txt")
-CompTime_Synthetic_Case1 = np.loadtxt("case0_norminal_ComptimeSytheticIEEE14bus.txt")
-CompTime_Synthetic_Case2 = np.loadtxt("case1_line_13_failure_ComptimeSytheticIEEE14bus.txt")
+# CompTime_Synthetic_Case1 = np.loadtxt("case0_norminal_ComptimeSytheticIEEE14bus.txt")
+# CompTime_Synthetic_Case2 = np.loadtxt("case1_line_13_failure_ComptimeSytheticIEEE14bus.txt")
 
-fig = plt.figure()
-ax = fig.add_subplot(1, 1, 1)
+# fig = plt.figure()
+# ax = fig.add_subplot(1, 1, 1)
 
-ax.semilogy(Time_Synthetic, CompTime_Synthetic_Case1,'-', color='b', lw=1.5, alpha=0.7,label=r"Case I")
-ax.semilogy(Time_Synthetic, CompTime_Synthetic_Case2,':', color='r', lw=1.5, alpha=0.7,label=r"Case II")
+# ax.semilogy(Time_Synthetic, CompTime_Synthetic_Case1,'-', color='b', lw=1.5, alpha=0.7,label=r"Case I")
+# ax.semilogy(Time_Synthetic, CompTime_Synthetic_Case2,':', color='r', lw=1.5, alpha=0.7,label=r"Case II")
 
-# ax.set_ylim(10**-3, 6*10**-2)
+# # ax.set_ylim(10**-3, 6*10**-2)
 
-ax.tick_params(direction='in',which='both')
-#  # axx.xaxis.tick_top()
-ax.grid(True,which="both",ls="-", color='0.75')
-# # axx2.grid(True,which="both",ls="-", color='0.75')
-ax.tick_params(axis='both', labelsize=18)
[…]

### 2022-02-28 code/ThreeCases_IEEE14_prox_simulation.m

@@ -1,4 +1,4 @@
-close all; clear all; clc;
+close all; clear; clc;
 set(groot,'defaultAxesTickLabelInterpreter','latex');  
 set(groot,'defaulttextinterpreter','latex');
 set(groot,'defaultLegendInterpreter','latex');
@@ -17,12 +17,12 @@ bus_init = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertain
 % case2_time_series_
 
 % parameters for the IEEE 14 bus system
-Y_reduced_imag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_Y_reduced_imag_part.csv');
-Y_reduced_real = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_Y_reduced_real_part.csv');
+Y_reduced_imag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_Y_reduced_imag_part.csv');
+Y_reduced_real = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_Y_reduced_real_part.csv');
 % reduced admittance matrix
 Y_reduced = complex(Y_reduced_real,Y_reduced_imag);
 
-gen_param = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_gen_parameters.csv');
+gen_param = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_gen_parameters.csv');
 gen_bus_idx = gen_param(:,1);
 
 num_Oscillator = numel(gen_bus_idx); dim  = 2*num_Oscillator;
@@ -35,17 +35,17 @@ E = E_mag.*exp(1i*E_phase);
 %I_reduced_polar = readmatrix('I_red.csv');
 %I_reduced_mag = I_reduced_polar(:,1);
 %I_reduced_phase = I_reduced_polar(:,2);
[…]

### 2022-03-03 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
polishing rebuttal round 1

@@
 \caption{\small{Boxplots for the rotor angles $\theta_{i}(t)$, $i=1,\hdots,5$, for the IEEE 14 bus system simulation in Sec.
-\ref{subsec:NumSimIEEEBus} in the (a) nominal \textbf{Case I}, and in the post-fault \textbf{Case II} scenarios, respectively.
+\ref{subsec:NumSimIEEEBus} in the (a) nominal \textbf{Case I}, and in the (b) post-fault \textbf{Case II} scenarios, respectively.
 As usual, the boxes show the respective interquartile ranges, and the whiskers show the respective minimum and maximum values.
@@
 \ref{subsec:NumSimSynthetic}.}}
+\vspace*{-0.15in}
+\label{fig:RelErrSynthetic}
+\end{figure}
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.95\linewidth]{RelErrCovMCvsProxSythetic50Gen.png}
+\caption{\small{The normalized Bures-Wasserstein distance $d_{\rm{BW}}(\cdot,\cdot)$ between the empirical covariance  (i.e., Monte Carlo estimate) ${\rm{Cov}}^{\rm{MC}}\in\mathbb{R}^{100\times 100}$, and the ensemble covariance ${\rm{Cov}}^{\rm{Prox}}\in\mathbb{R}^{100\times 100}$ obtained using the proximal updates of the joint state PDFs, for the simulation set up in Sec.
+\ref{subsec:NumSimSynthetic}.
+The normalization is w.r.t.
+$\sqrt{{\rm{trace}}\left({\rm{Cov}}^{\rm{MC}}\right)}$.}}
 \vspace*{-0.2in}
-\label{fig:RelErrSynthetic}
+\label{fig:RelErrCovSynthetic}
+\end{figure}
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.95\linewidth]{WassMCvsProxSythetic50Gen.png}
+\caption{\small{The 2-Wasserstein distance between the empirical joint PDF (i.e., Monte Carlo estimate) $\rho^{\rm{MC}}$, and the proximal update $\varrho^{k}$ obtained using the proposed proximal algorithm, for the simulation set up in Sec.
+\ref{subsec:NumSimSynthetic}.}}
+\vspace*{-0.2in}
+\label{fig:WassJointPDFSynthetic}
 \end{figure}
@@
 The latter was computed using the proximal updates $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
-Fig.
-\ref{fig:RelErrSynthetic} shows that the time-varying statistics from these two computation match very well at all times.
+{\blue{Likewise, Fig.
+\ref{fig:RelErrCovSynthetic} plots the time-varying normalized Bures-Wasserstein distance $d_{\rm{BW}}(\cdot,\cdot)$ between the empirical covariance matrix (i.e., Monte Carlo estimate) ${\rm{Cov}}^{\rm{MC}}\in\mathbb{R}^{100\times 100}$, and the ensemble covariance matrix ${\rm{Cov}}^{\rm{Prox}}\in\mathbb{R}^{100\times 100}$ obtained using the proximal updates.
+The unnormalized Bures-Wasserstein distance \cite{bhatia2019bures} given by
+\begin{align*}
+&d_{\rm{BW}}(\rm{Cov}^{\rm{MC}},\rm{Cov}^{\rm{Prox}}) := \left[\operatorname{trace}\left(\rm{Cov}^{\rm{MC}}+\rm{Cov}^{\rm{Prox}}\right)\right.\\
+&\qquad\qquad \left.-2\operatorname{trace}\left(\left(\sqrt{\rm{Cov}^{\rm{MC}}}\:\rm{Cov}^{\rm{Prox}}\:\sqrt{\rm{Cov}^{\rm{MC}}}\right)^{\frac{1}{2}}\right)\right]^{\frac{1}{2}},	
+\end{align*}
+is a metric on the cone of symmetric positive definite matrices such as the covariance matrices, and can be seen as an absolute error between the covariances.
+Thus, Fig.
+\ref{fig:RelErrCovSynthetic} can be viewed as a relative error curve between the covariances.
+In summary, Figs.
+\ref{fig:RelErrSynthetic} and \ref{fig:RelErrCovSynthetic} show that the time-varying statistics between the Monte Carlo and the proposed proximal computation remain close at all times.}}
+
+{\blue{Since it is not possible to \emph{a priori} determine the dimension of the sufficient statistic of the transient joint state PDFs, Fig.
+\ref{fig:WassJointPDFSynthetic} shows the 2-Wasserstein metric between the empirical joint PDF (i.e., Monte Carlo estimate) $\rho^{\rm{MC}}$, and the proximal update $\varrho^{k}$.
+In other words, Fig.
+\ref{fig:WassJointPDFSynthetic} can be viewed as a nonparametric absolute error curve between $\rho^{\rm{MC}}$ and $\varrho^{k}$.
+Each snapshot value in Fig.
+\ref{fig:WassJointPDFSynthetic} is computed by solving the well-known optimal transport linear program (see e.g., \cite[Sec.
+4.2-4.3]{halder2014probabilistic}, \cite[Ch.
+3]{peyre2019computational}) using the corresponding time-varying weighted point clouds.}}
 
@@
 \bibitem{cui2021andes}
-H.~Cui, F.~Li, and K.~Tomsovic, ``Hybrid Symbolic-Numeric Framework for Power System Modeling and Analysis,'' \emph{IEEE Transactions on Power Systems}, vol. 36, no. 2, pp. 1373-1384, 2021.
+H.~Cui, F.~Li, and K.~Tomsovic, ``Hybrid Symbolic-Numeric Framework for Power System Modeling and Analysis'', \emph{IEEE Transactions on Power Systems}, Vol. 36, No. 2, pp. 1373-1384, 2021.
+
+% == Bures-Wasserstein ====
+
+{\blue{\bibitem{bhatia2019bures}
+R.~Bhatia, T.~Jain, and Y.~Lim, ``On the Bures--Wasserstein Distance between Positive Definite Matrices", \emph{Expositiones Mathematicae}, Vol. 37, No. 2, pp. 165--191, 2019.
+
+\bibitem{halder2014probabilistic}
+A.~Halder, and R.~Bhattacharya, ``Probabilistic Model Validation for Uncertain Nonlinear Systems", \emph{Automatica}, Vol. 50, No. 8, pp. 2038--2050, 2014.
+
+\bibitem{peyre2019computational}
+G.~Peyr{\'e}, and M.~Cuturi, ``Computational Optimal Transport: With Applications to Data Science", \emph{Foundations and Trends{\textregistered} in Machine Learning}, Vol.
+11, No.
+5-6, pp.
+355--607, 2019.
+
+}}
 

### 2022-03-03 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py

@@ -124,33 +124,79 @@ Time_Synthetic = np.loadtxt("TimeSyntheticIEEE14bus.txt")
 # plt.savefig('ComputationalTimeSyntheticIEEE14bus.png', dpi=400)
 
 # =============================================================================
-# Relative error in mean vector: MC versus Prox plot
+# Relative error in mean vector and covariance matrix: MC versus Prox plot
 # =============================================================================
 # RelErrMeanVectorMCvsProx_Synthetic = np.loadtxt("RelErrMeanVectorMCvsProxSythetic50Gen.txt")
+
+# DistCovMatrixMCvsProx_Synthetic = np.loadtxt("/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/code/RelErrCovMCvsProxSythetic50Gen.txt")
 
 # fig = plt.figure()
 # ax = fig.add_subplot(1, 1, 1)
 
-# ax.semilogy(Time_Synthetic[3:-1], RelErrMeanVectorMCvsProx_Synthetic[4:-1],'-', color='k', lw=1.5)
+# # ax.semilogy(Time_Synthetic[3:-1], RelErrMeanVectorMCvsProx_Synthetic[4:-1],'-', color='k', lw=1.5)
+# ax.semilogy(Time_Synthetic[3:-1], DistCovMatrixMCvsProx_Synthetic[3:-1],'-', color='k', lw=1.5)
 
-# # ax.set_ylim(1*10**-3, 1.2*10**-2)
+# # ax.set_ylim(3*10**-2, 2*10**-1)
 
 # ax.tick_params(direction='in',which='both')
-#  # axx.xaxis.tick_top()
+# #  # axx.xaxis.tick_top()
 
 # ax.grid(True,which="both",ls="-", color='0.75')
-# # axx2.grid(True,which="both",ls="-", color='0.75')
+# # # axx2.grid(True,which="both",ls="-", color='0.75')
 # ax.tick_params(axis='both', labelsize=18)
 
 
-# ax.set_ylabel(r"Realtive error $\frac{\|\boldsymbol{\mu}_{k}^{\rm{MC}}-\boldsymbol{\mu}_{k}^{\rm{Prox}}\|_{2}}{\|\boldsymbol{\mu}_{k}^{\rm{MC}}\|_{2}}$")
-# ax.set_xlabel(r"Physical time $t=kh$ [s]")
-# # # axx.yaxis.set_label_coords(-0.125,-0.05)
[…]

### 2022-03-03 code/Gen_50_prox_simulation.m

@@ -1,4 +1,4 @@
-cclose all; clear; clc;
+close all; clear; clc;
 %% Parameters
 beta = 2; % inverse temperature
 % golbal parameters 
@@ -30,7 +30,7 @@ K = K - diag(diag(K)); % zero the diagonal entries
 theta0_a = 0 ; theta0_b = 2*pi; omega0_a = -.1 ; omega0_b = .1;
 
 % parameters for proximal recursion
-nSample = 2000;                      % number of samples                                                           
+nSample = 1000;                      % number of samples                                                           
 epsilon = 0.5;                      % regularizing coefficient                                      
 h = 1e-3;                         % time step
 numSteps= 1000;                    % number of steps k, in discretization t=kh
@@ -94,6 +94,9 @@ mean_prox_omega(1,:) = sum(omega_upd(:,:,1).*rho_theta_omega_upd(:,1))/sum(rho_t
     
 mean_prox_theta(1,:) = weighted_angle_mean(theta_upd(:,:,1),rho_theta_omega_upd(:,1));
 
+numRun = 100;
+
+for r=1:numRun
 tic;
 for j=1:numSteps 
     
@@ -105,8 +108,10 @@ for j=1:numSteps
     theta_upd(:,:,j+1) = wrapTo2Pi((psi_upper_diag\xi_eta_upd(:,1:num_Oscillator,j+1)')');
     
     omega_upd(:,:,j+1) = (psi_lower_diag\xi_eta_upd(:,num_Oscillator+1:dim,j+1)')';
-        
-  %proximal update for joint PDF
+    
+    theta_omega_upd(:,:,j+1) = [theta_upd(:,:,j+1), omega_upd(:,:,j+1)];
+
+   % proximal update for joint PDF
[…]

### 2022-03-03 code/PowerDrift.m

@@ -1,10 +1,10 @@
 function [driftk,GradU] = PowerDrift(xi,eta,nSample)
 
-global num_Oscillator phi Gamma M Sigma P K sigma  m
+global num_Oscillator dim Gamma M Sigma K Pmech cc sigma m phi
  
 GradF_eta = ((Gamma/M)*eta')';
 
-GradU_2 = -repmat((diag(P/Sigma))',nSample,1);
+GradU_2 = -repmat((diag(Pmech/Sigma))',nSample,1);
 % 
 % for l=1:nSample
 % 

### 2022-03-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
polishing rebuttal round 1

@@
 
-{\red{Regarding the usage of simple model, please see our response to comment \textbf{R3.1} and \textbf{1.3, sub-comment (1)}.
+Regarding the usage of simple model, please see our response to comment \textbf{R3.1} and \textbf{1.3, sub-comment (1)}.
 In summary, the current state-of-the-art in power system literature for this problem is ref.
@@
 This computation is online and approximates neither the statistics nor the dynamics.
-\ul{We clarify two important points here}:\\(i) The proposed proximal recursion framework on the space of probability measures is not conceptually limited by the second order dynamics, and we expect follow up research to design custom proximal recursions for higher order synchronous machine models.
-In this sense, the present paper should be seen as the stepping stone laying the foundation for handling more complex dynamical models.
-The authors have been completely up front by explicitly mentioning this in the Conclusions section.
-We have also reiterated the same in this rebuttal.\\(ii) Deriving proximal recursion requires understanding the generalized gradient flow structure over probability measures.
-Our research disposition is that one cannot bypass understanding the generalized gradient flow structure for the simpler second order sample path dynamics before attempting the same for more complex sample path models.
-In fact, a significant contribution of our article, in the authors' opinion, is that it distills out what exactly is mathematically challenging.}}
+We believe the technical contribution of this paper is a necessary stepping stone to handle more complex models in follow up research.
 }}
@@
 
-{\red{New plots}}
+{In addition to the above, \ul{we have included the following new figures in the revised manuscript}: Figs.
+8, 10, 13, 14.
+In particular, Fig.
+8 addresses comment \textbf{R2.5}.
+Fig.
+10 addresses comment \textbf{R1.7}.
+Figs.
+13 and 14 address comments \textbf{R1.4, R4.2 and R4.7}.
 }}
@@
 \ul{Our research disposition is that one cannot bypass understanding the generalized gradient flow structure for the simpler second order sample path dynamics before attempting the same for more complex sample path models}.
-In fact, a significant contribution of our article, in the authors' opinion, is that it distills out what exactly is technically challenging and what are the peripheral issues.
 
@@
 
-{\nib{ \blue {\red{TBD}}.}}
+{\nib{ \blue \ul{Regarding the comparison with existing methods}, we have provided comparison with Monte Carlo in the revised manuscript's Figs.
+12, 13 and 14; \ul{among these Figs.
+13 and 14 are new} (in response to comment \textbf{R4.7}).
+Also, the \ul{new Fig.
+8 in the revised manuscript shows the consistency for the rotor angle stochastic predictions w.r.t.
+Monte Carlo for the simulation reported in Sec.
+V-A}, as requested by comment \textbf{R2.5}.
+Please also see our response to comment \textbf{R4.2} where we explain why propagating the joint PDF for even a modest 5 generator system with finite element simulation is challenging with general purpose computers.}}
 
@@
 V-A and V-B are only used for generating example data to perform our illustrative numerical simulations.
-\ul{This again underscores how looking into numerical simulations without contemplative reading and understanding of the earlier Sections may lead to misleading conclusions}.
+%\ul{This again underscores how looking into numerical simulations without contemplative reading and understanding of the earlier Sections may lead to misleading conclusions}.
 
@@
 
-{\nib{ \blue {\red{TBD.}}}}
+{\nib{ \blue Please note that our joint PDF computation has no truncation error accumulation issue (as in for example, moment based propagation techniques where such errors arise from a priori assuming finite dimensionality of the sufficient statistic).
+Also, the new Fig.
+10 reports the total computational time for 1 minute time horizon.
+Please also see the additional error statistics included in the revised version's Fig.
+13 and Fig.
+14.}}
 
@@
 
-{\nib{ \blue While the depiction of the per-step computational times in Figs.
-8 and 9 in the revised manuscript are somewhat unconventional, we explain why we choose to plot that way.
-\ul{To aid clarity, in the revised manuscript, we have added an excerpt of the following at the end of Sec.
-V-A.}
+{\nib{ \blue Thanks.
+\ul{We have included new Fig.
+10 which reports the total computational time for 1 minute time horizon} for all the three simulation cases in Sec.
+V-A and V-B.
+For each of these cases, Fig.
+10 shows the total computational time across 100 simulation instances.
+This Fig.
+10 is referred to in text in Sec.
+V-A (last paragraph), and also in Sec.
+V-B (third paragraph, last sentence).
 
-At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1.
-This proximal update has no analytical solution and instead requires execution of a contractive fixed point recursion.
-The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update.
-\ul{To claim that the proposed method is practical, we therefore, need to demonstrate that the computational timescale is faster than the physical timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time}.
-On the other hand, theoretical consistency (see the unnumbered equation after (26), before Sec.
-IV) requires that $h$ itself should be small.
-Thus, even if one has a provably convergent algorithm that can perform the one step proximal update, it is not clear \emph{a priori} if the computational time needed to do so can be much faster than $h$.
-If not, then the method would be statistically consistent but practically not useful.
-The Figs.
-8 and 9 in the revised manuscript show that the computational time needed to perform the proximal updates, for all $t_{k}$, are indeed orders-of-magnitude faster than $h$.}}
+While showing the per-step computational times in Figs.
+9 and 11 in the revised manuscript are somewhat unconventional, in the revised manuscript's Sec.
+V-A end, \ul{we have added detailed explanation (highlighted in blue)} of why such depictions are meaningful in our computation.}}
 
@@
 It is upon the reader to parse that the two high probability regions on the left and right of that subplot are actually referring to the same angular high probability region around $\theta_{2}=0$ rad, thanks to the periodic boundary condition on the angular variable.
-Our philosophy of technical communication is that the onus is on the authors, not on the readers, to clarify the scientific information as correctly as possible.
+%Our philosophy of technical communication is that the onus is on the authors, not on the readers, to clarify the scientific information as correctly as possible.
 %We are aware that many papers in the power system literature make plots in the rectangular domain ignoring the manifold nature of the support, which in our experience, continue to add pain to the graduate students and beginning researchers trying to decipher the statistical meaning conveyed in those plots.
@@
 
-{\nib \blue{\ul{Regarding the comparison with existing methods}, we have provided comparison with Monte Carlo in the revised manuscript's Fig.
-11 {\red{and another??
-(in response to comment \textbf{R4.7})}}.
-The new Fig.
-8 in the revised manuscript also shows the consistency for the rotor angle stochastic predictions w.r.t.
+{\nib \blue{\ul{Regarding the comparison with existing methods}, we have provided comparison with Monte Carlo in the revised manuscript's Figs.
+12, 13 and 14; \ul{among these Figs.
+13 and 14 are new} (in response to comment \textbf{R4.7}).
+Also, the \ul{new Fig.
+8 in the revised manuscript shows the consistency for the rotor angle stochastic predictions w.r.t.
 Monte Carlo for the simulation reported in Sec.
-V-A, as requested by comment \textbf{R2.5}.
+V-A}, as requested by comment \textbf{R2.5}.
+
 Please note that performing finite element simulation for 5-generator system is a very tall order since the state space then is 10 dimensional, and even if we do a very coarse 10 discretizations per dimension, then such a simulation would require us to evolve a joint PDF over $10^{10}$ grid points.
@@
 
-\ul{Regarding the computational time for longer horizon}, {\red{TBD}}}}
+\ul{Regarding the computational time for longer horizon}, we have included new Fig.
+10 which reports the total computational time for 1 minute time horizon for all the three simulation cases in Sec.
+V-A and V-B.
+For each of these cases, Fig.
+10 shows the total computational time across 100 simulation instances.
+This Fig.
+10 is referred to in text in Sec.
+V-A (last paragraph), and also in Sec.
+V-B (third paragraph, last sentence).}}
 
@@
 
-{\nib \blue{ {\red{TBD.}}}}
+{\nib \blue{\ul{In the revised manuscript, we have incorporated new Fig.
+13 and Fig.
+14}.
+Fig.
+13 shows the relative error between the covariances of the Monte Carlo and proximal joint state PDFs, as measured w.r.t.
+the normalized Bures-Wasserstein distance.
+\ul{We have explained the same in text in the last but one paragraph of Sec.
+V-B (highlighted in blue)}.
+Fig.
+14 shows the absolute error between the Monte Carlo and proximal joint state PDFs, as measured w.r.t.
+the 2-Wasserstein metric.
+\ul{We have explained the same in text in the last paragraph of Sec.
+V-B (highlighted in blue, with added inline citations)}.}}
 

### 2022-03-09 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
polishing rebuttal round 1

@@
 The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update.
-To claim that the proposed method is practical, we therefore, need to demonstrate that the computational timescale is faster than the physical timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time.
+For the proposed method to be practical, we therefore, need to demonstrate that the computational timescale is faster than the physical dynamics timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time.
 On the other hand, theoretical consistency (see the last paragraph of Sec.
@@
 The Fig \ref{fig:CompTimeIEEE14bus} shows that the computational time needed to perform the proximal updates, for all $t_{k}$, are indeed orders-of-magnitude faster than $h$.
+
+Fig.
+\ref{fig:TotalCompTime} (circle markers for the \textbf{Case I}, diamond markers for the \textbf{Case II}) shows the total computational times needed for propagating the transient joint state PDFs over a longer horizon $[0, 1\;\text{min}]$ for 100 different simulation instances.
 }}
+
+\begin{figure}[htpb]
+\centering
+\includegraphics[width=0.95\linewidth]{TotalCompTimeForSimulating1MinHist.png}
+\caption{\small{The total computational times for propagating the transient joint state PDFs over $[0, 1\;\text{min}]$ for the simulation set up in Sec.
+\ref{subsec:NumSimIEEEBus} and Sec.
+\ref{subsec:NumSimSynthetic}.
+In each case, we show the total computational time for 100 different simulation instances.}}
+\vspace*{-0.1in}
+\label{fig:TotalCompTime}
+\end{figure}
 
@@
 Notice that constructing histograms or other conventional density estimators for the joint state PDF by making a grid over the 100 dimensional state space, is computationally prohibitive in a general purpose computer.
+{\blue{Fig.
+\ref{fig:TotalCompTime} (square markers) shows the total computational times needed for propagating the transient joint state PDFs over the time horizon $[0, 1\;\text{min}]$ for 100 different simulation instances.}}
 

### 2022-03-09 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py

@@ -873,57 +873,26 @@ Time_Synthetic = np.loadtxt("TimeSyntheticIEEE14bus.txt")
 # PLOT hist of loger horizon total computational time
 ################################################################################ 
 
-# alpha_param = np.loadtxt('alpha.txt')
-# beta_param = np.loadtxt('beta.txt')
+TotalComptime_IEEE14busCaseI = np.loadtxt('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/code/total_comptime_IEEE14bus_case0_nominal.txt')
+TotalComptime_IEEE14busCaseII = np.loadtxt('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/code/total_comptime_IEEE14bus_case1_line_13_failure.txt')
+TotalComptime_50gen = np.loadtxt('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/code/total_comptime_50gen.txt')
 
-# fig, (ax1, ax2) = plt.subplots(1,2, sharey=True)
 
-# wt_alpha = np.ones_like(alpha_param)/float(len(alpha_param))
-# ax1.hist(alpha_param, 50, weights=wt_alpha, normed=0, facecolor='dodgerblue', alpha=0.6)
-# ax1.get_xaxis().tick_bottom()
-# ax1.get_yaxis().tick_left()
-# ax1.tick_params(axis='x', direction='out')
-# ax1.yaxis.set_ticklabels([])
-# ax1.spines['top'].set_visible(False)
-# ax1.spines['right'].set_visible(False)
-# ax1.tick_params(
-#     axis='x',          # changes apply to the x-axis
-#     which='both',      # both major and minor ticks are affected
-#     bottom='on',      # ticks along the bottom edge are off
-#     top='off',         # ticks along the top edge are off
-#     labelbottom='off') # labels along the bottom edge are off
-# ax1.tick_params(
-#     axis='y',          # changes apply to the x-axis
-#     which='both',      # both major and minor ticks are affected
-#     left='on',      # ticks along the bottom edge are off
-#     right='off',         # ticks along the top edge are off
-#     labelbottom='off') # labels along the bottom edge are off
-# ax1.spines['bottom'].set_position(('outward', 5))
+fig = plt.figure()
+ax = fig.add_subplot(1, 1, 1)
[…]

### 2022-03-09 code/Gen_50_prox_simulation.m

@@ -30,10 +30,11 @@ K = K - diag(diag(K)); % zero the diagonal entries
 theta0_a = 0 ; theta0_b = 2*pi; omega0_a = -.1 ; omega0_b = .1;
 
 % parameters for proximal recursion
-nSample = 1000;                      % number of samples                                                           
+nSample = 2000;                      % number of samples                                                           
 epsilon = 0.5;                      % regularizing coefficient                                      
 h = 1e-3;                         % time step
-numSteps= 1000;                    % number of steps k, in discretization t=kh
+tfinal = 1;                        % final time
+numSteps= tfinal/h;                    % number of steps k, in discretization t=kh
 cc = 1e7;
 t_vec = h*(1:1:numSteps);
 %% propagate joint PDF
@@ -95,8 +96,10 @@ mean_prox_omega(1,:) = sum(omega_upd(:,:,1).*rho_theta_omega_upd(:,1))/sum(rho_t
 mean_prox_theta(1,:) = weighted_angle_mean(theta_upd(:,:,1),rho_theta_omega_upd(:,1));
 
 numRun = 100;
+toc_50gen = zeros(numRun,1);
 
 for r=1:numRun
+disp(['Now doing run ' num2str(r)])
 tic;
 for j=1:numSteps 
     
@@ -112,16 +115,17 @@ for j=1:numSteps
     theta_omega_upd(:,:,j+1) = [theta_upd(:,:,j+1), omega_upd(:,:,j+1)];
 
    % proximal update for joint PDF
-   [rho_xi_eta_upd(:,j+1),comptime(j),niter(j)] = FixedPointIteration(beta,epsilon,h,rho_xi_eta_upd(:,j),xi_eta_upd(:,:,j),xi_eta_upd(:,:,j+1),PowerFraction(beta,xi_eta_upd(:,num_Oscillator+1:dim,j)),GradU,dim);  
+   [rho_xi_eta_upd(:,j+1),comptime(r,j),niter(j)] = FixedPointIteration(beta,epsilon,h,rho_xi_eta_upd(:,j),xi_eta_upd(:,:,j),xi_eta_upd(:,:,j+1),PowerFraction(beta,xi_eta_upd(:,num_Oscillator+1:dim,j)),GradU,dim);  
     
[…]

### 2022-03-09 code/ThreeCases_IEEE14_prox_simulation.m

@@ -17,12 +17,12 @@ bus_init = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertain
 % case2_time_series_
 
 % parameters for the IEEE 14 bus system
-Y_reduced_imag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_Y_reduced_imag_part.csv');
-Y_reduced_real = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_Y_reduced_real_part.csv');
+Y_reduced_imag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_Y_reduced_imag_part.csv');
+Y_reduced_real = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_Y_reduced_real_part.csv');
 % reduced admittance matrix
 Y_reduced = complex(Y_reduced_real,Y_reduced_imag);
 
-gen_param = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_gen_parameters.csv');
+gen_param = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_gen_parameters.csv');
 gen_bus_idx = gen_param(:,1);
 
 num_Oscillator = numel(gen_bus_idx); dim  = 2*num_Oscillator;
@@ -35,17 +35,17 @@ E = E_mag.*exp(1i*E_phase);
 %I_reduced_polar = readmatrix('I_red.csv');
 %I_reduced_mag = I_reduced_polar(:,1);
 %I_reduced_phase = I_reduced_polar(:,2);
-I_reduced_mag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case1_line_13_failure_I_red_magnitude.csv');
+I_reduced_mag = readmatrix('/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/Data/case0_norminal_I_red_magnitude.csv');
[…]

### 2022-03-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
polishing rebuttal round 1

@@
 \ul{Furthermore, Fig.
-7 in the revised manuscript has also been expanded to include the univariate angular velocity transient marginal plots for all the five generators of the IEEE 14 bus system; previous version had this plot only for the generator 4 (buss 6)}.
+7 in the revised manuscript has also been expanded to include the univariate angular velocity transient marginal plots for all the five generators of the IEEE 14 bus system; previous version had this plot only for the generator 4 (bus 6)}.
 

### 2022-03-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
polishing rebuttal round 1

@@
 
-(2) \textbf{Regarding randomness in active and reactive power:} {\red{TBD}}
+(2) \textbf{Regarding randomness in active and reactive power:} Please note that the randomness mentioned by the reviewer is not ignored.
+In the simulation, their effects are aggregated as parametric uncertainty in the randomness of $P_{i}$.
+Also notice that to say this parametric uncertainty in $P_i$ manifests in the Kron reduced form, is equivalent to assuming first order stochastic dynamic uncertainties in the unreduced load dynamics.
+If one has an explicit (possibly higher order) stochastic load dynamics model, then the proposed framework applies as is by first computing the law of the stochastic process $P_{i}$ in the network reduced model, and then applying the proximal computation.
 
@@
 V-A}, as requested by comment \textbf{R2.5}.
-Please also see our response to comment \textbf{R4.2} where we explain why propagating the joint PDF for even a modest 5 generator system with finite element simulation is challenging with general purpose computers.}}
+Please also see our response to comment \textbf{R4.2} where we explain why propagating the joint PDF for even a modest 5 generator system with discretization-based PDE solvers such as finite element simulation is challenging with general purpose computers.}}
 
@@
 {\nib{ \blue Please note that our joint PDF computation has no truncation error accumulation issue (as in for example, moment based propagation techniques where such errors arise from a priori assuming finite dimensionality of the sufficient statistic).
-Also, the new Fig.
-10 reports the total computational time for 1 minute time horizon.
+\ul{Furthermore, the new Fig.
+10 reports the total computational time for 1 minute time horizon}.
 Please also see the additional error statistics included in the revised version's Fig.
@@
 \ul{In fact, all stochastic uncertainties are considered to be ``joint", i.e., statistically correlated}.
-Please note that the initial joint PDFs (28) and (30) are taken to be in that form for the sake of concrete example simulations.
+Please note that the \ul{initial joint PDFs} (29) and (31) are taken to be in that form for the sake of concrete example simulations.
 Their only purpose, in the simulations reported in Sec V, is to generate joint samples, and to evaluate the initial joint PDF values at the joint samples thus generated.
@@
 {\nib \blue{Please note that the synthetic system in Sec.
-V-B is actually simulating a higher dimensional system than the IEEE 118-bus system where the 19 generators would have resulted in the transient joint PDFs with 38 dimensional supports.
+V-B is actually simulating a much higher dimensional system than the IEEE 118-bus system where the 19 generators would have resulted in the transient joint PDFs with 38 dimensional supports.
 In comparison, the synthetic system is evolving joint state PDFs having 100 dimensional supports, i.e., in $\mathbb{T}^{50}\times\mathbb{R}^{50}$.

### 2022-03-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex
polishing rebuttal round 1

@@
 
-\emph{First}, \ul{in our opinion, this kind of plot is scientifically misleading as they hide the fact that these bivariate marginal PDFs are not really supported on the plane, or the subsets thereof, but instead have cylindrical supports}.
+\emph{First}, \ul{this kind of plot is scientifically misleading as they hide the fact that these bivariate marginal PDFs are not really supported on the plane, or the subsets thereof, but instead have cylindrical supports}.
 Our message that the proposed method explicitly accounts this manifold geometry in the computation, would be missed had we chosen to make the plots the other way.%as per populist cognitive bias on behalf of the readers.
@@
 
-Please note that performing finite element simulation for 5-generator system is a very tall order since the state space then is 10 dimensional, and even if we do a very coarse 10 discretizations per dimension, then such a simulation would require us to evolve a joint PDF over $10^{10}$ grid points.
+\ul{Please note that performing finite element simulation for 5-generator system is a very tall order since the state space then is 10 dimensional, and even if we do a very coarse 10 discretizations per dimension, then such a simulation would require us to evolve a joint PDF over $10^{10}$ grid points}.
 This requires dedicated high-performance computing architecture and an undertaking well beyond the scope of this paper.

### 2022-08-12 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound2-TPWRS-01391-2021R1.tex
Rebuttal round 2

\documentclass[12pt,onecolumn]{IEEEtran}
\usepackage{amsmath ,amssymb,euscript ,yfonts,psfrag,latexsym,dsfont,graphicx,bbm,color,amstext,wasysym,subfig,flushend,parskip}
\usepackage{bm,cite,soul,url}
\usepackage{verbatim}
\usepackage[normalem]{ulem}
\usepackage{amsthm}
%\usepackage{refcheck}
%\usepackage{showkeys}
%\usepackage{epstopdf,hyperref,pdfsync,url}
\graphicspath{{./},{./figures/}}
\newcommand{\mR}{{\mathbb R}}
\newcommand{\D}{{\mathbb D}}
\newcommand{\E}{{\mathbb E}}
\newcommand{\mcN}{{\mathcal N}}
\newcommand{\mcR}{{\mathcal R}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\tr}{\operatorname{trace}}
\newcommand{\ignore}[1]{}
\newcommand{\mike}{\color{magenta}}

\newcommand{\blue}{\color{blue}}
\newcommand{\red}{\textcolor{red}}

\newcommand{\bQ}{{\bm{Q}}}
\newcommand{\bPi}{\bm{\Pi}}
\newcommand{\bv}{{\bm{v}}}
\newcommand{\balpha}{{\bm{\alpha}}}
\newcommand{\bbeta}{{\bm{\beta}}}
\newcommand{\bq}{{\bm{q}}}
\newcommand{\bp}{{\bm{p}}}
\newcommand{\bbf}{{\bm{f}}}
\newcommand{\differential}{{\mathrm{d}}}

\newtheorem{claim}{Claim}

\def\spacingset#1{\def\baselinestretch{#1}\small\normalsize}
\setlength{\parskip}{10pt}
\setlength{\parindent}{20pt}
\spacingset{1}

\definecolor{grey}{rgb}{0.6,0.6,0.6}
\definecolor{lg}{rgb}{0.6,.6,0.6}
\newcommand{\nib}{\noindent  {\bf Response:} }

\begin{document}
%\newcommand{\bp}{{\bm{p}}}
\noindent
{\large {\bf Revision of
Manuscript Number}: TPWRS-01391-2021.R1\hfill \today\\[.1in]
{\bf Journal}: IEEE Transactions on Power Systems \\[.1in]
{\bf Author}: Abhishek Halder, Kenneth F. Caluya, Pegah Ojaghi, Xinbo Geng}\\

\section*{\large \bf Response to decision:}

{\noindent\blue
The perceptive input from all reviewers and the AE are much appreciated.

Please find below itemized responses to the individual comments addressing the concerns and detailing the revisions. All the edits in the revised manuscript are marked in blue.\\

\noindent
Sincerely,\\
The authors
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\newpage
\spacingset{1}

\section*{\large \bf Response to AE:}

\noindent
{\bf AE.1. Comments by AE}:\\
{\em Authors are requested to revise and resubmit the manuscript after addressing all reviewer comments that persist.
}

{\nib {\blue We. 
}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\spacingset{1}

\section*{\large \bf Response to Reviewer \#1:}

\noindent
{\bf R1.1. Comment by reviewer}:\\
{\em No further comments.}

{\nib {\blue Thank you for your time and efforts to help improve the manuscript.}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{\large \bf Response to Reviewer \#2:}

\noindent
{\bf R2.1. Comments by reviewer}:\\
{\em This reviewer appreciates the authors' effort in revising the manuscript based on the feedback in the previous round while clarifying and addressing the remarks with due diligence. The revised version reads well and adequately takes into account several relevant facets. The authors are kindly requested to proofread the article prior to final submission.}

{\nib \blue{Thank you, we have thoroughly proofread the article. We appreciate your comments from previous rounds to help us improve the manuscript.}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section*{\large \bf Response to Reviewer \#3:}

\noindent
{\bf R3.1. Comments by reviewer}:\\
{\em First and very importantly, the authors are suggested to readdress my previous two questions.\\
\noindent(1) Regarding my previous question about the classical model, I need to reemphasize here that the model is too simplistic that does not sufficiently reflect the responses from the exciter and the governor. Therefore, for a full paper on IEEE transactions on power systems, this model is incomplete, and cannot reflect reality. Also, for such an over-simplistic model, I still doubt the scalability of the method. Unfortunately, this problem is not addressed and improved after the revision.\\
\noindent(2) Regarding my previous questions about the comparison studies with the state-of-the-art method, indeed, I am very disappointed to find that the revised version still has added no comparison study with the state-of-the-art method at all. The ONLY comparison is still with Monte Carlo.}

{\nib \blue{The 
}}

\noindent
{\bf R3.2. Comments by reviewer}:\\
{\em Second and most importantly, I disagree with the theory and the claim behind the SDE model used in this paper.1.     Here, I am very familiar with the Ito SDE model in power systems that you cited here in your refs [16], [17], [45].

Although the authors claim the model you used here in equation (7) is the same as the ones used in ref[16] and ref[17] and extended from ref[45], I would like to tell the authors: NO, you are not using the same one. 16][17][45] are all right, but the one used in this article is incorrect. Please pay very special attention to [16][17], they only add uncertainties directly on differential equations, namely wi here, but not from P and Q.  This is because transferring the power system model from DAE to Ito-SDE is still an unsolved problem where P and is contained in algebraic equations while Ito-SDE only considers differential equations. Also, [45] is one machine system that does not even have a network. That is different from your problem. Here, the authors might argue that you are using the Kron reductions to remove algebraic equations. However, I have to tell the authors the Kron reductions to apply to the SDE model here. It is because the Kron reduction is relying on constant P and Q to derive a constant admittance matrix.  However, the P and Q in this manuscript are also assumed to be uncertain, and also claimed in the response letter to be time-varying stochastic processes. Then, its assumptions do not hold.  Then, its Ito SDE form is inconsistent.  I highly suggest to the authors study the following article carefully and learn more about the basics of Kron reduction.

Time-domain generalization of Kron Reduction, MK Singh, S Dhople, F Dorfler, GB Giannakis.
}

{\nib \blue{\ul{The.}}
}

\noindent
{\bf R3.3. Comments by reviewer}:\\
{\em The simulation results are weird. Just a simple check at fig 8. How a power system dynamic simulation with line 13 tripped can has no oscillations at all within the first 1 second?  If so, it has no practical meaning at all.}

{\nib \blue{The.}}

\noindent
{\bf R3.4. Comments by reviewer}:\\
{\em Third, my engineering motivation for me is not clear. The authors reclaim multiple times about the joint pdf for the output system state.  I understand that we consider joint pdf in the model inputs and understand its practical meaning. However, among the current literature, to my best knowledge, I saw no previous efforts in obtaining the ``joint" pdf of outputs. They only seed for marginal pdf for output states. So an important question is what motivates us to get joint pdf, instead of individual pdf, for model output? For which exactly applications, you must know the joint pdf of the output states?  What is its practical meaning and why it is important?}

{\nib \blue{Thanks for this comment. The}}

\noindent
{\bf R3.5. Comments by reviewer}:\\
{\em Also, the authors should put a very clear position about some of the references that you are making use to argue with the reviewer. For example,  [45] is a 10-year ago paper. Why do you claim it to the state-of-the-art?  How can this argument convince the reviewer?}

{\nib \blue{The }}

\noindent
{\bf R3.5. Comments by reviewer}:\\
{\em Regarding Ref [23]F. Milano, and R. Zarate-Minano, ``A systematic method to model power systems as stochastic differential-algebraic equations", IEEE transactions on Power Systems, Vol. 28, No. 4, pp. 4537--4544, 2013. I mentioned this article in the first-round review, not for the purpose of a simple cite, but trying to help you to improve the model and make your work realistic. However, a suggestion is ignored. This is indeed a pity. Besides, please do not argue again that this is a different work because this ref is well-known in modeling realistic stochastic power system dynamics, which should be the foundation of your model.}

{\nib \blue{The }}

\noindent
{\bf R3.6. Comments by reviewer}:\\
{\em In general, I need to claim here that math is fine in itself, but only publishable when connected to or interpreted correctly in a convincing engineering application.}

{\nib \blue{The }}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section*{\large \bf Response to Reviewer \#4:}

\noindent
{\bf R4.1. Comments by reviewer}:\\
{\em The authors have successfully addressed most of the reviewer's previous concerns. However, the reviewer still has some questions, mostly regarding the numerical simulations.}

{\nib \blue{Thank you. Please find our itemized responses to the questions related to numerical simulations.}}

\noindent
{\bf R4.2. Comments by reviewer}:\\
{\em It is stated that ``The Fig 9 shows that the computational time needed to perform the proximal updates, for all $t_k$, are indeed orders-of-magnitude faster than $h$."
This statement doesn't seem to hold, since $h=10^{-3}$ and the computational time lies within the range of around $10^{-3}$ to $10^{-2}$. How does the computational time needed to perform the proximal updates relate to time step $h$? If they are not strongly related, does that mean the step size $h$ is constrained to be greater than that computational time?}

{\nib \blue{The 
}}

\noindent
{\bf R4.3. Comments by reviewer}:\\
{\em In Fig. 10, could the authors please explain why the computational time of Case I is less than that of Case II?
The computational cost for the 5-generator system is roughly 4 times that of Case II. Could the authors comment if such a scale makes sense intuitively considering the state dimension of the two systems?}

{\nib \blue{Thanks for this comment. The ...}}

\noindent
{\bf R4.4. Comments by reviewer}:\\
{\em Both Fig. 12 and Fig. 13 demonstrate a decreasing trend as k increases. Could the authors comment on this and link this to the property of the proposed proximal algorithm? Similarly, why is the trend in Fig. 14? Comparing Figs. 12, 13 and Fig. 14, did the authors omit the initial time instant in Figs. 12, 13?}

{\nib \blue{Thanks for ...}}

\end{document}

### 2022-08-12 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev2.tex
Rebuttal round 2

--- 2022-03-09 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev1.tex
@@
 However, this leads to moment closure problems since the dimensions of the time-varying sufficient statistics associated with the corresponding transient joint state PDFs are not known in general.
-{\blue{We also note the usage of \emph{stochastic} differential algebraic equations (DAEs) for power system dynamic simulation \cite{wang2011numerical,milano2013systematic}.
-These studies, however, simulate the sample paths and do not directly propagate the joint state PDFs.}}
+We also note the usage of \emph{stochastic} differential algebraic equations (DAEs) for power system dynamic simulation \cite{wang2011numerical,milano2013systematic}.
+These studies, however, simulate the sample paths and do not directly propagate the joint state PDFs.
 
@@
 \end{align}
-We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in {\blue{terms}} of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous.
+We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in terms of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous.
 Fig.
@@
 A point cloud-based proximal algorithm was proposed in \cite{caluya2019ACC,caluya2019TAC} which was reported to have very fast runtime.
-{\blue{The main idea in these references was to \emph{co-evolve} the time-varying state samples (via SDE discretization schemes such as the Euler-Maruyama scheme) as well as the joint PDF values (via \eqref{ProxRecursionInfiniteDim}) evaluated at those samples.
+The main idea in these references was to \emph{co-evolve} the time-varying state samples (via SDE discretization schemes such as the Euler-Maruyama scheme) as well as the joint PDF values (via \eqref{ProxRecursionInfiniteDim}) evaluated at those samples.
 The resulting computation is an \emph{online} propagation of the joint PDFs as opposed to the \emph{offline} computation in Monte Carlo or density estimation methods \cite{silverman1998density}.
-The latter methods only propagate the state samples and then approximate the joint PDFs as post-processing.}}
+The latter methods only propagate the state samples and then approximate the joint PDFs as post-processing.
 
@@
 \end{align}
-where {\blue{the diagonal matrix $\bm{\Upsilon}:= \left(\prod_{i=1}^{n}\sigma_{i}^{2}m_{i}^{-2}\right)\bm{M\Sigma}^{-2}$ has positive entries along the main diagonal, and}} the potentials
+where the diagonal matrix $\bm{\Upsilon}:= \left(\prod_{i=1}^{n}\sigma_{i}^{2}m_{i}^{-2}\right)\bm{M\Sigma}^{-2}$ has positive entries along the main diagonal, and the potentials
 {\small{\begin{subequations}
@@
 \begin{align}
-\Phi(\tilde{\rho}) &:= \!\!\displaystyle\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)}}\times\mathbb{R}^{n}}\!\!\!\left(\!U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\!\right)\nonumber\\
+\Phi(\tilde{\rho}) &:= \!\!\displaystyle\int_{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)\times\mathbb{R}^{n}}\!\!\!\left(\!U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\!\right)\nonumber\\
 &\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta},
@@
 \begin{align}
-\widetilde{\Phi}(\tilde{\rho}) := \!\!\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)}} \times \mathbb{R}^{n}}\!\!\left(\!F(\boldsymbol{\eta})+\frac{1}{2} \log \tilde{\rho}\!\right) \tilde{\rho}\: \mathrm{d} \boldsymbol{\xi} \mathrm{d} \boldsymbol{\eta},
+\widetilde{\Phi}(\tilde{\rho}) := \!\!\int_{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right) \times \mathbb{R}^{n}}\!\!\left(\!F(\boldsymbol{\eta})+\frac{1}{2} \log \tilde{\rho}\!\right) \tilde{\rho}\: \mathrm{d} \boldsymbol{\xi} \mathrm{d} \boldsymbol{\eta},
 \label{Phitilde}	
@@
 \begin{align}
-&\widetilde{W}_{h,\bm{\Upsilon}}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) := \!\underset{\pi\in\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)}{\inf}\!\displaystyle\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)^{2}}} \times \mathbb{R}^{2n}}\nonumber\\
+&\widetilde{W}_{h,\bm{\Upsilon}}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) := \!\underset{\pi\in\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)}{\inf}\!\displaystyle\int_{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)^{2} \times \mathbb{R}^{2n}}\nonumber\\
 &\qquad\qquad\qquad\qquad s_{h,\bm{\Upsilon}}\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right)\differential\pi\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right),
@@
 \end{align}
-wherein $h>0$ denotes the step-size in proximal recursion, and $\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)$ denotes the set of joint probability measures over the product space ${\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)^{2}}} \times \mathbb{R}^{2n}$ that have finite second moments  with the first marginal $\tilde{\varrho}$, and the second marginal $\tilde{\varrho}_{k-1}$.
+wherein $h>0$ denotes the step-size in proximal recursion, and $\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)$ denotes the set of joint probability measures over the product space $\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)^{2} \times \mathbb{R}^{2n}$ that have finite second moments  with the first marginal $\tilde{\varrho}$, and the second marginal $\tilde{\varrho}_{k-1}$.
 The ``ground cost" $s_{h,\bm{\Upsilon}}$ in \eqref{Wtilde} is given by
@@
 converges to the flow generated by \eqref{TransformedPDEivp}, i.e., 
-$$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} {\blue{\tilde{\rho}}}(t=kh,\bm{\xi},\bm{\eta})\;\text{in}\;L^{1}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right),$$
+$$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} \tilde{\rho}(t=kh,\bm{\xi},\bm{\eta})\;\text{in}\;L^{1}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right),$$
 can be established following the arguments in \cite{duong2014conservative}.

### 2022-08-12 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev2.tex
Rebuttal round 2

@@
 The proposed computational framework is also applicable when $P_{i}^{\rm{mech}}, P_{i}^{\rm{load}}$, and thus $P_{i}$ in \eqref{defPi}, are bounded time-varying functions.
-This can be the case, for instance, with rapid fluctuations from renewables {\blue{or from loads}} in the short term.
+This can be the case, for instance, with rapid fluctuations from renewables or from loads in the short term.
 \end{remark}
@@
 
-\noindent\textbf{Case I.} \textbf{Nominal case} where the $P_{i}^{\text{mech}}$ and $P_{i}^{\text{load}}$ are from the steady state power flow solutions without the process noise, and {\blue{are assumed}} to remain constant over the simulation time horizon.
+\noindent\textbf{Case I.} \textbf{Nominal case} where the $P_{i}^{\text{mech}}$ and $P_{i}^{\text{load}}$ are from the steady state power flow solutions without the process noise, and are assumed to remain constant over the simulation time horizon.
 The transient stochastic simulations (i.e., with process noise) are then performed using the proposed framework.
@@
 Using the algorithm detailed in Sec.
-\ref{sec:ProxAlgo}, for both cases, we propagated $N=1000$ random {\blue{state}} samples from \eqref{rho0IEEE14}, and {\blue{corresponding joint PDF values}} evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
+\ref{sec:ProxAlgo}, for both cases, we propagated $N=1000$ random state samples from \eqref{rho0IEEE14}, and corresponding joint PDF values evaluated at \eqref{rho0IEEE14}, to update the scattered weighted point clouds $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$ over $\mathbb{T}^{5}\times\mathbb{R}^{5}$ at times $t_{k}:=kh$, $k\in\mathbb{N}$.
 In Algorithm \ref{algo:KineticProx}, we used the algorithmic parameters $\varepsilon=0.05$, $\delta=10^{-3}$, $\ell_{\max}=300$.
@@
 \ref{fig:IEEE14marginalsCase1}.
-{\blue{The statistical uncertainties in the rotor angles are shown in Fig.
-\ref{fig:IEEE14theta} for both \textbf{Case I} and \textbf{Case II}.}}
+The statistical uncertainties in the rotor angles are shown in Fig.
+\ref{fig:IEEE14theta} for both \textbf{Case I} and \textbf{Case II}.
 
@@
 \ref{sec:ProxAlgo}, for both \textbf{Cases I} and \textbf{II}.
-{\blue{While the depiction of the computational times in Fig.
+While the depiction of the computational times in Fig.
 \ref{fig:CompTimeIEEE14bus} appears unconventional, we next explain why that is meaningful in our framework.
@@
 \ref{fig:TotalCompTime} (circle markers for the \textbf{Case I}, diamond markers for the \textbf{Case II}) shows the total computational times needed for propagating the transient joint state PDFs over a longer horizon $[0, 1\;\text{min}]$ for 100 different simulation instances.
-}}
 
@@
 Notice that constructing histograms or other conventional density estimators for the joint state PDF by making a grid over the 100 dimensional state space, is computationally prohibitive in a general purpose computer.
-{\blue{Fig.
-\ref{fig:TotalCompTime} (square markers) shows the total computational times needed for propagating the transient joint state PDFs over the time horizon $[0, 1\;\text{min}]$ for 100 different simulation instances.}}
+Fig.
+\ref{fig:TotalCompTime} (square markers) shows the total computational times needed for propagating the transient joint state PDFs over the time horizon $[0, 1\;\text{min}]$ for 100 different simulation instances.
 
@@
 The latter was computed using the proximal updates $\{\bm{x}_{k}^{i},\varrho_{k}^{i}\}_{i=1}^{N}$.
-{\blue{Likewise, Fig.
+Likewise, Fig.
 \ref{fig:RelErrCovSynthetic} plots the time-varying normalized Bures-Wasserstein distance $d_{\rm{BW}}(\cdot,\cdot)$ between the empirical covariance matrix (i.e., Monte Carlo estimate) ${\rm{Cov}}^{\rm{MC}}\in\mathbb{R}^{100\times 100}$, and the ensemble covariance matrix ${\rm{Cov}}^{\rm{Prox}}\in\mathbb{R}^{100\times 100}$ obtained using the proximal updates.
@@
 In summary, Figs.
-\ref{fig:RelErrSynthetic} and \ref{fig:RelErrCovSynthetic} show that the time-varying statistics between the Monte Carlo and the proposed proximal computation remain close at all times.}}
-
-{\blue{Since it is not possible to \emph{a priori} determine the dimension of the sufficient statistic of the transient joint state PDFs, Fig.
+\ref{fig:RelErrSynthetic} and \ref{fig:RelErrCovSynthetic} show that the time-varying statistics between the Monte Carlo and the proposed proximal computation remain close at all times.
+
+Since it is not possible to \emph{a priori} determine the dimension of the sufficient statistic of the transient joint state PDFs, Fig.
 \ref{fig:WassJointPDFSynthetic} shows the 2-Wasserstein metric between the empirical joint PDF (i.e., Monte Carlo estimate) $\rho^{\rm{MC}}$, and the proximal update $\varrho^{k}$.
@@
 4.2-4.3]{halder2014probabilistic}, \cite[Ch.
-3]{peyre2019computational}) using the corresponding time-varying weighted point clouds.}}
+3]{peyre2019computational}) using the corresponding time-varying weighted point clouds.
 
@@
 
-{\blue{\bibitem{wang2011numerical}
+\bibitem{wang2011numerical}
 K.~Wang, and M.L.~Crow, ``Numerical simulation of stochastic differential algebraic equations for power system transient stability with random loads", \emph{2011 IEEE Power \& Energy Society General Meeting (PESGM)}, pp.
@@
 4, pp.
-4537--4544, 2013.}}
+4537--4544, 2013.
 
@@
 
-{\blue{\bibitem{geng2020reach}
+\bibitem{geng2020reach}
 S.~Geng, and I.A.~Hiskens, ``Reach-Set Estimation for DAE Systems under Uncertainty and Disturbances Using Trajectory Sensitivity and Logarithmic Norm", \emph{21st IFAC World Congress}, Vol.
@@
 2, pp.
-1955--1961, 2020.}}
+1955--1961, 2020.
 
@@
 \bibitem{silverman1998density}
-{\blue{B.W.~Silverman, \emph{Density Estimation for Statistics and Data Analysis}, 1st ed., Chapman \& Hall/CRC, 1998.}}
+B.W.~Silverman, \emph{Density Estimation for Statistics and Data Analysis}, 1st ed., Chapman \& Hall/CRC, 1998.
 
@@
 
-{\blue{\bibitem{bhatia2019bures}
+\bibitem{bhatia2019bures}
 R.~Bhatia, T.~Jain, and Y.~Lim, ``On the Bures--Wasserstein Distance between Positive Definite Matrices", \emph{Expositiones Mathematicae}, Vol. 37, No. 2, pp. 165--191, 2019.
@@
 355--607, 2019.
-
-}}
 

### 2022-08-12 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound2-TPWRS-01391-2021R1.tex
Rebuttal round 2

@@
 
-{\nib \blue{The 
-}}
+{\nib \blue{The reviewer is right, we should have been more careful in phrasing that sentence.
+We meant to say that the computational time needed to perform a proximal update is \ul{statistically} comparable (in the big O sense) to $h$.
+The reason why the precise computational time is probabilistic is twofold: random sampling, and randomness in computational latency, power and memory footprint.
+Fig.
+9 shows that $h=10^{-3}$ and the computational time needed to perform a single proximal update is $O(10^{-3})$ with high probability in the sense most time-steps incur $O(10^{-3})$ time.
+In Fig.
+9, while instances when computational times are $O(10^{-2})$ do occur for Case I, they are rare.
+\ul{In the present revision, we have now rephrased the corresponding text (highlighted in blue) toward the end of Sec.
+V-A}.
+
+The proximal updates involve certain block coordinate ascent recursions in the dual co-ordinates $\bm{y},\bm{z}$; see lines 13-21 in Algorithm 1.
+This recursion is guaranteed to be contractive (hence existence, uniqueness and convergence are guaranteed) in the product orthant cone: $(\bm{y},\bm{z})\in\mathbb{R}^{N}_{>0}\times \mathbb{R}^{N}_{>0}$, see ref.
+[41, Sec.
+III-B, III-C] and especially Corollary 4 in that reference.
+We have cited that reference in this manuscript's Sec.
+IV.}}
 

### 2022-08-12 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev2.tex
Rebuttal round 2

@@
 The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update.
-For the proposed method to be practical, we therefore, need to demonstrate that the computational timescale is faster than the physical dynamics timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time.
+For the proposed method to be practical, we therefore, need to demonstrate that {\blue{the computational timescale is at least statistically comparable (in the order of magnitude sense) to the physical dynamics timescale.
+In other words, the computational times needed to complete the zero-order-hold fixed point recursions are, with high probability, the same orders-of-magnitude as $h$, the step-size of physical time.
+The precise computational time is probabilistic, not just due to random sampling, but also due to randomness in power and memory footprint.}}
+
+%is faster than the physical dynamics timescale, i.e., computational times needed to complete the zero-order-hold fixed point recursions are orders-of-magnitude faster than $h$, the step-size of physical time.
 On the other hand, theoretical consistency (see the last paragraph of Sec.
@@
 If not, then the method would be statistically consistent but practically not useful.
-The Fig \ref{fig:CompTimeIEEE14bus} shows that the computational time needed to perform the proximal updates, for all $t_{k}$, are indeed orders-of-magnitude faster than $h$.
+
+{\blue{In Fig \ref{fig:CompTimeIEEE14bus}, $h=10^{-3}$ and the computational time needed to perform a single proximal update, \emph{with high probability}, is $O(10^{-3})$ in the sense most time-steps incur $O(10^{-3})$ time.
+While instances when computational times are $O(10^{-2})$ do occur for Case I in Fig \ref{fig:CompTimeIEEE14bus}, but are rare.}}
 

### 2022-08-13 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound2-TPWRS-01391-2021R1.tex
Rebuttal round 2

@@
 We have cited that reference in this manuscript's Sec.
-IV.}}
+IV.
+\ul{While $h$ appears as parameter in this recursion (lines 5 and 14 in Algorithm 1), we do not have analytical relation between the computational time and $h$}.
+We did not constrain or adapt $h$ in Algorithm 1.
+We simply set a small value $h=10^{-3}$ to promote statistical consistency (see the unnumbered equation after (27)).
+Depending on the random sampling, dynamics and computational platform, it is possible for the single-step proximal update computational time to fall below $h$ (see e.g., ref.
+[41, Fig.
+9] which used the same $h=10^{-3}$).
+\ul{The standard desktop computer/processor we used for our simulation is reported in Sec.
+V, first paragraph, last sentence}.}}
 
@@
 
-{\nib \blue{Thanks for this comment. The ...}}
+{\nib \blue{Thanks for this comment.
+The state dimensions for both Cases I and II in Fig.
+10 are the same, i.e., $5\times2 = 10$ dimensions.
+Other than the difference in dynamics due to line failure in Case II, we do not think there is any specific physical insight to be gleaned on why the computational times for Case II are below Case I; they both use identical set of random initial samples although the Wiener process noise realizations are different.
+
+Regarding the scaling comment, perhaps the reviewer meant to say that the total computational times for the 50 generator (100 state) system are roughly 4 times that of Case II (5 generator, 10 state).
+The possible explanation of why the scaling is not so bad is because the number of samples for the 50 generator case is $2000$, which is twice the number of samples used in Cases I and II.
+This is made possible by our usage of scattered point clouds with explicit proximal computation of the time-varying probability weights, as opposed to gridding the state space and then approximating the joint PDF via histogram (which would have caused exponential in dimension scaling in the computational cost).
+Being able to directly evaluate the joint PDF values via proximal computation allows us to bypass construction of function approximation oracle over the high dimensional state space (which in general scales poorly with dimensions), and then to evaluate it.
+A direct comparison between the two cases is less meaningful due to difference in parameter values, initial joint PDF etc.
+\ul{However, the general trend that the proposed scattered point cloud-type proximal framework, similar to the one proposed here, can incur computational times which scale sublinearly with dimensions while resolving the joint PDF well, were reported in our earlier works in different application contexts: ref.
+[41, Sec.
+V-B and Fig.
+14], ref.
+[42, Fig.
+5]}.
+
+}}
 

### 2022-08-13 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev2.tex
Rebuttal round 2

@@
 In other words, the computational times needed to complete the zero-order-hold fixed point recursions are, with high probability, the same orders-of-magnitude as $h$, the step-size of physical time.
-The precise computational time is probabilistic, not just due to random sampling, but also due to randomness in power and memory footprint.}}
+The precise computational time is probabilistic, not just due to random sampling, but also due to randomness in latency, power and memory footprint of the hardware platform.}}
 
@@
 {\blue{In Fig \ref{fig:CompTimeIEEE14bus}, $h=10^{-3}$ and the computational time needed to perform a single proximal update, \emph{with high probability}, is $O(10^{-3})$ in the sense most time-steps incur $O(10^{-3})$ time.
-While instances when computational times are $O(10^{-2})$ do occur for Case I in Fig \ref{fig:CompTimeIEEE14bus}, but are rare.}}
+Few instances when computational times are $O(10^{-2})$ do occur for Case I in Fig \ref{fig:CompTimeIEEE14bus}, but are rare.}}
 

### 2022-08-15 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound2-TPWRS-01391-2021R1.tex
Rebuttal round 2

@@
 
-{\nib {\blue We. 
+{\nib {\blue We have addressed all the reviewers' comments as detailed itemized responses below.
+We are thankful to the AE and the reviewers for their helpful comments and efforts contributing to elevate the quality of our manuscript.
 }}
@@
 
-{\nib \blue{The 
+{\nib \blue{(1) We fully acknowledge the reviewer's comment about the simplicity of our SDE model.
+In the revised manuscript's Conclusion (last paragraph) and in the previous round's rebuttal document, we had explicitly mentioned about incorporating more complex models in our future work.
+As we explained in previous round, our contribution is nonparametric joint PDF propagation methodology and even for our simple model, presents significant technical challenge as we explain in Sec.
+III-A and III-B.
+We do not believe it is possible to design a proximal recursion for more complex models if one does not understand how to do so in simpler setting, as there is no ``one-size-fits-all" proximal algorithm in the manifold of PDFs for arbitrary nonlinear SDEs.
+In particular, it is impossible for us to design such an algorithm for more complex SDE models by simply revising our paper--this requires completely new line of research and is out-of-scope of the present work.
+
+(2) Our ground truth state-of-the-art needs to be non-parametric.
+We had explained that Monte Carlo remains the state-of-the-art ground truth since e.g., it is impossible to perform an FEM simulation on a general purpose computer solving the Kolmogorov's forward PDE even for the ``simpler" 5 generator model in Sec.
+V-A as it requires us to grid a 10 dimensional space and 1 dimensional time, which is what the transient joint PDFs are supported on.
+This is why ref.
+[45] did a 2D state + 1D time simulation, and the authors are not aware of any further attempts in the power systems literature on solving a higher dimensional Kolmogorov's forward PDE subject to the nonlinear power system dynamics.
 }}
@@
 
-{\nib \blue{\ul{The.}}
+{\nib \blue{The reviewer may have misunderstood our computational framework.
+We are considering $P,Q$ to be uncertain as in epistemic/parametric uncertainty, but not as in aleatoric uncertainty (i.e., not as in considering $P,Q$ to have their own SDEs).
+This is definitely a simplifying assumption, as the reviewer aptly points out and we find the reviewer's suggested reference very insightful.
+\ul{Please note that the suggested reference is published on IEEE Xplore on 23rd June, 2022, and we were not aware of this work as this manuscript has been under review since August, 2021}.
+Our future work on more sophisticated dynamics models will definitely use the suggested framework.
+We are indebted to the reviewer for this suggestion.}
 }
@@
 
-{\nib \blue{The.}}
+{\nib \blue{That is not exactly what Fig.
+8 says since it is plotting the ensemble mean and box-whisker plots but not the individual 1000 sample paths.
+\ul{In fact, the whiskers show significant spread of probability mass indicating large excursions made by the stochastic sample paths}.
+\ul{Please note that this Fig.
+8 was included by the authors in the first round of revision in response to the specific request made by the reviewer's comment \textbf{R2.5} in that round}.
+Furthermore, in the revised manuscript's p.
+8, right column, \ul{we specifically discuss how Fig.
+7(c) depict the post-fault marginals having larger spread or dispersion of the probability mass.
+There, even though the means stay around zero, the spreading of probability mass implies larger probability of having sample paths taking excursions away from zero, compared to the nominal Case I}.
+In the same paragraph, we also point out how similar information can be gleaned by comparing the fourth row in Fig.
+5 with the same in Fig.
+6.}}
 
@@
 
-{\nib \blue{Thanks for this comment. The}}
+{\nib \blue{The reviewer is spot on: the existing power system dynamics literature mostly do not compute the joint PDFs, and directly try to approximate statistics of derived quantities such as marginal PDFs or moments.
+The fact that our proposed methodology enables joint computation, is a feature not a bug.
+Here is why:
+
+(i) Even if one is only interested in univariate or bivariate marginals, their computation is more accurate if the joint is first computed and then the marginals are obtained from the joint PDF by numerical integration (following the definition of the corresponding marginal) as we have done in Figs.
+5, 6, 7.
+This is because the coupling in the nonlinear power system dynamics makes the states correlated in a non-trivial manner, and this correlation evolves over time, which can only be resolved accurately by directly computing the joint PDFs.
+Please note that joint PDFs are information-theoretically complete i.e., encode all the statistical information such as the statistical moments and marginals.
+In fact, unlike the joint PDF generator given by the forward Kolmogorov operator, one cannot derive a PDE for the marginal generator for general nonlinear It\^{o} SDEs.
+For application-oriented references focusing on the importance and usage of the joint PDF prediction subject to nonlinear state dynamics, we suggest the following references:
+
+S. Haddad, A. Halder, B. Singh, ``Density-based stochastic reachability computation for occupancy prediction in automated driving", \emph{IEEE Transactions on Control Systems Technology}, 2022.
+
+A.
+Halder, R.
+Bhattacharya, ``Dispersion analysis in hypersonic flight during planetary entry using stochastic Liouville equation", \emph{Journal of Guidance, Control, and Dynamics}, 34(2), 459-474, 2011.
+
+(ii) Related to the above: given the joint PDF, the marginals are unique.
+However, the converse is false, i.e., looking at the marginals one cannot reconstruct the complete statistics in the state space.
+In fact, there are infinitely many joints consistent with the given marginals.
+This means that every time some different statistics such as higher order moments or other integral functionals of the joint are desired, in the absence of the availability of transient joint PDFs, separate algorithms need to be designed in a case-by-case manner.
+
+(iii) The transient joint PDFs are also useful if one is interested in model (in)validation or controller verification in the probabilistic sense, see e.g.,
+
+A. Halder, R. Bhatacharya, ``Probabilistic model validation for uncertain nonlinear systems", \emph{Automatica}, 50(8), 2038--2050, 2014.
+
+A. Halder, K. Lee, R. Bhatacharya,``Optimal transport approach for probabilistic robustness analysis of F-16 controllers", \emph{Journal of Guidance, Control, and Dynamics}, 38(10), 1935--1946, 2015.
+
+(iv) If one is interested in nonlinear estimation of the states using noisy sensor data and known stochastic process dynamics, then the predicted transient joint state PDF serves as the prior on which the Bayesian update is to be made to compute the posterior.
+Directly propagating the prior joint PDFs can alleviate well-known ``particle degeneracy" issue suffered by  sequential Monte Carlo-based methods such as the particle filter; see e.g.,
+
+P.
+Dutta, A.
+Halder, R.
+Bhatacharya, ``Nonlinear estimation with Perron-Frobenius operator and Karhunen-Loeve expansion", \emph{IEEE Transactions on Aerospace and Electronic Systems}, 51(4), 3210--3225, 2015.
+
+}}
 
@@
 
-{\nib \blue{The }}
+{\nib \blue{\ul{The reason why [45] is the state-of-the-art is because that is the only reference to our knowledge, which attempted to directly and non-parametrically solve the Kolmogorov's forward PDE subject to the power system dynamics, albeit in even simpler setting (single-machine-infinite-bus, i.e., independent variables being two dimensional state and one dimensional time) than ours}.
+The technical topic of our interest is \emph{not} about what SDE/SDAE model is being used for the sample path dynamics, it is about how to solve the transient joint PDFs by directly solving the generator of the corresponding stochastic process while approximating neither the dynamics nor the statistics.
+In our \ul{previous round's rebuttal's response to comment \textbf{R3.1}}, we clearly explained that there is good reason why no progress were made on this specific technical issue--which is that finite element/volume or Monte Carlo based function approximation schemes cannot be scaled to the type of examples we consider (e.g., 100 dimensional state space in Sec.
+V-B).
+\ul{This is not a matter of doing research to find intelligent discretization schemes.
+Instead, it requires to  re-think what does it mathematically mean to solve the generator, and fundamentally break away from ``solve PDE as a PDE" philosophy.} This is what motivated our contribution.}}
 
 \noindent
-{\bf R3.5. Comments by reviewer}:\\
+{\bf R3.6. Comments by reviewer}:\\
 {\em Regarding Ref [23]F.
@@
 
-{\nib \blue{The }}
+{\nib \blue{We have no doubt that ref. [23] is important and well-known for modeling realistic stochastic power system dynamics. However, please notice that:
+
+(i) Our present work is not about modeling stochastic power system dynamics.
+
+(ii) Our citation of [23] in pg.
+2, left column, top paragraph, last two sentences, is not mere perfunctory.
+For the readers' understanding, we clarified that both refs.
+[22] and [23] simulate the sample paths and do not directly propagate the joint state PDFs.
+The latter is what this manuscript is about.
+
+(iii) We have no doubt that more sophisticated models as in [23] should be investigated in future for designing the proximal recursion algorithms such as the current manuscript.
+\ul{However, as we discuss in detail in Sec.
+III-A and Sec.
+III-B, deriving the proximal recursions is not a matter of simply swapping one SDE/SDAE model with another}.
+Exploiting the sub-Riemannian-like geometry induced by the base SDE/SDAE model on the ground (finite dimensional) manifold, requires a significant technical tour-de-force (see e.g., refs.
+[39, 48]) for the specific model under consideration unless the base manifold dynamics has a gradient or mixed conservative-dissipative drift with isotropic diffusion--\ul{none of which occurs in stochastic power system dynamics even for the simple model we considered}.
+The proximal updates are essentially trying to construct a ``steepest-descent-like" scheme on the manifold of probability measures (which is not even a vector space).
+Despite being a highly active research topic for the last two decades in the PDE and geometry community, there still is no systematic mathematical way to derive the distance and functional pair as in (27) to rigorously establish that the associated PDE is a steepest descent of which functional measured w.r.t.
+which distance.
+\ul{This situation is significantly different from the finite dimensional Euclidean or Bregman proximal algorithms as in refs.
+[32, 34]}.
+The current results in the PDF setting have focused on deriving proximal schemes in case-by-case basis.
+Therefore, we believe our contribution, though restricted to a simple model, is a significant first stepping stone necessary to consider more realistic models such as in [23], and should also be of interest to mathematicians outside the power systems community.
+
+}}
 
 \noindent
-{\bf R3.6. Comments by reviewer}:\\
+{\bf R3.7. Comments by reviewer}:\\
 {\em In general, I need to claim here that math is fine in itself, but only publishable when connected to or interpreted correctly in a convincing engineering application.}
 
-{\nib \blue{The }}
+{\nib \blue{Thank you. We have done our best to clarify these connections and interpretations. We hope you approve.}}
 
@@
 A direct comparison between the two cases is less meaningful due to difference in parameter values, initial joint PDF etc.
-\ul{However, the general trend that the proposed scattered point cloud-type proximal framework, similar to the one proposed here, can incur computational times which scale sublinearly with dimensions while resolving the joint PDF well, were reported in our earlier works in different application contexts: ref.
+\ul{However, the general trend that the proposed scattered point cloud-type proximal framework, similar to the one proposed here, can incur computational times which scale conveniently with dimensions while resolving the joint PDF well, were reported in our earlier works in different application contexts: ref.
 [41, Sec.
@@
 
-{\nib \blue{Thanks for ...}}
+{\nib \blue{We appreciate these helpful comments.
+
+(i) As per the reviewer's suggestion, in the current revised version, \ul{we have added two sentences (highlighted in blue) at the end of Sec. V explaining the trends in Fig. 12, 13 and 14}.
+
+(ii) \ul{In the revised version, we have updated Figs.
+12 and 13 to include the initial time instance $k=0$}.
+To be consistent with the computation of the errors for all $k$, we did the empirical versus ensemble computation in all the three Figs.
+12, 13, 14 at $k=0$ too, even though theoretically these curves are known to be zero at $k=0$.
+}}
 

### 2022-08-15 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev2.tex
Rebuttal round 2

@@
 The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update.
-For the proposed method to be practical, we therefore, need to demonstrate that {\blue{the computational timescale is at least statistically comparable (in the order of magnitude sense) to the physical dynamics timescale.
+For the proposed method to be practical, we therefore, need to demonstrate that {\blue{the computational timescale is at least \emph{statistically} comparable (in the order of magnitude sense) to the physical dynamics timescale.
 In other words, the computational times needed to complete the zero-order-hold fixed point recursions are, with high probability, the same orders-of-magnitude as $h$, the step-size of physical time.
@@
 
-{\blue{In Fig \ref{fig:CompTimeIEEE14bus}, $h=10^{-3}$ and the computational time needed to perform a single proximal update, \emph{with high probability}, is $O(10^{-3})$ in the sense most time-steps incur $O(10^{-3})$ time.
-Few instances when computational times are $O(10^{-2})$ do occur for Case I in Fig \ref{fig:CompTimeIEEE14bus}, but are rare.}}
+{\blue{In Fig \ref{fig:CompTimeIEEE14bus}, $h=10^{-3}$ and the computational time needed to perform a single proximal update, with high probability, is $O(10^{-3})$ in the sense most time-steps incur $O(10^{-3})$ time.
+For the Case I in Fig \ref{fig:CompTimeIEEE14bus}, few instances when computational times are $O(10^{-2})$ do occur but are statistically rare.}}
 
@@
 3]{peyre2019computational}) using the corresponding time-varying weighted point clouds.
+{\blue{The dynamics helps guide the concentration of the time-varying point clouds in regions of high and low likelihoods in the state space, and hence as time progresses, the empirical Monte Carlo approximants get closer to the proximal updates.
+Consequently, Figs.
+\ref{fig:RelErrSynthetic}, \ref{fig:RelErrCovSynthetic} and \ref{fig:WassJointPDFSynthetic} show that after initial transients, the distances between the Monte Carlo and the proximal statistics decrease with time.}}
 

### 2022-08-15 Papers/Stochastic_Uncertainty_Propagation_TPS/Plots-TPS-UncPropProx.py

@@ -126,39 +126,41 @@ Time_Synthetic = np.loadtxt("TimeSyntheticIEEE14bus.txt")
 # =============================================================================
 # Relative error in mean vector and covariance matrix: MC versus Prox plot
 # =============================================================================
-# RelErrMeanVectorMCvsProx_Synthetic = np.loadtxt("RelErrMeanVectorMCvsProxSythetic50Gen.txt")
+RelErrMeanVectorMCvsProx_Synthetic = np.loadtxt("RelErrMeanVectorMCvsProxSythetic50Gen.txt")
 
-# DistCovMatrixMCvsProx_Synthetic = np.loadtxt("/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/code/RelErrCovMCvsProxSythetic50Gen.txt")
+DistCovMatrixMCvsProx_Synthetic = np.loadtxt("/Users/abhishekhaldermac/NSF-1923278-Stochastic-Uncertainty-Management/code/RelErrCovMCvsProxSythetic50Gen.txt")
 
-# fig = plt.figure()
-# ax = fig.add_subplot(1, 1, 1)
+fig = plt.figure()
+ax = fig.add_subplot(1, 1, 1)
 
-# # ax.semilogy(Time_Synthetic[3:-1], RelErrMeanVectorMCvsProx_Synthetic[4:-1],'-', color='k', lw=1.5)
-# ax.semilogy(Time_Synthetic[3:-1], DistCovMatrixMCvsProx_Synthetic[3:-1],'-', color='k', lw=1.5)
+# DistCovMatrixMCvsProx_Synthetic[0] = 0
+
+#ax.semilogy(Time_Synthetic[0:-1], RelErrMeanVectorMCvsProx_Synthetic[1:-1],'-', color='k', lw=1.5)
+ax.semilogy(Time_Synthetic[0:-1], DistCovMatrixMCvsProx_Synthetic[0:-1],'-', color='k', lw=1.5)
 
 # # ax.set_ylim(3*10**-2, 2*10**-1)
 
-# ax.tick_params(direction='in',which='both')
+ax.tick_params(direction='in',which='both')
 # #  # axx.xaxis.tick_top()
 
-# ax.grid(True,which="both",ls="-", color='0.75')
+ax.grid(True,which="both",ls="-", color='0.75')
[…]

### 2022-08-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_arXiv.tex
adding arxiv version

--- 2022-08-12 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev2.tex
@@
 The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update.
-For the proposed method to be practical, we therefore, need to demonstrate that {\blue{the computational timescale is at least statistically comparable (in the order of magnitude sense) to the physical dynamics timescale.
+For the proposed method to be practical, we therefore, need to demonstrate that the computational timescale is at least \emph{statistically} comparable (in the order of magnitude sense) to the physical dynamics timescale.
 In other words, the computational times needed to complete the zero-order-hold fixed point recursions are, with high probability, the same orders-of-magnitude as $h$, the step-size of physical time.
-The precise computational time is probabilistic, not just due to random sampling, but also due to randomness in power and memory footprint.}}
+The precise computational time is probabilistic, not just due to random sampling, but also due to randomness in latency, power and memory footprint of the hardware platform.
 
@@
 
-{\blue{In Fig \ref{fig:CompTimeIEEE14bus}, $h=10^{-3}$ and the computational time needed to perform a single proximal update, \emph{with high probability}, is $O(10^{-3})$ in the sense most time-steps incur $O(10^{-3})$ time.
-While instances when computational times are $O(10^{-2})$ do occur for Case I in Fig \ref{fig:CompTimeIEEE14bus}, but are rare.}}
+In Fig \ref{fig:CompTimeIEEE14bus}, $h=10^{-3}$ and the computational time needed to perform a single proximal update, with high probability, is $O(10^{-3})$ in the sense most time-steps incur $O(10^{-3})$ time.
+For the Case I in Fig \ref{fig:CompTimeIEEE14bus}, few instances when computational times are $O(10^{-2})$ do occur but are statistically rare.
 
@@
 3]{peyre2019computational}) using the corresponding time-varying weighted point clouds.
+The dynamics helps guide the concentration of the time-varying point clouds in regions of high and low likelihoods in the state space, and hence as time progresses, the empirical Monte Carlo approximants get closer to the proximal updates.
+Consequently, Figs.
+\ref{fig:RelErrSynthetic}, \ref{fig:RelErrCovSynthetic} and \ref{fig:WassJointPDFSynthetic} show that after initial transients, the distances between the Monte Carlo and the proximal statistics decrease with time.
 

### 2022-10-24 Papers/Stochastic_Uncertainty_Propagation_TPS/FinalSubmissionSourceTPS/FINAL VERSION.tex
final submission files

--- 2022-08-24 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_arXiv.tex
@@
 
-%\begin{IEEEbiography}[{\includegraphics[height=1.25in]{Abhishek-Halder.jpg}}]{Abhishek
-%Halder}
-%(S'10-M'14-SM'21) is an Assistant Professor in the Department of Applied Mathematics,
-%and is an affiliated faculty in the Department of Electrical and Computer Engineering at
-%University of California, Santa Cruz. Before that he held postdoctoral
-%positions in the Department of Mechanical and Aerospace Engineering at
-%University of California, Irvine, and in the Department of Electrical
-%and Computer Engineering at Texas A\&M University. He obtained his
-%Bachelors and Masters from Indian Institute of Technology Kharagpur in
-%2008, and Ph.D. from Texas A\&M University in 2014, all in Aerospace
-%Engineering. His research interests are in stochastic systems, control, learning and optimization with application focus on large scale cyber-physical
-%systems.
-%\end{IEEEbiography}
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{XinboGeng.jpg}}]{Xinbo Geng} (S'13-M'20)
-%received his B.E.
-degree in Electrical Engineering from Tsinghua University, Beijing, China, in 2013.
-He received the M.S.
-and Ph.D.
-degrees from Department of Electrical and Computer Engineering, Texas A\&M University, College Station, TX, in 2015 and 2019, respectively.
-%He is a postdoctoral researcher jointly affiliated with Cornell University and Texas A\&M University since 2020.
-%He was a power system design and studies intern at National Renewable Energy Lab in 2016, a data analyst intern at ISO New England in 2017, a visiting student at UC Berkeley and MIT in 2018, and a visiting student at University of Wisconsin Madison in 2019.
-%His main research interest is decision making under uncertainties.
-%\end{IEEEbiography}
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\begin{IEEEbiography}[{\includegraphics[height=1.25in]{kenneth.jpg}}]{Kenneth F. Caluya}
-%is a Ph.D.
+\begin{IEEEbiography}[{\includegraphics[height=1.25in]{Abhishek-Halder.jpg}}]{Abhishek
+Halder}
+(S'10-M'14-SM'21) is an Assistant Professor in the Department of Applied Mathematics,
+and is an affiliated faculty in the Department of Electrical and Computer Engineering at
+University of California, Santa Cruz. Before that he held postdoctoral
+positions in the Department of Mechanical and Aerospace Engineering at
+University of California, Irvine, and in the Department of Electrical
+and Computer Engineering at Texas A\&M University. He obtained his
+Bachelors and Masters from Indian Institute of Technology Kharagpur in
+2008, and Ph.D. from Texas A\&M University in 2014, all in Aerospace
+Engineering. His research interests are in stochastic systems, control, learning and optimization with application focus on large scale cyber-physical
+systems.
+\end{IEEEbiography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{IEEEbiography}[{\includegraphics[height=1.25in]{kenneth.jpg}}]{Kenneth F. Caluya}
+is a Ph.D.
 student in the Department of Applied Mathematics at University of California, Santa Cruz.
@@
 His research interests lie at the intersection of probability, convex optimization, partial differential equations, and dynamical systems.
-%\end{IEEEbiography}
-%
+\end{IEEEbiography}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\begin{IEEEbiography}[{\includegraphics[height=1.25in]{parmita.png}}]{Pegah Ojaghi}
-%is a Ph.D.
+
+\begin{IEEEbiography}[{\includegraphics[height=1.25in]{parmita.png}}]{Pegah Ojaghi}
+is a Ph.D.
 student in the Department of Computer Science and Engineering at University of California, Santa Cruz.
@@
 in Control Engineering from the Hamedan University of Technology, Hamedan, Iran in 2012.
-%\end{IEEEbiography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{IEEEbiography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{XinboGeng.jpg}}]{Xinbo Geng} (S'13-M'20)
+received his B.E.
+degree in Electrical Engineering from Tsinghua University, Beijing, China, in 2013.
+He received the M.S.
+and Ph.D.
+degrees from Department of Electrical and Computer Engineering, Texas A\&M University, College Station, TX, in 2015 and 2019, respectively.
+He is a postdoctoral researcher jointly affiliated with Cornell University and Texas A\&M University since 2020.
+He was a power system design and studies intern at National Renewable Energy Lab in 2016, a data analyst intern at ISO New England in 2017, a visiting student at UC Berkeley and MIT in 2018, and a visiting student at University of Wisconsin Madison in 2019.
+His main research interest is decision making under uncertainties.
+\end{IEEEbiography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
